### 2019-09-11 hw1/results/Exploratory Data Analysis.Rmd
Initial commit of EDA

---
title: "Exploratory Data Analysis"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(dplyr)
library(ggplot2)
library(kableExtra)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r read_data}
training <- read.csv("../data/moneyball-training-data.csv")
```

### Summary Statistics

The following is a summary of the data.

```{r}
training %>% 
  select(-INDEX) %>%
  summary(.) %>%
  kable() %>%
  kable_styling()
```

### Variable Distributions

The following density plots show the spread of the data.  The red verticle line is the mean and the blue verticle line is the median.

```{r}
density_plot <- function(d, title, xlab, ylab){
  nas <- d %>% filter(is.na(x)) %>% nrow()
  p <- ggplot(d, aes(x, stat(count))) +
    geom_density() +
    labs(title = title, 
       x = xlab,
       y = ylab) +
    theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray95")) +
    geom_vline(xintercept = mean(d$x, na.rm = TRUE), color = "red") +
    geom_vline(xintercept = median(d$x, na.rm = TRUE), color = "blue")
  # Add annotation for NA's
  if(nas > 0){
    p <- p + ggtitle(title, subtitle = paste(nas, "Null Observations")) 
  }
  p
}
```

```{r}
training %>%
  mutate(x = TARGET_WINS) %>%
  density_plot(., "Wins", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_H) %>%
  density_plot(., "Base Hits", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_2B) %>%
  density_plot(., "Doubles", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_3B) %>%
  density_plot(., "Triples", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_HR) %>%
  density_plot(., "Homeruns", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_BB) %>%
  density_plot(., "Walks", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_HBP) %>%
  density_plot(., "Batters Hit by Pitch", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_SO) %>%
  density_plot(., "Strikeouts", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BASERUN_SB) %>%
  density_plot(., "Stolen Bases", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BASERUN_CS) %>%
  density_plot(., "Caught Stealing Bases", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_FIELDING_E) %>%
  density_plot(., "Errors", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_FIELDING_DP) %>%
  density_plot(., "Double Plays", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_BB) %>%
  density_plot(., "Walks Allowed", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_H) %>%
  density_plot(., "Hits Allowed", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_HR) %>%
  density_plot(., "Homeruns Allowed", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_SO) %>%
  density_plot(., "Strikouts by Pitcher", element_blank(), element_blank())
```

### 2019-09-11 hw1/results/Exploratory Data Analysis.Rmd
Removed cached files

@@
 library(kableExtra)
+library(gridExtra)
 
@@
 
+## Exploration
+
 ### Summary Statistics
@@
 
-### Variable Distributions
+### Plots
 
-The following density plots show the spread of the data.  The red verticle line is the mean and the blue verticle line is the median.
+The following density plots show the spread of the data.
+The red verticle line is the mean and the blue verticle line is the median.
+The scatter plot shows the relationship between wins and the variable
 
 ```{r}
-density_plot <- function(d, title, xlab, ylab){
+plot_me <- function(d, title, xlab, ylab){
   nas <- d %>% filter(is.na(x)) %>% nrow()
-  p <- ggplot(d, aes(x, stat(count))) +
+  density <- d %>%
+    filter(!is.na(x)) %>%
+    ggplot(., aes(x, stat(count))) +
     geom_density() +
@@
         panel.grid.major = element_line(color = "gray95")) +
-    geom_vline(xintercept = mean(d$x, na.rm = TRUE), color = "red") +
+    geom_vline(xintercept = mean(d$x, nna.rm = TRUE), color = "red") +
     geom_vline(xintercept = median(d$x, na.rm = TRUE), color = "blue")
@@
   if(nas > 0){
-    p <- p + ggtitle(title, subtitle = paste(nas, "Null Observations")) 
+    density <- density + ggtitle(title, subtitle = paste(nas, "Null Observations")) 
   }
-  p
+
+  # Create a scatter plot
+  scatter <- ggplot(d, aes(x, TARGET_WINS)) +
+    geom_point() + 
+    labs(x = xlab,
+         y = ylab) +
+    expand_limits(y=0) +
+    theme(panel.background = element_blank(),
+          panel.grid.major = element_line(color = "gray95"))
+  # Put them side by side
+  grid.arrange(density, scatter, ncol=2)
 }
@@
   mutate(x = TARGET_WINS) %>%
-  density_plot(., "Wins", element_blank(), element_blank())
+  plot_me(., "Wins", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_H) %>%
-  density_plot(., "Base Hits", element_blank(), element_blank())
+  plot_me(., "Base Hits", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_2B) %>%
-  density_plot(., "Doubles", element_blank(), element_blank())
+  plot_me(., "Doubles", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_3B) %>%
-  density_plot(., "Triples", element_blank(), element_blank())
+  plot_me(., "Triples", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_HR) %>%
-  density_plot(., "Homeruns", element_blank(), element_blank())
+  plot_me(., "Homeruns", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_BB) %>%
-  density_plot(., "Walks", element_blank(), element_blank())
+  plot_me(., "Walks", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_HBP) %>%
-  density_plot(., "Batters Hit by Pitch", element_blank(), element_blank())
+  plot_me(., "Batters Hit by Pitch", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_SO) %>%
-  density_plot(., "Strikeouts", element_blank(), element_blank())
+  plot_me(., "Strikeouts", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BASERUN_SB) %>%
-  density_plot(., "Stolen Bases", element_blank(), element_blank())
+  plot_me(., "Stolen Bases", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BASERUN_CS) %>%
-  density_plot(., "Caught Stealing Bases", element_blank(), element_blank())
+  plot_me(., "Caught Stealing Bases", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_FIELDING_E) %>%
-  density_plot(., "Errors", element_blank(), element_blank())
+  plot_me(., "Errors", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_FIELDING_DP) %>%
-  density_plot(., "Double Plays", element_blank(), element_blank())
+  plot_me(., "Double Plays", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_BB) %>%
-  density_plot(., "Walks Allowed", element_blank(), element_blank())
+  plot_me(., "Walks Allowed", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_H) %>%
-  density_plot(., "Hits Allowed", element_blank(), element_blank())
+  plot_me(., "Hits Allowed", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_HR) %>%
-  density_plot(., "Homeruns Allowed", element_blank(), element_blank())
+  plot_me(., "Homeruns Allowed", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_SO) %>%
-  density_plot(., "Strikouts by Pitcher", element_blank(), element_blank())
+  plot_me(., "Strikouts by Pitcher", element_blank(), element_blank())
 ```

### 2019-09-11 hw1/results/Exploratory Data Analysis.Rmd
More work on EDA

@@
 opts_chunk$set(echo=FALSE,
-	             cache=TRUE,
+	             cache=FALSE,
                prompt=FALSE,
@@
         panel.grid.major = element_line(color = "gray95")) +
-    geom_vline(xintercept = mean(d$x, nna.rm = TRUE), color = "red") +
+    geom_vline(xintercept = mean(d$x, na.rm = TRUE), color = "red") +
     geom_vline(xintercept = median(d$x, na.rm = TRUE), color = "blue")
@@
     geom_point() + 
+    geom_smooth(method = lm) + 
     labs(x = xlab,

### 2019-09-12 hw1/results/Exploratory Data Analysis.Rmd
More work on HW 1 EDA

@@
 title: "Exploratory Data Analysis"
+author: "Critical Thinking Group 3"
 date: "`r Sys.Date()`"

### 2019-09-14 hw1/results/Additional Exploratory Analysis.Rmd

---
title: "Additional Exploratory Data Analysis"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)
library(corrplot)
```

```{r, include=FALSE}
# Read in the training data
training <- read_csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw1/data/moneyball-training-data.csv")

# Ensure we have explicit NAs for missing data
training <- complete(training)

# Remove redundant "TEAM_" moniker
names(training) <- names(training) %>% str_remove("TEAM_")
```

## Correlation

Below is a visual showing every variable and their correlations within the training data set (including only complete observations).

```{r correlation plot}
train.cor <- training %>% select(-INDEX) %>% cor(., use = "complete.obs")
corrplot(train.cor, method="color",type="upper", tl.col="black", tl.cex = 0.7,
         diag=F)
```

```{r}
# Correlation with response variable specifically
data.frame(cor(training$TARGET_WINS, training, use="complete.obs")) %>%
  gather(key="Variable", value="Correlation",-INDEX) %>% select(-INDEX) %>%
  mutate(Correlation = cell_spec(round(Correlation,4),
                                 "html",
                                 color = ifelse(abs(Correlation) > .4,
                                                "red", "black"))) %>%
  kable(format = "html", escape=F) %>% 
  kable_styling(full_width = F, position = "left") %>%
  add_header_above(c("Correlation with 'TARGET_WINS'"=2)) %>%
  footnote(general = "Red highlight indicates correlation > 0.4 (pos or neg).")
```

We see some correlation between a few explanatory variables and our target variable:

`BATTING_H`,`BATTING_HR`,`BATTING_BB`,`PITCHING_H`,`PITCHING_HR`, and `PITCHING_BB`.

However there is some correlation between them that we may have to account for:

```{r smaller correlation plot}
train.cor2 <- training %>%
  select(BATTING_H,BATTING_HR,BATTING_BB,PITCHING_H,PITCHING_HR,PITCHING_BB) %>%
  cor(., use = "complete.obs")

corrplot(train.cor2, method="color",type="lower", tl.col="black", tl.cex = 0.7,
         diag=F)
```

## Missing Data
Of the `r comma(nrow(training))` observations in the training data set, only `r comma(sum(complete.cases(training)))` are complete (no missing values in any variable). So, some imputation may be required to ensure we have enough training data to work with.

Looking at which variables are missing:

```{r}
# Get variable names
vars <- names(training)[3:17]
missing <- data.frame(variable=NA,missingVals=0)

for(i in 1:length(vars)){
  missing <- rbind(missing,
                   data.frame(variable = paste(vars[i]),
                              missingVals = sum(is.na(training[,paste(vars[i])])))
                  )
}

missing %>% filter(!is.na(variable)) %>%
  select(`Variable Name` = variable, 
         `# Missing Values` = missingVals) %>%
  arrange(-`# Missing Values`) %>% kable()
```

We see that `BATTING_HBP`, `BASERUN_CS`, `FIELDING_DP`, and `BASERUN_SB` have the most missing values. None of those variables were well correlated with our outcome variable so maybe it is ok to leave them alone.

Next we have `BATTING_SO` and `PITCHING_SO`. These variables had correlations close to 0.22. So, if we plan on using those in our model, we may want to exclude those cases where they are missing (which leaves us approximately `r training %>% filter(!is.na(BATTING_SO),!is.na(PITCHING_SO)) %>% nrow %>%
comma()` complete observations).

### 2019-09-15 hw1/results/Exploratory Data Analysis.Rmd
More work on EDA

@@
 
-```{r}
+```{r plot_me}
 plot_me <- function(d, title, xlab, ylab){
@@
   mutate(x = TEAM_BATTING_H) %>%
-  plot_me(., "Base Hits", element_blank(), element_blank())
+  plot_me(., "Base Hits (TEAM_BATTING_H)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_2B) %>%
-  plot_me(., "Doubles", element_blank(), element_blank())
+  plot_me(., "Doubles (TEAM_BATTING_2B)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_3B) %>%
-  plot_me(., "Triples", element_blank(), element_blank())
+  plot_me(., "Triples (TEAM_BATTING_3B)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_HR) %>%
-  plot_me(., "Homeruns", element_blank(), element_blank())
+  plot_me(., "Homeruns (TEAM_BATTING_HR)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_BB) %>%
-  plot_me(., "Walks", element_blank(), element_blank())
+  plot_me(., "Walks (TEAM_BATTING_BB)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_HBP) %>%
-  plot_me(., "Batters Hit by Pitch", element_blank(), element_blank())
+  plot_me(., "Hit by Pitch (TEAM_BATTING_HBP)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BATTING_SO) %>%
-  plot_me(., "Strikeouts", element_blank(), element_blank())
+  plot_me(., "Strikeouts (TEAM_BATTING_SO)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BASERUN_SB) %>%
-  plot_me(., "Stolen Bases", element_blank(), element_blank())
+  plot_me(., "Stolen Bases (TEAM_BASERUN_SB)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_BASERUN_CS) %>%
-  plot_me(., "Caught Stealing Bases", element_blank(), element_blank())
+  plot_me(., "Caught Stealing Bases (TEAM_BASERUN_CS)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_FIELDING_E) %>%
-  plot_me(., "Errors", element_blank(), element_blank())
+  plot_me(., "Errors (TEAM_FIELDING_E)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_FIELDING_DP) %>%
-  plot_me(., "Double Plays", element_blank(), element_blank())
+  plot_me(., "Double Plays (TEAM_FIELDING_DP)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_BB) %>%
-  plot_me(., "Walks Allowed", element_blank(), element_blank())
+  plot_me(., "Walks Allowed (TEAM_PITCHING_BB)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_H) %>%
-  plot_me(., "Hits Allowed", element_blank(), element_blank())
+  plot_me(., "Hits Allowed (TEAM_PITCHING_H)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_HR) %>%
-  plot_me(., "Homeruns Allowed", element_blank(), element_blank())
+  plot_me(., "Homeruns Allowed (TEAM_PITCHING_HR)", element_blank(), element_blank())
 ```
@@
   mutate(x = TEAM_PITCHING_SO) %>%
-  plot_me(., "Strikouts by Pitcher", element_blank(), element_blank())
+  plot_me(., "Strikouts by Pitcher (TEAM_PITCHING_SO)", element_blank(), element_blank())
 ```
+
+## Missing Data
+
+### Batting Strike Outs
+
+To fill the missing in the missing data we will alternate between the two modes (578 and 909)
+
+```{r}
+## How did we know that? We looked at this table
+# D <- training[!is.na(training$TEAM_BATTING_SO),]$TEAM_BATTING_SO
+# data.frame(x = density(D)$x, y = density(D)$y) %>%
+#   arrange(desc(y)) %>%
+#   kable() %>%
+#   kable_styling()
+# # Modes at
+# # 578 and 909
+```
+
+```{r}
+set.seed(42)
+
+training <- training %>%
+  mutate(TEAM_BATTING_SO  = ifelse(is.na(TEAM_BATTING_SO), sample(c(578, 909), 1), TEAM_BATTING_SO))
+```
+
+## Scaled and Combined
+
+The idea behind this model is that teams that are better than the average will win more games and teams worse than the average will win less.
+The way we determine if a team is better than average is by looking at how well they preform at batting, pitching, and fielding.
+
+Since there are more than one way to win a baseball game (i.e.
+have some power sluggers that hit home runs, vs have really good single batters.) we need to combine the various batting measures.
+Now since getting a strikout at bat is bad, we need to change the sign of this variable.
+That way it can be combined and will fit the better teams win more and worse teams less model.
+
+We are going to scale all variables .
+That centers them at 0 and gives them a standard deviation of 1.
+We can then combine almost all the batting variables into one measure (hit by pitcher is excluded).
+
+```{r}
+training <- training %>%
+  mutate(TEAM_BATTING_SO = -1 * TEAM_BATTING_SO) %>%
+  mutate_at(scale, .vars = vars(-TARGET_WINS)) %>%
+  mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO)
+```
+
+```{r}
+training %>%
+  mutate(x = TEAM_BATTING) %>%
+  plot_me(., "Batting", element_blank(), element_blank())
+fit <- lm(TARGET_WINS ~ TEAM_BATTING, training)
+summary(fit)
+```
+
+This model says that the average baseball team will win about `r round(coef(fit)[1],0)` games.
+If their batting is one standard deviation better than the average they will win `r round(coef(fit)[2],0)` more games.

### 2019-09-16 hw1/results/Exploratory Data Analysis.Rmd
Updated EDA

@@
 
-To fill the missing in the missing data we will alternate between the two modes (578 and 909)
+To fill the missing in the `r nrow(training[is.na(training$TEAM_BATTING_SO),])` missing data we will alternate between the two modes (578 and 909)
 
@@
 training <- training %>%
+  mutate(TEAM_BATTING_SO_flag  = ifelse(is.na(TEAM_BATTING_SO), 1, 0)) %>%
   mutate(TEAM_BATTING_SO  = ifelse(is.na(TEAM_BATTING_SO), sample(c(578, 909), 1), TEAM_BATTING_SO))
+```
+
+### Pitching Strike Outs
+
+To fill the the `r nrow(training[is.na(training$TEAM_PITCHING_SO),])` missing values with the mean.
+
+```{r}
+training <- training %>%
+  mutate(TEAM_PITCHING_SO_flag = ifelse(is.na(TEAM_PITCHING_SO), 1, 0)) %>%
+  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_PITCHING_SO), mean(TEAM_PITCHING_SO), TEAM_PITCHING_SO))
+```
+
+### Double Plays
+
+To fill the the `r nrow(training[is.na(training$TEAM_FIELDING_DP),])` missing values with the mean.
+
+```{r}
+training <- training %>%
+  mutate(TEAM_FIELDING_DP_flag = ifelse(is.na(TEAM_FIELDING_DP), 1, 0)) %>%
+  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_FIELDING_DP), mean(TEAM_FIELDING_DP), TEAM_FIELDING_DP))
 ```
@@
   mutate(TEAM_BATTING_SO = -1 * TEAM_BATTING_SO) %>%
+  mutate(TEAM_PITCHING_SO = -1 * TEAM_PITCHING_SO) %>%
+  mutate(TEAM_PITCHING_H = -1 * TEAM_PITCHING_H) %>%
+  mutate(TEAM_FIELDING_E = -1 * TEAM_FIELDING_E) %>%
   mutate_at(scale, .vars = vars(-TARGET_WINS)) %>%
-  mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO)
+  #mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
+  mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
+  mutate(TEAM_PITCHING = TEAM_PITCHING_BB + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) %>%
+  mutate(TEAM_FIELDING = TEAM_FIELDING_DP + TEAM_FIELDING_E)
 ```
@@
   plot_me(., "Batting", element_blank(), element_blank())
-fit <- lm(TARGET_WINS ~ TEAM_BATTING, training)
+```
+
+```{r}
+training %>%
+  mutate(x = TEAM_PITCHING) %>%
+  plot_me(., "Pitching", element_blank(), element_blank())
+```
+
+```{r}
+training %>%
+  mutate(x = TEAM_FIELDING) %>%
+  plot_me(., "Fielding", element_blank(), element_blank())
+```
+
+```{r}
+fit <- lm(TARGET_WINS ~ TEAM_BATTING + TEAM_PITCHING + TEAM_FIELDING, training)
 summary(fit)
@@
 If their batting is one standard deviation better than the average they will win `r round(coef(fit)[2],0)` more games.
+They will win `r round(coef(fit)[3],0)` more games for if their pitching and `r round(coef(fit)[4],0)` if their fielding is better than average.

### 2019-09-20 MoneyBallR_RMD.Rmd
EDA and preliminary feature engineering

---
title: "MoneyBallR_RMD"
author: "Corey Arnouts"
date: "9/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Decision Trees to Find Variable Interactions
```{r}
library(tidyverse)

df_train <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-training-data.csv")

df_test <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-evaluation-data.csv")

source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")

colnames(df_train) <- sub("TEAM_", "", colnames(df_train))

colnames(df_test) <- sub("TEAM_", "", colnames(df_test))

df_test <- RF_with_Nulls(df_train,df_test,"TARGET_WINS",.5,10,25,.02,5,1)

```

# See if we can impute some misssing values Using a Random Forest
```{r}
library(randomForest)
df_train.na <- df_train
df_train.imputed <-rfImpute(TARGET_WINS ~.,df_train.na)

str(df_train)
str(df_train.imputed)

df_train <- df_train.imputed
```

#Correlation Matrix
```{r}

na_count <-sapply(df_train, function(y) sum(length(which(is.na(y)))))

na_count <- data.frame(na_count)

library(corrplot)

M <- cor(df_train)
corrplot(M, method = "circle") #plot matrix

library("PerformanceAnalytics")

chart.Correlation(df_train, histogram=TRUE, pch=19)
```

#Principal Components
```{r}
#PCA Analysis in R DataCamp Article
library(tidyverse)

mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE) #exclude the categorical variables

mtcars <- mtcars

summary(mtcars.pca)

str(mtcars.pca)

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(mtcars.pca) # this shows you what attributes contribute to principal component 1 and 2

ggbiplot(mtcars.pca, labels=rownames(mtcars)) # add in the sample information

mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),rep("US",3), "Europe", rep("Japan", 3), rep("US",4), rep("Europe", 3), "US", rep("Europe", 3))

ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country) #represent circles around groupings I want to do this binary classifiers

ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components

################################################################################
##################### Same thing but with MoneyBallR data ######################
################################################################################

df_train.pca <- prcomp(df_train[,c(3:17)], center = TRUE,scale. = TRUE) #exclude the categorical variables

df_train.target <-

summary(df_train.pca)

str(df_train.pca)

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)

ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components

library(tidyverse)

test <- df_train %>% dplyr::select(BATTING_H)

df_train.pca <- df_train %>% dplyr::select(PITCHING_HR,BATTING_HR) %>%  prcomp(center = TRUE,scale. = TRUE) #exclude the categorical variables

summary(df_train.pca)

str(df_train.pca)

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)

ggplot(df_train, aes(x=PITCHING_HR, y=BATTING_HR, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles

df_train$HR_RATIO <- df_train$BATTING_HR/df_train$PITCHING_HR

library(plotly)
ggplotly(ggplot(df_train, aes(x=PITCHING_HR, y=HR_RATIO, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles
)

mid <- mean(df_train$TARGET_WINS)

ggplotly(ggplot(df_train, aes(x=BATTING_H, y=PITCHING_H, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" )     # Use hollow circles
)

## look at batting hits to pitching hits ratio
df_train$Hitting_Ratio <- df_train$BATTING_H/df_train$PITCHING_H

ggplotly(ggplot(df_train, aes(x=BATTING_H, y=Hitting_Ratio, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" )     # Use hollow circles
)

for (i in 1:ncol(df_train)){
ggplotly(ggplot(df_train, aes(x=df_train[[i]], y=TARGET_WINS, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(low = "blue",
                            high = "red", space = "Lab" )     # Use hollow circles
)

}

ggplotly(ggplot(df_train, aes(x=BATTING_2B , y=BATTING_H, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" ) )

```

#Predict Hits Batted by Hits Pitched and then see what the difference is so like hits batted above expectation
And then use this as a predictive feature

```{r}
#First Fit Target Wins just by Batted_H and Pitched_H

#model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = df_train)

#df_train$Prediction <- predict(model,df_train)

#summary(model)
## .45 correlation 
##try on it on out of sample

### Add my meta features ###
model <- lm(BATTING_H ~ PITCHING_H, data = df_train)

df_train$BATTING_H_EXP <- predict(model,df_train)
df_train$BATTING_H_ABV_EXP <- df_train$BATTING_H/df_train$BATTING_H_EXP

train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]

model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = train_data)

hold_out_data$Prediction <- predict(model,hold_out_data)

cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)

summary(model)

############# FITTING THE MODEL WITH EVERYTHING ###################
############# BASELINE ############################################
df_train$INDEX <- NULL

train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]

model <- lm(TARGET_WINS ~ ., data = train_data)

hold_out_data$Prediction <- predict(model,hold_out_data)

cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)

summary(model)

ggplotly(ggplot(hold_out_data, aes(x=Prediction, y=TARGET_WINS, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(low = "blue",
                            high = "red", space = "Lab" )     # Use hollow circles
)

toselect.x <- summary(model)$coeff[-1,4] < 0.05 #

relevant.x <- names(toselect.x)[toselect.x == TRUE]

sig.formula <- as.formula(paste("TARGET_WINS ~",paste(relevant.x, collapse= "+")))

model <- lm(formula = sig.formula, data = train_data)

summary(model)

```

##Look at Variance Inflation Factors

### 2019-09-20 hw1/results/MoneyBallR_RMD.Rmd
Add files via upload

--- 2019-09-20 MoneyBallR_RMD.Rmd
@@
 ```{r}
-#PCA Analysis in R DataCamp Article
-library(tidyverse)
-
-mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE) #exclude the categorical variables
-
-mtcars <- mtcars
-
-summary(mtcars.pca)
-
-str(mtcars.pca)
-
-library(devtools)
-install_github("vqv/ggbiplot")
-
-library(ggbiplot)
-
-ggbiplot(mtcars.pca) # this shows you what attributes contribute to principal component 1 and 2
-
-ggbiplot(mtcars.pca, labels=rownames(mtcars)) # add in the sample information
-
-mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),rep("US",3), "Europe", rep("Japan", 3), rep("US",4), rep("Europe", 3), "US", rep("Europe", 3))
-
-ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country) #represent circles around groupings I want to do this binary classifiers
-
-ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components
 

### 2019-09-21 hw1/results/Cleanse Data.Rmd
Added data cleanse
Cleansing data RMD added

---
title: "Additional Exploratory Data Analysis"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)
library(DMwR)

set.seed(621)
```

## Data Cleansing
```{r}
# Read in the training data
training <- read_csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw1/data/moneyball-training-data.csv")

# Create a clean dataset to make changes to
cleaned <- training
```

### Missing and Invalid Values

Looking at our training data set, here are the variables with 0 or NA values that we need to look at:

```{r}
# Find columns with 0's and NAs
mins <- data.frame(sapply(training,FUN = min)) %>% rownames_to_column()
colnames(mins) <- c("variable","min")

mins %>% filter(is.na(min) | min==0) %>% kable()
```

#### TARGET_WINS

```{r}
training %>% ggplot(aes(x=TARGET_WINS)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TARGET_WINS)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TARGET_WINS)),col="green",lty=2) +
  labs(title="Distribution of Wins",
       caption="* Red line is the mean value and green is the median")
```

The range of data looks good except for a single zero value. Since there are no recorded seasons with zero wins in MLB, we should impute a value in it's place. Here the mean seems as good as any value, so we will replace it.

```{r}
# Replace 0 with the mean value
cleaned[which(cleaned$TARGET_WINS==0),"TARGET_WINS"] <- mean(cleaned$TARGET_WINS)
```

#### TEAM_BATTING_3B

This field represents triples hit by the team. Triples aren't very common because the ball is still in the field of play (unlike a homerun) but the batter still has enough time to get 3 bases.

```{r}
training %>% ggplot(aes(x=TEAM_BATTING_3B)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BATTING_3B)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BATTING_3B)),col="green",lty=2) +
  labs(title="Distribution of Triples",
       caption="* Red line is the mean value and green is the median")
```

Looking at the distribution, the value of zero doesn't look too unusual. Even if it were, the value is not likely to have a large impact.

#### TEAM_BATTING_HR

Although homeruns are more common in modern baseball (thank you PDEs!), there are some low values in the data. So zero doesn't seem too unusual here either.

```{r}
training %>% ggplot(aes(x=TEAM_BATTING_HR)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BATTING_HR)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BATTING_HR)),col="green",lty=2) +
  labs(title="Distribution of Homeruns",
       caption="* Red line is the mean value and green is the median")
```

#### TEAM_BATTING_BB

This variable represents when the batter is "walked" by the pitcher (also known as Base on Balls):

```{r}
training %>% ggplot(aes(x=TEAM_BATTING_BB)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BATTING_BB)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BATTING_BB)),col="green",lty=2) +
  labs(title="Distribution of Walks (Base on Balls)",
       caption="* Red line is the mean value and green is the median")
```

Four balls will walk a batter in modern baseball, however that wasn't always the case. A century or more ago (within the date range of this data set) walks took as many as 9 balls to happen[^1]. Knowing this, and looking at the left-tail of the values above, it is not unreasonable that there might be a season with no walks. Like triples above, leaving the one zero data point in is unlikely to adversely impact any regression, since there are valid values nearby.

#### TEAM_BATTING_SO

Here we saw some NA values, `r training %>% filter(is.na(TEAM_BATTING_SO)) %>% nrow(.)` of them to be specific. Plus we have some zero values as well.

```{r}
training %>% ggplot(aes(x=TEAM_BATTING_SO)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BATTING_SO,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BATTING_SO,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Strikeouts (Batter)",
       caption="* Red line is the mean value and green is the median")
```

First, the zero values seem nigh-impossible. An entire season (100+ games) without a single batter being struck out seems highly suspect, let alone 20 of them in the dataset.

We can replace these values with imputed values, but the distribution looks to be bimodal, so using a mean or median (which is squarely between those peaks) may cause some issues with the model. So, instead, we will impute values using regression.

We can impute a value for this variable by looking at it's nearest neighbors (based on other variables) and taking a weighted average of their values.

```{r}
# Change 0's to NA so they too can be imputed
cleaned$TEAM_BATTING_SO <- 
  ifelse(cleaned$TEAM_BATTING_SO == 0,NA,cleaned$TEAM_BATTING_SO)

temp <- cleaned %>% select(-TARGET_WINS, -INDEX) %>% data.frame() %>%
  knnImputation()

cleaned$TEAM_BATTING_SO <- temp$TEAM_BATTING_SO
```

Looking at our distribution, the shape hasn't changed radically and the means and medians ony shifted slightly:

```{r}
cleaned %>% ggplot(aes(x=TEAM_BATTING_SO)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BATTING_SO,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BATTING_SO,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Strikeouts (Batter)",
       subtitle="After Imputation",
       caption="* Red line is the mean value and green is the median")
```

#### TEAM_BASERUN_SB

With this variable, we have a good number of NA values, and 2 zeroes:

```{r}
training %>% ggplot(aes(x=TEAM_BASERUN_SB)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BASERUN_SB,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BASERUN_SB,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Stolen Bases",
       caption="* Red line is the mean value and green is the median")
```

The zeroes may be legitimate here so we will leave them alone. For the NAs, we can use the same KNN imputation we used above for strikeouts

```{r}
cleaned$TEAM_BASERUN_SB <- temp$TEAM_BASERUN_SB

cleaned %>% ggplot(aes(x=TEAM_BASERUN_SB)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BASERUN_SB,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BASERUN_SB,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Stolen Bases",
       subtitle="After Imputation",
       caption="* Red line is the mean value and green is the median")
```

Again we see little structural change in our distribution.

#### TEAM_BASERUN_CS

This variable is NA for nearly a third of records and only 2 zero values (which could be legitimate values):

```{r}
training %>% ggplot(aes(x=TEAM_BASERUN_CS)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BASERUN_CS,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BASERUN_CS,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Caught Stealing Bases",
       caption="* Red line is the mean value and green is the median")
```

Despite the high number of missing values (and a potential for increased error), we will use the KNN imputed values.

```{r}
cleaned$TEAM_BASERUN_CS <- temp$TEAM_BASERUN_CS
```

#### TEAM_BATTING_HBP

With this variable, we see nearly all entries are missing:

```{r}
training %>% ggplot(aes(x=TEAM_BATTING_HBP)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_BATTING_HBP,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_BATTING_HBP,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Being Hit By a Pitch",
       caption="* Red line is the mean value and green is the median")
```

We *could* make an assumption that these are all cases where there were no batters hit by a pitch, but that seems ill-advised given the distribution above. In this case, we might be wise to leave them as NA for now.

#### TEAM_PITCHING_HR

This variable has no NA values, but there are a few zero values. However, the zero values seem to be legitimate given the distribution:

```{r}
training %>% ggplot(aes(x=TEAM_PITCHING_HR)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_PITCHING_HR,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_PITCHING_HR,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Homeruns Pitched",
       caption="* Red line is the mean value and green is the median")
```

#### TEAM_PITCHING_BB

Here we have no NA values and a single zero:

```{r}
training %>% ggplot(aes(x=TEAM_PITCHING_BB)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_PITCHING_BB,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_PITCHING_BB,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Walks Pitched",
       caption="* Red line is the mean value and green is the median")
```

As we did with walks above, we can assume that is is possible to have no walks (and therefore pitch no walks). So, we will leave the zero alone.

However, there are some **really** high values in the data, which strains reality a little. We could take anything defined as an outlier ($1.5 \cdot \text{IQR}$) and set it to NA so those records will be excluded from any model we build with this variable. But, when you do the math it seems extreme, but plausible. For example, the most number of games in a season in MLB is 162 (currently). With a max value or 3,645 walks pitched you get 22.5 walks per game on average. Divided equally amongst 9 innings, it comes out to 2.5 walks per inning.

I'd be surprised that any pitcher wouldn't be removed after an inning or two of 2-3 walks, but neither can we rule it out as a possibility.

#### TEAM_PITCHING_SO

This variable represents strikeouts pitched. We see that there are 102 NA values and a *lot* of extremely high values:

```{r}
training %>% ggplot(aes(x=TEAM_PITCHING_SO)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_PITCHING_SO,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_PITCHING_SO,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Strikeouts Pitched",
       caption="* Red line is the mean value and green is the median")
```

We can use the KNN imputation to help with the NA values

```{r}
cleaned$TEAM_PITCHING_SO <- temp$TEAM_PITCHING_SO
```

With the NA's fixed, we can turn our attention now to the extreme values.

The extreme values need to be handled. First, a typical game will be 9 innings in length, and in each inning you can only pitch 3 strikeouts (because then your part of the inning is over). Those 27 potential strikeouts multiplied by 162 games means an upper limit near 4,374 a season.

Games can go beyond 9 innings, but even if every game in a season was as long as the longest ever MLB game (26 innings) you can only have 12,636 strikeouts. So, the max value of `r max(training$TEAM_PITCHING_SO,na.rm=T)` is invalid.

We'll make a high-yet-reasonable assumption of a mean 11 innings per game, and call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be disregarded in any modeling.

```{r}
cleaned[which(cleaned$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
```

```{r}
cleaned %>% ggplot(aes(x=TEAM_PITCHING_SO)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_PITCHING_SO,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_PITCHING_SO,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Strikeouts Pitched",
       subtitle="After Removing Invalid Data and Imputation",
       caption="* Red line is the mean value and green is the median")
```

#### TEAM_FIELDING_DP

The values in this variable seem reasonable, however we do have some NA values.

```{r}
training %>% ggplot(aes(x=TEAM_FIELDING_DP)) + geom_histogram(bins = 50) +
  geom_vline(aes(xintercept=mean(TEAM_FIELDING_DP,na.rm = T)),col="red",lty=2) +
  geom_vline(aes(xintercept=median(TEAM_FIELDING_DP,na.rm=T)),col="green",lty=2) +
  labs(title="Distribution of Double Plays",
       caption="* Red line is the mean value and green is the median")
```

Again, we use the KNN imputation from earlier to fill in NAs with imputed values.

```{r}
cleaned$TEAM_FIELDING_DP <- temp$TEAM_FIELDING_DP
```

## Linear Models

### Simple Models

#### Model 1

Now we can begin to put together a linear model and see how well it fits.

We will start with a few obvious variables: `TEAM_BATTING_HR`, `TEAM_PITCHING_HR`, `TEAM_BATTING_H`, and `TEAM_PITCHING_H`:

```{r}
model1 <- lm(TARGET_WINS ~ TEAM_BATTING_HR + TEAM_PITCHING_HR +
               TEAM_BATTING_H + TEAM_PITCHING_H, data=cleaned)
model1.summary <- summary(model1)
model1.summary
```

Interestingly, only the "hit" variables seem to be statistically significant in this model. The adjusted $R^2$ value is only `r round(model1.summary$adj.r.squared,3)`.

#### Model 2

If we remove the "HR" variables, we have a decrease in adjusted $R^2$:

```{r}
model2 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_PITCHING_H, data=cleaned)
model2.summary <- summary(model2)
model2.summary
```

[^1]: https://en.wikipedia.org/wiki/Base_on_balls

### 2019-09-24 hw1/doc/HW1.Rmd
Initial commit of HW1.Rmd

---
title: "DATA 621 Homework #1"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
```

## Introduction

```{r read_data, echo=FALSE}
# Read in the training data
training <- read.csv("../data/moneyball-training-data.csv") %>%
  select(-INDEX) # Dropping meaningless index
# Read in the evaluation data
evaluation <- read.csv("../data/moneyball-evaluation-data.csv")
```

We have been given a dataset with `r nrow(training)` records summarizing a major league baseball team's season. The records span 1871 to 2006 inclusive.  All statistics have been adjusted to match the performance of a 162 game season.  The objective is to build a linear regression model to predict the number of wins for a team.

### Working Theory

We are working on the premise that there are "good" teams and there are "bad" teams.  The good teams win more than the bad teams.  We are assuming that some of the predictors will be higher for the good teams than for the bad teams.  Consequently we can use these variables to predict how many times a team will win in a season.

### Notes About the Data

There are some difficulties with this dataset.  First it covers such a wide time period.  We know there are different "eras" of baseball.  This data will span multiple eras.  Has the fundamental relationships between wining and these predictors change over time?  We think it has.  If so this will be a challenge.

## Data Exploration

### First Look at the Data

We will first look at the data to get a sense of what we have.

```{r small_multiples_density, warning=FALSE}
training %>%
  gather(variable, value, TARGET_WINS:TEAM_FIELDING_DP) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

```{r}
quick_summary <- function(df){
  df %>%
    summary() %>%
    kable() %>%
    kable_styling()
}

quick_summary(training)
```

Some initial observations:

* The response variable (`TARGET_WINS`) looks to be normally distributed.  This supports the working theory that there are good teams and bad teams.  There are also a lot of average teams.
* There are also quite a few variables with missing values.  We may need to deal with these in order to have the largest data set possible for modeling.
* A couple variables are bimodal (`TEAM_BATTING_HR`, `TEAM_BATTING_SO` `TEAM_PITCHING_HR`).  This may be a challenge as some of them are missing values and that may be a challenge in filling in missing values.
* Some variables are right skewed (`TEAM_BASERUN_CS`, `TEAM_BASERUN_SB`, etc.).  This might support the good team theory.  It may also introduce non-normally distributed residuals in the model.  We shall see.

### Correlations

Let's take a look at the correlations.  The following is the correlations from the complete cases only:

```{r correlation plot}
training %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

```{r}
temp <- training %>% 
  cor(., use = "complete.obs") #%>%

temp[lower.tri(temp, diag=TRUE)] <- ""
temp <- temp %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  gather(Variable, Correlation, -rowname) %>%
  filter(Variable != rowname) %>%
  filter(Correlation != "") %>%
  mutate(Correlation = as.numeric(Correlation)) %>%
  rename(` Variable` = rowname) %>%
  arrange(desc(abs(Correlation))) 
```

#### Correlations with Response Variable

Let's take a look at how the predictors are correlated with the response variable:

```{r warning=FALSE}
training %>%
  gather(variable, value, -TARGET_WINS) %>%
  ggplot(., aes(value, TARGET_WINS)) + 
  geom_point(fill = "indianred4", color="indianred4") + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = "Wins")
```

```{r}
temp %>%
  filter(` Variable` == "TARGET_WINS") %>%
  kable() %>%
  kable_styling()
```

It looks like the hits, walks, home runs, and errors have the strongest correlations with wins.  None of these correlations are particularly strong.  This suggests there is a lot of 'noise' in these relationships.

It is interesting to note allowing hits is positively correlated with wins. How strange! It is also noteworthy that pitching strikeouts is negatively correlated with winning.  That does not make any sense.  When one examines the scatter plots above it becomes apparent that these correlations are being effected by some outliers.

#### Strong Correlations (Absolute Value > 0.5)

Are any predictors are correlated with each other?  We will only look for "strong" correlations:

```{r}
temp %>%
  filter(abs(Correlation) > .5) %>%
  kable() %>%
  kable_styling()
```

There are `r temp %>% filter(Correlation > .99) %>% nrow(.)` variables that have a correlation that is almost 1!  We will need to be careful to prevent adding autocorrelation errors to our model.

### Missing Values

During our first look at the data it was noted that there were variables that are missing data.  Here's a look at what variables are missing data and how big of a problem is it:

```{r}
training %>% 
  gather(variable, value) %>%
  filter(is.na(value)) %>%
  group_by(variable) %>%
  tally() %>%
  mutate(percent = n / nrow(training) * 100) %>%
  mutate(percent = paste0(round(percent, ifelse(percent < 10, 1, 0)), "%")) %>%
  arrange(desc(n)) %>%
  rename(`Variable Missing Data` = variable,
         `Number of Records` = n,
         `Share of Total` = percent) %>%
  kable() %>%
  kable_styling()
```

The hit by pitcher varriable is missing over 90% of it's data.  Based on the weak correlation with the response variable we will exclude it from consideration in our model.

Caught stealling a base (`TEAM_BASERUN_CS`) is next on the list.  It may be possible to predict it using `TEAM_BASERUN_SB` since they are strongly correlated, but there are `r training %>% filter(is.na(TEAM_BASERUN_SB) & is.na(TEAM_BASERUN_SB)) %>% nrow()` times they both are missing data.

The strike outs are going to be a little tricky because of their bimodal distribution.  All `r training %>% filter(is.na(TEAM_BATTING_SO) & is.na(TEAM_PITCHING_SO)) %>% nrow()`

### Zero Values

There are also variables that have verly low values.  Let's see how big of a problem this is:

```{r}
training %>% 
  gather(variable, value) %>%
  filter(value == 0) %>%
  group_by(variable) %>%
  tally() %>%
  mutate(percent = n / nrow(training) * 100) %>%
  mutate(percent = paste0(round(percent, ifelse(percent < 10, 1, 0)), "%")) %>%
  arrange(desc(n)) %>%
  rename(`Variable With Zeros` = variable,
         `Number of Records` = n,
         `Share of Total` = percent) %>%
  kable() %>%
  kable_styling()
```

This isn't nearly as large of a problem as the missing values.  Now it's time to fix the data issues.

## Data Preparation

### Fixing Missing/Zero Values

#### TARGET_WINS

```{r}
training %>% 
  ggplot(aes(TARGET_WINS)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TARGET_WINS)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TARGET_WINS)), col = "green", lty = 2) +
  labs(x = element_blank(), 
       y = "Count",
       title = "Distribution of Wins",
       caption = "* Red line is the mean value and green is the median")
```

The range of data looks good except for a single zero value. Since there are no recorded seasons with zero wins in MLB, we should impute a value in it's place. Here the mean (`r round(mean(training$TARGET_WINS), 0)`) seems as good as any value, so we will replace it.

```{r}
# Replace 0 with the mean value
training[which(training$TARGET_WINS == 0), "TARGET_WINS"] <- round(mean(training$TARGET_WINS), 0)
```

#### TEAM_BATTING_3B

This field represents triples hit by the team. Triples aren't very common because the ball is still in the field of play (unlike a homerun) but the batter still has enough time to get 3 bases.

```{r}
training %>% 
  ggplot(aes(TEAM_BATTING_3B)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BATTING_3B)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BATTING_3B)), col = "green", lty = 2) +
  labs(x = element_blank(), 
       y = "Count",
       title = "Distribution of Triples",
       caption = "* Red line is the mean value and green is the median")
```

Looking at the distribution, the value of zero doesn't look too unusual. Even if it were, the value is not likely to have a large impact.

#### TEAM_BATTING_HR

Although homeruns are more common in modern baseball (thank you PDEs!), there are some low values in the data. So zero doesn't seem too unusual here either.

```{r}
training %>% 
  ggplot(aes(TEAM_BATTING_HR)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BATTING_HR)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BATTING_HR)), col = "green", lty = 2) +
  labs(x = element_blank(), 
       y = "Count",
       title = "Distribution of Homeruns",
       caption = "* Red line is the mean value and green is the median")
```

#### TEAM_BATTING_BB

This variable represents when the batter is "walked" by the pitcher (also known as Base on Balls):

```{r}
training %>% 
  ggplot(aes(TEAM_BATTING_BB)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BATTING_BB)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BATTING_BB)), col = "green", lty = 2) +
  labs(x = element_blank(), 
       y = "Count",
       title = "Distribution of Walks (Base on Balls)",
       caption = "* Red line is the mean value and green is the median")
```

Four balls will walk a batter in modern baseball, however that wasn't always the case. A century or more ago (within the date range of this data set) walks took as many as 9 balls to happen[^1]. Knowing this, and looking at the left-tail of the values above, it is not unreasonable that there might be a season with no walks. Like triples above, leaving the one zero data point in is unlikely to adversely impact any regression, since there are valid values nearby.

#### TEAM_BATTING_SO

Here we saw some NA values, `r training %>% filter(is.na(TEAM_BATTING_SO)) %>% nrow(.)` of them to be specific. Plus we have `r training %>% filter(TEAM_BATTING_SO == 0) %>% nrow(.)` zero values as well.

```{r}
training %>% 
  ggplot(aes(TEAM_BATTING_SO)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BATTING_SO,na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BATTING_SO, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(), 
       y = "Count",
       title = "Distribution of Strikeouts (Batter)",
       caption = "* Red line is the mean value and green is the median")
```

First, the zero values seem nigh-impossible. An entire season (162 games) without a single batter being struck out seems highly suspect, let alone 20 of them in the dataset.

We can replace these values with imputed values, but the distribution looks to be bimodal, so using a mean or median (which is squarely between those peaks) may cause some issues with the model. So, instead, we will impute values using regression.

We can impute a value for this variable by looking at it's nearest neighbors (based on other variables) and taking a weighted average of their values.

```{r}
# Change 0's to NA so they too can be imputed
training <- training %>%
  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
```

Looking at our distribution, the shape hasn't changed radically and the means and medians ony shifted slightly:

```{r}
training %>% 
  ggplot(aes(TEAM_BATTING_SO)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BATTING_SO, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BATTING_SO, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Strikeouts (Batter)",
       subtitle = "After Imputation",
       caption = "* Red line is the mean value and green is the median")
```

#### TEAM_BASERUN_SB

With this variable, we have a good number of NA values, and 2 zeroes:

```{r}
training %>% 
  ggplot(aes(TEAM_BASERUN_SB)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BASERUN_SB, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BASERUN_SB, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Stolen Bases",
       caption = "* Red line is the mean value and green is the median")
```

The zeroes may be legitimate here so we will leave them alone. For the NAs, we can use the same KNN imputation we used above for strikeouts

```{r}
#cleaned$TEAM_BASERUN_SB <- temp$TEAM_BASERUN_SB

training %>% 
  ggplot(aes(TEAM_BASERUN_SB)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BASERUN_SB, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BASERUN_SB, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Stolen Bases",
       subtitle = "After Imputation",
       caption = "* Red line is the mean value and green is the median")
```

Again we see little structural change in our distribution.

#### TEAM_BASERUN_CS

This variable is NA for nearly a third of records and only 2 zero values (which could be legitimate values):

```{r}
training %>% 
  ggplot(aes(TEAM_BASERUN_CS)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BASERUN_CS, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BASERUN_CS, na.rm = T)),col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Caught Stealing Bases",
       caption = "* Red line is the mean value and green is the median")
```

Despite the high number of missing values (and a potential for increased error), we will use the KNN imputed values.

```{r}
#cleaned$TEAM_BASERUN_CS <- temp$TEAM_BASERUN_CS
```

#### TEAM_BATTING_HBP

With this variable, we see nearly all entries are missing:

```{r}
training %>% 
  ggplot(aes(TEAM_BATTING_HBP)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_BATTING_HBP, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BATTING_HBP, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Being Hit By a Pitch",
       caption = "* Red line is the mean value and green is the median")
```

We *could* make an assumption that these are all cases where there were no batters hit by a pitch, but that seems ill-advised given the distribution above. In this case, we might be wise to leave them as NA for now.

#### TEAM_PITCHING_HR

This variable has no NA values, but there are a few zero values. However, the zero values seem to be legitimate given the distribution:

```{r}
training %>% 
  ggplot(aes(TEAM_PITCHING_HR)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_PITCHING_HR, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_PITCHING_HR, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Homeruns Pitched",
       caption = "* Red line is the mean value and green is the median")
```

#### TEAM_PITCHING_BB

Here we have no NA values and a single zero:

```{r}
training %>% 
  ggplot(aes(TEAM_PITCHING_BB)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_PITCHING_BB, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_PITCHING_BB, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Walks Pitched",
       caption = "* Red line is the mean value and green is the median")
```

As we did with walks above, we can assume that is is possible to have no walks (and therefore pitch no walks). So, we will leave the zero alone.

However, there are some **really** high values in the data, which strains reality a little. We could take anything defined as an outlier ($1.5 \cdot \text{IQR}$) and set it to NA so those records will be excluded from any model we build with this variable. But, when you do the math it seems extreme, but plausible. For example, the most number of games in a season in MLB is 162 (currently). With a max value or 3,645 walks pitched you get 22.5 walks per game on average. Divided equally amongst 9 innings, it comes out to 2.5 walks per inning.

I'd be surprised that any pitcher wouldn't be removed after an inning or two of 2-3 walks, but neither can we rule it out as a possibility.

#### TEAM_PITCHING_SO

This variable represents strikeouts pitched. We see that there are 102 NA values and a *lot* of extremely high values:

```{r}
training %>% 
  ggplot(aes(TEAM_PITCHING_SO)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_PITCHING_SO, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_PITCHING_SO, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Strikeouts Pitched",
       caption = "* Red line is the mean value and green is the median")
```

We can use the KNN imputation to help with the NA values

```{r}
#cleaned$TEAM_PITCHING_SO <- temp$TEAM_PITCHING_SO
```

With the NA's fixed, we can turn our attention now to the extreme values.

The extreme values need to be handled. First, a typical game will be 9 innings in length, and in each inning you can only pitch 3 strikeouts (because then your part of the inning is over). Those 27 potential strikeouts multiplied by 162 games means an upper limit near 4,374 a season.

Games can go beyond 9 innings, but even if every game in a season was as long as the longest ever MLB game (26 innings) you can only have 12,636 strikeouts. So, the max value of `r max(training$TEAM_PITCHING_SO,na.rm=T)` is invalid.

We'll make a high-yet-reasonable assumption of a mean 11 innings per game, and call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be disregarded in any modeling.

```{r}
#cleaned[which(cleaned$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
```

```{r}
training %>% 
  ggplot(aes(TEAM_PITCHING_SO)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_PITCHING_SO, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_PITCHING_SO, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Strikeouts Pitched",
       subtitle = "After Removing Invalid Data and Imputation",
       caption = "* Red line is the mean value and green is the median")
```

#### TEAM_FIELDING_DP

The values in this variable seem reasonable, however we do have some NA values.

```{r}
training %>% 
  ggplot(aes(TEAM_FIELDING_DP)) + 
  geom_histogram(bins = 50) +
  geom_vline(aes(xintercept = mean(TEAM_FIELDING_DP, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_FIELDING_DP, na.rm = T)), col = "green", lty = 2) +
  labs(x = element_blank(),
       y = "Counts",
       title = "Distribution of Double Plays",
       caption = "* Red line is the mean value and green is the median")
```

Again, we use the KNN imputation from earlier to fill in NAs with imputed values.

```{r}
#cleaned$TEAM_FIELDING_DP <- temp$TEAM_FIELDING_DP
```

```{r}
#training  <- training %>%
#  na.omit()
  #select(-TEAM_BATTING_HBP)
# set.seed(1)
# fix_missing_vals <- training %>%
#   na.omit() %>%
#   preProcess(., "knnImpute")

# training_mean <- mean(training$TARGET_WINS)
# 
# training <- predict(fix_missing_vals, training)
# 
# training$TEAM_FIP <- (training$TEAM_PITCHING_HR * 13) + (3 * training$TEAM_PITCHING_BB) - (2 * training$TEAM_PITCHING_SO)
# 
# pp <- preProcess(training, method = c("BoxCox", "center", "scale"))
# training <- predict(pp, training)
```

```{r eval=FALSE, echo=FALSE}
library(rpart)
library(rpart.plot)
m1 <- rpart(TARGET_WINS ~ ., data = training, method = "anova")
rpart.plot(m1)
plotcp(m1)
```

```{r eval=FALSE, echo=FALSE}
# Specify 10-fold cross validation
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- training %>%
  na.omit() %>%
  train(
    TARGET_WINS ~ ., 
    data = .,
    method = "treebag",
    trControl = ctrl,
    importance = TRUE
  )

# assess results
bagged_cv

plot(varImp(bagged_cv))

imp <- varImp(bagged_cv)$importance %>% 
  data.frame() %>%
  rownames_to_column()
```

## Model Building

## Model Selection

```{r}
set.seed(42)
train_index <- createDataPartition(training$TARGET_WINS, p = .8, list = FALSE, times = 1)

moneyball_train <- training[train_index,]
moneyball_test <- training[-train_index,]
```

```{r}
fit <- lm(TARGET_WINS ~ ., moneyball_train)
summary(fit)
```

```{r}
moneyball_test$y_hat <- predict(fit, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(error = TARGET_WINS - y_hat) %>%
  mutate(squared_error = error ^ 2)

ggplot(moneyball_test, aes(error))+
  geom_histogram(bins = 50)

summary(moneyball_test$squared_error)
```

## Appendix

### 2019-09-24 hw1/doc/HW1.Rmd
Updated HW1.Rmd

@@
 library(caret)
+library(DMwR)
 ```
@@
   mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
+
+knn <- training %>% knnImputation()
+impute_me <- is.na(training$TEAM_BATTING_SO)
+training[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
 ```
@@
 ```{r}
-#cleaned$TEAM_BASERUN_SB <- temp$TEAM_BASERUN_SB
+impute_me <- is.na(training$TEAM_BASERUN_SB)
+training[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"]
 
@@
 ```{r}
-#cleaned$TEAM_BASERUN_CS <- temp$TEAM_BASERUN_CS
+impute_me <- is.na(training$TEAM_BASERUN_CS)
+training[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
 ```
@@
 ```{r}
-#cleaned$TEAM_PITCHING_SO <- temp$TEAM_PITCHING_SO
+impute_me <- is.na(training$TEAM_PITCHING_SO)
+training[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"] 
 ```
@@
 ```{r}
-#cleaned[which(cleaned$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
-```
+training[which(training$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
 
-```{r}
 training %>% 
@@
 ```{r}
-#cleaned$TEAM_FIELDING_DP <- temp$TEAM_FIELDING_DP
-```
-
-```{r}
-#training  <- training %>%
-#  na.omit()
-  #select(-TEAM_BATTING_HBP)
-# set.seed(1)
-# fix_missing_vals <- training %>%
-#   na.omit() %>%
-#   preProcess(., "knnImpute")
-
-# training_mean <- mean(training$TARGET_WINS)
-# 
-# training <- predict(fix_missing_vals, training)
-# 
-# training$TEAM_FIP <- (training$TEAM_PITCHING_HR * 13) + (3 * training$TEAM_PITCHING_BB) - (2 * training$TEAM_PITCHING_SO)
-# 
-# pp <- preProcess(training, method = c("BoxCox", "center", "scale"))
-# training <- predict(pp, training)
-```
-
-```{r eval=FALSE, echo=FALSE}
-library(rpart)
-library(rpart.plot)
-m1 <- rpart(TARGET_WINS ~ ., data = training, method = "anova")
-rpart.plot(m1)
-plotcp(m1)
-```
-
-```{r eval=FALSE, echo=FALSE}
-# Specify 10-fold cross validation
-ctrl <- trainControl(method = "cv",  number = 10)
-
-# CV bagged model
-bagged_cv <- training %>%
-  na.omit() %>%
-  train(
-    TARGET_WINS ~ ., 
-    data = .,
-    method = "treebag",
-    trControl = ctrl,
-    importance = TRUE
-  )
-
-# assess results
-bagged_cv
-
-plot(varImp(bagged_cv))
-
-imp <- varImp(bagged_cv)$importance %>% 
-  data.frame() %>%
-  rownames_to_column()
+impute_me <- is.na(training$TEAM_FIELDING_DP)
+training[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"] 
 ```
@@
 ## Model Building
-
-## Model Selection
 
@@
 ```
+
+## Model Selection
 
@@
 
-ggplot(moneyball_test, aes(error))+
+ggplot(moneyball_test, aes(error)) +
   geom_histogram(bins = 50)

### 2019-09-24 hw1/results/MoneyBallR_Stepwise_Model_Critical_Thinking_Group_3.Rmd
Add files via upload

---
title: "MoneyBallR_Modeling"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
  html_document:
    df_print: paged
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
```

# Decision Trees to Find Variable Interactions
```{r}
library(tidyverse)
df_train <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-training-data.csv")
df_test <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-evaluation-data.csv")
source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")
colnames(df_train) <- sub("TEAM_", "", colnames(df_train))
colnames(df_test) <- sub("TEAM_", "", colnames(df_test))
df_test <- RF_with_Nulls(df_train,df_test,"TARGET_WINS",.5,10,25,.02,5,1)
```

# See if we can impute some misssing values Using a Random Forest
```{r}
library(randomForest)
df_train.na <- df_train
df_train.imputed <-rfImpute(TARGET_WINS ~.,df_train.na)
str(df_train)
str(df_train.imputed)
df_train <- df_train.imputed
```

#Correlation Matrix
```{r}
na_count <-sapply(df_train, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
library(corrplot)
M <- cor(df_train)
corrplot(M, method = "circle") #plot matrix
library("PerformanceAnalytics")
chart.Correlation(df_train, histogram=TRUE, pch=19)
```

#Principal Components
```{r}
#PCA Analysis in R DataCamp Article
library(tidyverse)
mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE) #exclude the categorical variables
mtcars <- mtcars 
summary(mtcars.pca)
str(mtcars.pca)
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
ggbiplot(mtcars.pca) # this shows you what attributes contribute to principal component 1 and 2
ggbiplot(mtcars.pca, labels=rownames(mtcars)) # add in the sample information
mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),rep("US",3), "Europe", rep("Japan", 3), rep("US",4), rep("Europe", 3), "US", rep("Europe", 3))
ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country) #represent circles around groupings I want to do this binary classifiers
ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components
################################################################################
##################### Same thing but with MoneyBallR data ######################
################################################################################
df_train.pca <- prcomp(df_train[,c(3:17)], center = TRUE,scale. = TRUE) #exclude the categorical variables
df_train.target <- 
summary(df_train.pca)
str(df_train.pca)
ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)
ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components
library(tidyverse)
test <- df_train %>% dplyr::select(BATTING_H)
df_train.pca <- df_train %>% dplyr::select(PITCHING_HR,BATTING_HR) %>%  prcomp(center = TRUE,scale. = TRUE) #exclude the categorical variables
summary(df_train.pca)
str(df_train.pca)
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)
ggplot(df_train, aes(x=PITCHING_HR, y=BATTING_HR, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles
df_train$HR_RATIO <- df_train$BATTING_HR/df_train$PITCHING_HR
library(plotly)
ggplotly(ggplot(df_train, aes(x=PITCHING_HR, y=HR_RATIO, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles
)
mid <- mean(df_train$TARGET_WINS)
ggplotly(ggplot(df_train, aes(x=BATTING_H, y=PITCHING_H, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" )     # Use hollow circles
)
## look at batting hits to pitching hits ratio
df_train$Hitting_Ratio <- df_train$BATTING_H/df_train$PITCHING_H
ggplotly(ggplot(df_train, aes(x=Hitting_Ratio, y=TARGET_WINS, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" )     # Use hollow circles
)
for (i in 1:ncol(df_train)){
ggplotly(ggplot(df_train, aes(x=df_train[[i]], y=TARGET_WINS, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(low = "blue",
                            high = "red", space = "Lab" )     # Use hollow circles
)
}
ggplotly(ggplot(df_train, aes(x=BATTING_2B , y=BATTING_H, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" ) )
```

#Predict Hits Batted by Hits Pitched and then see what the difference is so like hits batted above expectation
And then use this as a predictive feature

```{r}
#First Fit Target Wins just by Batted_H and Pitched_H
#model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = df_train)
#df_train$Prediction <- predict(model,df_train)
#summary(model)
## .45 correlation 
##try on it on out of sample
### Add my meta features ###
df_train <- dplyr::filter(df_train,df_train$TARGET_WINS != 0) %>% filter(BATTING_H != 0) %>% filter(BATTING_3B != 0) %>% 
filter(BATTING_2B != 0)%>% filter(BATTING_HR != 0)%>% filter(BATTING_BB != 0)%>% filter(BATTING_SO != 0)
# ## Adding in some new features ##
# df_train$Pitch_HR_under_2 <- ifelse(df_train$PITCHING_HR <2,1,0)
# df_train$Errors_over_1215 <- ifelse(df_train$FIELDING_E > 1215,1,0)
# df_train$HR_RATIO_2 <- (df_train$HR_RATIO)^2
### Add in some new feature engineering 
df_train$BATTING_1B <- df_train$BATTING_H - df_train$BATTING_2B - df_train$BATTING_3B - df_train$BATTING_HR
df_train$BASES_ACQUIRED <- df_train$BATTING_1B + 2*df_train$BATTING_2B + 3*df_train$BATTING_3B + 4*df_train$BATTING_H + df_train$BATTING_BB
train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]
model <- lm(TARGET_WINS ~ ., data = train_data)
hold_out_data$Prediction <- predict(model,hold_out_data)
cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
summary(model)
## Non linear benchmark random forest
train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]
hold_out_data <- RF_with_Nulls(df_train,hold_out_data,"TARGET_WINS",.5,4,500,.005,5,10)
cor(hold_out_data$prediction_overall,hold_out_data$TARGET_WINS)
# 
# 
# ############# FITTING THE MODEL WITH EVERYTHING ###################
# ############# BASELINE ############################################
# df_train$INDEX <- NULL
# train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
# train_data<- df_train[ train.index1,]
# hold_out_data  <- df_train[-train.index1,]
# model <- lm(TARGET_WINS ~ ., data = train_data)
# hold_out_data$Prediction <- predict(model,hold_out_data)
# cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
# summary(model)
ggplotly(ggplot(hold_out_data, aes(x=Prediction, y=TARGET_WINS, colour = TARGET_WINS)) +
     geom_point(shape=1) + 
  scale_color_gradient2(low = "blue",
                             high = "red", space = "Lab" )     # Use hollow circles
 )
ggplot(data = hold_out_data, aes(x = Prediction, y = TARGET_WINS)) + 
  geom_point(color='blue') +
  geom_smooth(method = "lm", se = FALSE)
toselect.x <- summary(model)$coeff[-1,4] < 0.05 #
relevant.x <- names(toselect.x)[toselect.x == TRUE]

sig.formula <- as.formula(paste("TARGET_WINS ~",paste(relevant.x, collapse= "+")))

model <- lm(formula = sig.formula, data = train_data)
summary(model)
```

##Look at Variance Inflation Factors 
```{r}
df_train1 <- df_train
df_train1$BATTING_H <- NULL
df_train1$BATTING_SO<- NULL
df_train1$HR_RATIO <- NULL
df_train1$BASES_ACQUIRED <- NULL
df_train1$Hitting_Ratio <- NULL
df_train1$PITCHING_HR <- NULL
for (i in 3:ncol(df_train1)){
col <- noquote(paste(colnames(df_train1)[i],"~ ."))
model <- lm(col, data = df_train1)
r_squared <- summary(model)$adj.r.squared
VIF <- 1/(1-(r_squared))
print(colnames(df_train1)[i])
print(VIF)
#print(summary(model))
}
```

```{r}
train.index1 <- createDataPartition(df_train1$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]
model <- lm(TARGET_WINS ~ BATTING_BB + BATTING_SO + BASERUN_CS + PITCHING_HR + 
    PITCHING_BB + PITCHING_SO + FIELDING_E + FIELDING_DP + I(BATTING_2B^2) + 
    I(BATTING_3B^2) + I(BATTING_BB^2) + I(BATTING_SO^2) + I(BASERUN_SB^2) + 
    I(BASERUN_CS^2) + I(PITCHING_H^2) + I(PITCHING_BB^2) + I(PITCHING_SO^2) + 
    I(FIELDING_E^2) + I(FIELDING_DP^2) + I(BATTING_1B^2) + 
    I(BATTING_2B^3) + I(BATTING_3B^3) + I(BATTING_HR^3) + I(BATTING_SO^3) + 
    I(BASERUN_CS^3) + I(PITCHING_H^3) + I(PITCHING_BB^3) + I(FIELDING_E^3) + 
    I(FIELDING_DP^3) + I(BATTING_2B^4) + I(BATTING_3B^4) + I(BATTING_HR^4) + 
    I(BATTING_BB^4) + I(BATTING_SO^4) + I(BASERUN_CS^4) + I(PITCHING_H^4) + 
    I(PITCHING_BB^4) + I(FIELDING_E^4) + I(FIELDING_DP^4) + I(BATTING_1B^4), data = df_train)
hold_out_data$Prediction <- predict(model,hold_out_data)
cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
summary(model)
```

#Poisson GLMS
```{r}
df_train$INDEX <- NULL
train.index1 <- createDataPartition(df_train1$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]
# Poisson Regression
# where count is a count and
# x1-x3 are continuous predictors
fit <- glm(TARGET_WINS ~ ., data = train_data, family=poisson())
summary(fit)
hold_out_data$Prediction <- predict(fit,hold_out_data)
cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
```

#Stepwise Regression
```{r}
train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]

full_formula <- "TARGET_WINS ~ BATTING_2B + BATTING_3B + BATTING_HR + BATTING_BB + BATTING_SO + BASERUN_SB + BASERUN_CS + PITCHING_H + PITCHING_HR + PITCHING_BB + PITCHING_SO + FIELDING_E + FIELDING_DP + BATTING_1B + I(BATTING_2B^2) + I(BATTING_3B^2) + I(BATTING_HR^2) + I(BATTING_BB^2) + I(BATTING_SO^2) + I(BASERUN_SB^2) + I(BASERUN_CS^2) + I(PITCHING_H^2) + I(PITCHING_HR^2) + I(PITCHING_BB^2) + I(PITCHING_SO^2) + I(FIELDING_E^2) + I(FIELDING_DP^2) + I(BATTING_1B^2) + I(BATTING_2B^3) + I(BATTING_3B^3) + I(BATTING_HR^3) + I(BATTING_BB^3) + I(BATTING_SO^3) + I(BASERUN_SB^3) + I(BASERUN_CS^3) + I(PITCHING_H^3) + I(PITCHING_HR^3) + I(PITCHING_BB^3) + I(PITCHING_SO^3) + I(FIELDING_E^3) + I(FIELDING_DP^3) + I(BATTING_1B^3) + I(BATTING_2B^4) + I(BATTING_3B^4) + I(BATTING_HR^4) + I(BATTING_BB^4) + I(BATTING_SO^4) + I(BASERUN_SB^4) + I(BASERUN_CS^4) + I(PITCHING_H^4) + I(PITCHING_HR^4) + I(PITCHING_BB^4) + I(PITCHING_SO^4) + I(FIELDING_E^4) + I(FIELDING_DP^4) + I(BATTING_1B^4)"

full_model <- lm(full_formula, train_data)
step_back_model <- stepAIC(full_model, direction="backward", trsace = F)
poly.call <- summary(step_back_model)$call
final_model <- lm(poly.call[2], train_data)
summary(final_model)

hold_out_data$Prediction <- predict(final_model,hold_out_data)
cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)

```

### 2019-09-24 hw1/results/moneyballr_stepwise_model_corey.rmd
Initial commit

--- 2019-09-24 hw1/results/MoneyBallR_Stepwise_Model_Critical_Thinking_Group_3.Rmd
@@
 title: "MoneyBallR_Modeling"
-author: "Critical Thinking Group 3"
+author: "Corey Arnouts"
 date: "`r Sys.Date()`"
@@
     toc_depth: 3
+    code_folding: "hide"
   html_document:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = NA)
 ```
@@
 df_test <- RF_with_Nulls(df_train,df_test,"TARGET_WINS",.5,10,25,.02,5,1)
+## Mike Silva Added this line
+df_train$BATTING_1B <- df_train$BATTING_H - df_train$BATTING_2B - df_train$BATTING_3B - df_train$BATTING_HR
 ```
@@
 
-#Correlation Matrix
+# Correlation Matrix
 ```{r}
@@
 
-#Principal Components
+# Principal Components
 ```{r}
@@
 
-#Predict Hits Batted by Hits Pitched and then see what the difference is so like hits batted above expectation
+# Predict Hits Batted by Hits Pitched and then see what the difference is so like hits batted above expectation
 And then use this as a predictive feature
 
-```{r}
+```{r eval=FALSE}
 #First Fit Target Wins just by Batted_H and Pitched_H
-#model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = df_train)
-#df_train$Prediction <- predict(model,df_train)
-#summary(model)
+model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = df_train)
+df_train$Prediction <- predict(model,df_train)
+summary(model)
 ## .45 correlation 
@@
 # ## Adding in some new features ##
-# df_train$Pitch_HR_under_2 <- ifelse(df_train$PITCHING_HR <2,1,0)
-# df_train$Errors_over_1215 <- ifelse(df_train$FIELDING_E > 1215,1,0)
-# df_train$HR_RATIO_2 <- (df_train$HR_RATIO)^2
+ df_train$Pitch_HR_under_2 <- ifelse(df_train$PITCHING_HR <2,1,0)
+ df_train$Errors_over_1215 <- ifelse(df_train$FIELDING_E > 1215,1,0)
+ df_train$HR_RATIO_2 <- (df_train$HR_RATIO)^2
 ### Add in some new feature engineering 
@@
 
-##Look at Variance Inflation Factors 
+## Look at Variance Inflation Factors 
 ```{r}
@@
 
-#Poisson GLMS
+# Poisson GLMS
 ```{r}
@@
 
-#Stepwise Regression
+# Stepwise Regression
 ```{r}

### 2019-09-24 hw1/results/moneyballr_stepwise_model_corey.rmd
Added set.seed to make it reproducible

@@
 ```{r}
+set.seed(1234)
 library(devtools)
@@
 ```{r}
+set.seed(1234)
 train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)

### 2019-09-25 hw1/doc/HW1.Rmd

@@
 
-We *could* make an assumption that these are all cases where there were no batters hit by a pitch, but that seems ill-advised given the distribution above.
-In this case, we might be wise to leave them as NA for now.
+We will fill in the values with KNN
+
+```{r}
+impute_me <- is.na(training$TEAM_BATTING_HBP)
+training[impute_me,"TEAM_BATTING_HBP"] <- knn[impute_me,"TEAM_BATTING_HBP"] 
+```
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BATTING_HBP)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BATTING_HBP, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BATTING_HBP, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Being Hit By a Pitch",
+       subtitle = "After Imputation",
+       caption = "* Red line is the mean value and green is the median")
+```
 
@@
 
-We can use the KNN imputation to help with the NA values
-
-```{r}
-impute_me <- is.na(training$TEAM_PITCHING_SO)
-training[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"] 
-```
-
-With the NA's fixed, we can turn our attention now to the extreme values.
+We can use the KNN imputation to help with the NA values.  We can turn our attention now to the extreme values.
 
@@
 
-We'll make a high-yet-reasonable assumption of a mean 11 innings per game, and call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be disregarded in any modeling.
+We'll make a high-yet-reasonable assumption of a mean 11 innings per game, and call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be imputed prior to modeling.
 
@@
 training[which(training$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
+
+impute_me <- is.na(training$TEAM_PITCHING_SO)
+training[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
 
@@
 
-## Model Selection
+### Kitchen Sink Model
+
+We will begin with a "kitchen sink" model.
 
@@
 ```
+
+### Select Variables
+
+We will select a subset of variables.
+
+```{r}
+fit_final <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_HBP + TEAM_FIELDING_E, data = moneyball_train)
+summary(fit_final)
+```
+
+## Model Selection
+
+In order to select which model is the "best" we will test it against a test set.  We will examine the difference between the predicted and actual.
+
+### Kitchen Sink Selection
 
@@
 
-summary(moneyball_test$squared_error)
+summary(moneyball_test$error)
+```
+
+### Select Model
+
+```{r}
+moneyball_test$y_hat <- predict(fit_final, moneyball_test)
+moneyball_test <- moneyball_test %>%
+  mutate(error = TARGET_WINS - y_hat) %>%
+  mutate(squared_error = error ^ 2)
+
+ggplot(moneyball_test, aes(error)) +
+  geom_histogram(bins = 50)
+
+summary(moneyball_test$error)
 ```

### 2019-09-25 hw1/results/Exploratory Data Analysis.Rmd

@@
   mutate(TEAM_PITCHING_SO_flag = ifelse(is.na(TEAM_PITCHING_SO), 1, 0)) %>%
-  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_PITCHING_SO), mean(TEAM_PITCHING_SO), TEAM_PITCHING_SO))
+  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_PITCHING_SO), mean(TEAM_PITCHING_SO, na.rm = TRUE), TEAM_PITCHING_SO))
 ```
@@
 
-To fill the the `r nrow(training[is.na(training$TEAM_FIELDING_DP),])` missing values with the mean.
+We will fill the the `r nrow(training[is.na(training$TEAM_FIELDING_DP),])` missing values with the mean.
 
@@
   mutate(TEAM_FIELDING_DP_flag = ifelse(is.na(TEAM_FIELDING_DP), 1, 0)) %>%
-  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_FIELDING_DP), mean(TEAM_FIELDING_DP), TEAM_FIELDING_DP))
+  mutate(TEAM_FIELDING_DP = ifelse(is.na(TEAM_FIELDING_DP), mean(TEAM_FIELDING_DP, na.rm = TRUE), TEAM_FIELDING_DP))
+```
+
+### Stolen Bases
+
+Wee will fill the the `r nrow(training[is.na(training$TEAM_BASERUN_SB),])` missing values with the mean.
+
+```{r}
+training <- training %>%
+  mutate(TEAM_BASERUN_SB_flag = ifelse(is.na(TEAM_BASERUN_SB), 1, 0)) %>%
+  mutate(TEAM_BASERUN_SB = ifelse(is.na(TEAM_BASERUN_SB), mean(TEAM_BASERUN_SB, na.rm = TRUE), TEAM_BASERUN_SB))
+```
+
+### Caught Stealing Bases
+
+We will fill the the `r nrow(training[is.na(training$TEAM_BASERUN_CS),])` missing values with the mean.
+
+```{r}
+training <- training %>%
+  mutate(TEAM_BASERUN_CS_flag = ifelse(is.na(TEAM_BASERUN_CS), 1, 0)) %>%
+  mutate(TEAM_BASERUN_CS = ifelse(is.na(TEAM_BASERUN_CS), mean(TEAM_BASERUN_CS, na.rm = TRUE), TEAM_BASERUN_CS))
 ```
@@
   mutate(TEAM_BATTING_SO = -1 * TEAM_BATTING_SO) %>%
-  mutate(TEAM_PITCHING_SO = -1 * TEAM_PITCHING_SO) %>%
-  mutate(TEAM_PITCHING_H = -1 * TEAM_PITCHING_H) %>%
-  mutate(TEAM_FIELDING_E = -1 * TEAM_FIELDING_E) %>%
+  #mutate(TEAM_PITCHING_SO = -1 * TEAM_PITCHING_SO) %>%
+  #mutate(TEAM_PITCHING_H = -1 * TEAM_PITCHING_H) %>%
+  #mutate(TEAM_BASERUN_CS = -1 * TEAM_BASERUN_CS) %>%
+  #mutate(TEAM_FIELDING_E = -1 * TEAM_FIELDING_E) %>%
   mutate_at(scale, .vars = vars(-TARGET_WINS)) %>%
-  #mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
-  mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
+  #mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_SO, TEAM_BATTING_HR) %>%
+  mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
   mutate(TEAM_PITCHING = TEAM_PITCHING_BB + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) %>%
+  mutate(TEAM_BASERUN = TEAM_BASERUN_SB + TEAM_BASERUN_CS) %>%
   mutate(TEAM_FIELDING = TEAM_FIELDING_DP + TEAM_FIELDING_E)
@@
   plot_me(., "Batting", element_blank(), element_blank())
+```
+
+```{r}
+training %>%
+  mutate(x = TEAM_BASERUN) %>%
+  plot_me(., "Stealing Bases", element_blank(), element_blank())
 ```
@@
 ```{r}
-fit <- lm(TARGET_WINS ~ TEAM_BATTING + TEAM_PITCHING + TEAM_FIELDING, training)
+fit <- lm(TARGET_WINS ~ TEAM_BATTING, training)
 summary(fit)
@@
 If their batting is one standard deviation better than the average they will win `r round(coef(fit)[2],0)` more games.
-They will win `r round(coef(fit)[3],0)` more games for if their pitching and `r round(coef(fit)[4],0)` if their fielding is better than average.
+They will win `r round(coef(fit)[3],0)` more games for if their pitching is better than average.

### 2019-09-25 hw1/results/MoneyBallR_Stepwise_Model_Critical_Thinking_Group_3.Rmd

@@
 
-# Decision Trees to Find Variable Interactions
+#Load in the Data
 ```{r}
 library(tidyverse)
+library(tidyverse)
+library(kableExtra)
+library(corrplot)
+library(caret)
+library(DMwR)
 df_train <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-training-data.csv")
+df_train$INDEX <- NULL
 df_test <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-evaluation-data.csv")
+df_test$INDEX <- NULL
 source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")
@@
 colnames(df_test) <- sub("TEAM_", "", colnames(df_test))
-df_test <- RF_with_Nulls(df_train,df_test,"TARGET_WINS",.5,10,25,.02,5,1)
+
+```
+
+#Missing Variables 
+```{r}
+df_train %>% 
+  gather(variable, value) %>%
+  filter(is.na(value)) %>%
+  group_by(variable) %>%
+  tally() %>%
+  mutate(percent = n / nrow(df_train) * 100) %>%
+  mutate(percent = paste0(round(percent, ifelse(percent < 10, 1, 0)), "%")) %>%
+  arrange(desc(n)) %>%
+  rename(`Variable Missing Data` = variable,
+         `Number of Records` = n,
+         `Share of Total` = percent) %>%
+  kable() %>%
+  kable_styling()
 ```
@@
 ```{r}
-library(randomForest)
-df_train.na <- df_train
-df_train.imputed <-rfImpute(TARGET_WINS ~.,df_train.na)
-str(df_train)
-str(df_train.imputed)
-df_train <- df_train.imputed
+
+df_train[which(df_train$TARGET_WINS == 0), "TARGET_WINS"] <- round(mean(df_train$TARGET_WINS), 0)
+
+# Change 0's to NA so they too can be imputed
+df_train <- df_train %>%
+  mutate(BATTING_SO = ifelse(BATTING_SO == 0, NA, BATTING_SO))
+knn <- df_train %>% knnImputation()
+impute_me <- is.na(df_train$BATTING_SO)
+df_train[impute_me,"BATTING_SO"] <- knn[impute_me,"BATTING_SO"]
+
+impute_me <- is.na(df_train$BASERUN_SB)
+df_train[impute_me,"BASERUN_SB"] <- knn[impute_me,"BASERUN_SB"]
+
+impute_me <- is.na(df_train$BASERUN_CS)
+df_train[impute_me,"BASERUN_CS"] <- knn[impute_me,"BASERUN_CS"]
+
+impute_me <- is.na(df_train$BATTING_HBP)
+df_train[impute_me,"BATTING_HBP"] <- knn[impute_me,"BATTING_HBP"]
+
+df_train[which(df_train$PITCHING_SO > 5346),"PITCHING_SO"] <- NA
+impute_me <- is.na(df_train$PITCHING_SO)
+df_train[impute_me,"PITCHING_SO"] <- knn[impute_me,"PITCHING_SO"]
+
+impute_me <- is.na(df_train$FIELDING_DP)
+df_train[impute_me,"FIELDING_DP"] <- knn[impute_me,"FIELDING_DP"]
+
+# Change 0's to NA so they too can be imputed
+df_test <- df_test %>%
+  mutate(BATTING_SO = ifelse(BATTING_SO == 0, NA, BATTING_SO))
+knn <- df_test %>% knnImputation()
+impute_me <- is.na(df_test$BATTING_SO)
+df_test[impute_me,"BATTING_SO"] <- knn[impute_me,"BATTING_SO"]
+
+impute_me <- is.na(df_test$BASERUN_SB)
+df_test[impute_me,"BASERUN_SB"] <- knn[impute_me,"BASERUN_SB"]
+
+impute_me <- is.na(df_test$BASERUN_CS)
+df_test[impute_me,"BASERUN_CS"] <- knn[impute_me,"BASERUN_CS"]
+
+impute_me <- is.na(df_test$BATTING_HBP)
+df_test[impute_me,"BATTING_HBP"] <- knn[impute_me,"BATTING_HBP"]
+
+df_test[which(df_test$PITCHING_SO > 5346),"PITCHING_SO"] <- NA
+impute_me <- is.na(df_test$PITCHING_SO)
+df_test[impute_me,"PITCHING_SO"] <- knn[impute_me,"PITCHING_SO"]
+
+impute_me <- is.na(df_test$FIELDING_DP)
+df_test[impute_me,"FIELDING_DP"] <- knn[impute_me,"FIELDING_DP"]
+
 ```
@@
 
-#Principal Components
+#Feature Creation and Data Cleaning
+
 ```{r}
-#PCA Analysis in R DataCamp Article
-library(tidyverse)
-mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE) #exclude the categorical variables
-mtcars <- mtcars 
-summary(mtcars.pca)
-str(mtcars.pca)
-library(devtools)
-install_github("vqv/ggbiplot")
-library(ggbiplot)
-ggbiplot(mtcars.pca) # this shows you what attributes contribute to principal component 1 and 2
-ggbiplot(mtcars.pca, labels=rownames(mtcars)) # add in the sample information
-mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),rep("US",3), "Europe", rep("Japan", 3), rep("US",4), rep("Europe", 3), "US", rep("Europe", 3))
-ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country) #represent circles around groupings I want to do this binary classifiers
-ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components
-################################################################################
-##################### Same thing but with MoneyBallR data ######################
-################################################################################
-df_train.pca <- prcomp(df_train[,c(3:17)], center = TRUE,scale. = TRUE) #exclude the categorical variables
-df_train.target <- 
-summary(df_train.pca)
-str(df_train.pca)
-ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)
-ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components
-library(tidyverse)
-test <- df_train %>% dplyr::select(BATTING_H)
-df_train.pca <- df_train %>% dplyr::select(PITCHING_HR,BATTING_HR) %>%  prcomp(center = TRUE,scale. = TRUE) #exclude the categorical variables
-summary(df_train.pca)
-str(df_train.pca)
-library(devtools)
-install_github("vqv/ggbiplot")
-library(ggbiplot)
-ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)
-ggplot(df_train, aes(x=PITCHING_HR, y=BATTING_HR, colour = TARGET_WINS)) +
-    geom_point(shape=1) + 
-scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles
-df_train$HR_RATIO <- df_train$BATTING_HR/df_train$PITCHING_HR
-library(plotly)
-ggplotly(ggplot(df_train, aes(x=PITCHING_HR, y=HR_RATIO, colour = TARGET_WINS)) +
-    geom_point(shape=1) + 
-scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles
-)
-mid <- mean(df_train$TARGET_WINS)
-ggplotly(ggplot(df_train, aes(x=BATTING_H, y=PITCHING_H, colour = TARGET_WINS)) +
-    geom_point(shape=1) + 
- scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
-                            high = "red", space = "Lab" )     # Use hollow circles
-)
-## look at batting hits to pitching hits ratio
-df_train$Hitting_Ratio <- df_train$BATTING_H/df_train$PITCHING_H
-ggplotly(ggplot(df_train, aes(x=Hitting_Ratio, y=TARGET_WINS, colour = TARGET_WINS)) +
-    geom_point(shape=1) + 
- scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
-                            high = "red", space = "Lab" )     # Use hollow circles
-)
-for (i in 1:ncol(df_train)){
-ggplotly(ggplot(df_train, aes(x=df_train[[i]], y=TARGET_WINS, colour = TARGET_WINS)) +
-    geom_point(shape=1) + 
- scale_color_gradient2(low = "blue",
-                            high = "red", space = "Lab" )     # Use hollow circles
-)
-}
-ggplotly(ggplot(df_train, aes(x=BATTING_2B , y=BATTING_H, colour = TARGET_WINS)) +
-    geom_point(shape=1) + 
- scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
-                            high = "red", space = "Lab" ) )
+
+df_train$BATTING_1B <- df_train$BATTING_H - df_train$BATTING_2B - df_train$BATTING_3B - df_train$BATTING_HR
+df_test$BATTING_1B <- df_test$BATTING_H - df_test$BATTING_2B - df_test$BATTING_3B - df_test$BATTING_HR
+
 ```
 
-#Predict Hits Batted by Hits Pitched and then see what the difference is so like hits batted above expectation
-And then use this as a predictive feature
-
+#Look at Linear Model with all First Order Predictors
 ```{r}
-#First Fit Target Wins just by Batted_H and Pitched_H
-#model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = df_train)
-#df_train$Prediction <- predict(model,df_train)
-#summary(model)
-## .45 correlation 
-##try on it on out of sample
-### Add my meta features ###
-df_train <- dplyr::filter(df_train,df_train$TARGET_WINS != 0) %>% filter(BATTING_H != 0) %>% filter(BATTING_3B != 0) %>% 
-filter(BATTING_2B != 0)%>% filter(BATTING_HR != 0)%>% filter(BATTING_BB != 0)%>% filter(BATTING_SO != 0)
-# ## Adding in some new features ##
-# df_train$Pitch_HR_under_2 <- ifelse(df_train$PITCHING_HR <2,1,0)
-# df_train$Errors_over_1215 <- ifelse(df_train$FIELDING_E > 1215,1,0)
-# df_train$HR_RATIO_2 <- (df_train$HR_RATIO)^2
-### Add in some new feature engineering 
-df_train$BATTING_1B <- df_train$BATTING_H - df_train$BATTING_2B - df_train$BATTING_3B - df_train$BATTING_HR
-df_train$BASES_ACQUIRED <- df_train$BATTING_1B + 2*df_train$BATTING_2B + 3*df_train$BATTING_3B + 4*df_train$BATTING_H + df_train$BATTING_BB
 train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
@@
 hold_out_data$Prediction <- predict(model,hold_out_data)
-cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
+oos_cor <- paste(round(cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS),digits = 2))
 summary(model)
-## Non linear benchmark random forest
-train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
-train_data<- df_train[ train.index1,]
-hold_out_data  <- df_train[-train.index1,]
-hold_out_data <- RF_with_Nulls(df_train,hold_out_data,"TARGET_WINS",.5,4,500,.005,5,10)
-cor(hold_out_data$prediction_overall,hold_out_data$TARGET_WINS)
-# 
-# 
-# ############# FITTING THE MODEL WITH EVERYTHING ###################
-# ############# BASELINE ############################################
-# df_train$INDEX <- NULL
-# train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
-# train_data<- df_train[ train.index1,]
-# hold_out_data  <- df_train[-train.index1,]
-# model <- lm(TARGET_WINS ~ ., data = train_data)
-# hold_out_data$Prediction <- predict(model,hold_out_data)
-# cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
-# summary(model)
-ggplotly(ggplot(hold_out_data, aes(x=Prediction, y=TARGET_WINS, colour = TARGET_WINS)) +
-     geom_point(shape=1) + 
-  scale_color_gradient2(low = "blue",
-                             high = "red", space = "Lab" )     # Use hollow circles
- )
-ggplot(data = hold_out_data, aes(x = Prediction, y = TARGET_WINS)) + 
+
+first_order_linear <- ggplot(data = hold_out_data, aes(x = Prediction, y = TARGET_WINS)) + 
   geom_point(color='blue') +
-  geom_smooth(method = "lm", se = FALSE)
-toselect.x <- summary(model)$coeff[-1,4] < 0.05 #
-relevant.x <- names(toselect.x)[toselect.x == TRUE]
+  geom_smooth(method = "lm", se = FALSE)+ggtitle(paste("First Order Linear Model with Out of Sample Correlation of: ",oos_cor))
 
-sig.formula <- as.formula(paste("TARGET_WINS ~",paste(relevant.x, collapse= "+")))
+```
 
-model <- lm(formula = sig.formula, data = train_data)
-summary(model)
+#Look at some Decision Trees for Variable Interactions
+```{r}
+hold_out_data <- RF_with_Nulls(train_data,hold_out_data,"TARGET_WINS",1,10,5,.01,5,1)
 ```
@@
 df_train1 <- df_train
-df_train1$BATTING_H <- NULL
-df_train1$BATTING_SO<- NULL
-df_train1$HR_RATIO <- NULL
-df_train1$BASES_ACQUIRED <- NULL
-df_train1$Hitting_Ratio <- NULL
-df_train1$PITCHING_HR <- NULL
+# df_train1$BATTING_H <- NULL
+# df_train1$BATTING_SO<- NULL
+# df_train1$HR_RATIO <- NULL
+# df_train1$BASES_ACQUIRED <- NULL
+# df_train1$Hitting_Ratio <- NULL
+# df_train1$PITCHING_HR <- NULL
 for (i in 3:ncol(df_train1)){
@@
 
-```{r}
-train.index1 <- createDataPartition(df_train1$TARGET_WINS, p = .7, list = FALSE)
-train_data<- df_train[ train.index1,]
-hold_out_data  <- df_train[-train.index1,]
-model <- lm(TARGET_WINS ~ BATTING_BB + BATTING_SO + BASERUN_CS + PITCHING_HR + 
-    PITCHING_BB + PITCHING_SO + FIELDING_E + FIELDING_DP + I(BATTING_2B^2) + 
-    I(BATTING_3B^2) + I(BATTING_BB^2) + I(BATTING_SO^2) + I(BASERUN_SB^2) + 
-    I(BASERUN_CS^2) + I(PITCHING_H^2) + I(PITCHING_BB^2) + I(PITCHING_SO^2) + 
-    I(FIELDING_E^2) + I(FIELDING_DP^2) + I(BATTING_1B^2) + 
-    I(BATTING_2B^3) + I(BATTING_3B^3) + I(BATTING_HR^3) + I(BATTING_SO^3) + 
-    I(BASERUN_CS^3) + I(PITCHING_H^3) + I(PITCHING_BB^3) + I(FIELDING_E^3) + 
-    I(FIELDING_DP^3) + I(BATTING_2B^4) + I(BATTING_3B^4) + I(BATTING_HR^4) + 
-    I(BATTING_BB^4) + I(BATTING_SO^4) + I(BASERUN_CS^4) + I(PITCHING_H^4) + 
-    I(PITCHING_BB^4) + I(FIELDING_E^4) + I(FIELDING_DP^4) + I(BATTING_1B^4), data = df_train)
-hold_out_data$Prediction <- predict(model,hold_out_data)
-cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
-summary(model)
-```
-
-#Poisson GLMS
-```{r}
-df_train$INDEX <- NULL
-train.index1 <- createDataPartition(df_train1$TARGET_WINS, p = .7, list = FALSE)
-train_data<- df_train[ train.index1,]
-hold_out_data  <- df_train[-train.index1,]
-# Poisson Regression
-# where count is a count and
-# x1-x3 are continuous predictors
-fit <- glm(TARGET_WINS ~ ., data = train_data, family=poisson())
-summary(fit)
-hold_out_data$Prediction <- predict(fit,hold_out_data)
-cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
-```
-
 #Stepwise Regression
 ```{r}
+
 train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
@@
 full_model <- lm(full_formula, train_data)
-step_back_model <- stepAIC(full_model, direction="backward", trsace = F)
+step_back_model <- stepAIC(full_model, direction="backward", trace = F)
 poly.call <- summary(step_back_model)$call
@@
 
+```
+
+#Model Results on the Hold Out Dataset
+```{r}
+
 hold_out_data$Prediction <- predict(final_model,hold_out_data)
+
+hold_out_data <- hold_out_data %>% filter(Prediction < 182)%>% filter(Prediction > 0)
+
 cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)
 
+oos_cor <- paste(round(cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS),digits = 2))
+
+ggplotly(ggplot(hold_out_data, aes(x=Prediction, y=TARGET_WINS, colour = TARGET_WINS)) +
+     geom_point(shape=1) + 
+  scale_color_gradient2(low = "blue",
+                             high = "red", space = "Lab" ) +
+  geom_smooth(method = "lm", se = FALSE, colour = "orange")+ggtitle(paste("Polynomial Stepwise Model with Out of Sample Correlation of: ",oos_cor))    # Use hollow circles
+ )
+
+ggplotly(first_order_linear)
 ```

### 2019-09-25 hw1/doc/HW1.Rmd

--- 2019-09-24 hw1/doc/HW1.Rmd
@@
 
-### Missing Values
+### Strange Data Values
 
-During our first look at the data it was noted that there were variables that are missing data.  Here's a look at what variables are missing data and how big of a problem is it:
+#### Missing Values
+
+During our first look at the data it was noted that there were variables that are missing data.  Here's a look at what variables are missing data and how big of a problem it is:
 
@@
 
-The hit by pitcher varriable is missing over 90% of it's data.  Based on the weak correlation with the response variable we will exclude it from consideration in our model.
+The hit by pitcher varriable is missing over 90% of it's data.  We will exclude it from consideration in our model.
 
@@
 
-### Zero Values
+#### Zero Values
 
@@
 
-This isn't nearly as large of a problem as the missing values.  Now it's time to fix the data issues.
+This isn't nearly as large of a problem as the missing values.
 
-## Data Preparation
-
-### Fixing Missing/Zero Values
+### Deeper Dive into the Variables
 
@@
 
-The range of data looks good except for a single zero value.
-Since there are no recorded seasons with zero wins in MLB, we should impute a value in it's place.
-Here the mean (`r round(mean(training$TARGET_WINS), 0)`) seems as good as any value, so we will replace it.
-
-```{r}
-# Replace 0 with the mean value
-training[which(training$TARGET_WINS == 0), "TARGET_WINS"] <- round(mean(training$TARGET_WINS), 0)
-```
+The range of data looks good.  There is a single zero value, which is the Washington Nationals 1872 season.
 
@@
 
-We can replace these values with imputed values, but the distribution looks to be bimodal, so using a mean or median (which is squarely between those peaks) may cause some issues with the model.
+We will replace these values with imputed values, but the distribution looks to be bimodal, so using a mean or median (which is squarely between those peaks) may cause some issues with the model.
 So, instead, we will impute values using regression.
 
-We can impute a value for this variable by looking at it's nearest neighbors (based on other variables) and taking a weighted average of their values.
-
-```{r}
-# Change 0's to NA so they too can be imputed
-training <- training %>%
-  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
-
-knn <- training %>% knnImputation()
-impute_me <- is.na(training$TEAM_BATTING_SO)
-training[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
-```
-
-Looking at our distribution, the shape hasn't changed radically and the means and medians ony shifted slightly:
-
-```{r}
-training %>% 
-  ggplot(aes(TEAM_BATTING_SO)) + 
-  geom_histogram(bins = 50) +
-  geom_vline(aes(xintercept = mean(TEAM_BATTING_SO, na.rm = T)), col = "red", lty = 2) +
-  geom_vline(aes(xintercept = median(TEAM_BATTING_SO, na.rm = T)), col = "green", lty = 2) +
-  labs(x = element_blank(),
-       y = "Count",
-       title = "Distribution of Strikeouts (Batter)",
-       subtitle = "After Imputation",
-       caption = "* Red line is the mean value and green is the median")
-```
+We will impute values for this variable by looking at it's nearest neighbors (based on other variables) and taking a weighted average of their values.
 
@@
 
-The zeroes may be legitimate here so we will leave them alone. For the NAs, we can use the same KNN imputation we used above for strikeouts
-
-```{r}
-impute_me <- is.na(training$TEAM_BASERUN_SB)
-training[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"]
-
-training %>% 
-  ggplot(aes(TEAM_BASERUN_SB)) + 
-  geom_histogram(bins = 50) +
-  geom_vline(aes(xintercept = mean(TEAM_BASERUN_SB, na.rm = T)), col = "red", lty = 2) +
-  geom_vline(aes(xintercept = median(TEAM_BASERUN_SB, na.rm = T)), col = "green", lty = 2) +
-  labs(x = element_blank(),
-       y = "Count",
-       title = "Distribution of Stolen Bases",
-       subtitle = "After Imputation",
-       caption = "* Red line is the mean value and green is the median")
-```
-
-Again we see little structural change in our distribution.
+The zeroes may be legitimate here so we will leave them alone. For the NAs, we will use the same KNN imputation we used above for strikeouts to impute values.
 
@@
 
-```{r}
-impute_me <- is.na(training$TEAM_BASERUN_CS)
-training[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
-```
-
 #### TEAM_BATTING_HBP
@@
 
-We *could* make an assumption that these are all cases where there were no batters hit by a pitch, but that seems ill-advised given the distribution above.
-In this case, we might be wise to leave them as NA for now.
+We will drop this variable as we cannot responsibly impute values with such thin data.
 
@@
 
-We can use the KNN imputation to help with the NA values
-
-```{r}
-impute_me <- is.na(training$TEAM_PITCHING_SO)
-training[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"] 
-```
-
-With the NA's fixed, we can turn our attention now to the extreme values.
-
 The extreme values need to be handled.
@@
 
-We'll make a high-yet-reasonable assumption of a mean 11 innings per game, and call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be disregarded in any modeling.
-
-```{r}
-training[which(training$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
-
-training %>% 
-  ggplot(aes(TEAM_PITCHING_SO)) + 
-  geom_histogram(bins = 50) +
-  geom_vline(aes(xintercept = mean(TEAM_PITCHING_SO, na.rm = T)), col = "red", lty = 2) +
-  geom_vline(aes(xintercept = median(TEAM_PITCHING_SO, na.rm = T)), col = "green", lty = 2) +
-  labs(x = element_blank(),
-       y = "Count",
-       title = "Distribution of Strikeouts Pitched",
-       subtitle = "After Removing Invalid Data and Imputation",
-       caption = "* Red line is the mean value and green is the median")
-```
+We'll make a high-yet-reasonable assumption of a mean 11 innings per game.
+We will call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be imputed prior to modeling.
 
@@
 
-Again, we use the KNN imputation from earlier to fill in NAs with imputed values.
+Again, we will use the KNN imputation from earlier to fill in NAs with imputed values.
+
+## Data Preparation
+
+### Fixing Missing/Zero Values
+
+First we will remove the invalid data and prep it for imputation.  We will drop the hit by pitcher variable from the dataset.
 
 ```{r}
-impute_me <- is.na(training$TEAM_FIELDING_DP)
-training[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"] 
+remove_bad_values <- function(df){
+  # Change 0's to NA so they too can be imputed
+  df <- df %>%
+  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
+  # Remove the high pitching strikeout values
+  df[which(df$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
+  # Drop the hit by pitcher variable
+  df %>%
+    select(-TEAM_BATTING_HBP)
+}
+
+training <- remove_bad_values(training)
+evaluation <- remove_bad_values(evaluation)
+```
+
+```{r}
+set.seed(42)
+knn <- training %>% knnImputation()
+
+impute_values <- function(df, knn){
+  impute_me <- is.na(df$TEAM_BATTING_SO)
+  df[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
+  impute_me <- is.na(df$TEAM_BASERUN_SB)
+  df[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"] 
+  impute_me <- is.na(df$TEAM_BASERUN_CS)
+  df[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
+  impute_me <- is.na(df$TEAM_PITCHING_SO)
+  df[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
+  impute_me <- is.na(df$TEAM_FIELDING_DP)
+  df[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"]
+  return(df)
+}
+
+training <- impute_values(training, knn)
+```
+
+### Feature Engineering
+
+The batting singles is not included but we can back it out of the hits.  We will do this.
+
+```{r}
+add_features <- function(df){
+  df %>%
+    mutate(TEAM_BATTING_1B = TEAM_BATTING_H - TEAM_BATTING_2B - TEAM_BATTING_3B - TEAM_BATTING_HR)
+}
+
+training <- add_features(training)
+evaluation <- add_features(evaluation)
+```
+
+### Results
+
+Here's what the data look like after imputation and correction:
+
+```{r warning=FALSE}
+training %>%
+  gather(variable, value) %>%
+  ggplot(., aes(value)) + 
+  geom_density(fill = "indianred4", color="indianred4") + 
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+```{r}
+quick_summary <- function(df){
+  df %>%
+    summary() %>%
+    kable() %>%
+    kable_styling()
+}
+
+quick_summary(training)
 ```
@@
 
+We will divide the data into training and test sets using a 70/30 split.  We will build our models on the training set and evaluate it on the test set.
+
 ```{r}
 set.seed(42)
-train_index <- createDataPartition(training$TARGET_WINS, p = .8, list = FALSE, times = 1)
+train_index <- createDataPartition(training$TARGET_WINS, p = .7, list = FALSE, times = 1)
 
@@
 
+### Kitchen Sink Model
+
+We will begin with a "kitchen sink" model.
+
+```{r}
+kitchen_sink <- lm(TARGET_WINS ~ ., moneyball_train)
+summary(kitchen_sink)
+```
+
+It does a fairly good job predicting, but there are a lot of variables that are not statistically significant.
+
+### Simple Model
+
+Let's try to create a simplier model. We will pick variables that had high correlations and exclude the pitching variables which would introduce autocorrelation issues.
+
+```{r}
+simple_fit <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_FIELDING_E, data = moneyball_train)
+summary(simple_fit)
+```
+
+This model isn't as fitted to the data as well but the variables are statistically significant.
+
+### Higher Order Stepwise Regression
+
+For the third model we will use a stepwise regression method using a backwards elimination process.  We also introduce some higher order polynomial variables.
+
+```{r}
+full_formula <- "TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP + TEAM_BATTING_1B + I(TEAM_BATTING_2B^2) + I(TEAM_BATTING_3B^2) + I(TEAM_BATTING_HR^2) + I(TEAM_BATTING_BB^2) + I(TEAM_BATTING_SO^2) + I(TEAM_BASERUN_SB^2) + I(TEAM_BASERUN_CS^2) + I(TEAM_PITCHING_H^2) + I(TEAM_PITCHING_HR^2) + I(TEAM_PITCHING_BB^2) + I(TEAM_PITCHING_SO^2) + I(TEAM_FIELDING_E^2) + I(TEAM_FIELDING_DP^2) + I(TEAM_BATTING_1B^2) + I(TEAM_BATTING_2B^3) + I(TEAM_BATTING_3B^3) + I(TEAM_BATTING_HR^3) + I(TEAM_BATTING_BB^3) + I(TEAM_BATTING_SO^3) + I(TEAM_BASERUN_SB^3) + I(TEAM_BASERUN_CS^3) + I(TEAM_PITCHING_H^3) + I(TEAM_PITCHING_HR^3) + I(TEAM_PITCHING_BB^3) + I(TEAM_PITCHING_SO^3) + I(TEAM_FIELDING_E^3) + I(TEAM_FIELDING_DP^3) + I(TEAM_BATTING_1B^3) + I(TEAM_BATTING_2B^4) + I(TEAM_BATTING_3B^4) + I(TEAM_BATTING_HR^4) + I(TEAM_BATTING_BB^4) + I(TEAM_BATTING_SO^4) + I(TEAM_BASERUN_SB^4) + I(TEAM_BASERUN_CS^4) + I(TEAM_PITCHING_H^4) + I(TEAM_PITCHING_HR^4) + I(TEAM_PITCHING_BB^4) + I(TEAM_PITCHING_SO^4) + I(TEAM_FIELDING_E^4) + I(TEAM_FIELDING_DP^4) + I(TEAM_BATTING_1B^4)"
+
+full_model <- lm(full_formula, moneyball_train)
+step_back <- MASS::stepAIC(full_model, direction="backward", trace = F)
+poly_call <- summary(step_back)$call
+step_back <- lm(poly_call[2], moneyball_train)
+summary(step_back)
+```
+
+This model has the highest adjusted R-squared but has some variables that are statistically insignificant.
+
 ## Model Selection
 
+In order to select which model is the "best" we will test it against a test set.  We will examine the difference between the predicted and actual values.
+
+### Kitchen Sink Results
+
 ```{r}
-fit <- lm(TARGET_WINS ~ ., moneyball_train)
-summary(fit)
+moneyball_test$kitchen_sink <- predict(kitchen_sink, moneyball_test)
+moneyball_test <- moneyball_test %>%
+  mutate(kitchen_sink_error = TARGET_WINS - kitchen_sink)
+
+ggplot(moneyball_test, aes(kitchen_sink_error)) +
+  geom_histogram(bins = 50)
+
+summary(moneyball_test$kitchen_sink_error)
 ```
 
+### Stepwise Results
+
 ```{r}
-moneyball_test$y_hat <- predict(fit, moneyball_test)
+moneyball_test$step_back <- predict(step_back, moneyball_test)
 moneyball_test <- moneyball_test %>%
-  mutate(error = TARGET_WINS - y_hat) %>%
-  mutate(squared_error = error ^ 2)
+  mutate(step_back_error = TARGET_WINS - step_back)
 
-ggplot(moneyball_test, aes(error)) +
+moneyball_test %>%
+  filter(step_back_error > -100) %>%
+  ggplot(., aes(step_back_error)) +
+  geom_histogram(bins = 50) +
+  labs(caption = "Outlier removed")
+
+summary(moneyball_test$step_back_error)
+```
+
+This model did not preform very well on the out of sample data.
+
+### Simple Model
+
+```{r}
+moneyball_test$simple <- predict(simple_fit, moneyball_test)
+moneyball_test <- moneyball_test %>%
+  mutate(simple_error = TARGET_WINS - simple)
+
+ggplot(moneyball_test, aes(simple_error)) +
   geom_histogram(bins = 50)
 
-summary(moneyball_test$squared_error)
+summary(moneyball_test$simple_error)
 ```
 
-## Appendix
+[^1]: https://en.wikipedia.org/wiki/Base_on_balls

### 2019-09-25 hw1/doc/HW1.Rmd

@@
 # Read in the training data
-training <- read.csv("../data/moneyball-training-data.csv") %>%
+training <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-training-data.csv") %>%
   select(-INDEX) # Dropping meaningless index
 # Read in the evaluation data
-evaluation <- read.csv("../data/moneyball-evaluation-data.csv")
+evaluation <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-evaluation-data.csv")
 ```
@@
 }
-
 quick_summary(training)
@@
 }
-
 training <- remove_bad_values(training)
@@
 knn <- training %>% knnImputation()
-
 impute_values <- function(df, knn){
@@
 }
-
 training <- impute_values(training, knn)
@@
 }
-
 training <- add_features(training)
@@
 train_index <- createDataPartition(training$TARGET_WINS, p = .7, list = FALSE, times = 1)
-
 moneyball_train <- training[train_index,]
@@
 full_formula <- "TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP + TEAM_BATTING_1B + I(TEAM_BATTING_2B^2) + I(TEAM_BATTING_3B^2) + I(TEAM_BATTING_HR^2) + I(TEAM_BATTING_BB^2) + I(TEAM_BATTING_SO^2) + I(TEAM_BASERUN_SB^2) + I(TEAM_BASERUN_CS^2) + I(TEAM_PITCHING_H^2) + I(TEAM_PITCHING_HR^2) + I(TEAM_PITCHING_BB^2) + I(TEAM_PITCHING_SO^2) + I(TEAM_FIELDING_E^2) + I(TEAM_FIELDING_DP^2) + I(TEAM_BATTING_1B^2) + I(TEAM_BATTING_2B^3) + I(TEAM_BATTING_3B^3) + I(TEAM_BATTING_HR^3) + I(TEAM_BATTING_BB^3) + I(TEAM_BATTING_SO^3) + I(TEAM_BASERUN_SB^3) + I(TEAM_BASERUN_CS^3) + I(TEAM_PITCHING_H^3) + I(TEAM_PITCHING_HR^3) + I(TEAM_PITCHING_BB^3) + I(TEAM_PITCHING_SO^3) + I(TEAM_FIELDING_E^3) + I(TEAM_FIELDING_DP^3) + I(TEAM_BATTING_1B^3) + I(TEAM_BATTING_2B^4) + I(TEAM_BATTING_3B^4) + I(TEAM_BATTING_HR^4) + I(TEAM_BATTING_BB^4) + I(TEAM_BATTING_SO^4) + I(TEAM_BASERUN_SB^4) + I(TEAM_BASERUN_CS^4) + I(TEAM_PITCHING_H^4) + I(TEAM_PITCHING_HR^4) + I(TEAM_PITCHING_BB^4) + I(TEAM_PITCHING_SO^4) + I(TEAM_FIELDING_E^4) + I(TEAM_FIELDING_DP^4) + I(TEAM_BATTING_1B^4)"
-
 full_model <- lm(full_formula, moneyball_train)
@@
   mutate(kitchen_sink_error = TARGET_WINS - kitchen_sink)
-
 ggplot(moneyball_test, aes(kitchen_sink_error)) +
   geom_histogram(bins = 50)
-
 summary(moneyball_test$kitchen_sink_error)
@@
   mutate(step_back_error = TARGET_WINS - step_back)
-
 moneyball_test %>%
-  filter(step_back_error > -100) %>%
+  filter(step_back >0) %>%
+  filter(step_back<182) %>% 
   ggplot(., aes(step_back_error)) +
@@
   mutate(simple_error = TARGET_WINS - simple)
-
 ggplot(moneyball_test, aes(simple_error)) +
   geom_histogram(bins = 50)
-
 summary(moneyball_test$simple_error)
@@
 
+### Regression Ensemble 
+```{r}
+moneyball_test <- moneyball_test %>% mutate(ensemble_score = (kitchen_sink + step_back + simple)/3)
+moneyball_test <- moneyball_test %>%
+  mutate(ensemble_error = TARGET_WINS - ensemble_score)
+ggplot(moneyball_test, aes(ensemble_error)) +
+  geom_histogram(bins = 50)
+summary(moneyball_test$ensemble_error)
+
+```
+
 [^1]: https://en.wikipedia.org/wiki/Base_on_balls

### 2019-09-25 hw1/doc/HW1.Rmd
Update HW1.Rmd
Added RMSE annotations to OOS error graphs

@@
 ggplot(moneyball_test, aes(kitchen_sink_error)) +
-  geom_histogram(bins = 50)
+  geom_histogram(bins = 50) +
+  annotate("text",x=0,y=10,
+           label = paste("RMSE = ",
+                         round(sqrt(mean(moneyball_test$kitchen_sink_error^2)),2)
+                         ),
+           color="white"
+           )
 summary(moneyball_test$kitchen_sink_error)
@@
   geom_histogram(bins = 50) +
+  annotate("text",x=0,y=10,
+           label = paste("RMSE = ",
+                         round(sqrt(mean(moneyball_test$step_back_error^2)),2)
+                         ),
+           color="white"
+           )
   labs(caption = "Outlier removed")
@@
 ggplot(moneyball_test, aes(simple_error)) +
-  geom_histogram(bins = 50)
+  geom_histogram(bins = 50) +
+  annotate("text",x=0,y=10,
+           label = paste("RMSE = ",
+                         round(sqrt(mean(moneyball_test$simple_error^2)),2)
+                         ),
+           color="white"
+           )
 summary(moneyball_test$simple_error)
@@
 ggplot(moneyball_test, aes(ensemble_error)) +
-  geom_histogram(bins = 50)
+  geom_histogram(bins = 50) +
+  annotate("text",x=0,y=10,
+           label = paste("RMSE = ",
+                         round(sqrt(mean(moneyball_test$ensemble_error^2)),2)
+                         ),
+           color="white"
+           )
 summary(moneyball_test$ensemble_error)

### 2019-09-27 hw2/hw2.R
Initial commit of HW2 files

@@ -1,0 +1,63 @@
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the accuracy of the 
+# predictions.
+
+accuracy <- function(df, actual, predicted){
+  
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the classification error 
+# rate of the predictions.
+# Verify that you get an accuracy and an error rate that sums to one.
+
+classification_error_rate <- function(df, actual, predicted){
+  
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the precision of the 
+# predictions.
+
+precision <- function(df, actual, predicted){
+  
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the sensitivity of the 
+# predictions. Sensitivity is also known as recall.
+
+sensitivity <- function(df, actual, predicted){
+  
+}
+
+recall <- function(df, actual, predicted){
+  sensitivity(df, actual, predicted)
+}
+
+# Write a function that takes the data set as a dataframe, with actual and
+# predicted classifications identified, and returns the specificity of the
+# predictions.
+
+specificity <- function(df, actual, predicted){
+  
+}
+
+# Write a function that takes the data set as a dataframe, with actual and
+# predicted classifications identified, and returns the F1 score of the
+# predictions.
+
+f1_score <- function(df, actual, predicted){
+  
+}
+
+# Write a function that generates an ROC curve from a data set with a true
+# classification column (class in our example) and a probability column
+# (scored.probability in our example). Your function should return a list
+# that includes the plot of the ROC curve and a vector that contains the
+# calculated area under the curve (AUC). Note that I recommend using a
+# sequence of thresholds ranging from 0 to 1 at 0.01 intervals.
+
+roc_curve <- function(df, true_classification, probabilty){
+  
+}

### 2019-09-27 hw2/HW2.Rmd
Initial commit

---
title: "DATA 621 Homework #2"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(pROC)
```

```{r}
source("hw2.R")
df <- read.csv("./data/classification-output-data.csv")
actual <- "class"
predicted <- "scored.class"

cm <- confusion_matrix(df, actual, predicted)
```

### 2019-09-27 hw2/hw2.R
Added confusion_matrix

@@ -1,3 +1,11 @@
+confusion_matrix <- function(df, actual, predicted){
+  table(df[[actual]], df[[predicted]])
+}
+
+confusion_matrix_counts <- function(cm){
+  
+}
+
 # Write a function that takes the data set as a dataframe, with actual and 
 # predicted classifications identified, and returns the accuracy of the 
 # predictions.

### 2019-09-27 hw2/HW2.Rmd
Additional work on HW2

@@
 library(pROC)
+# Thank you Stack Overflow!
+# A Prefix nulling hook.
+
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set( output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if( is.na(comment) ) comment <- ""
+  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+                     x, perl = TRUE)
+  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
+  if( can_null && do_null ) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr( "\\]", x )[1] - 1
+    # Two cases: start or newline
+    re <- paste0( "^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub( re, rep,  x )
+    re <- paste0( "\\\n.{", align_index, "}\\]")
+    rep <- paste0( "\n", comment )
+    x <- gsub( re, rep,  x )
+  }
+
+  default_output_hook( x, options )
+
+})
+knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
 ```
+
+## Confusion Matrix
 
@@
 
-cm <- confusion_matrix(df, actual, predicted)
+confusion_matrix(df, actual, predicted) %>%
+  kable() %>%
+  kable_styling()
 ```
+
+## Accuracy
+
+```{r, null_prefix=TRUE}
+accuracy(df, actual, predicted)
+```
+
+## Classification Error Rate
+
+```{r, null_prefix=TRUE}
+classification_error_rate(df, actual, predicted)
+```
+
+### Verify Accuracy and Error Rate Sum to 1
+
+```{r, null_prefix=TRUE}
+accuracy(df, actual, predicted) + classification_error_rate(df, actual, predicted)
+```
+
+## Precision
+
+```{r, null_prefix=TRUE}
+precision(df, actual, predicted)
+```
+
+## Sensitivity
+
+```{r, null_prefix=TRUE}
+sensitivity(df, actual, predicted)
+```
+
+## Specificity
+
+```{r, null_prefix=TRUE}
+specificity(df, actual, predicted)
+```
+
+## F1 Score
+
+```{r, null_prefix=TRUE}
+f1_score(df, actual, predicted)
+```
+### F1 Score Will Always be Between 0 and 1

### 2019-09-27 hw2/hw2.R

@@ -2,8 +2,13 @@ confusion_matrix <- function(df, actual, predicted){
   table(df[[actual]], df[[predicted]])
 }
 
-confusion_matrix_counts <- function(cm){
-  
+confusion_matrix_outcomes <- function(cm){
+  # This will obviously need some work to handle more than a 2x2 matrix
+  TN <- cm[1]
+  FP <- cm[2]
+  FN <- cm[3]
+  TP <- cm[4]
+  list("TN" = TN, "FP" = FP, "FN" = FN, "TP" = TP)
 }
 
 # Write a function that takes the data set as a dataframe, with actual and 
@@ -11,7 +16,8 @@ confusion_matrix_counts <- function(cm){
 # predictions.
 
 accuracy <- function(df, actual, predicted){
-  
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  (cm$TP + cm$TN) / (cm$TP + cm$FP + cm$TN + cm$FN)
 }
 
 # Write a function that takes the data set as a dataframe, with actual and 
@@ -20,7 +26,8 @@ accuracy <- function(df, actual, predicted){
 # Verify that you get an accuracy and an error rate that sums to one.
 
 classification_error_rate <- function(df, actual, predicted){
-  
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  (cm$FP + cm$FN) / (cm$TP + cm$FP + cm$TN + cm$FN)
 }
 
 # Write a function that takes the data set as a dataframe, with actual and 
@@ -28,7 +35,8 @@ classification_error_rate <- function(df, actual, predicted){
 # predictions.
 
 precision <- function(df, actual, predicted){
-  
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  cm$TP / (cm$TP + cm$FP)
 }
 
 # Write a function that takes the data set as a dataframe, with actual and 
@@ -36,7 +44,8 @@ precision <- function(df, actual, predicted){
 # predictions. Sensitivity is also known as recall.
 
 sensitivity <- function(df, actual, predicted){
-  
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  cm$TP / (cm$TP + cm$FN)
 }
 
 recall <- function(df, actual, predicted){
@@ -48,7 +57,8 @@ recall <- function(df, actual, predicted){
 # predictions.
 
[…]

### 2019-09-27 hw2/hw2.R
Minor updates

@@ -78,6 +78,6 @@ f1_score <- function(df, actual, predicted){
 # calculated area under the curve (AUC). Note that I recommend using a
 # sequence of thresholds ranging from 0 to 1 at 0.01 intervals.
 
-roc_curve <- function(df, true_classification, probabilty){
+roc_curve <- function(df, actual, probabilty){
   
 }

### 2019-10-05 hw2 2/hw2_Adam.R
Beats me

@@ -1,0 +1,76 @@
+##################################################
+# Functions to compute classification measures.
+##################################################
+# I know you have these already, but I thought
+# it may be good practice to do my own.
+#
+
+accuracy <- function(df, pred, act){
+  sum(pred == act)/nrow(df)
+}
+
+errors <- function(df, pred,act){
+  sum(pred != act)/nrow(df)
+}
+
+precision <- function(df, pred, act){
+  sum(pred == 1 & act == 1) /
+    (sum(pred == 1 & act == 1) + sum(pred == 1 & act == 0))
+}
+
+sensitivity <- function(df, pred, act){
+  sum(pred == 1 & act == 1) /
+    (sum(act == 1))
+}
+
+specificity <- function(df, pred, act){
+  sum(pred == 0 & act == 0) /
+    (sum(pred == 0 & act == 0) + sum(pred == 1 & act == 0))
+}
+
+false_positive <- function(df, pred, act){
+  sum(pred == 1 & act == 0) /
+    (sum(act == 0))
+}
+
+f1 <- function(df, pred, act){
+  p <- precision(df, pred, act)
+  s <- sensitivity(df, pred, act)
+  2 * (p * s)/
+    (p + s)
+}
+
+########################
+# ROC Curve generation
+########################
+# I think this will also work with your functions, Mike.
+#
+ROC <- function(df, prob, act){
+  require(dplyr)
+  outcome <- data.frame(matrix(ncol = 4, nrow = 0))
+  colnames(outcome) <- c("prob","TPR","FPR")
+  
+  for (thresh in seq(0,1,0.01)){
+    
+    outcome <- rbind(outcome,
+                     data.frame(
+                       prob=thresh,
+                       TPR = sensitivity(df,ifelse(prob >= thresh,1,0),act),
+                       FPR = false_positive(df,ifelse(prob >= thresh,1,0),act),
+                       area = 0
+                       )
+    )
+  }
+  
+  # Get AUC
+  outcome$area <- (dplyr::lag(outcome$FPR) - outcome$FPR) * outcome$TPR
+
+    plot.new()
+    plot(x=outcome$FPR, y=outcome$TPR, type="l",main="ROC Curve",
+         xlab="False Positive Rate (FPR)",
+         ylab="True Positive Rate (TPR)")
+    abline(a=0,b=1,lty=2)
+    text(0.6,0.1,paste("AUC = ",round(sum(outcome$area, na.rm = T),3)))
+}
+
+ROC(df,df$scored.probability,df$class)

### 2019-10-06 hw2/HW2.Rmd
Minor updates
More work on HW2

@@
 
-## Confusion Matrix
-
 ```{r}
-source("hw2.R")
 df <- read.csv("./data/classification-output-data.csv")
@@
 predicted <- "scored.class"
+probability <- "scored.probability"
+# Load our functions
+source("hw2.R")
+```
 
+## Caret's Output
+
+```{r}
+confusionMatrix(factor(df[[predicted]]), factor(df[[actual]]))
+```
+
+## Our Function's Output
+
+**Confusion Matrix**
+
+```{r}
 confusion_matrix(df, actual, predicted) %>%
@@
 
-## Accuracy
+**Accuracy**: `r accuracy(df, actual, predicted)`
 
-```{r, null_prefix=TRUE}
-accuracy(df, actual, predicted)
-```
+**Classification Error Rate**: `r classification_error_rate(df, actual, predicted)`
 
-## Classification Error Rate
+**Precision**: `r precision(df, actual, predicted)`
 
-```{r, null_prefix=TRUE}
-classification_error_rate(df, actual, predicted)
-```
+**Sensitivity (Recall)**: `r sensitivity(df, actual, predicted)`
 
-### Verify Accuracy and Error Rate Sum to 1
+**Specificity**: `r specificity(df, actual, predicted)`
 
-```{r, null_prefix=TRUE}
-accuracy(df, actual, predicted) + classification_error_rate(df, actual, predicted)
-```
-
-## Precision
-
-```{r, null_prefix=TRUE}
-precision(df, actual, predicted)
-```
-
-## Sensitivity
-
-```{r, null_prefix=TRUE}
-sensitivity(df, actual, predicted)
-```
-
-## Specificity
-
-```{r, null_prefix=TRUE}
-specificity(df, actual, predicted)
-```
-
-## F1 Score
-
-```{r, null_prefix=TRUE}
-f1_score(df, actual, predicted)
-```
-### F1 Score Will Always be Between 0 and 1
+**F1 Score**: `r f1_score(df, actual, predicted)`

### 2019-10-06 hw2/hw2.R

@@ -1,13 +1,13 @@
 confusion_matrix <- function(df, actual, predicted){
-  table(df[[actual]], df[[predicted]])
+  table(df[[predicted]], df[[actual]])
 }
 
 confusion_matrix_outcomes <- function(cm){
   # This will obviously need some work to handle more than a 2x2 matrix
-  TN <- cm[1]
-  FP <- cm[2]
-  FN <- cm[3]
-  TP <- cm[4]
+  TP <- cm[1]
+  FN <- cm[2]
+  FP <- cm[3]
+  TN <- cm[4]
   list("TN" = TN, "FP" = FP, "FN" = FN, "TP" = TP)
 }
 
@@ -67,7 +67,7 @@ specificity <- function(df, actual, predicted){
 
 f1_score <- function(df, actual, predicted){
   p <- precision(df, actual, predicted)
-  s <- specificity(df, actual, predicted)
+  s <- sensitivity(df, actual, predicted)
   (2 * p * s) / (p + s)
 }
 
@@ -78,6 +78,5 @@ f1_score <- function(df, actual, predicted){
 # calculated area under the curve (AUC). Note that I recommend using a
 # sequence of thresholds ranging from 0 to 1 at 0.01 intervals.
 
-roc_curve <- function(df, actual, probabilty){
-  
+roc_curve <- function(df, actual, probability, interval = 0.01){
 }

### 2019-10-06 hw2/HW2.Rmd
Added pROC Output
Work on ROC

@@
 **F1 Score**: `r f1_score(df, actual, predicted)`
+
+## pROC Output
+
+```{r}
+par(pty = "s")
+roc(df[[actual]], df[[predicted]], plot = TRUE, legacy.axes = TRUE, percent = TRUE, print.auc = TRUE)
+par(pty = "m")
+```
+
+## Our ROC Curve
+
+```{r}
+# This is for testing only
+roc_curve(df, actual, probability) %>%
+  kable() %>%
+  kable_styling()
+```
+
+```{r}
+df$test <- ifelse(df[[probability]] > 0.02, 1, 0)
+confusion_matrix(df, actual, "test")
+```
+
+```{r}
+df$test <- ifelse(df[[probability]] > 0.95, 1, 0)
+confusion_matrix(df, actual, "test")
+```

### 2019-10-06 hw2/hw2.R

@@ -79,4 +79,16 @@ f1_score <- function(df, actual, predicted){
 # sequence of thresholds ranging from 0 to 1 at 0.01 intervals.
 
 roc_curve <- function(df, actual, probability, interval = 0.01){
+  outcome <- data.frame(matrix(ncol = 4, nrow = 0))
+  names(outcome) <- c("prob", "TPR", "FPR")
+  for (threshold in seq(0, 1, interval)){
+    df$roc_prediction <- ifelse(df[[probability]] > threshold, 1, 0)
+    cm <- confusion_matrix(df, actual, "roc_prediction")
+    cmo <- confusion_matrix_outcomes(cm)
+    s <- sensitivity(df, actual, "roc_prediction")
+    f <- 1 - specificity(df, actual, "roc_prediction")
+    row <- data.frame(prob = threshold, TPR = s, FPR = f, area = 0)
+    outcome <- rbind(outcome, row)
+  }
+  return(outcome)
 }

### 2019-10-06 hw2/hw2.R
Fixed confusion matrix issue and added ROC curve

@@ -1,5 +1,22 @@
+library(dplyr)
+library(ggplot2)
+
 confusion_matrix <- function(df, actual, predicted){
-  table(df[[predicted]], df[[actual]])
+  cm <- table(df[[predicted]], df[[actual]])
+  if(nrow(cm) == ncol(cm)){
+    return(cm)
+  }
+  # Make sure all missing values are present in the confusion matrix
+  fixed_cm <- matrix(0, ncol(cm), ncol(cm))
+  colnames(fixed_cm) <- colnames(cm)
+  rownames(fixed_cm) <- colnames(cm)
+  # Horribly inefficient at large scale but since we're dealing with a confusion matrix...
+  for(r in rownames(cm)){
+    for(c in colnames(cm)){
+      fixed_cm[r, c] <- cm[r, c]
+    }
+  }
+  return(fixed_cm)
 }
 
 confusion_matrix_outcomes <- function(cm){
@@ -79,7 +96,7 @@ f1_score <- function(df, actual, predicted){
 # sequence of thresholds ranging from 0 to 1 at 0.01 intervals.
 
 roc_curve <- function(df, actual, probability, interval = 0.01){
-  outcome <- data.frame(matrix(ncol = 4, nrow = 0))
+  outcome <- data.frame(matrix(ncol = 3, nrow = 0))
   names(outcome) <- c("prob", "TPR", "FPR")
   for (threshold in seq(0, 1, interval)){
     df$roc_prediction <- ifelse(df[[probability]] > threshold, 1, 0)
@@ -87,8 +104,24 @@ roc_curve <- function(df, actual, probability, interval = 0.01){
     cmo <- confusion_matrix_outcomes(cm)
     s <- sensitivity(df, actual, "roc_prediction")
     f <- 1 - specificity(df, actual, "roc_prediction")
-    row <- data.frame(prob = threshold, TPR = s, FPR = f, area = 0)
+    row <- data.frame(prob = threshold, TPR = s, FPR = f)
     outcome <- rbind(outcome, row)
   }
+  
+  outcome$area <- (outcome$FPR - dplyr::lag(outcome$FPR)) * outcome$TPR
+  
+  auc <- sum(outcome$area, na.rm = TRUE)
+  
+  roc_plot <- ggplot(outcome) +
+    geom_line(aes(FPR, TPR), color="dodger blue") +
+    xlab("False Positive Rate (FPR)") +
[…]

### 2019-10-07 hw2/HW2.Rmd
Updated HW2.Rmd

--- 2019-10-06 hw2/HW2.Rmd
@@
 **F1 Score**: `r f1_score(df, actual, predicted)`
+
+## pROC Output
+
+```{r}
+par(pty = "s")
+roc(df[[actual]], df[[predicted]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+```
+
+## Our ROC Curve
+
+```{r}
+our_roc <- roc_curve(df, actual, probability)
+
+our_roc$plot
+```
+
+### F1 Score Will Always be Between 0 and 1
+Before we move on, let’s consider a question that was asked: What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0<a<1 and 0<b<1 then ab<a).
+
+**Proving the range of the F1 Score must be between 0 and 1**
+
+**Let's expand the F1 equation using the definitions of precision and sensitivity.**
+
+$$\textrm{F1 Score}= \frac{2 \times\mathrm{Precision}\times\mathrm{Sensitivity}}{\mathrm{Precision}+\mathrm{Sensitivity}}$$ 
+**The numerator is equal to**
+$$=\frac{2 \times TP \times TP}{(TP + FP)(TP+FN)}$$ 
+**The denominator is equal to**
+$$= \frac{1}{\frac{TP}{TP+FP}+\frac{TP}{TP+FN}}$$  
+**Combine the Numerator and the denominator:**
+$$F1=\frac{2 \times TP \times TP}{(TP + FP)(TP+FN)} \times \frac{1}{\frac{TP}{TP+FP}+\frac{TP}{TP+FN}}$$  
+**Multiply the denominator by (TP + FP)(TP + FN):**
+
+$$denominator = \frac{(TP+FP)(TP+FN)}{TP(TP+FN)+TP(TP+FP)}$$  
+$$F1=\frac{2(TP)^2}{(TP+FP)(TP+FN)}\times \frac{(TP+FP)(TP+FN)}{TP(TP+FN)+TP(TP+FP)}$$  
+**Cancel the denominator of the first term with the numerator of the second term:**
+
+$$F1 =\frac{2(TP)^2}{TP(TP+FN)+TP(TP+FP)}$$
+**Simplify the Denominator:**
+$$F1 =\frac{2(TP)^2}{TP^2+FNTP+TP^2+TPFP}$$  
+$$F1 =\frac{2(TP)^2}{TP(TP+FN+TP+FP)}$$  
+$$F1 =\frac{2(TP)^2}{TP(TP+FN+TP+FP)}$$  
+$$F1 =\frac{2(TP)^2}{TP(2TP +FN +FP)}$$ 
+**Cancel the TP terms and get final answer**
+$$F1=\frac{2TP}{2TP+FN+FP}$$
+<br>  
+In the equation above, $TP,FN,FP \in \mathbb{N}$ where $\mathbb{N} = \{0,1,2,3...\}$
+
+$2TP \le2TP+FN+FP$ this equation shows that the numerator is at most equal to the denominator (i.e $F1\le1$).
+
+There would be a Maximum value of $1$ when $FN=FP=0$ and $TP > 0$
+
+The fraction $=\frac{2TP}{2TP+FN+FP}$ has a minimum value of zero when the numerator is minimized (i.e. $TP=0$).
+
+We have shown that $0\le F1\le1$
+<br>

### 2019-10-08 hw2/hw2.R
Added AUC vector

@@ -97,20 +97,26 @@ f1_score <- function(df, actual, predicted){
 
 roc_curve <- function(df, actual, probability, interval = 0.01){
   outcome <- data.frame(matrix(ncol = 3, nrow = 0))
-  names(outcome) <- c("prob", "TPR", "FPR")
+  names(outcome) <- c("probability", "TPR", "FPR")
   for (threshold in seq(0, 1, interval)){
     df$roc_prediction <- ifelse(df[[probability]] > threshold, 1, 0)
     cm <- confusion_matrix(df, actual, "roc_prediction")
     cmo <- confusion_matrix_outcomes(cm)
     s <- sensitivity(df, actual, "roc_prediction")
     f <- 1 - specificity(df, actual, "roc_prediction")
-    row <- data.frame(prob = threshold, TPR = s, FPR = f)
+    row <- data.frame(probability = threshold, TPR = s, FPR = f)
     outcome <- rbind(outcome, row)
   }
   
   outcome$area <- (outcome$FPR - dplyr::lag(outcome$FPR)) * outcome$TPR
+  # Fill in missing values with zeros
+  outcome <- outcome %>%
+    mutate(area = ifelse(is.na(area), 0, area))
   
-  auc <- sum(outcome$area, na.rm = TRUE)
+  auc_vector <- outcome$area
+  names(auc_vector) <- outcome$probability
+  
+  auc <- sum(auc_vector)
   
   roc_plot <- ggplot(outcome) +
     geom_line(aes(FPR, TPR), color="dodger blue") +
@@ -121,7 +127,5 @@ roc_curve <- function(df, actual, probability, interval = 0.01){
     geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
     coord_equal(ratio=1)
   
-  return(list(plot = roc_plot, auc = auc))
-  
-  return(outcome)
+  return(list(plot = roc_plot, auc = auc, auc_vector = auc_vector))
 }

### 2019-10-08 hw2/hw2.R
Code cleanup

@@ -3,16 +3,17 @@ library(ggplot2)
 
 confusion_matrix <- function(df, actual, predicted){
   cm <- table(df[[predicted]], df[[actual]])
-  if(nrow(cm) == ncol(cm)){
+  if (nrow(cm) == ncol(cm)){
     return(cm)
   }
   # Make sure all missing values are present in the confusion matrix
   fixed_cm <- matrix(0, ncol(cm), ncol(cm))
   colnames(fixed_cm) <- colnames(cm)
   rownames(fixed_cm) <- colnames(cm)
-  # Horribly inefficient at large scale but since we're dealing with a confusion matrix...
-  for(r in rownames(cm)){
-    for(c in colnames(cm)){
+  # Horribly inefficient at large scale but since we're dealing with a
+  # confusion matrix...
+  for (r in rownames(cm)){
+    for (c in colnames(cm)){
       fixed_cm[r, c] <- cm[r, c]
     }
   }
@@ -40,7 +41,6 @@ accuracy <- function(df, actual, predicted){
 # Write a function that takes the data set as a dataframe, with actual and 
 # predicted classifications identified, and returns the classification error 
 # rate of the predictions.
-# Verify that you get an accuracy and an error rate that sums to one.
 
 classification_error_rate <- function(df, actual, predicted){
   cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
@@ -100,32 +100,31 @@ roc_curve <- function(df, actual, probability, interval = 0.01){
   names(outcome) <- c("probability", "TPR", "FPR")
   for (threshold in seq(0, 1, interval)){
     df$roc_prediction <- ifelse(df[[probability]] > threshold, 1, 0)
-    cm <- confusion_matrix(df, actual, "roc_prediction")
-    cmo <- confusion_matrix_outcomes(cm)
-    s <- sensitivity(df, actual, "roc_prediction")
-    f <- 1 - specificity(df, actual, "roc_prediction")
-    row <- data.frame(probability = threshold, TPR = s, FPR = f)
+    tpr <- sensitivity(df, actual, "roc_prediction")
+    fpr <- 1 - specificity(df, actual, "roc_prediction")
+    row <- data.frame(probability = threshold, TPR = tpr, FPR = fpr)
     outcome <- rbind(outcome, row)
   }
-  
+  # Compute the area
[…]

### 2019-10-23 hw3/HW3.Rmd
Initial commit of HW3

---
title: "DATA 621 Homework #3"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
```

```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .8, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

## Data Exploration

```{r}
train %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

```{r}
for (var in names(train)){
  if(var != "target"){
    plot_df <- train
    plot_df$y <- plot_df[,var]
    p <- ggplot(plot_df, aes(target, y, fill = factor(target))) +
      geom_boxplot() +
      theme_light() +
      ggtitle(var) +
      theme(legend.position = "none",
            axis.title.y = element_blank(),
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank())
    print(p)
  }
}
```

## Data Preparation

## Model Building

## Model Selection

```{r}
library(FFTrees)
fit <- FFTrees(target ~ ., train, test, main = "Crime Rate", decision.labels = c("Below Median", "Above Median"))

for (i in 1:6){
  p <- plot(fit, data = "test", tree=i)
  print(p)
}

```

### 2019-10-23 hw3/HW3.Rmd
Updated HW3.Rmd

@@
 evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
-```
-
-```{r}
 set.seed(42)
@@
 
+The following show the how predictors are distributed between the areas with crime rates higher than the median *blue) and below the median (red).
+
 ```{r}
@@
     plot_df <- train
-    plot_df$y <- plot_df[,var]
-    p <- ggplot(plot_df, aes(target, y, fill = factor(target))) +
-      geom_boxplot() +
+    plot_df$x <- plot_df[,var]
+    p <- ggplot(plot_df, aes(x, color = factor(target))) +
+      geom_density() +
       theme_light() +
       ggtitle(var) +
+      scale_color_brewer(palette = "Set1") +
       theme(legend.position = "none",
             axis.title.y = element_blank(),
-            axis.title.x = element_blank(),
-            axis.text.x = element_blank(),
-            axis.ticks.x = element_blank())
+            axis.title.x = element_blank())
     print(p)

### 2019-10-25 hw3/HW3.Rmd
More work on HW3

@@
 library(caret)
+library(Amelia)
 # Thank you Stack Overflow!
@@
 evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
+```
+
+## Data Exploration
+
+### Are There Missing Values?
+
+```{r}
+missmap(df, main = "Missing vs Observed Values")
+```
+
+It looks like we have a complete data set.  No need to impute values.
+
+### Splitting the Data
+
+```{r}
 set.seed(42)
-train_index <- createDataPartition(df$target, p = .8, list = FALSE, times = 1)
+train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
 train <- df[train_index,]
@@
 
-## Data Exploration
+### Exploratory Data Analysis
 
@@
 train %>% 
-  cor(., use = "complete.obs") %>%
+  cor(.) %>%
   corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
@@
 The following show the how predictors are distributed between the areas with crime rates higher than the median *blue) and below the median (red).
+What we are looking for is variables that could split data into two groups.
 
@@
 
+NOX seems to be the best variable to divide the data into the two groups
+
 ## Data Preparation
@@
 ## Model Building
+
+Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.
+
+```{r}
+baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
+summary(baseline)
+test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
+test$baseline2 <- predict(baseline, test)
+confusionMatrix(factor(test$baseline), factor(test$target))
+```
 
@@
 library(FFTrees)
+
+test <- test[names(train)]
+
 fit <- FFTrees(target ~ ., train, test, main = "Crime Rate", decision.labels = c("Below Median", "Above Median"))

### 2019-10-29 hw3/HW3.Rmd
More work on HW3

@@
 library(caret)
+library(rpart)
+library(rpart.plot)
 library(Amelia)
@@
 ## Data Preparation
-
-## Model Building
-
-Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.
-
 ```{r}
-baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
-summary(baseline)
-test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
-test$baseline2 <- predict(baseline, test)
-confusionMatrix(factor(test$baseline), factor(test$target))
+ggplot(train, aes(rad , nox, color=as.factor(target))) +
+  geom_point()
 ```
-
-## Model Selection
 
@@
 }
+```
 
+```{r}
+fit <- rpart(target ~ ., method="class", data=train)
+rpart.plot(fit)
 ```
+
+```{r}
+get_high_nox <- function(x){
+  if(x > 0.53){
+    return("High")
+  } else {
+    return("Low")
+  }
+}
+
+get_divider <- function(x, y){
+  if(x < 10 & y < 0.6){
+    return(1)
+  } else {
+    return(0)
+  }
+}
+
+train <- train %>%
+  rowwise() %>%
+  mutate(high_nox = get_high_nox(nox),
+         divider = get_divider(rad, nox))
+
+test <- test %>%
+  rowwise() %>%
+  mutate(high_nox = get_high_nox(nox),
+         divider = get_divider(rad, nox))
+```
+
+## Model Building
+
+Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.
+
+```{r}
+baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
+summary(baseline)
+test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
+test$baseline_yhat <- predict(baseline, test)
+confusionMatrix(factor(test$baseline), factor(test$target))
+```
+
+```{r}
+model1 <- glm(target ~ high_nox + indus, family = binomial(link = "logit"), train)
+summary(model1)
+test$model1 <- ifelse(predict(model1, test) < 0.5, 0, 1)
+test$model1_yhat <- predict(model1, test)
+confusionMatrix(factor(test$model1), factor(test$target))
+```
+
+```{r}
+model2 <- glm(target ~ divider, family = binomial(link = "logit"), train)
+summary(model2)
+test$model2 <- ifelse(predict(model2, test) < 0.5, 0, 1)
+test$model2_yhat <- predict(model2, test)
+confusionMatrix(factor(test$model2), factor(test$target))
+```
+
+## Model Selection

### 2019-10-29 hw3/HW3_Adam.Rmd

--- 2019-10-25 hw3/HW3.Rmd
@@
 summary(baseline)
-test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
-test$baseline2 <- predict(baseline, test)
-confusionMatrix(factor(test$baseline), factor(test$target))
+test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
+cm <- confusionMatrix(factor(test$baseline2), factor(test$target),"1")
+results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
+                  deviance=baseline$deviance, 
+                  r2 = 1 - baseline$deviance/baseline$null.deviance,
+                  aic=baseline$aic)
+cm
+```
+
+Our baseline model is ok with an F1 score of `r cm$byClass[7]`.
+
+Now we will try adding all other variables and work backwards to eliminate non-significant predictors:
+
+```{r}
+fullmodel <- glm(target ~ nox + dis, family = binomial(link = "logit"), train)
+summary(fullmodel)
+test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
+results <- rbind(results,tibble(model = "fullmodel",
+                                predictors = 12,F1 = cm$byClass[7],
+                                deviance=fullmodel$deviance, 
+                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
+                                aic=fullmodel$aic))
+cm
+```
+
+The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, the new variable doesn't seem to be significant.
+
+After some backward elimination of non-significant predictor variables, we arrive at the following model:
+
+```{r}
+model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
+              family = binomial(link = "logit"), 
+              train)
+summary(model1)
+test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
+results <- rbind(results,tibble(model = "model1",
+                                predictors = 6,F1 = cm$byClass[7],
+                                deviance=model1$deviance, 
+                                r2 = 1-model1$deviance/model1$null.deviance,
+                                aic=model1$aic))
+cm
+```
+
+Looking at our models together, `model1` looks like the best so far based on highest F1 score and lowest deviance & AIC.
+
+```{r}
+kable(results)
 ```

### 2019-10-30 hw3/HW3.Rmd
testing commit
Folded in Adam's work

--- 2019-10-25 hw3/HW3.Rmd
@@
 
-The following show the how predictors are distributed between the areas with crime rates higher than the median *blue) and below the median (red).
+The following show the how predictors are distributed between the areas with crime rates higher than the median (blue) and below the median (red).
 What we are looking for is variables that could split data into two groups.
@@
 summary(baseline)
-test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
-test$baseline2 <- predict(baseline, test)
-confusionMatrix(factor(test$baseline), factor(test$target))
+test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
+cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
+results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
+                  deviance=baseline$deviance, 
+                  r2 = 1 - baseline$deviance/baseline$null.deviance,
+                  aic=baseline$aic)
+cm
+```
+
+Our baseline model is ok with an F1 score of `r cm$byClass[7]`.
+
+Now we will try adding all other variables and work backwards to eliminate non-significant predictors:
+
+```{r}
+fullmodel <- glm(target ~ nox + dis, family = binomial(link = "logit"), train)
+summary(fullmodel)
+test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
+results <- rbind(results,tibble(model = "fullmodel",
+                                predictors = 12,F1 = cm$byClass[7],
+                                deviance=fullmodel$deviance, 
+                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
+                                aic=fullmodel$aic))
+cm
+```
+
+The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, the new variable doesn't seem to be significant.
+
+After some backward elimination of non-significant predictor variables, we arrive at the following model:
+
+```{r}
+model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
+              family = binomial(link = "logit"), 
+              train)
+summary(model1)
+test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
+results <- rbind(results,tibble(model = "model1",
+                                predictors = 6,F1 = cm$byClass[7],
+                                deviance=model1$deviance, 
+                                r2 = 1-model1$deviance/model1$null.deviance,
+                                aic=model1$aic))
+cm
+```
+
+Looking at our models together, `model1` looks like the best so far based on highest F1 score and lowest deviance & AIC.
+
+```{r}
+kable(results)
 ```
@@
 ## Model Selection
-
-```{r}
-library(FFTrees)
-
-test <- test[names(train)]
-
-fit <- FFTrees(target ~ ., train, test, main = "Crime Rate", decision.labels = c("Below Median", "Above Median"))
-
-for (i in 1:6){
-  p <- plot(fit, data = "test", tree=i)
-  print(p)
-}
-
-```

### 2019-10-30 hw3/HW3.Rmd
Update HW3.Rmd
Added some description

@@
 ### Are There Missing Values?
+First we look at the data to see if any variables have missing data:
 
@@
 ### Splitting the Data
-
+Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
 ```{r}
@@
 ### Exploratory Data Analysis
+Now, let's look at our training data.
+By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables.
+This will help us later during the model selection process.
 
@@
 
-The following show the how predictors are distributed between the areas with crime rates higher than the median (blue) and below the median (red).
-What we are looking for is variables that could split data into two groups.
+Next we will look at each potential predictor and how it is distributed across the target variable.
+
+The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e.
+blue) and a negative target variable (areas with crime rates below the median, i.e.
+red).
+What we are looking for is variables that show way to split data into two groups.
 
@@
 
-NOX seems to be the best variable to divide the data into the two groups
+Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.
 
@@
 
-Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.
+We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.
 
@@
 
-Now we will try adding all other variables and work backwards to eliminate non-significant predictors:
+Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:
 
 ```{r}
-fullmodel <- glm(target ~ nox + dis, family = binomial(link = "logit"), train)
+fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
 summary(fullmodel)
@@
 
-The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, the new variable doesn't seem to be significant.
+The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.
 
@@
 
-Looking at our models together, `model1` looks like the best so far based on highest F1 score and lowest deviance & AIC.
+Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.
 
@@
 
-## Model Selection
+Looking these three models together, `model1` looks like the best one.
+Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.

### 2019-10-30 hw3/HW3_Corey.Rmd
committing feature engineering using my density difference function

---
title: "DATA 621 Homework #3"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
library(rpart)
library(rpart.plot)
library(Amelia)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
```

## Data Exploration

### Are There Missing Values?

```{r}
missmap(df, main = "Missing vs Observed Values")
```

It looks like we have a complete data set.  No need to impute values.

### Splitting the Data

```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Exploratory Data Analysis

```{r}
train %>% 
  cor(.) %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

The following show the how predictors are distributed between the areas with crime rates higher than the median *blue) and below the median (red).  What we are looking for is variables that could split data into two groups.

```{r}
for (var in names(train)){
  if(var != "target"){
    plot_df <- train
    plot_df$x <- plot_df[,var]
    p <- ggplot(plot_df, aes(x, color = factor(target))) +
      geom_density() +
      theme_light() +
      ggtitle(var) +
      scale_color_brewer(palette = "Set1") +
      theme(legend.position = "none",
            axis.title.y = element_blank(),
            axis.title.x = element_blank())
    print(p)
  }
}
```

NOX seems to be the best variable to divide the data into the two groups

## Data Preparation
```{r}
ggplot(train, aes(rad , nox, color=as.factor(target))) +
  geom_point()
```

```{r}
library(FFTrees)

test <- test[names(train)]

fit <- FFTrees(target ~ ., train, test, main = "Crime Rate", decision.labels = c("Below Median", "Above Median"))

for (i in 1:6){
  p <- plot(fit, data = "test", tree=i)
  print(p)
}
```

```{r}
fit <- rpart(target ~ ., method="class", data=train)
rpart.plot(fit)
```

```{r}
get_high_nox <- function(x){
  if(x > 0.53){
    return("High")
  } else {
    return("Low")
  }
}

get_divider <- function(x, y){
  if(x < 10 & y < 0.6){
    return(1)
  } else {
    return(0)
  }
}

train <- train %>%
  rowwise() %>%
  mutate(high_nox = get_high_nox(nox),
         divider = get_divider(rad, nox))

test <- test %>%
  rowwise() %>%
  mutate(high_nox = get_high_nox(nox),
         divider = get_divider(rad, nox))
```

## Feature Engineering
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(rpart)				        # Popular decision tree algorithm
library(rattle)					# Fancy tree plot
library(rpart.plot)				# Enhanced tree plots
library(RColorBrewer)				# Color selection for fancy tree plot
library(party)					# Alternative decision tree algorithm
library(partykit)				# Convert rpart object to BinaryTree
library(caret)
library(stringr)
library(data.table)
library(plyr);
library(dplyr)
library(randomForest)
library(memisc)
library(plotly)
library(grid)
library(fuzzyjoin)

df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")

density_diff <- function(data,var,target,new_column){
#standardize variable between 0 and 1
var <- (var - min(var))/(max(var) - min(var))
## Calculate density estimates
g1 <- ggplot(data, aes(x=var, group=target, colour=target)) +
  geom_density(data = data) + xlim(min(var), max(var))
gg1 <- ggplot_build(g1)
# construct the dataset
x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), 
                  side=(y1<y2), ydiff = y2-y1)
##creating the second graph object
g3 <- ggplot(df2) +
  geom_line(aes(x = x, y = ydiff, colour = side)) +
  geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
  guides(alpha = FALSE, fill = FALSE)

data$join <- var
df2$join <- df2$x
temp <- difference_left_join(data,df2, by="join", max_dist =.001)
means <- aggregate(ydiff ~ temp$join.x, temp, mean)

##new stuff
colnames(means) <- c("std_var","density_diff")
data["key"] <- (var - min(var))/(max(var) - min(var))

new_data <- merge(means,data, by.x ="std_var", by.y ="key")
new_data$std_var <- NULL
new_data$join <- NULL
new_data[new_column]<- new_data$density_diff
new_data$density_diff <- NULL
##new stuff

mylist <- list(g1,g3,new_data)
names(mylist)<- c("dist","dist_diff","new_data_training")
return(mylist)

}

new_data2 <- density_diff(df,df$nox,df$target,"nox_density")$new_data_training

### do it now with a way to include evaluation data

df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")

density_diff <- function(data,var,target,new_column){
#standardize variable between 0 and 1
data[var] <- (data[var] - min(data[var]))/(max(data[var]) - min(data[var]))
## Calculate density estimates
g1 <- ggplot(data, aes(x=data[var], group=target, colour=target)) +
  geom_density(data = data) + xlim(min(data[var]), max(data[var]))
gg1 <- ggplot_build(g1)
# construct the dataset
x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), 
                  side=(y1<y2), ydiff = y2-y1)
##creating the second graph object
g3 <- ggplot(df2) +
  geom_line(aes(x = x, y = ydiff, colour = side)) +
  geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
  guides(alpha = FALSE, fill = FALSE)

data$join <- data[var]
df2$join <- df2$x
temp <- difference_left_join(data,df2, by="join", max_dist =.001)
means <- aggregate(ydiff ~ temp$join.x, temp, mean)

##new stuff
colnames(means) <- c("std_var","density_diff")
data["key"] <- (var - min(var))/(max(var) - min(var))

new_data <- merge(means,data, by.x ="std_var", by.y ="key")
new_data$std_var <- NULL
new_data$join <- NULL
new_data[new_column]<- new_data$density_diff
new_data$density_diff <- NULL
##new stuff

mylist <- list(g1,g3,new_data)
names(mylist)<- c("dist","dist_diff","new_data_training")
return(mylist)

}

new_data2 <- density_diff(df,df$nox,df$target,"nox_density")$new_data_training

```

## Model Building

Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.

```{r}
baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
test$baseline_yhat <- predict(baseline, test)
confusionMatrix(factor(test$baseline), factor(test$target))
```

```{r}
model1 <- glm(target ~ high_nox + indus, family = binomial(link = "logit"), train)
summary(model1)
test$model1 <- ifelse(predict(model1, test) < 0.5, 0, 1)
test$model1_yhat <- predict(model1, test)
confusionMatrix(factor(test$model1), factor(test$target))
```

```{r}
model2 <- glm(target ~ divider, family = binomial(link = "logit"), train)
summary(model2)
test$model2 <- ifelse(predict(model2, test) < 0.5, 0, 1)
test$model2_yhat <- predict(model2, test)
confusionMatrix(factor(test$model2), factor(test$target))
```

## Model Selection

### 2019-10-30 hw3/HW3.Rmd
added in my new feature engineering and tested out some models

--- 2019-10-29 hw3/HW3.Rmd
@@
 
+## Feature Engineering
+```{r}
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(rpart)				        # Popular decision tree algorithm
+library(rattle)					# Fancy tree plot
+library(rpart.plot)				# Enhanced tree plots
+library(RColorBrewer)				# Color selection for fancy tree plot
+library(party)					# Alternative decision tree algorithm
+library(partykit)				# Convert rpart object to BinaryTree
+library(caret)
+library(stringr)
+library(data.table)
+library(plyr);
+library(dplyr)
+library(randomForest)
+library(memisc)
+library(plotly)
+library(grid)
+library(fuzzyjoin)
+
+df <- read.csv('data/crime-training-data_modified.csv')
+evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
+
+density_diff <- function(data,eval_data,var,target,new_column){
+
+data[[paste0(var,"temp")]]<- data[[var]]
+
+eval_data[[paste0(var,"temp")]]<- eval_data[[var]]
+
+#standardize variable between 0 and 1
+data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
+
+eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))
+
+## Calculate density estimates
+g1 <- ggplot(data, aes(x=data[[var]], group=target, colour=target)) +
+  geom_density(data = data) + xlim(min(data[[var]]), max(data[[var]]))
+gg1 <- ggplot_build(g1)
+# construct the dataset
+x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
+y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
+y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
+df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), 
+                  side=(y1<y2), ydiff = y2-y1)
+##creating the second graph object
+g3 <- ggplot(df2) +
+  geom_line(aes(x = x, y = ydiff, colour = side)) +
+  geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
+  guides(alpha = FALSE, fill = FALSE)
+
+data$join <- data[[var]]
+df2$join <- df2$x
+temp <- difference_left_join(data,df2, by="join", max_dist =.001)
+means <- aggregate(ydiff ~ temp$join.x, temp, mean)
+
+colnames(means) <- c("std_var","density_diff")
+
+new_data <- merge(means,data, by.x ="std_var", by.y =var)
+new_data$std_var <- NULL
+new_data$join <- NULL
+new_data[new_column]<- new_data$density_diff
+new_data$density_diff <- NULL
+###################same thing with eval data ############################
+
+eval_data$join <- eval_data[[var]]
+df2$join <- df2$x
+temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
+means <- aggregate(ydiff ~ temp$join.x, temp, mean)
+
+##new stuff
+colnames(means) <- c("std_var","density_diff")
+#data["key"] <- (var - min(var))/(max(var) - min(var))
+
+eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
+eval_data2$std_var <- NULL
+eval_data2$join <- NULL
+eval_data2[new_column]<- eval_data2$density_diff
+eval_data2$density_diff <- NULL
+
+eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
+eval_data2[[paste0(var,"temp")]]<- NULL
+
+new_data[[var]] <- new_data[[paste0(var,"temp")]]
+new_data[[paste0(var,"temp")]]<- NULL
+
+mylist <- list(g1,g3,new_data,eval_data2)
+names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
+return(mylist)
+
+}
+
+```
+
+```{r}
+baseline <- glm(target ~ ., family = binomial(link = "logit"), train)
+summary(baseline)
+test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
+test$baseline_yhat <- predict(baseline, test)
+confusionMatrix(factor(test$baseline), factor(test$target))
+```
+
+#Creating Some New Variables
+
+```{r}
+str(train)
+str(test)
+## try it out 
+train <- density_diff(train,test,"nox",train$target,"nox_density")$new_data_training
+train <- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
+train <- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
+train <- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
+train <- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
+train <- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
+train <- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
+train <- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
+train <- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
+
+test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
+test<- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
+test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
+test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
+test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
+test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
+test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
+test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
+test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
+
+str(train)
+str(test)
+
+```
+
 ## Model Building
@@
 
+```{r}
+
+density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
+summary(density_models)
+test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
+test$density_models_yhat <- predict(density_models, test)
+confusionMatrix(factor(test$density_models), factor(test$target))
+```
+
 ## Model Selection
+
+## pROC Output
+
+```{r}
+library(pROC)
+par(pty = "s")
+roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+
+library(pROC)
+par(pty = "s")
+roc(test[["target"]], test[["model2_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+
+library(pROC)
+par(pty = "s")
+roc(test[["target"]], test[["model1_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+```

### 2019-10-30 hw3/HW3.Rmd
cleaned some things up

@@
 ```{r}
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(rpart)				        # Popular decision tree algorithm
-library(rattle)					# Fancy tree plot
-library(rpart.plot)				# Enhanced tree plots
-library(RColorBrewer)				# Color selection for fancy tree plot
-library(party)					# Alternative decision tree algorithm
-library(partykit)				# Convert rpart object to BinaryTree
-library(caret)
-library(stringr)
-library(data.table)
-library(plyr);
-library(dplyr)
-library(randomForest)
-library(memisc)
-library(plotly)
-library(grid)
+
 library(fuzzyjoin)
-
-df <- read.csv('data/crime-training-data_modified.csv')
-evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
 

### 2019-10-30 hw3/HW3.Rmd
Fixed a mistake

@@
 train <- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
+train <- density_diff(train,test,"rm",train$target,"rm_density")$new_data_training
+train <- density_diff(train,test,"zn",train$target,"zn_density")$new_data_training
 
 test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
-test<- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
-test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
-test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
-test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
-test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
-test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
-test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
-test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
+test<- density_diff(train,test,"age",train$target,"age_density")$new_data_eval
+test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_eval
+test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_eval
+test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_eval
+test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_eval
+test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_eval
+test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_eval
+test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_eval
+test<- density_diff(train,test,"rm",train$target,"rm_density")$new_data_eval
+test<- density_diff(train,test,"zn",train$target,"zn_density")$new_data_eval
+
+str(train)
+str(test)
 

### 2019-10-30 hw3/HW3.Rmd
updated H3.rmd
Updated my model

@@
 
-density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
+density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density+zn_density+rm_density+divider, family = binomial(link = "logit"), train)
 summary(density_models)
-test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
+test$density_models <- ifelse(predict(density_models, test) < mean(train$target), 0, 1)
 test$density_models_yhat <- predict(density_models, test)

### 2019-10-30 hw3/HW3.Rmd
Folded in Corey's work

--- 2019-10-30 hw3/HW3.Rmd
@@
 library(caret)
-library(rpart)
-library(rpart.plot)
 library(Amelia)
+library(pROC)
 # Thank you Stack Overflow!
@@
 
-The following show the how predictors are distributed between the areas with crime rates higher than the median *blue) and below the median (red).
-What we are looking for is variables that could split data into two groups.
+Next we will look at each potential predictor and how it is distributed across the target variable.
+
+The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e.
+blue) and a negative target variable (areas with crime rates below the median, i.e.
+red).
+What we are looking for is variables that show way to split data into two groups.
 
@@
 
-NOX seems to be the best variable to divide the data into the two groups
+Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.
 
 ## Data Preparation
+
+### Feature Engineering
 ```{r}
-ggplot(train, aes(rad , nox, color=as.factor(target))) +
-  geom_point()
-```
-
-```{r}
-library(FFTrees)
-
-test <- test[names(train)]
-
-fit <- FFTrees(target ~ ., train, test, main = "Crime Rate", decision.labels = c("Below Median", "Above Median"))
-
-for (i in 1:6){
-  p <- plot(fit, data = "test", tree=i)
-  print(p)
-}
-```
-
-```{r}
-fit <- rpart(target ~ ., method="class", data=train)
-rpart.plot(fit)
-```
-
-```{r}
-get_high_nox <- function(x){
-  if(x > 0.53){
-    return("High")
-  } else {
-    return("Low")
-  }
-}
-
-get_divider <- function(x, y){
-  if(x < 10 & y < 0.6){
-    return(1)
-  } else {
-    return(0)
-  }
-}
-
-train <- train %>%
-  rowwise() %>%
-  mutate(high_nox = get_high_nox(nox),
-         divider = get_divider(rad, nox))
-
-test <- test %>%
-  rowwise() %>%
-  mutate(high_nox = get_high_nox(nox),
-         divider = get_divider(rad, nox))
-```
-
-## Feature Engineering
-```{r}
-
 library(fuzzyjoin)
-
 density_diff <- function(data,eval_data,var,target,new_column){
-
-data[[paste0(var,"temp")]]<- data[[var]]
-
-eval_data[[paste0(var,"temp")]]<- eval_data[[var]]
+  data[[paste0(var,"temp")]] <- data[[var]]
+  eval_data[[paste0(var,"temp")]] <- eval_data[[var]]
 
@@
   data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
-
   eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))
@@
   y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
-df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), 
-                  side=(y1<y2), ydiff = y2-y1)
+  df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), side=(y1<y2), ydiff = y2-y1)
   ##creating the second graph object
@@
   new_data$join <- NULL
-new_data[new_column]<- new_data$density_diff
+  new_data[new_column] <- new_data$density_diff
   new_data$density_diff <- NULL
   ###################same thing with eval data ############################
-
   eval_data$join <- eval_data[[var]]
@@
   eval_data2$join <- NULL
-eval_data2[new_column]<- eval_data2$density_diff
+  eval_data2[new_column] <- eval_data2$density_diff
   eval_data2$density_diff <- NULL
@@
   eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
-eval_data2[[paste0(var,"temp")]]<- NULL
+  eval_data2[[paste0(var,"temp")]] <- NULL
 
   new_data[[var]] <- new_data[[paste0(var,"temp")]]
-new_data[[paste0(var,"temp")]]<- NULL
+  new_data[[paste0(var,"temp")]] <- NULL
 
@@
   return(mylist)
-
 }
-
 ```
 
-```{r}
-baseline <- glm(target ~ ., family = binomial(link = "logit"), train)
-summary(baseline)
-test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
-test$baseline_yhat <- predict(baseline, test)
-confusionMatrix(factor(test$baseline), factor(test$target))
-```
-
-#Creating Some New Variables
+### Creating Some New Variables
 
@@
 str(test)
-
 ```
@@
 summary(baseline)
-test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
-test$baseline_yhat <- predict(baseline, test)
-confusionMatrix(factor(test$baseline), factor(test$target))
+test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
+cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
+results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
+                  deviance=baseline$deviance, 
+                  r2 = 1 - baseline$deviance/baseline$null.deviance,
+                  aic=baseline$aic)
+cm
+```
+
+Our baseline model is ok with an F1 score of `r cm$byClass[7]`.
+
+Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:
+
+```{r}
+fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
+summary(fullmodel)
+test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
+results <- rbind(results,tibble(model = "fullmodel",
+                                predictors = 12,F1 = cm$byClass[7],
+                                deviance=fullmodel$deviance, 
+                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
+                                aic=fullmodel$aic))
+cm
+```
+
+The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.
+
+After some backward elimination of non-significant predictor variables, we arrive at the following model:
+
+```{r}
+model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
+              family = binomial(link = "logit"), 
+              train)
+summary(model1)
+test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
+results <- rbind(results,tibble(model = "model1",
+                                predictors = 6,F1 = cm$byClass[7],
+                                deviance=model1$deviance, 
+                                r2 = 1-model1$deviance/model1$null.deviance,
+                                aic=model1$aic))
+cm
 ```
@@
 ```{r}
-model1 <- glm(target ~ high_nox + indus, family = binomial(link = "logit"), train)
-summary(model1)
-test$model1 <- ifelse(predict(model1, test) < 0.5, 0, 1)
-test$model1_yhat <- predict(model1, test)
-confusionMatrix(factor(test$model1), factor(test$target))
-```
-
-```{r}
-model2 <- glm(target ~ divider, family = binomial(link = "logit"), train)
-summary(model2)
-test$model2 <- ifelse(predict(model2, test) < 0.5, 0, 1)
-test$model2_yhat <- predict(model2, test)
-confusionMatrix(factor(test$model2), factor(test$target))
-```
-
-```{r}
-
 density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
@@
 test$density_models_yhat <- predict(density_models, test)
-confusionMatrix(factor(test$density_models), factor(test$target))
+cm <- confusionMatrix(factor(test$density_models), factor(test$target))
+results <- rbind(results,tibble(model = "density models",
+                                predictors = 9,
+                                F1 = cm$byClass[7],
+                                deviance = density_models$deviance,
+                                r2 = 1 - density_models$deviance / density_models$null.deviance,
+                                aic = density_models$aic))
+cm
 ```
+
+Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.
+
+```{r}
+kable(results)
+```
+
+Looking these three models together, `model1` looks like the best one.
+Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.
 
@@
 
-## pROC Output
+### pROC Output
 
+#### Baseline
 ```{r}
-library(pROC)
+par(pty = "s")
+roc(test[["target"]], test[["baseline"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+```
+
+#### Full Model
+```{r}
+par(pty = "s")
+roc(test[["target"]], test[["fullmodel"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+```
+
+#### Model1
+```{r}
+par(pty = "s")
+roc(test[["target"]], test[["model1"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
+par(pty = "m")
+```
+
+#### Density Model
+```{r}
 par(pty = "s")
@@
 par(pty = "m")
-
-library(pROC)
-par(pty = "s")
-roc(test[["target"]], test[["model2_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
-par(pty = "m")
-
-library(pROC)
-par(pty = "s")
-roc(test[["target"]], test[["model1_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
-par(pty = "m")
 ```

### 2019-10-31 hw3/HW3.Rmd
Fixes and added description

@@
 ### Feature Engineering
+In this portion below we explored using the difference between the class distributions as a feature itself.
+So instead of the raw feature itself the we created density features that are essientially the differences in probability of the positve and negative classes at that certain point for the feature.
+
 ```{r}
@@
 test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
-test<- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
-test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
-test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
-test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
-test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
-test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
-test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
-test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
+test<- density_diff(train,test,"age",train$target,"age_density")$new_data_eval
+test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_eval
+test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_eval
+test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_eval
+test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_eval
+test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_eval
+test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_eval
+test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_eval
 
@@
 str(test)
+```
+
+### Example of the new meta feature
+The first graph below is the difference between the distributions and the second graph is the new derived predictive feature
+```{r}
+
+density_diff(train,test,"age",train$target,"age_density")$dist
+density_diff(train,test,"age",train$target,"age_density")$dist_diff
 ```

### 2019-10-31 hw3/HW3.Rmd
Update HW3

@@
 
-## Data Preparation
+## Basic Model Building
+
+We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.
+
+```{r}
+baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
+summary(baseline)
+test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
+cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
+results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
+                  deviance=baseline$deviance, 
+                  r2 = 1 - baseline$deviance/baseline$null.deviance,
+                  aic=baseline$aic)
+cm
+```
+
+Our baseline model is ok with an F1 score of `r cm$byClass[7]`.
+
+Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:
+
+```{r}
+fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
+summary(fullmodel)
+test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
+results <- rbind(results,tibble(model = "fullmodel",
+                                predictors = 12,F1 = cm$byClass[7],
+                                deviance=fullmodel$deviance, 
+                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
+                                aic=fullmodel$aic))
+cm
+```
+
+The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.
+
+After some backward elimination of non-significant predictor variables, we arrive at the following model:
+
+```{r}
+model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
+              family = binomial(link = "logit"), 
+              train)
+summary(model1)
+test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
+cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
+results <- rbind(results,tibble(model = "model1",
+                                predictors = 6,F1 = cm$byClass[7],
+                                deviance=model1$deviance, 
+                                r2 = 1-model1$deviance/model1$null.deviance,
+                                aic=model1$aic))
+cm
+```
+
+## Alternative Models
 
@@
 ```{r}
-
 density_diff(train,test,"age",train$target,"age_density")$dist
@@
 
-## Model Building
-
-We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.
-
-```{r}
-baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
-summary(baseline)
-test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
-cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
-results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
-                  deviance=baseline$deviance, 
-                  r2 = 1 - baseline$deviance/baseline$null.deviance,
-                  aic=baseline$aic)
-cm
-```
-
-Our baseline model is ok with an F1 score of `r cm$byClass[7]`.
-
-Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:
-
-```{r}
-fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
-summary(fullmodel)
-test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
-cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
-results <- rbind(results,tibble(model = "fullmodel",
-                                predictors = 12,F1 = cm$byClass[7],
-                                deviance=fullmodel$deviance, 
-                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
-                                aic=fullmodel$aic))
-cm
-```
-
-The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.
-
-After some backward elimination of non-significant predictor variables, we arrive at the following model:
-
-```{r}
-model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
-              family = binomial(link = "logit"), 
-              train)
-summary(model1)
-test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
-cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
-results <- rbind(results,tibble(model = "model1",
-                                predictors = 6,F1 = cm$byClass[7],
-                                deviance=model1$deviance, 
-                                r2 = 1-model1$deviance/model1$null.deviance,
-                                aic=model1$aic))
-cm
-```
+With our new density variables, we can run another model:
 
@@
 
+## Model Selection
+
 Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.
@@
 
-Looking these three models together, `model1` looks like the best one.
+Looking the three models together, `model1` looks like the best one.
 Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.
 
-## Model Selection
+### pROC Output
 
-### pROC Output
+ROC curves can give us another look at which model might be better suited for prediction.
 

### 2019-11-02 hw4/HW4.Rmd
Initial commit

--- 2019-10-30 hw3/HW3.Rmd
@@
 ---
-title: "DATA 621 Homework #3"
+title: "DATA 621 Homework #4"
 author: "Critical Thinking Group 3"
@@
 knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
-library(tidyverse)
+library(dplyr)
+library(Amelia)
 library(kableExtra)
-library(corrplot)
 library(caret)
-library(Amelia)
-library(pROC)
+library(DMwR)
 # Thank you Stack Overflow!
@@
 ```{r}
-df <- read.csv('data/crime-training-data_modified.csv')
-evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
+df <- read.csv('./data/insurance_training_data.csv') %>%
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG))
+evaluation <- read.csv("./data/insurance-evaluation-data.csv")
 ```
@@
 
-It looks like we have a complete data set.  No need to impute values.
+There are missing values for the `CAR_AGE` and `YOJ`(Years on Job) variables.  We will fill in the missing variables using KNN.
+
+```{r}
+set.seed(42)
+knn <- knnImputation(df)
+impute_me <- is.na(df$CAR_AGE)
+df[impute_me,"CAR_AGE"] <- knn[impute_me,"CAR_AGE"] 
+impute_me <- is.na(df$YOJ)
+df[impute_me,"YOJ"] <- knn[impute_me,"YOJ"] 
+```
 
@@
 Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
+
 ```{r}
 set.seed(42)
-train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
+train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
 train <- df[train_index,]
@@
 
-### Exploratory Data Analysis
-Now, let's look at our training data.
-By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables.
-This will help us later during the model selection process.
+### First Look at the Data
+
+We will now take our first look at the data to get a sense of what we have.
 
@@
 train %>%
-  cor(.) %>%
-  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
+  summary() %>%
+  kable() %>%
+  kable_styling()
 ```
-
-Next we will look at each potential predictor and how it is distributed across the target variable.
-
-The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e.
-blue) and a negative target variable (areas with crime rates below the median, i.e.
-red).
-What we are looking for is variables that show way to split data into two groups.
-
-```{r}
-for (var in names(train)){
-  if(var != "target"){
-    plot_df <- train
-    plot_df$x <- plot_df[,var]
-    p <- ggplot(plot_df, aes(x, color = factor(target))) +
-      geom_density() +
-      theme_light() +
-      ggtitle(var) +
-      scale_color_brewer(palette = "Set1") +
-      theme(legend.position = "none",
-            axis.title.y = element_blank(),
-            axis.title.x = element_blank())
-    print(p)
-  }
-}
-```
-
-Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.
-
-## Data Preparation
-
-### Feature Engineering
-```{r}
-library(fuzzyjoin)
-density_diff <- function(data,eval_data,var,target,new_column){
-  data[[paste0(var,"temp")]] <- data[[var]]
-  eval_data[[paste0(var,"temp")]] <- eval_data[[var]]
-
-  #standardize variable between 0 and 1
-  data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
-  eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))
-
-  ## Calculate density estimates
-  g1 <- ggplot(data, aes(x=data[[var]], group=target, colour=target)) +
-    geom_density(data = data) + xlim(min(data[[var]]), max(data[[var]]))
-  gg1 <- ggplot_build(g1)
-  # construct the dataset
-  x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
-  y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
-  y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
-  df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), side=(y1<y2), ydiff = y2-y1)
-  ##creating the second graph object
-  g3 <- ggplot(df2) +
-    geom_line(aes(x = x, y = ydiff, colour = side)) +
-    geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
-    guides(alpha = FALSE, fill = FALSE)
-
-  data$join <- data[[var]]
-  df2$join <- df2$x
-  temp <- difference_left_join(data,df2, by="join", max_dist =.001)
-  means <- aggregate(ydiff ~ temp$join.x, temp, mean)
-
-  colnames(means) <- c("std_var","density_diff")
-
-  new_data <- merge(means,data, by.x ="std_var", by.y =var)
-  new_data$std_var <- NULL
-  new_data$join <- NULL
-  new_data[new_column] <- new_data$density_diff
-  new_data$density_diff <- NULL
-  ###################same thing with eval data ############################
-  eval_data$join <- eval_data[[var]]
-  df2$join <- df2$x
-  temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
-  means <- aggregate(ydiff ~ temp$join.x, temp, mean)
-
-  ##new stuff
-  colnames(means) <- c("std_var","density_diff")
-  #data["key"] <- (var - min(var))/(max(var) - min(var))
-
-  eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
-  eval_data2$std_var <- NULL
-  eval_data2$join <- NULL
-  eval_data2[new_column] <- eval_data2$density_diff
-  eval_data2$density_diff <- NULL
-
-  eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
-  eval_data2[[paste0(var,"temp")]] <- NULL
-
-  new_data[[var]] <- new_data[[paste0(var,"temp")]]
-  new_data[[paste0(var,"temp")]] <- NULL
-
-  mylist <- list(g1,g3,new_data,eval_data2)
-  names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
-  return(mylist)
-}
-```
-
-### Creating Some New Variables
-
-```{r}
-str(train)
-str(test)
-## try it out 
-train <- density_diff(train,test,"nox",train$target,"nox_density")$new_data_training
-train <- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
-train <- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
-train <- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
-train <- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
-train <- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
-train <- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
-train <- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
-train <- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
-
-test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
-test<- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
-test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
-test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
-test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
-test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
-test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
-test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
-test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training
-
-str(train)
-str(test)
-```
-
-## Model Building
-
-We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.
-
-```{r}
-baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
-summary(baseline)
-test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
-cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
-results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
-                  deviance=baseline$deviance, 
-                  r2 = 1 - baseline$deviance/baseline$null.deviance,
-                  aic=baseline$aic)
-cm
-```
-
-Our baseline model is ok with an F1 score of `r cm$byClass[7]`.
-
-Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:
-
-```{r}
-fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
-summary(fullmodel)
-test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
-cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
-results <- rbind(results,tibble(model = "fullmodel",
-                                predictors = 12,F1 = cm$byClass[7],
-                                deviance=fullmodel$deviance, 
-                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
-                                aic=fullmodel$aic))
-cm
-```
-
-The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.
-
-After some backward elimination of non-significant predictor variables, we arrive at the following model:
-
-```{r}
-model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
-              family = binomial(link = "logit"), 
-              train)
-summary(model1)
-test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
-cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
-results <- rbind(results,tibble(model = "model1",
-                                predictors = 6,F1 = cm$byClass[7],
-                                deviance=model1$deviance, 
-                                r2 = 1-model1$deviance/model1$null.deviance,
-                                aic=model1$aic))
-cm
-```
-
-```{r}
-density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
-summary(density_models)
-test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
-test$density_models_yhat <- predict(density_models, test)
-cm <- confusionMatrix(factor(test$density_models), factor(test$target))
-results <- rbind(results,tibble(model = "density models",
-                                predictors = 9,
-                                F1 = cm$byClass[7],
-                                deviance = density_models$deviance,
-                                r2 = 1 - density_models$deviance / density_models$null.deviance,
-                                aic = density_models$aic))
-cm
-```
-
-Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.
-
-```{r}
-kable(results)
-```
-
-Looking these three models together, `model1` looks like the best one.
-Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.
-
-## Model Selection
-
-### pROC Output
-
-#### Baseline
-```{r}
-par(pty = "s")
-roc(test[["target"]], test[["baseline"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
-par(pty = "m")
-```
-
-#### Full Model
-```{r}
-par(pty = "s")
-roc(test[["target"]], test[["fullmodel"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
-par(pty = "m")
-```
-
-#### Model1
-```{r}
-par(pty = "s")
-roc(test[["target"]], test[["model1"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
-par(pty = "m")
-```
-
-#### Density Model
-```{r}
-par(pty = "s")
-roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
-par(pty = "m")
-```

### 2019-11-04 hw4/HW4.Rmd
Updated HW4

@@
 knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
+library(gridExtra)
 library(dplyr)
@@
 
-There are missing values for the `CAR_AGE` and `YOJ`(Years on Job) variables.  We will fill in the missing variables using KNN.
+There are missing values for the `CAR_AGE`, `YOJ`(Years on Job), and `AGE` variables.  We will fill in the missing variables using KNN.
 
@@
 df[impute_me,"YOJ"] <- knn[impute_me,"YOJ"]
+impute_me <- is.na(df$AGE)
+df[impute_me,"AGE"] <- knn[impute_me,"AGE"]
 ```

### 2019-11-04 hw4/HW4.Rmd
Update HW4

@@
 knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
-library(gridExtra)
 library(dplyr)
@@
 df <- read.csv('./data/insurance_training_data.csv') %>%
-  mutate(TARGET_FLAG = as.factor(TARGET_FLAG))
-evaluation <- read.csv("./data/insurance-evaluation-data.csv")
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
+  select(-INDEX)
+evaluation <- read.csv("./data/insurance-evaluation-data.csv") %>%
+  select(-INDEX)
 ```

### 2019-11-05 hw4/HW4.Rmd
Updated HW4

--- 2019-11-02 hw4/HW4.Rmd
@@
 ---
-
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
-library(dplyr)
+library(tidyverse)
 library(Amelia)
@@
 ```{r}
-df <- read.csv('./data/insurance_training_data.csv') %>%
-  mutate(TARGET_FLAG = as.factor(TARGET_FLAG))
-evaluation <- read.csv("./data/insurance-evaluation-data.csv")
+df <- read.csv("./data/insurance_training_data.csv") %>%
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
+  select(-INDEX)
+evaluation <- read.csv("./data/insurance-evaluation-data.csv") %>%
+  select(-INDEX)
 ```
+
+## Introduction
+
+We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.
+Each record has two response variables.
+The first is a binary flag where one means a person was in a car crash and zero means they were not.
+There are `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` rows with a one flag and `r df %>% filter(TARGET_FLAG == 0) %>% nrow()` with a zero flag.
+The first objective is to logistic regression classifier to predict if a person was in a car crash.
+The second sreponse variable is the amount it will cost if the person crashes their car.
+The value is zero if the person did not crash their car.
 
@@
 
-### Are There Missing Values?
-First we look at the data to see if any variables have missing data:
+## Data Preparation
+
+```{r}
+df %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Fix Data Types
+
+There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.
+
+```{r}
+strip_dollars <- function(x){
+  x <- as.character(x)
+  x <- gsub(",", "", x)
+  x <- gsub("\\$", "", x)
+  as.numeric(x)
+}
+
+fix_data_types <- function(messy_df){
+  messy_df %>%
+    rowwise() %>%
+    mutate(INCOME = strip_dollars(INCOME),
+           HOME_VAL = strip_dollars(HOME_VAL),
+           BLUEBOOK = strip_dollars(BLUEBOOK),
+           OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
+    ungroup()
+}
+
+df <- fix_data_types(df)
+evaluation <- fix_data_types(evaluation)
+```
 
@@
 
-There are missing values for the `CAR_AGE` and `YOJ`(Years on Job) variables.  We will fill in the missing variables using KNN.
+```{r}
+df %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Fix Missing Values
+There are missing values.  We will fill in the missing variables using KNN.
+
+```{r, null_prefix = TRUE}
+set.seed(42)
+#knn <- df %>%
+#  select(-TARGET_FLAG, -TARGET_AMT) %>%
+#  na.omit() %>%
+#  knnImputation()
+
+for(var in names(df)){
+  impute_me <- is.na(df[[var]])
+  if (nrow(df[impute_me,]) > 0){
+    print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
+    #df[impute_me, var] <- knn[impute_me, var] 
+  }
+}
+```
+
+### Feature Creation
+
+We will create a log transformed income and home value feature.  We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.
 
 ```{r}
-set.seed(42)
-knn <- knnImputation(df)
-impute_me <- is.na(df$CAR_AGE)
-df[impute_me,"CAR_AGE"] <- knn[impute_me,"CAR_AGE"] 
-impute_me <- is.na(df$YOJ)
-df[impute_me,"YOJ"] <- knn[impute_me,"YOJ"] 
+feature_creation <- function(d){
+  d %>%
+    rowwise() %>%
+    mutate(LOG_INCOME = log(INCOME + 1),
+           LOG_HOME_VAL = log(HOME_VAL + 1),
+           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0)) %>%
+    ungroup()
+}
+
+df <- feature_creation(df)
+evaluation <- feature_creation(evaluation)
 ```
 
-### Splitting the Data
-Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
+### Creating Training/Test Data Sets
+
+Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`). We'll use a 70-30 split between train and test, respectively.
 
@@
 ```
-
-### First Look at the Data
-
-We will now take our first look at the data to get a sense of what we have.
-
-```{r}
-train %>%
-  summary() %>%
-  kable() %>%
-  kable_styling()
-```

### 2019-11-05 final project/Final_Project.Rmd
Initial commit of final project framework

---
title: "DATA 621 - Final Project"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Use 250 words or less to summarize your problem, methodology, and major outcomes.

**Key Words: Regression**

## Introduction

Describe the background and motivation of your problem.

## Literature Review

Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-theart. Please cite the relevant papers where appropriate.

## Methodology

Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

## Experimentation and Results

Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).

## Discussion and Conclusions

Conclude your findings, limitations, and suggest areas for future work.

## References

Be sure to cite all references used in the report (APA format).

## Appendices

- Supplemental tables and/or figures.
- R statistical programming code.

### 2019-11-05 final project/Final_Project.Rmd
Update final

@@
 title: "DATA 621 - Final Project"
-author: "Critical Thinking Group 3"
+author:
+  - name: Corey Arnouts
+    affiliation: CUNY School of Professional Studies
+  - name: Adam Douglas
+    affiliation: CUNY School of Professional Studies
+  - name: Jason Givens-Doyle
+    affiliation: CUNY School of Professional Studies
+  - name: Michael Silva
+    affiliation: CUNY School of Professional Studies
+    email: Michael.Silva39@spsmail.cuny.edu
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  html_document: default
+  rticles::asa_article:
+    citation_package: natbib
+    keep_tex: yes
+    number_sections: no
+    toc: no
+keywords: keywordA, keywordB
+abstract: Use 250 words or less to summarize your problem, methodology, and major outcomes. \newline This is the second line of abstract.
 ---
@@
 ```
-
-## Abstract
-
-Use 250 words or less to summarize your problem, methodology, and major outcomes.
-
-**Key Words: Regression**
 

### 2019-11-05 final project/Final_Project.tex
Initial commit

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf DATA 621 - Final Project}

  \author{
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf DATA 621 - Final Project}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Use 250 words or less to summarize your problem, methodology, and major
outcomes. \newline This is the second line of abstract.
\end{abstract}

\noindent%
{\it Keywords:} keywordA, keywordB
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Describe the background and motivation of your problem.

\hypertarget{literature-review}{%
\subsection{Literature Review}\label{literature-review}}

Discuss how other researchers have addressed similar problems, what
their achievements are, and what the advantage and drawbacks of each
reviewed approach are. Explain how your investigation is similar or
different to the state-of-theart. Please cite the relevant papers where
appropriate.

\hypertarget{methodology}{%
\subsection{Methodology}\label{methodology}}

Discuss the key aspects of your problem, data set and regression
model(s). Given that you are working on real-world data, explain at a
high-level your exploratory data analysis, how you prepared the data for
regression modeling, your process for building regression models, and
your model selection.

\hypertarget{experimentation-and-results}{%
\subsection{Experimentation and
Results}\label{experimentation-and-results}}

Describe the specifics of what you did (data exploration, data
preparation, model building, model selection, model evaluation, etc.),
and what you found out (statistical analyses, interpretation and
discussion of the results, etc.).

\hypertarget{discussion-and-conclusions}{%
\subsection{Discussion and
Conclusions}\label{discussion-and-conclusions}}

Conclude your findings, limitations, and suggest areas for future work.

\hypertarget{references}{%
\subsection{References}\label{references}}

Be sure to cite all references used in the report (APA format).

\hypertarget{appendices}{%
\subsection{Appendices}\label{appendices}}

\begin{itemize}
\tightlist
\item
  Supplemental tables and/or figures.
\item
  R statistical programming code.
\end{itemize}

\bibliographystyle{agsm}
\bibliography{}

\end{document}

### 2019-11-05 final project/Manuscript.Rmd
Initial commit of final project manuscript files

---
title             : "Fall 2019 DATA 621 Final Project"
shorttitle        : "Arnouts et al. Final Project"
author: 
  - name          : "Corey Arnouts"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "119 W 31st St., New York, NY 10001"
    email         : "Corey.Arnouts@spsmail.cuny.edu"
  - name          : "Adam Douglas"
    affiliation   : "1"
  - name          : "Jason Givens-Doyle"
    affiliation   : "1"
  - name          : "Michael Silva"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "CUNY School of Professional Studies"
authornote: |
  MS in Data Science Students
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->
keywords          : "CoIL Challenge, Logistic Regression"
wordcount         : "X"
bibliography      : ["Manuscript.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output:
  papaja::apa6_pdf
---

```{r setup, include = FALSE}
# Documentation at https://crsh.github.io/papaja_man/
# If you have no use for TeX beyond rendering R Markdown documents, I recommend you use TinyTex.
# install.packages("tinytex")
# tinytex::install_tinytex()
if(!"papaja" %in% rownames(installed.packages())) devtools::install_github("crsh/papaja")
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction
```{r eval=FALSE, echo=FALSE}
#Describe the background and motivation of your problem.
```
Busineses use data science to extract insight from data.  It has many practical business applications.  The CoIL Challenge showcases the power data science can bring to bear on fundamental business problems.

The CoIL Challenge was a datamining competition organized by the the Computational Intelligence and Learning Cluster. It was held in the period of March-May 2000.  The COIL Challenge had two tasks:

- Predict which customers are potentially interested in a caravan insurance policy; and
- Describe the actual or potential customers; and possibly explain why these customers buy a caravan policy.

In total 43 solutions were submitted. The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.

In this paper we set out to complete the COIL Challenge ourselves using a logistic regression classifier. **SUMARISE FINDINGS**

# Literature Review

Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-theart. Please cite the relevant papers where appropriate.

# Methodology

Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

# Experimentation and Results

Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).

We used `r cite_r("Manuscript.bib")` for all our analyses.

# Discussion and Conclusions

Conclude your findings, limitations, and suggest areas for future work.

\newpage

# References
```{r create_references}
r_refs(file = "Manuscript.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

\newpage

# Appendices

- Supplemental tables and/or figures.
- R statistical programming code.

### 2019-11-05 final project/Manuscript.tex
Initial commit of final project manuscript files

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Fall 2019 DATA 621 Final Project},
            pdfauthor={Corey Arnouts, Adam Douglas, Jason Givens-Doyle, \& Michael Silva},
            pdfkeywords={CoIL Challenge, Logistic Regression},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Fall 2019 DATA 621 Final Project}
    \author{Corey Arnouts\textsuperscript{1}, Adam Douglas\textsuperscript{1}, Jason Givens-Doyle\textsuperscript{1}, \& Michael Silva\textsuperscript{1}}
    \date{}

\shorttitle{Arnouts et al. Final Project}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} CUNY School of Professional Studies}
\keywords{CoIL Challenge, Logistic Regression\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

\authornote{MS in Data Science Students

Correspondence concerning this article should be addressed to Corey Arnouts, 119 W 31st St., New York, NY 10001. E-mail: \href{mailto:Corey.Arnouts@spsmail.cuny.edu}{\nolinkurl{Corey.Arnouts@spsmail.cuny.edu}}}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.

}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Busineses use data science to extract insight from data. It has many practical business applications. The CoIL Challenge showcases the power data science can bring to bear on fundamental business problems.

The CoIL Challenge was a datamining competition organized by the the Computational Intelligence and Learning Cluster. It was held in the period of March-May 2000. The COIL Challenge had two tasks:

\begin{itemize}
\tightlist
\item
  Predict which customers are potentially interested in a caravan insurance policy; and
\item
  Describe the actual or potential customers; and possibly explain why these customers buy a caravan policy.
\end{itemize}

In total 43 solutions were submitted. The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.

In this paper we set out to complete the COIL Challenge ourselves using a logistic regression classifier. \textbf{SUMARISE FINDINGS}

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-theart. Please cite the relevant papers where appropriate.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

\hypertarget{experimentation-and-results}{%
\section{Experimentation and Results}\label{experimentation-and-results}}

Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).

We used R (Version 3.6.1; R Core Team, 2019) and the R-package \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018) for all our analyses.

\hypertarget{discussion-and-conclusions}{%
\section{Discussion and Conclusions}\label{discussion-and-conclusions}}

Conclude your findings, limitations, and suggest areas for future work.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\endgroup

\newpage

\hypertarget{appendices}{%
\section{Appendices}\label{appendices}}

\begin{itemize}
\tightlist
\item
  Supplemental tables and/or figures.
\item
  R statistical programming code.
\end{itemize}

\end{document}

### 2019-11-05 final project/Manuscript.Rmd
Updated final - Added HTML output

@@
 output:
-  papaja::apa6_pdf
+  papaja::apa6_pdf:
+    citation_package: biblatex
+  html_document: default
 ---

### 2019-11-05 final project/Manuscript.Rmd
Updated final

--- 2019-11-05 final project/Manuscript.Rmd
@@
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : no
@@
 output:
-  papaja::apa6_pdf
+  papaja::apa6_pdf:
+    #citation_package: natbib #biblatex
+  #html_document: default
 ---

### 2019-11-06 hw4/HW4.Rmd
Update HW4.Rmd

@@
 library(DMwR)
+library(scales)
+library(RColorBrewer)
 # Thank you Stack Overflow!
@@
 There are `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` rows with a one flag and `r df %>% filter(TARGET_FLAG == 0) %>% nrow()` with a zero flag.
-The first objective is to logistic regression classifier to predict if a person was in a car crash.
-The second sreponse variable is the amount it will cost if the person crashes their car.
+
+```{r, echo=FALSE}
+df %>% 
+  ggplot(aes(x=TARGET_FLAG,fill=TARGET_FLAG)) +
+  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  labs(x="TARGET_FLAG", y="# Observations")
+```
+
+The first objective is to train a logistic regression classifier to predict if a person was in a car crash.
+The second reponse variable is the amount it will cost if the person crashes their car.
 The value is zero if the person did not crash their car.
 
+The second objective will be to train a regression model to predict the cost of a craash, if one occurred.
+
+```{r, echo=FALSE}
+df %>% filter(TARGET_FLAG == 1) %>%
+  ggplot(aes(x=TARGET_AMT)) + geom_density() +
+  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
+  geom_label(aes(x=mean(TARGET_AMT),y=1,label="mu"),parse=T) +
+  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
+  geom_label(aes(x=median(TARGET_AMT),y=.5,label="median")) +
+  scale_x_log10(labels=comma) + theme_light() +
+  labs(title="TARGET_AMT Density Plot", caption="x-axis is log 10 scale",
+       y="Density", x="LOG(TARGET_AMT)")
+```
+
+Looking at the distribution of the `TARGET_AMT` variable, we can see that the variable is considerably right-skewed. Thus, a LOG transform might be best here.
+
 ## Data Exploration
+
+Looking at the data, we can see that there are some missing values:
+
+```{r}
+missmap(df, main = "Missing vs Observed Values")
+```
+
+Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, but only `r df %>% filter(is.na(YOJ),is.na(CAR_AGE)) %>% nrow()` observations missing both of these varables.
+So, if imputation is not viable here, we will lose as many as 488 observations out of the 8161 in our training set (or about 6%).
 
@@
 
-```{r}
-missmap(df, main = "Missing vs Observed Values")
-```
+Now that we have fixed the variables, we can look at a sumamry of the data:
 
@@
 set.seed(42)
-#knn <- df %>%
-#  select(-TARGET_FLAG, -TARGET_AMT) %>%
-#  na.omit() %>%
-#  knnImputation()
+knn <- df %>%
+  select(-TARGET_FLAG, -TARGET_AMT, is.numeric()) %>%
+  #na.omit() %>%
+  knnImputation()
 

### 2019-11-06 hw4/HW4.Rmd
More work on HW4

--- 2019-11-05 hw4/HW4.Rmd
@@
 library(tidyverse)
+require(gridExtra)
 library(Amelia)
@@
   select(-INDEX)
+
+strip_dollars <- function(x){
+  x <- as.character(x)
+  x <- gsub(",", "", x)
+  x <- gsub("\\$", "", x)
+  as.numeric(x)
+}
+
+hw4_plot <- function(df, var, strip_dollar_signs = FALSE){
+  if(strip_dollar_signs){
+    df$X <- strip_dollars(df[[var]])
+  } else {
+    df$X <- df[[var]]
+  }
+  plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
+    geom_jitter() +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+
+  plot_2 <- ggplot(df, aes(X, TARGET_AMT, color = TARGET_FLAG)) +
+    geom_point() +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+
+  grid.arrange(plot_1, plot_2, ncol=2)
+}
+
+hw4_categorical_plot <- function(df, var, strip_dollar_signs = FALSE){
+  if(strip_dollar_signs){
+    df$X <- strip_dollars(df[[var]])
+  } else {
+    df$X <- df[[var]]
+  }
+  plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
+    geom_jitter() +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+
+  plot_2 <- df %>%
+    filter(TARGET_FLAG == 1) %>%
+    ggplot(., aes(X, TARGET_AMT)) +
+      geom_boxplot() +
+      theme_minimal() +
+      theme(legend.position = "none",
+            axis.title.x = element_blank(),
+            axis.title.y = element_blank())
+
+  grid.arrange(plot_1, plot_2, ncol=2)
+}
 ```
@@
 
-## Data Preparation
+We will first look at the summary statistics for the data
 
@@
 
+There are some missing values that we will need to deal with.
+Let's look at the predictors to see how well they predict if there is an accident (plot on the left) and how much it would cost (plot on the right).
+Numeric variables will get scatterplots, while categorical will get box plots
+
+### AGE
+
+**Theoretical Effect**
+Very young people tend to be risky.  Maybe very old people also.
+
+```{r} 
+hw4_plot(df, "AGE")
+```
+
+**Observed Effect**
+There is very little information encapsulated in the age of the person.
+
+### BLUEBOOK
+
+**Theoretical Effect**
+Unknown effect on probability of collision, but probably effects the payout if there is a crash.
+
+```{r} 
+hw4_plot(df, "BLUEBOOK", TRUE)
+```
+
+**Observed Effect**
+The plot suggests that people with low value and high value vehicles are equally likely to be in a crash.  It also suggests the payout is not strongly correlated with the vehicle's value.
+
+### CAR_AGE
+
+**Theoretical Effect**
+Unknown effect on probability of collision, but probably effects the payout if there is a crash.
+
+```{r} 
+hw4_plot(df, "CAR_AGE")
+```
+
+**Observed Effect**
+The plot suggests that people with old or new vehicles are equally likely to be in a crash.
+It also suggests the payout is not strongly correlated with the vehicle's age.
+*There is an observation with a negative value that will need to be cleaned up.*
+
+### CAR_TYPE
+
+**Theoretical Effect**
+Unknown effect on probability of collision, but probably effects the payout if there is a crash.
+
+```{r} 
+hw4_categorical_plot(df, "CAR_TYPE")
+```
+
+**Observed Effect**
+The plot suggests that the type of vehicle has no effect on the probability of being in an accident, and little effect of the cost.
+
+## Data Preparation
+
 ### Fix Data Types
@@
 ```{r}
-strip_dollars <- function(x){
-  x <- as.character(x)
-  x <- gsub(",", "", x)
-  x <- gsub("\\$", "", x)
-  as.numeric(x)
-}
-
 fix_data_types <- function(messy_df){

### 2019-11-06 hw4/HW4.Rmd
Folded in some visualizations

@@
 There are `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` rows with a one flag and `r df %>% filter(TARGET_FLAG == 0) %>% nrow()` with a zero flag.
-The first objective is to logistic regression classifier to predict if a person was in a car crash.
-The second sreponse variable is the amount it will cost if the person crashes their car.
+
+```{r, echo=FALSE}
+df %>% 
+  ggplot(aes(x=TARGET_FLAG,fill=TARGET_FLAG)) +
+  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  labs(x="TARGET_FLAG", y="# Observations")
+```
+
+The first objective is to train a logistic regression classifier to predict if a person was in a car crash.
+The second reponse variable is the amount it will cost if the person crashes their car.
 The value is zero if the person did not crash their car.
+
+The second objective will be to train a regression model to predict the cost of a crash, if one occurred.
+
+```{r, echo=FALSE}
+df %>% filter(TARGET_FLAG == 1) %>%
+  ggplot(aes(x=TARGET_AMT)) + geom_density() +
+  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
+  geom_label(aes(x=mean(TARGET_AMT),y=1,label="mu"),parse=T) +
+  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
+  geom_label(aes(x=median(TARGET_AMT),y=.5,label="median")) +
+  scale_x_log10(labels=comma) + theme_light() +
+  labs(title="TARGET_AMT Density Plot", caption="x-axis is log 10 scale",
+       y="Density", x="LOG(TARGET_AMT)")
+```
+
+Looking at the distribution of the `TARGET_AMT` variable, we can see that the variable is considerably right-skewed. Thus, a LOG transform might be best here.
 
@@
 
+Looking at the data, we can see that there are some missing values:
+
+```{r}
+missmap(df, main = "Missing vs Observed Values")
+```
+
+Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, but only `r df %>% filter(is.na(YOJ),is.na(CAR_AGE)) %>% nrow()` observations missing both of these varables.
+So, if imputation is not viable here, we will lose as many as 488 observations out of the 8161 in our training set (or about 6%).
+
 ## Data Preparation
+
+```{r}
+df %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
 
@@
 ```{r}
+strip_dollars <- function(x){
+  x <- as.character(x)
+  x <- gsub(",", "", x)
+  x <- gsub("\\$", "", x)
+  as.numeric(x)
+}
+
 fix_data_types <- function(messy_df){
@@
 ```{r, null_prefix = TRUE}
-set.seed(42)
-#knn <- df %>%
-#  select(-TARGET_FLAG, -TARGET_AMT) %>%
-#  na.omit() %>%
+# set.seed(42)
+# knn <- df %>%
+#   select(-TARGET_FLAG, -TARGET_AMT, is.numeric()) %>%
+#   #na.omit() %>%
 #   knnImputation()
-
-for(var in names(df)){
-  impute_me <- is.na(df[[var]])
-  if (nrow(df[impute_me,]) > 0){
-    print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
-    #df[impute_me, var] <- knn[impute_me, var] 
-  }
-}
+# 
+# for(var in names(df)){
+#   impute_me <- is.na(df[[var]])
+#   if (nrow(df[impute_me,]) > 0){
+#     print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
+#     #df[impute_me, var] <- knn[impute_me, var] 
+#   }
+# }
 ```

### 2019-11-06 final project/Manuscript.Rmd
More work on final

--- 2019-11-05 final project/Manuscript.Rmd
@@
 ---
-title             : "Fall 2019 DATA 621 Final Project"
-shorttitle        : "Arnouts et al. Final Project"
+title             : "CoIL Challenge Submission"
+shorttitle        : "CoIL Challenge Submission"
 author: 
@@
 output:
-  papaja::apa6_pdf
+  papaja::apa6_pdf:
+    #citation_package: natbib #biblatex
+  html_document: default
 ---
@@
 # Introduction
+
 ```{r eval=FALSE, echo=FALSE}
@@
 It has many practical business applications.
-The CoIL Challenge showcases the power data science can bring to bear on fundamental business problems.
+The Computational Intelligence and Learning (CoIL) Challenge showcases the power data science can bring to bear on fundamental business problems.
 
-The CoIL Challenge was a datamining competition organized by the the Computational Intelligence and Learning Cluster. It was held in the period of March-May 2000.  The COIL Challenge had two tasks:
+The CoIL Challenge was a datamining competition held from March 17 to May 8 in 2000.  Participants could:
 
@@
 
-In total 43 solutions were submitted. The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.
+In total 147 participants registered and 43 submitted solutions [@Putten-2000].
+The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.
 
-In this paper we set out to complete the COIL Challenge ourselves using a logistic regression classifier. **SUMARISE FINDINGS**
+In this paper we set out to complete the prediction part of the COIL Challenge ourselves using a logistic regression classifier. **SUMARISE FINDINGS**
 
@@
 Please cite the relevant papers where appropriate.
+
+The maximum number of policyowners that could be found was 238, the winning model selected 121 policy owners.
+Random selection results in 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks)and 118 (linear!) policy owners.
+[@Putten-2000]
 
@@
 - Supplemental tables and/or figures.
-- R statistical programming code.
+
+## R statistical programming code.
+
+```{r source, code=readLines("CoIL.R")}
+```

### 2019-11-06 final project/Manuscript.tex
More work on final

@@
 \hypersetup{unicode=true,
-            pdftitle={Fall 2019 DATA 621 Final Project},
+            pdftitle={CoIL Challenge Submission},
             pdfauthor={Corey Arnouts, Adam Douglas, Jason Givens-Doyle, \& Michael Silva},
@@
 
-  \title{Fall 2019 DATA 621 Final Project}
+  \title{CoIL Challenge Submission}
     \author{Corey Arnouts\textsuperscript{1}, Adam Douglas\textsuperscript{1}, Jason Givens-Doyle\textsuperscript{1}, \& Michael Silva\textsuperscript{1}}
@@
 
-\shorttitle{Arnouts et al. Final Project}
+\shorttitle{CoIL Challenge Submission}
 \affiliation{
@@
 \makeatother
-\usepackage{lineno}
-
-\linenumbers
 
@@
 It has many practical business applications.
-The CoIL Challenge showcases the power data science can bring to bear on fundamental business problems.
+The Computational Intelligence and Learning (CoIL) Challenge showcases the power data science can bring to bear on fundamental business problems.
 
-The CoIL Challenge was a datamining competition organized by the the Computational Intelligence and Learning Cluster. It was held in the period of March-May 2000. The COIL Challenge had two tasks:
+The CoIL Challenge was a datamining competition held from March 17 to May 8 in 2000. Participants could:
 
@@
 
-In total 43 solutions were submitted. The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.
+In total 147 participants registered and 43 submitted solutions (Putten, Ruiter, \& Someren, 2000).
+The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.
 
-In this paper we set out to complete the COIL Challenge ourselves using a logistic regression classifier. \textbf{SUMARISE FINDINGS}
+In this paper we set out to complete the prediction part of the COIL Challenge ourselves using a logistic regression classifier. \textbf{SUMARISE FINDINGS}
 
@@
 Please cite the relevant papers where appropriate.
+
+The maximum number of policyowners that could be found was 238, the winning model selected 121 policy owners.
+Random selection results in 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks)and 118 (linear!) policy owners.
+(Putten et al., 2000)
 
@@
 
-We used R (Version 3.6.1; R Core Team, 2019) and the R-package \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018) for all our analyses.
+We used R (Version 3.6.1; R Core Team, 2019) and the R-package \emph{\}papaja} {[}@\}R-papaja{]} for all our analyses.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-R-papaja}{}%
-Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
+\leavevmode\hypertarget{ref-Putten-2000}{}%
+Putten, P., Ruiter, M., \& Someren, M. (2000). CoIL challenge 2000 tasks and results: Predicting and explaining caravan policy ownership.
 
@@
   Supplemental tables and/or figures.
-\item
-  R statistical programming code.
 \end{itemize}
 
+\hypertarget{r-statistical-programming-code.}{%
+\subsection{R statistical programming code.}\label{r-statistical-programming-code.}}
+
 \end{document}

### 2019-11-06 final project/CoIL.R

@@ -1,0 +1,34 @@
+# CoIL Challenge Source Code
+library(tidyverse)
+# Download the data sets from UCI if they are not present
+url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/tic-mld/"
+for (file_name in c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt")) {
+  file_path <- paste0("data/", file_name)
+  file_url <- paste0(url, file_name)
+  if (!file.exists(file_path)) {
+    message(paste("Downloading", file_name))
+    download.file(file_url, file_path)
+  }
+}
+# Read in the data
+df <- read.delim("data/ticdata2000.txt", header = FALSE)
+names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
+               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
+               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
+               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
+               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
+               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
+               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
+               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
+               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
+               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
+               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
+               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
+               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
+               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
+               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
[…]

### 2019-11-06 hw4/HW4.Rmd
Added some color to the visualizations

@@
     geom_jitter() +
+    scale_color_brewer(palette="Set1") +
     theme_minimal() +
@@
     geom_point() +
+    scale_color_brewer(palette="Set1") +
     theme_minimal() +
@@
   } else {
-    df$X <- df[[var]]
+    df$X <- as.factor(as.character(df[[var]]))
   }
@@
     filter(TARGET_FLAG == 1) %>%
-    ggplot(., aes(X, TARGET_AMT)) +
+    ggplot(., aes(X, TARGET_AMT, color = X)) +
       geom_boxplot() +
+      scale_color_brewer(palette="Dark2") + #Paired
       theme_minimal() +
@@
 
+### CAR_USE
+
+**Theoretical Effect**
+Commercian vehicles are driven more, so might increase probability of collision.
+
+```{r} 
+hw4_categorical_plot(df, "CAR_USE")
+```
+
+**Observed Effect**
+The plot suggests that the use does not increase or decrease the probability of being in an accident or the amount of the claim.
+
+### CLM_FREQ
+
+**Theoretical Effect**
+The more claims you filed in the past, the more you are likely to file in the future.
+
+```{r} 
+hw4_categorical_plot(df, "CLM_FREQ")
+```
+
+**Observed Effect**
+Looks like there are few people who signed up for the frequent accident club.  All jokes aside, the theoretical effect is not observed in the data.
+
+### EDUCATION
+
+**Theoretical Effect**
+Unknown effect, but in theory more educated people tend to drive more safely
+```{r} 
+hw4_categorical_plot(df, "EDUCATION")
+```
+
+**Observed Effect**
+The theoretical effect has no basis and is somewhat elitist.
+
 Looking at the data, we can see that there are some missing values:

### 2019-11-07 hw4/HW4.Rmd
More work on HW4

@@
   }
-  plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
+  plot_1_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
     geom_jitter() +
+    scale_color_brewer(palette="Set1") +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+
+  plot_1 <- ggplot(df, aes(X, color = TARGET_FLAG)) +
+    geom_density() + 
     scale_color_brewer(palette="Set1") +
@@
   }
-  plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
+  plot_1_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
     geom_jitter() +
@@
           axis.title.y = element_blank())
-
+  plot_1 <- df %>%
+    group_by(X, TARGET_FLAG) %>%
+    tally() %>%
+    ggplot(.) +
+    geom_bar(aes(X, n, fill = TARGET_FLAG), stat="identity") +
+    scale_fill_brewer(palette="Set1") +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
   plot_2 <- df %>%
@@
   grid.arrange(plot_1, plot_2, ncol=2)
+}
+
+hw4_target_table <- function(df, var, strip_dollar_signs = FALSE){
+  if(strip_dollar_signs){
+    df$X <- strip_dollars(df[[var]])
+  } else {
+    df$X <- df[[var]]
+  }
+  df %>%
+    group_by(X, TARGET_FLAG) %>%
+    tally() %>%
+    spread(TARGET_FLAG, n) %>%
+    rename(Variable = X) %>%
+    kable() %>%
+    kable_styling()
 }
@@
 hw4_categorical_plot(df, "CAR_TYPE")
+hw4_target_table(df, "CAR_TYPE")
 ```
@@
 hw4_categorical_plot(df, "CAR_USE")
+hw4_target_table(df, "CAR_USE")
 ```
@@
 hw4_categorical_plot(df, "CLM_FREQ")
+hw4_target_table(df, "CLM_FREQ")
 ```
@@
 Unknown effect, but in theory more educated people tend to drive more safely
+
 ```{r} 
 hw4_categorical_plot(df, "EDUCATION")
+hw4_target_table(df, "EDUCATION")
 ```
@@
 The theoretical effect has no basis and is somewhat elitist.
+
+### HOMEKIDS
+
+**Theoretical Effect**
+Unknown effect
+
+```{r} 
+hw4_categorical_plot(df, "HOMEKIDS")
+hw4_target_table(df, "HOMEKIDS")
+```
+
+**Observed Effect**
+It's interesting to note that those with kids have a 0.34 probability of being in an accident.  Those without kids have a 0.22 probability of being in an accident.  This may be useful.
+
+### HOME_VAL
+
+**Theoretical Effect**
+In theory, home owners tend to drive more responsibly.
+
+```{r} 
+hw4_plot(df, "HOME_VAL", TRUE)
+```
+
+**Observed Effect**
+There are a lot of zero values.  I am assuming that it is a zero because they are renters.
+
+### INCOME
+
+**Theoretical Effect**
+In theory, rich people trend to get into fewer crashes.
+
+```{r} 
+hw4_plot(df, "INCOME", TRUE)
+```
+
+**Observed Effect**
+There are a lot of zero values.  I am assuming that it is a zero because they are renters.
+
+### JOB
+
+**Theoretical Effect**
+In theory, white collar jobs tend to be safer
+
+```{r} 
+hw4_categorical_plot(df, "JOB")
+hw4_target_table(df, "JOB")
+```
+
+**Observed Effect**
+There is a 0.21 probability for white collar job holders being in an accident.  there's a 0.34 probability for the blue collar jobs.  There may be some truth to the theoretical effect.
+
+### KIDSDRIV
+
+**Theoretical Effect**
+When teenagers drive your car, ou are more likely to get into crashes
+
+```{r} 
+hw4_categorical_plot(df, "KIDSDRIV")
+hw4_target_table(df, "KIDSDRIV")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
 

### 2019-11-07 hw4/HW4.Rmd
A bit more work

@@
 
+### MSTATUS
+
+**Theoretical Effect**
+In theory, married people drive more safely
+
+```{r} 
+hw4_categorical_plot(df, "MSTATUS")
+hw4_target_table(df, "MSTATUS")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### MVR_PTS
+
+**Theoretical Effect**
+If you get lots of traffic tickets, you trend to get into more crashes
+
+```{r} 
+hw4_categorical_plot(df, "MVR_PTS")
+hw4_target_table(df, "MVR_PTS")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
 Looking at the data, we can see that there are some missing values:

### 2019-11-08 hw4/HW4.Rmd
Imputing values function added and added some density variables

@@
 # 
-# for(var in names(df)){
-#   impute_me <- is.na(df[[var]])
-#   if (nrow(df[impute_me,]) > 0){
-#     print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
-#     #df[impute_me, var] <- knn[impute_me, var] 
-#   }
-# }
+for(var in names(df)){
+  impute_me <- is.na(df[[var]])
+  if (nrow(df[impute_me,]) > 0){
+    print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
+    #df[impute_me, var] <- knn[impute_me, var] 
+  }
+}
+
+## adding in some imputing logic
+library(VIM)
+library(laeken)
+
+df[,3:ncol(df)] <- df[,3:ncol(df)] %>% kNN(numFun = weightedMean, weightDist=TRUE)
 ```
@@
 evaluation <- feature_creation(evaluation)
+
+df$TARGET_FLAG_BOOL <- ifelse(df$TARGET_FLAG=="1",1,0)
 ```
@@
 ```
+
+```{r}
+library(lme4)
+glm <- glm(TARGET_AMT ~ .-TARGET_FLAG-TARGET_FLAG_BOOL , data= train, family=poisson) 
+summary(glm)
+
+test$Prediction =predict.glm(glm, newdata=test, type="response")
+
+cor(test$TARGET_AMT, test$Prediction)
+
+glm2 <- glm(TARGET_AMT ~ .-TARGET_FLAG-TARGET_FLAG_BOOL , data= train) 
+summary(glm2)
+
+test$Prediction2 =predict.glm(glm2, newdata=test, type="response")
+
+cor(test$TARGET_AMT, test$Prediction)
+
+glm3 <- glm(TARGET_AMT ~ AGE+PARENT1+MSTATUS+JOB+BLUEBOOK+TIF+CAR_AGE_density+URBANICITY+MVR_PTS+REVOKED+CLM_FREQ+OLDCLAIM , data= train, family=poisson) 
+summary(glm3)
+
+test$Prediction =predict.glm(glm3, newdata=test, type="response")
+
+cor(test$TARGET_AMT, test$Prediction)
+```
+
+```{r}
+source("https://raw.githubusercontent.com/crarnouts/Corey-s_Scripts_For_Reference/master/11_5_19_Density_Diff_Function")
+```
+
+### Creating Some New Variables
+
+```{r}
+
+## try it out 
+train <- density_diff(train,test,"INCOME",train$TARGET_FLAG_BOOL,"income_density")$new_data_training
+test <- density_diff(train,test,"INCOME",train$TARGET_FLAG_BOOL,"income_density")$new_data_eval
+
+density_diff(train,test,"INCOME",train$TARGET_FLAG_BOOL,"income_density")$dist_diff
+density_diff(train,test,"INCOME",train$TARGET_FLAG_BOOL,"income_density")$dist
+
+train <- density_diff(train,test,"CAR_AGE",train$TARGET_FLAG_BOOL,"CAR_AGE_density")$new_data_training
+test <- density_diff(train,test,"CAR_AGE",train$TARGET_FLAG_BOOL,"CAR_AGE_density")$new_data_eval
+
+density_diff(train,test,"CAR_AGE",train$TARGET_FLAG_BOOL,"CAR_AGE_density")$dist_diff
+density_diff(train,test,"CAR_AGE",train$TARGET_FLAG_BOOL,"CAR_AGE_density")$dist
+
+train <- density_diff(train,test,"TRAVTIME",train$TARGET_FLAG_BOOL,"TRAVTIME_density")$new_data_training
+test <- density_diff(train,test,"TRAVTIME",train$TARGET_FLAG_BOOL,"TRAVTIME_density")$new_data_eval
+
+density_diff(train,test,"TRAVTIME",train$TARGET_FLAG_BOOL,"TRAVTIME_density")$dist_diff
+density_diff(train,test,"TRAVTIME",train$TARGET_FLAG_BOOL,"TRAVTIME_density")$dist
+
+```
+
+##Test out Some Models
+
+```{r}
+
+```

### 2019-11-08 hw4/HW4.Rmd
added some predictions

@@
 ```{r}
-library(lme4)
-glm <- glm(TARGET_AMT ~ .-TARGET_FLAG-TARGET_FLAG_BOOL , data= train, family=poisson) 
-summary(glm)
-
-test$Prediction =predict.glm(glm, newdata=test, type="response")
-
-cor(test$TARGET_AMT, test$Prediction)
-
-glm2 <- glm(TARGET_AMT ~ .-TARGET_FLAG-TARGET_FLAG_BOOL , data= train) 
-summary(glm2)
-
-test$Prediction2 =predict.glm(glm2, newdata=test, type="response")
-
-cor(test$TARGET_AMT, test$Prediction)
-
-glm3 <- glm(TARGET_AMT ~ AGE+PARENT1+MSTATUS+JOB+BLUEBOOK+TIF+CAR_AGE_density+URBANICITY+MVR_PTS+REVOKED+CLM_FREQ+OLDCLAIM , data= train, family=poisson) 
-summary(glm3)
-
-test$Prediction =predict.glm(glm3, newdata=test, type="response")
-
-cor(test$TARGET_AMT, test$Prediction)
-```
-
-```{r}
 source("https://raw.githubusercontent.com/crarnouts/Corey-s_Scripts_For_Reference/master/11_5_19_Density_Diff_Function")
@@
 ```{r}
+library(lme4)
+glm <- glm(TARGET_AMT ~ .-TARGET_FLAG-TARGET_FLAG_BOOL , data= train, family=poisson) 
+summary(glm)
 
+test$Prediction =predict.glm(glm, newdata=test, type="response")
+
+cor(test$TARGET_AMT, test$Prediction)
+
+glm2 <- glm(TARGET_AMT ~ .-TARGET_FLAG-TARGET_FLAG_BOOL , data= train) 
+summary(glm2)
+
+test$Prediction2 =predict.glm(glm2, newdata=test, type="response",family=poisson)
+
+cor(test$TARGET_AMT, test$Prediction2)
+
+ggplot(test, aes(x=Prediction2, y=TARGET_AMT)) +
+     geom_point(shape=1) +    # Use hollow circles
+     geom_smooth()            # Add a loess smoothed fit curve with confidence region
+
+glm3 <- glm(TARGET_AMT ~ AGE+PARENT1+MSTATUS+JOB+BLUEBOOK+TIF+CAR_AGE_density+URBANICITY+MVR_PTS+REVOKED , data= train, family=poisson) 
+summary(glm3)
+
+test$Prediction3 =predict.glm(glm3, newdata=test, type="response")
+
+cor(test$TARGET_AMT, test$Prediction3)
 ```

### 2019-11-08 final project/Manuscript.Rmd
Updated Final Project

--- 2019-11-05 final project/Manuscript.Rmd
@@
 ---
-title             : "Fall 2019 DATA 621 Final Project"
-shorttitle        : "Arnouts et al. Final Project"
+title             : "A Logistic Regression Qpproach to CoIL Challenge 2000"
+shorttitle        : " "
 author: 
@@
   - id            : "1"
-    institution   : "CUNY School of Professional Studies"
-authornote: |
-  MS in Data Science Students
+    institution   : "MS in Data Science Students CUNY School of Professional Studies"
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
+  A logistic regression based solution to the CoIL Challenge 2000 is described.
+The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
+  <!-- https://tinyurl.com/ybremelq 
     One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
@@
   Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
+  -->
 keywords          : "CoIL Challenge, Logistic Regression"
@@
   papaja::apa6_pdf:
-    citation_package: biblatex
+    #citation_package: natbib #biblatex
   html_document: default
@@
 # Introduction
-```{r eval=FALSE, echo=FALSE}
-#Describe the background and motivation of your problem.
-```
-Busineses use data science to extract insight from data.
+<!-- Describe the background and motivation of your problem.-->
+
+Businesses use data science to extract insights from data.
 It has many practical business applications.
-The CoIL Challenge showcases the power data science can bring to bear on fundamental business problems.
+Identifying households to include in a marketing campaign is one application.
+One example using real world data is the Computational Intelligence and Learning (CoIL) Challenge.
+The CoIL Challenge competition was held from March 17 to May 8 in 2000.
+The challenge is to:
 
-The CoIL Challenge was a datamining competition organized by the the Computational Intelligence and Learning Cluster. It was held in the period of March-May 2000.  The COIL Challenge had two tasks:
+1.  Identify potential customers for an insurance policy; and
 
-- Predict which customers are potentially interested in a caravan insurance policy; and
-- Describe the actual or potential customers; and possibly explain why these customers buy a caravan policy.
+2.  Provide a description of this customer base.
 
-In total 43 solutions were submitted. The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.
-
-In this paper we set out to complete the COIL Challenge ourselves using a logistic regression classifier. **SUMARISE FINDINGS**
+In total 147 participants registered and 43 submitted solutions [@Putten]. In this paper we set out to complete the first part of the COIL Challenge. **SUMARISE FINDINGS?**
 
@@
 
-Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are.
+<!--Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are.
 Explain how your investigation is similar or different to the state-of-theart.
-Please cite the relevant papers where appropriate.
+Please cite the relevant papers where appropriate.-->
+
+Participants used a variety of approaches in formulating their submissions including: Artificial Neural Networks [@Kim], Naïve Bayes [@Elkan].
+
+The maximum number of policyowners that could be found was 238.
+The winning model [@Elkan] selected 121 policy owners.
+Random selection results in 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear!) policy owners.
+[@Putten]
+
+<!--
+Other solutions (in alphabetical order):
+
+3 Achim Lewandowski. How to detect potential customers. [ps.gz], [pdf]
+
+4 Alexander Seewald. CoIL Challenge 2000 Submitted Solution. [ps.gz], [pdf]
+
+5 A.P. White and W.Z. Liu. The CoIL Challenge: an Application of Classification Trees with Bootstrap Aggregation. [ps.gz], [pdf]
+
+6 Andrew Greenyer. Coil 2000 Competition. The use of a learning classifier system JXCS. [ps.gz], [pdf]
+
+7 Arnold Koudijs. CoIL Challenge 2000 Submission for the Description Task. [ps.gz], [pdf]
+
+8 William M. Crocoll. Artificial Neural Network Portion of Coil Study. [ps.gz], [pdf]
+
+9 Chris Leckie and Herman Ferra. COIL Challenge 2000 Description Task. [ps.gz], [pdf]
+
+10 Gustavo Denicolay and Gaston Morales. CoIL Challenge 2000 Submission for the Description Task. [ps.gz] [pdf]
+
+11 Dragan Gamberger. Solution based on ILLM confirmation rule. [ps.gz], [pdf]
+
+12 János Abonyi, Hans Roubos. A Simple Fuzzy Classifier based on Inconsistency Analysis of Labeled Data. [ps.gz],[pdf]
+
+13 Jonathan Carter. Coil 2000 Challenge Submission: Genetic Algorithms and Hill-climbers. [ps.gz], [pdf]
+
+14 Juha Vesanto and Janne Sinkkonen. Submission for the CoIL Challenge 2000. [ps.gz], [pdf]
+
+15 Mark-André Krogel. A Data Mining Case Study. [ps.gz],[pdf]
+
+16 Michel Bera and Bertrand Lamy. Kxen at CoIL Challenge 2000. [ps.gz], [pdf]
+
+17 Paul Rickets. CoIL Challenge 2000 submission. [ps.gz], [pdf]
+
+18 Petr Mikšovský, Jirí Klema. COIL Challenge 2000. [ps.gz], [pdf]
+
+19 Petri Kontkanen. CoIL 2000 Submission. [ps.gz], [pdf]
+
+20 Philip Brierley. COIL 2000 Challenge: Characteristics of caravan insurance policy owners. [ps.gz] [pdf]
+
+21 Robert M. Simmonds. ACT Study report using Classification and Regression Tree (CART) Analysis. [ps.gz], [pdf]
+
+22 S. Sathiya Keerthi and Chong Jin Ong. Solution of the CoIL Challenge 2000 Task Using Support Vector Machines. [ps.gz], [pdf]
+
+23 Serhiy Shtovba, Yakiv Mashnitskiy. The Backpropagation Multilayer Feedforward Neural Network Based Competition Task Solution.
+
+24 Serhiy Shtovba. Phase Intervals and Genetic Algorithms Based Competition Task Solution. [ps.gz], [pdf]
+
+25 Jurgen A. Doornik and Steve Moyle. LOGIT Modelling. [ps.gz], [pdf]
+
+26 Thomas Martini Jorgensen and Christian Linneberg. Subspace Projections – an approach to variable selection and modeling. [ps.gz], [pdf]
+
+27 Tom McKone and Curt Stenger. COIL CHALLENGE 2000 Submission. [ps.gz], [pdf]
+
+28 Tomislav Šmuc. COIL 2000 Challenge Solution based on ILLM-SG Methodology. [ps.gz], [pdf]
+
+29 Uzay Kaymak and Magne Setnes. Target Selection based on fuzzy clustering: a volume prototype approach to CoIL Challenge 2000. [ps.gz], [pdf]
+
+-->
 
@@
 
-Discuss the key aspects of your problem, data set and regression model(s).
+<!--Discuss the key aspects of your problem, data set and regression model(s).
 Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.
@@
 
-Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).
+<!--Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).
 
 We used `r cite_r("Manuscript.bib")` for all our analyses.
+-->
 
@@
 
-Conclude your findings, limitations, and suggest areas for future work.
+<!--Conclude your findings, limitations, and suggest areas for future work.-->
 

### 2019-11-08 final project/Manuscript.tex
Updated Final Project

@@
 \hypersetup{unicode=true,
-            pdftitle={CoIL Challenge Submission},
+            pdftitle={A Logistic Regression Qpproach to CoIL Challenge 2000},
             pdfauthor={Corey Arnouts, Adam Douglas, Jason Givens-Doyle, \& Michael Silva},
@@
 
-  \title{CoIL Challenge Submission}
+  \title{A Logistic Regression Qpproach to CoIL Challenge 2000}
     \author{Corey Arnouts\textsuperscript{1}, Adam Douglas\textsuperscript{1}, Jason Givens-Doyle\textsuperscript{1}, \& Michael Silva\textsuperscript{1}}
@@
 
-\shorttitle{CoIL Challenge Submission}
+\shorttitle{ }
 \affiliation{
 \vspace{0.5cm}
-\textsuperscript{1} CUNY School of Professional Studies}
+\textsuperscript{1} MS in Data Science Students CUNY School of Professional Studies}
 \keywords{CoIL Challenge, Logistic Regression\newline\indent Word count: X}
@@
 
-\authornote{MS in Data Science Students
+\authornote{
 
@@
 \abstract{
-One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.
-
-Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.
-
-One sentence clearly stating the \textbf{general problem} being addressed by this particular study.
-
-One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).
-
-Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
-
-One or two sentences to put the results into a more \textbf{general context}.
-
-Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
+A logistic regression based solution to the CoIL Challenge 2000 is described.
+The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
 
@@
 
-Busineses use data science to extract insight from data.
+Businesses use data science to extract insights from data.
 It has many practical business applications.
-The Computational Intelligence and Learning (CoIL) Challenge showcases the power data science can bring to bear on fundamental business problems.
+Identifying households to include in a marketing campaign is one application.
+One example using real world data is the Computational Intelligence and Learning (CoIL) Challenge.
+The CoIL Challenge competition was held from March 17 to May 8 in 2000.
+The challenge is to:
 
-The CoIL Challenge was a datamining competition held from March 17 to May 8 in 2000. Participants could:
-
-\begin{itemize}
-\tightlist
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
 \item
-  Predict which customers are potentially interested in a caravan insurance policy; and
+  Identify potential customers for an insurance policy; and
 \item
-  Describe the actual or potential customers; and possibly explain why these customers buy a caravan policy.
-\end{itemize}
+  Provide a description of this customer base.
+\end{enumerate}
 
 In total 147 participants registered and 43 submitted solutions (Putten, Ruiter, \& Someren, 2000).
-The winners of the challenge were Charles Elkan for the prediction task and Nick Street and YongSeog Kim for the description task.
-
-In this paper we set out to complete the prediction part of the COIL Challenge ourselves using a logistic regression classifier. \textbf{SUMARISE FINDINGS}
+In this paper we set out to complete the first part of the COIL Challenge.
+\textbf{SUMARISE FINDINGS?}
 
@@
 
-Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are.
-Explain how your investigation is similar or different to the state-of-theart.
-Please cite the relevant papers where appropriate.
+Researchers used a variety of approaches in formulating their submissions including: Artificial Neural Networks (Kim \& Street, 2000), Naïve Bayes (Elkan, 2000).
+
+\hypertarget{methodology}{%
+\section{Methodology}\label{methodology}}
+
+\hypertarget{experimentation-and-results}{%
+\section{Experimentation and Results}\label{experimentation-and-results}}
+
+\hypertarget{discussion-and-conclusions}{%
+\section{Discussion and Conclusions}\label{discussion-and-conclusions}}
 
@@
 (Putten et al., 2000)
-
-\hypertarget{methodology}{%
-\section{Methodology}\label{methodology}}
-
-Discuss the key aspects of your problem, data set and regression model(s).
-Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.
-
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-
-\hypertarget{experimentation-and-results}{%
-\section{Experimentation and Results}\label{experimentation-and-results}}
-
-Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).
-
-We used R (Version 3.6.1; R Core Team, 2019) and the R-package \emph{\}papaja} {[}@\}R-papaja{]} for all our analyses.
-
-\hypertarget{discussion-and-conclusions}{%
-\section{Discussion and Conclusions}\label{discussion-and-conclusions}}
-
-Conclude your findings, limitations, and suggest areas for future work.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Putten-2000}{}%
+\leavevmode\hypertarget{ref-Elkan}{}%
+Elkan, C. (2000). CoIL challenge 2000 entry. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/ELKANP~1.pdf}
+
+\leavevmode\hypertarget{ref-Kim}{}%
+Kim, Y., \& Street, W.
+N.
+(2000).
+CoIL challenge 2000: Choosing and explaining likely caravan insurance customers.
+Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/STREET~1.pdf}
+
+\leavevmode\hypertarget{ref-Putten}{}%
 Putten, P., Ruiter, M., \& Someren, M. (2000). CoIL challenge 2000 tasks and results: Predicting and explaining caravan policy ownership.
-
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
 

### 2019-11-08 hw4/HW4.Rmd
More work on HW4

--- 2019-11-07 hw4/HW4.Rmd
@@
 
-hw4_plot <- function(df, var, strip_dollar_signs = FALSE){
+hw4_plot <- function(df, var, strip_dollar_signs = FALSE, x_log_scale = FALSE){
   if(strip_dollar_signs){
@@
   }
-  plot_1_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
-    geom_jitter() +
-    scale_color_brewer(palette="Set1") +
-    theme_minimal() +
-    theme(legend.position = "none",
-          axis.title.x = element_blank(),
-          axis.title.y = element_blank())
+  if(x_log_scale){
+    df$X <- log10(df$X)
+  }
+  # plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
+  #   geom_jitter() +
+  #   scale_color_brewer(palette="Set1") +
+  #   theme_minimal() +
+  #   theme(legend.position = "none",
+  #         axis.title.x = element_blank(),
+  #         axis.title.y = element_blank())
 
@@
     scale_color_brewer(palette="Set1") +
+    scale_y_log10(labels=comma) +
     theme_minimal() +
@@
   }
-  plot_1_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
-    geom_jitter() +
-    scale_color_brewer(palette="Set1") +
-    theme_minimal() +
-    theme(legend.position = "none",
-          axis.title.x = element_blank(),
-          axis.title.y = element_blank())
-  plot_1 <- df %>%
+  # plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
+  #   geom_jitter() +
+  #   scale_color_brewer(palette="Set1") +
+  #   theme_minimal() +
+  #   theme(legend.position = "none",
+  #         axis.title.x = element_blank(),
+  #         axis.title.y = element_blank())
+  temp <- df %>%
     group_by(X, TARGET_FLAG) %>%
+    tally()
+  temp <- df %>%
+    group_by(X) %>%
     tally() %>%
+    rename(TOTAL = n) %>%
+    merge(temp) %>%
+    mutate(PERCENT = n / TOTAL)
+  plot_1 <-  temp %>%
     ggplot(.) +
-    geom_bar(aes(X, n, fill = TARGET_FLAG), stat="identity") +
+    geom_bar(aes(X, PERCENT, fill = TARGET_FLAG), stat="identity") +
     scale_fill_brewer(palette="Set1") +
@@
       geom_boxplot() +
+      scale_y_log10() +
       scale_color_brewer(palette="Dark2") + #Paired
@@
 ```{r} 
-hw4_plot(df, "BLUEBOOK", TRUE)
+hw4_plot(df, "BLUEBOOK", TRUE, TRUE)
 ```
@@
 ```{r} 
-hw4_plot(df, "HOME_VAL", TRUE)
+hw4_plot(df, "HOME_VAL", TRUE, TRUE)
 ```
@@
 ```{r} 
-hw4_plot(df, "INCOME", TRUE)
+hw4_plot(df, "INCOME", TRUE, TRUE)
 ```
@@
 ```{r} 
-hw4_categorical_plot(df, "MVR_PTS")
-hw4_target_table(df, "MVR_PTS")
+hw4_plot(df, "MVR_PTS")
 ```
@@
 
+### OLDCLAIM
+
+**Theoretical Effect**
+If your total payout over the past five years was high, you are more likely to get into crashes.
+
+```{r} 
+hw4_plot(df, "OLDCLAIM", TRUE, TRUE)
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### PARENT1
+
+**Theoretical Effect**
+Unknown effect
+
+```{r} 
+hw4_categorical_plot(df, "PARENT1")
+hw4_target_table(df, "PARENT1")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### RED_CAR
+
+**Theoretical Effect**
+Urban legend says that red cars (espcially red sports cars) are more risky. Is that true?
+
+```{r} 
+hw4_categorical_plot(df, "RED_CAR")
+hw4_target_table(df, "RED_CAR")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### REVOKED
+
+**Theoretical Effect**
+If your license was revoked in the pat 7 years, you probably are a more risky driver.
+
+```{r}
+hw4_categorical_plot(df, "REVOKED")
+hw4_target_table(df, "REVOKED")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### SEX
+
+**Theoretical Effect**
+Urban legend says that women have less crashes then men.  Is that true?
+
+```{r}
+hw4_categorical_plot(df, "SEX")
+hw4_target_table(df, "SEX")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### TIF
+
+**Theoretical Effect**
+People who have been customers for a long time are usually more safe.
+
+```{r}
+hw4_plot(df, "TIF")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### TRAVTIME
+
+**Theoretical Effect**
+Log drives to work usually suggest greater risk
+
+```{r}
+hw4_plot(df, "TRAVTIME")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### URBANICITY
+
+**Theoretical Effect**
+Unknown
+
+```{r}
+hw4_categorical_plot(df, "URBANICITY")
+hw4_target_table(df, "URBANICITY")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### YOJ
+
+**Theoretical Effect**
+People who stay at a job for a long time are usuaslly more safe
+
+```{r}
+hw4_plot(df, "YOJ")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
 Looking at the data, we can see that there are some missing values:

### 2019-11-08 hw4/HW4_Mike.Rmd
Initial commit of my scratchpad

---
title: "Mike's Scratchpad for DATA 621 Homework #4"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(RColorBrewer)
library(FFTrees)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

## Data Cleanup

The dollar values have dollar signs.  These will be stripped out and converted from a character to numeric.  LOG10 and LN transformations of all dollar values will be preformed.  Categorical data will be translated into dummy variables.

```{r}
df <- read.csv("./data/insurance_training_data.csv")
evaluation <- read.csv("./data/insurance-evaluation-data.csv")

strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}
```

```{r}
to_dummies <- function(df, var, drop = FALSE){
  temp <- data.frame(INDEX = df$INDEX, CAT = df[[var]]) %>%
    mutate(CAT = as.character(CAT),
           VAL = 1) %>%
    spread(CAT, VAL, fill = 0)
  new_names <- c()
  for (n in names(temp)){
    if(n != "INDEX"){
      n <- paste0(var,"_",n)
      n <- gsub(" ", "_", n)
    }
    new_names <- c(new_names, n)
  }
  names(temp) <- new_names

  if(drop){
    drops <- c(var)
    df <- df[ , !(names(df) %in% drops)]
  }

  df %>%
    merge(temp)
}

get_driver_type <- function(x, y){
  if(is.na(x) | is.na(y)){
    return(NA)
  }
  if(y < 40){
    driver_type <- "Young"
  } else if (y < 70){
    driver_type <- "Middle Age "
  } else {
    driver_type <- "Older"
  }
  if( x == "M"){
    return(paste(driver_type, "Male"))
  } else {
    return(paste(driver_type, "Female"))
  }
}

temp <- df %>%
  rowwise() %>%
  mutate(INCOME = strip_dollars(INCOME),
         HOME_VAL = strip_dollars(HOME_VAL),
         BLUEBOOK = strip_dollars(BLUEBOOK),
         OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
  mutate(JOB = ifelse(str_length(JOB) > 0, JOB, NA),
         CAR_AGE = ifelse(str_length(CAR_AGE) > 0, CAR_AGE, NA),
         EDUCATION = ifelse(EDUCATION == "<High_School", "LT_High_School", EDUCATION),
         URBANICITY = ifelse(URBANICITY == "Highly Urban/ Urban", "Urban", "Rural")) %>%
  mutate(DRIVER_TYPE = get_driver_type(SEX, AGE)) %>%
  ungroup() %>%
  na.omit() %>%
  to_dummies("PARENT1", T) %>%
  to_dummies("KIDSDRIV", T) %>%
  to_dummies("HOMEKIDS", T) %>%
  to_dummies("MSTATUS", T) %>%
  to_dummies("SEX", T) %>%
  to_dummies("EDUCATION", T) %>%
  to_dummies("JOB", T) %>%
  to_dummies("CAR_USE", T) %>%
  to_dummies("CAR_TYPE", T) %>%
  to_dummies("RED_CAR", T) %>%
  to_dummies("REVOKED", T) %>%
  to_dummies("MVR_PTS", T) %>%
  to_dummies("URBANICITY", T) %>%
  to_dummies("DRIVER_TYPE", T) %>%
  rowwise() %>%
  mutate(BLUEBOOK = ifelse(BLUEBOOK > 0, BLUEBOOK, 0)) %>%
  # mutate(LOG10_INCOME = log10(INCOME + 1),
  #        LN_INCOME = log(INCOME + 1),
  #        LOG10_HOME_VAL = log10(HOME_VAL + 1),
  #        LN_HOME_VAL = log(HOME_VAL + 1),
  #        LOG10_BLUEBOOK = log10(BLUEBOOK + 1),
  #        LN_BLUEBOOK = log(BLUEBOOK + 1),
  #        LOG10_OLDCLAIM = log10(OLDCLAIM + 1),
  #        LN_OLDCLAIM = log(OLDCLAIM + 1)) %>%
  mutate(PRIOR_CLAIM = ifelse(CLM_FREQ + OLDCLAIM > 0, 1,0)) %>%
  ungroup() 
```

```{r}
set.seed(42)
minority <- nrow(temp[temp$TARGET_FLAG == 1,])
majority <- nrow(temp[temp$TARGET_FLAG == 0,])
diff <- majority - minority
minority_index <- temp[temp$TARGET_FLAG == 1,]$INDEX
over_sample_df <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
  merge(temp, .) %>%
  bind_rows(temp)

train_index <- createDataPartition(over_sample_df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
train <- over_sample_df[train_index,]
test <- over_sample_df[-train_index,]
```

```{r}
fft_train <- train %>%
  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
fft_test <- test %>%
  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
fit <- FFTrees(TARGET_FLAG ~ ., fft_train, fft_test)
plot(fit)
```
The FFTree suggests that past history of filing claims is a good predictor if someone will file a claim in the future.  This could be our baseline model criteria.

Other variables that seems to come up is if the driver is in an urban location and possibly their home value and if it's a commercial vehicle

### 2019-11-09 final project/Manuscript.Rmd
Updated final with more lit review

@@
 ---
-title             : "A Logistic Regression Qpproach to CoIL Challenge 2000"
+title             : "A Logistic Regression Approach to CoIL Challenge 2000"
 shorttitle        : " "
@@
 
-Participants used a variety of approaches in formulating their submissions including: Artificial Neural Networks [@Kim], Naïve Bayes [@Elkan].
+Participants used a variety of approaches in formulating their submissions including:
+
+- Classification Trees with Bagging [@White];
+
+- C4.5 [@Seewald];
+
+- Evolutionary Algorithm [@Koudijs];
+
+- Inductive Learning by Logic Minimization (ILLM) [@Gamberger];
+
+- Mask Perceptron with Boosting [@Leckie];
+
+- Naïve Bayes [@Elkan]; and
+
+- Neural Networks[@Crocoll], [@Kim];
+
+- Scoring System [@Lewandowski];
+
+- Support Vector Machines; and
+
+- XCS [@Greenyer].
 
@@
 <!--
-Other solutions (in alphabetical order):
-
-3 Achim Lewandowski. How to detect potential customers. [ps.gz], [pdf]
-
-4 Alexander Seewald. CoIL Challenge 2000 Submitted Solution. [ps.gz], [pdf]
-
-5 A.P. White and W.Z. Liu. The CoIL Challenge: an Application of Classification Trees with Bootstrap Aggregation. [ps.gz], [pdf]
-
-6 Andrew Greenyer. Coil 2000 Competition. The use of a learning classifier system JXCS. [ps.gz], [pdf]
-
-7 Arnold Koudijs. CoIL Challenge 2000 Submission for the Description Task. [ps.gz], [pdf]
-
-8 William M. Crocoll. Artificial Neural Network Portion of Coil Study. [ps.gz], [pdf]
-
-9 Chris Leckie and Herman Ferra. COIL Challenge 2000 Description Task. [ps.gz], [pdf]
-
-10 Gustavo Denicolay and Gaston Morales. CoIL Challenge 2000 Submission for the Description Task. [ps.gz] [pdf]
-
-11 Dragan Gamberger. Solution based on ILLM confirmation rule. [ps.gz], [pdf]
+10 Gustavo Denicolay and Gaston Morales. CoIL Challenge 2000 Submission for the Description Task. [ps.gz] [pdf] - Not sure what they used
 

### 2019-11-09 final project/Manuscript.tex
Updated final with more lit review

@@
 \hypersetup{unicode=true,
-            pdftitle={A Logistic Regression Qpproach to CoIL Challenge 2000},
+            pdftitle={A Logistic Regression Approach to CoIL Challenge 2000},
             pdfauthor={Corey Arnouts, Adam Douglas, Jason Givens-Doyle, \& Michael Silva},
@@
 
-  \title{A Logistic Regression Qpproach to CoIL Challenge 2000}
+  \title{A Logistic Regression Approach to CoIL Challenge 2000}
     \author{Corey Arnouts\textsuperscript{1}, Adam Douglas\textsuperscript{1}, Jason Givens-Doyle\textsuperscript{1}, \& Michael Silva\textsuperscript{1}}
@@
 
-Researchers used a variety of approaches in formulating their submissions including: Artificial Neural Networks (Kim \& Street, 2000), Naïve Bayes (Elkan, 2000).
+Participants used a variety of approaches in formulating their submissions including:
+
+\begin{itemize}
+\item
+  Classification Trees with Bagging (White \& Liu, 2000);
+\item
+  C4.5 (Seewald, 2000);
+\item
+  Evolutionary Algorithm (Koudijs, 2000);
+\item
+  Inductive Learning by Logic Minimization (ILLM) (Gamberger, 2000);
+\item
+  Mask Perceptron with Boosting (Leckie \& Ferra, 2000);
+\item
+  Naïve Bayes (Elkan, 2000); and
+\item
+  Neural Networks(Crocoll, 2000), (Kim \& Street, 2000);
+\item
+  Scoring System (Lewandowski, 2000);
+\item
+  Support Vector Machines; and
+\item
+  XCS (Greenyer, 2000).
+\end{itemize}
+
+The maximum number of policyowners that could be found was 238.
+The winning model (Elkan, 2000) selected 121 policy owners.
+Random selection results in 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear!) policy owners.
+(Putten et al., 2000)
 
@@
 \section{Discussion and Conclusions}\label{discussion-and-conclusions}}
-
-The maximum number of policyowners that could be found was 238, the winning model selected 121 policy owners.
-Random selection results in 42 policy owners.
-The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks)and 118 (linear!) policy owners.
-(Putten et al., 2000)
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Crocoll}{}%
+Crocoll, W. M. (2000). Artificial neural network portion of coil study. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/CROCOL~1.pdf}
+
 \leavevmode\hypertarget{ref-Elkan}{}%
 Elkan, C. (2000). CoIL challenge 2000 entry. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/ELKANP~1.pdf}
+
+\leavevmode\hypertarget{ref-Gamberger}{}%
+Gamberger, D. (2000). Solution based on illm confirmation rule. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/GAMBER~1.pdf}
+
+\leavevmode\hypertarget{ref-Greenyer}{}%
+Greenyer, A. (2000). Coil 2000 competition. The use of a learning classifier system jxcs. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/GREENY~1.pdf}
 
@@
 
+\leavevmode\hypertarget{ref-Koudijs}{}%
+Koudijs, A. (2000). CoIL challenge 2000 submission for the description task. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/KOUDIJ~1.pdf}
+
+\leavevmode\hypertarget{ref-Leckie}{}%
+Leckie, C., \& Ferra, H. (2000). COIL challenge 2000 description task. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/LECKIE~1.pdf}
+
+\leavevmode\hypertarget{ref-Lewandowski}{}%
+Lewandowski, A. (2000). How to detect potential customers. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/LEWAND~1.pdf}
+
 \leavevmode\hypertarget{ref-Putten}{}%
 Putten, P., Ruiter, M., \& Someren, M. (2000). CoIL challenge 2000 tasks and results: Predicting and explaining caravan policy ownership.
+
+\leavevmode\hypertarget{ref-Seewald}{}%
+Seewald, A. (2000). CoIL challenge 2000 submitted solution. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/SEEWAL~1.pdf}
+
+\leavevmode\hypertarget{ref-White}{}%
+White, A. P., \& Liu, W. Z. (2000). The coil challenge: An application of classification trees with bootstrap aggregation. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/WHITEP~1.pdf}
 

### 2019-11-09 hw4/HW4_Mike.Rmd
Updated Mike's scratchpad

@@
   #        LN_OLDCLAIM = log(OLDCLAIM + 1)) %>%
-  mutate(PRIOR_CLAIM = ifelse(CLM_FREQ + OLDCLAIM > 0, 1,0)) %>%
+  #mutate(PRIOR_CLAIM = ifelse(CLM_FREQ + OLDCLAIM > 0, 1,0)) %>%
   ungroup()
+
+set.seed(42)
+# Split the data
+train_index <- createDataPartition(temp$TARGET_FLAG, p = .7, list = FALSE, times = 1)
+train <- temp[train_index,]
+test <- temp[-train_index,]
+
+# Ballance the data
+minority <- nrow(train[train$TARGET_FLAG == 1,])
+majority <- nrow(train[train$TARGET_FLAG == 0,])
+diff <- majority - minority
+minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
+over_sample_df <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
+  merge(train, .) %>%
+  bind_rows(train)
 ```
@@
 set.seed(42)
-minority <- nrow(temp[temp$TARGET_FLAG == 1,])
-majority <- nrow(temp[temp$TARGET_FLAG == 0,])
-diff <- majority - minority
-minority_index <- temp[temp$TARGET_FLAG == 1,]$INDEX
-over_sample_df <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
-  merge(temp, .) %>%
-  bind_rows(temp)
-
-train_index <- createDataPartition(over_sample_df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
-train <- over_sample_df[train_index,]
-test <- over_sample_df[-train_index,]
-```
-
-```{r}
-fft_train <- train %>%
+fft_train <- over_sample_df %>%
   select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
@@
   select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
-fit <- FFTrees(TARGET_FLAG ~ ., fft_train, fft_test)
+fit <- FFTrees(TARGET_FLAG ~ ., fft_train, fft_test, do.comp = FALSE)
 plot(fit)
+plot(fit, data="test")
+
+test_results = data.frame(yhat = ifelse(predict(fit, test),1,0), TARGET_FLAG = test$TARGET_FLAG)
+table(test_results$yhat, test_results$TARGET_FLAG)
 ```

### 2019-11-09 final project/Manuscript.Rmd
Figured out how to print the R file

@@
 library("papaja")
+
+library(tidyverse)
+
+# Thank you Stack Overflow!
+# A Prefix nulling hook.
+
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set( output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if( is.na(comment) ) comment <- ""
+  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+                     x, perl = TRUE)
+  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
+  if( can_null && do_null ) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr( "\\]", x )[1] - 1
+    # Two cases: start or newline
+    re <- paste0( "^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub( re, rep,  x )
+    re <- paste0( "\\\n.{", align_index, "}\\]")
+    rep <- paste0( "\n", comment )
+    x <- gsub( re, rep,  x )
+  }
+
+  default_output_hook( x, options )
+
+})
+knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
 ```
@@
 
-```{r source, code=readLines("CoIL.R")}
+<code>
+```{r, null_prefix = TRUE, comment=NA}
+conn <- file("CoIL.r", open="r")
+lines <- readLines(conn)
+for (i in 1:length(lines)){
+  cat(lines[i], "\n")
+}
+close(conn)
 ```
+</code>

### 2019-11-09 final project/Manuscript.tex
Figured out how to print the R file

@@
 
+\begin{verbatim}
+# CoIL Challenge Source Code 
+library(tidyverse) 
+# Download the data sets from UCI if they are not present 
+url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/tic-mld/" 
+files <- c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt") 
+for (file_name in files) { 
+  file_path <- paste0("data/", file_name) 
+  file_url <- paste0(url, file_name) 
+  if (!file.exists(file_path)) { 
+    message(paste("Downloading", file_name)) 
+    download.file(file_url, file_path) 
+  } 
+} 
+# Read in the data 
+df <- read.delim("data/ticdata2000.txt", header = FALSE) 
+names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", 
+               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", 
+               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", 
+               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", 
+               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", 
+               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", 
+               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575", 
+               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", 
+               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO", 
+               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", 
+               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL", 
+               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", 
+               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", 
+               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", 
+               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL", 
+               "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN") 
+eval <- read.delim("data/ticeval2000.txt", header = FALSE) 
+temp <- read.delim("data/tictgts2000.txt", header = FALSE) 
+eval$CARAVAN <- temp$V1 
+names(eval) <- names(df) 
+\end{verbatim}
+
 \end{document}

### 2019-11-09 final project/CoIL.R

@@ -2,7 +2,8 @@
 library(tidyverse)
 # Download the data sets from UCI if they are not present
 url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/tic-mld/"
-for (file_name in c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt")) {
+files <- c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt")
+for (file_name in files) {
   file_path <- paste0("data/", file_name)
   file_url <- paste0(url, file_name)
   if (!file.exists(file_path)) {

### 2019-11-09 hw4/HW4.Rmd
Update HW4

@@
 ```{r}
-df <- read.csv("./data/insurance_training_data.csv") %>%
-  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
-  select(-INDEX)
-evaluation <- read.csv("./data/insurance-evaluation-data.csv") %>%
-  select(-INDEX)
+df <- read.csv("./data/insurance_training_data.csv")
+evaluation <- read.csv("./data/insurance-evaluation-data.csv")
 
@@
 
-```{r}
-missmap(df, main = "Missing vs Observed Values")
-```
-
-Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, but only `r df %>% filter(is.na(YOJ),is.na(CAR_AGE)) %>% nrow()` observations missing both of these varables.
-So, if imputation is not viable here, we will lose as many as 488 observations out of the 8161 in our training set (or about 6%).
-
 ## Data Preparation
-
-```{r}
-df %>%
-  summary() %>%
-  kable() %>%
-  kable_styling()
-```
 
@@
 
-There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.
+There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.
+We will do this to both the `df` and `evaluation` data frames.
+There are also some invalid data that will be changed to NAs.
 
@@
 
-df <- fix_data_types(df)
-evaluation <- fix_data_types(evaluation)
+na_bad_values <- function(messy_df){
+  messy_df %>%
+    rowwise() %>%
+    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
+    ungroup()
+}
+
+df <- df %>%
+  fix_data_types() %>%
+  na_bad_values()
+evaluation <- evaluation %>%
+  fix_data_types() %>%
+  na_bad_values()
 ```
@@
 ### Fix Missing Values
-There are missing values.  We will fill in the missing variables using KNN.
 
-```{r, null_prefix = TRUE}
-# set.seed(42)
-# knn <- df %>%
-#   select(-TARGET_FLAG, -TARGET_AMT, is.numeric()) %>%
-#   #na.omit() %>%
-#   knnImputation()
-# 
-# for(var in names(df)){
-#   impute_me <- is.na(df[[var]])
-#   if (nrow(df[impute_me,]) > 0){
-#     print(paste("Fixing the", nrow(df[impute_me,]), "missing values in", var, "(df)"))
-#     #df[impute_me, var] <- knn[impute_me, var] 
-#   }
-# }
+```{r}
+missmap(df, main = "Missing vs Observed Values")
 ```
+
+Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.
+There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.
 
@@
 ```
+
+There are `r nrow(train[TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.
+We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.
+
+```{r}
+set.seed(42)
+minority <- nrow(train[train$TARGET_FLAG == 1,])
+majority <- nrow(train[train$TARGET_FLAG == 0,])
+diff <- majority - minority
+minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
+over_sample_train <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
+  merge(train, .) %>%
+  bind_rows(train)
+```
+The over sampled data frame has `r nrow(over_sample_train)` records, and as you can see in the following figure is now balanced:
+
+```{r, echo=FALSE}
+over_sample_train %>% 
+  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
+  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  labs(x="TARGET_FLAG", y="# Observations")
+```

### 2019-11-09 hw4/HW4.Rmd
Pulled visualizations out of HW4 and storedthem in HW4_Mike

@@
 }
-
-hw4_plot <- function(df, var, strip_dollar_signs = FALSE, x_log_scale = FALSE){
-  if(strip_dollar_signs){
-    df$X <- strip_dollars(df[[var]])
-  } else {
-    df$X <- df[[var]]
-  }
-  if(x_log_scale){
-    df$X <- log10(df$X)
-  }
-  # plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
-  #   geom_jitter() +
-  #   scale_color_brewer(palette="Set1") +
-  #   theme_minimal() +
-  #   theme(legend.position = "none",
-  #         axis.title.x = element_blank(),
-  #         axis.title.y = element_blank())
-
-  plot_1 <- ggplot(df, aes(X, color = TARGET_FLAG)) +
-    geom_density() + 
-    scale_color_brewer(palette="Set1") +
-    theme_minimal() +
-    theme(legend.position = "none",
-          axis.title.x = element_blank(),
-          axis.title.y = element_blank())
-
-  plot_2 <- ggplot(df, aes(X, TARGET_AMT, color = TARGET_FLAG)) +
-    geom_point() +
-    scale_color_brewer(palette="Set1") +
-    scale_y_log10(labels=comma) +
-    theme_minimal() +
-    theme(legend.position = "none",
-          axis.title.x = element_blank(),
-          axis.title.y = element_blank())
-
-  grid.arrange(plot_1, plot_2, ncol=2)
-}
-
-hw4_categorical_plot <- function(df, var, strip_dollar_signs = FALSE){
-  if(strip_dollar_signs){
-    df$X <- strip_dollars(df[[var]])
-  } else {
-    df$X <- as.factor(as.character(df[[var]]))
-  }
-  # plot_1 <- ggplot(df, aes(X, TARGET_FLAG, color = TARGET_FLAG)) +
-  #   geom_jitter() +
-  #   scale_color_brewer(palette="Set1") +
-  #   theme_minimal() +
-  #   theme(legend.position = "none",
-  #         axis.title.x = element_blank(),
-  #         axis.title.y = element_blank())
-  temp <- df %>%
-    group_by(X, TARGET_FLAG) %>%
-    tally()
-  temp <- df %>%
-    group_by(X) %>%
-    tally() %>%
-    rename(TOTAL = n) %>%
-    merge(temp) %>%
-    mutate(PERCENT = n / TOTAL)
-  plot_1 <-  temp %>%
-    ggplot(.) +
-    geom_bar(aes(X, PERCENT, fill = TARGET_FLAG), stat="identity") +
-    scale_fill_brewer(palette="Set1") +
-    theme_minimal() +
-    theme(legend.position = "none",
-          axis.title.x = element_blank(),
-          axis.title.y = element_blank())
-  plot_2 <- df %>%
-    filter(TARGET_FLAG == 1) %>%
-    ggplot(., aes(X, TARGET_AMT, color = X)) +
-      geom_boxplot() +
-      scale_y_log10() +
-      scale_color_brewer(palette="Dark2") + #Paired
-      theme_minimal() +
-      theme(legend.position = "none",
-            axis.title.x = element_blank(),
-            axis.title.y = element_blank())
-
-  grid.arrange(plot_1, plot_2, ncol=2)
-}
-
-hw4_target_table <- function(df, var, strip_dollar_signs = FALSE){
-  if(strip_dollar_signs){
-    df$X <- strip_dollars(df[[var]])
-  } else {
-    df$X <- df[[var]]
-  }
-  df %>%
-    group_by(X, TARGET_FLAG) %>%
-    tally() %>%
-    spread(TARGET_FLAG, n) %>%
-    rename(Variable = X) %>%
-    kable() %>%
-    kable_styling()
-}
 ```
@@
 df %>% 
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
   ggplot(aes(x=TARGET_FLAG,fill=TARGET_FLAG)) +
@@
 
-There are some missing values that we will need to deal with.
-Let's look at the predictors to see how well they predict if there is an accident (plot on the left) and how much it would cost (plot on the right).
-Numeric variables will get scatterplots, while categorical will get box plots
-
-### AGE
-
-**Theoretical Effect**
-Very young people tend to be risky.  Maybe very old people also.
-
-```{r} 
-hw4_plot(df, "AGE")
-```
-
-**Observed Effect**
-There is very little information encapsulated in the age of the person.
-
-### BLUEBOOK
-
-**Theoretical Effect**
-Unknown effect on probability of collision, but probably effects the payout if there is a crash.
-
-```{r} 
-hw4_plot(df, "BLUEBOOK", TRUE, TRUE)
-```
-
-**Observed Effect**
-The plot suggests that people with low value and high value vehicles are equally likely to be in a crash.  It also suggests the payout is not strongly correlated with the vehicle's value.
-
-### CAR_AGE
-
-**Theoretical Effect**
-Unknown effect on probability of collision, but probably effects the payout if there is a crash.
-
-```{r} 
-hw4_plot(df, "CAR_AGE")
-```
-
-**Observed Effect**
-The plot suggests that people with old or new vehicles are equally likely to be in a crash.
-It also suggests the payout is not strongly correlated with the vehicle's age.
-*There is an observation with a negative value that will need to be cleaned up.*
-
-### CAR_TYPE
-
-**Theoretical Effect**
-Unknown effect on probability of collision, but probably effects the payout if there is a crash.
-
-```{r} 
-hw4_categorical_plot(df, "CAR_TYPE")
-hw4_target_table(df, "CAR_TYPE")
-```
-
-**Observed Effect**
-The plot suggests that the type of vehicle has no effect on the probability of being in an accident, and little effect of the cost.
-
-### CAR_USE
-
-**Theoretical Effect**
-Commercian vehicles are driven more, so might increase probability of collision.
-
-```{r} 
-hw4_categorical_plot(df, "CAR_USE")
-hw4_target_table(df, "CAR_USE")
-```
-
-**Observed Effect**
-The plot suggests that the use does not increase or decrease the probability of being in an accident or the amount of the claim.
-
-### CLM_FREQ
-
-**Theoretical Effect**
-The more claims you filed in the past, the more you are likely to file in the future.
-
-```{r} 
-hw4_categorical_plot(df, "CLM_FREQ")
-hw4_target_table(df, "CLM_FREQ")
-```
-
-**Observed Effect**
-Looks like there are few people who signed up for the frequent accident club.  All jokes aside, the theoretical effect is not observed in the data.
-
-### EDUCATION
-
-**Theoretical Effect**
-Unknown effect, but in theory more educated people tend to drive more safely
-
-```{r} 
-hw4_categorical_plot(df, "EDUCATION")
-hw4_target_table(df, "EDUCATION")
-```
-
-**Observed Effect**
-The theoretical effect has no basis and is somewhat elitist.
-
-### HOMEKIDS
-
-**Theoretical Effect**
-Unknown effect
-
-```{r} 
-hw4_categorical_plot(df, "HOMEKIDS")
-hw4_target_table(df, "HOMEKIDS")
-```
-
-**Observed Effect**
-It's interesting to note that those with kids have a 0.34 probability of being in an accident.  Those without kids have a 0.22 probability of being in an accident.  This may be useful.
-
-### HOME_VAL
-
-**Theoretical Effect**
-In theory, home owners tend to drive more responsibly.
-
-```{r} 
-hw4_plot(df, "HOME_VAL", TRUE, TRUE)
-```
-
-**Observed Effect**
-There are a lot of zero values.  I am assuming that it is a zero because they are renters.
-
-### INCOME
-
-**Theoretical Effect**
-In theory, rich people trend to get into fewer crashes.
-
-```{r} 
-hw4_plot(df, "INCOME", TRUE, TRUE)
-```
-
-**Observed Effect**
-There are a lot of zero values.  I am assuming that it is a zero because they are renters.
-
-### JOB
-
-**Theoretical Effect**
-In theory, white collar jobs tend to be safer
-
-```{r} 
-hw4_categorical_plot(df, "JOB")
-hw4_target_table(df, "JOB")
-```
-
-**Observed Effect**
-There is a 0.21 probability for white collar job holders being in an accident.  there's a 0.34 probability for the blue collar jobs.  There may be some truth to the theoretical effect.
-
-### KIDSDRIV
-
-**Theoretical Effect**
-When teenagers drive your car, ou are more likely to get into crashes
-
-```{r} 
-hw4_categorical_plot(df, "KIDSDRIV")
-hw4_target_table(df, "KIDSDRIV")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### MSTATUS
-
-**Theoretical Effect**
-In theory, married people drive more safely
-
-```{r} 
-hw4_categorical_plot(df, "MSTATUS")
-hw4_target_table(df, "MSTATUS")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### MVR_PTS
-
-**Theoretical Effect**
-If you get lots of traffic tickets, you trend to get into more crashes
-
-```{r} 
-hw4_plot(df, "MVR_PTS")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### OLDCLAIM
-
-**Theoretical Effect**
-If your total payout over the past five years was high, you are more likely to get into crashes.
-
-```{r} 
-hw4_plot(df, "OLDCLAIM", TRUE, TRUE)
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### PARENT1
-
-**Theoretical Effect**
-Unknown effect
-
-```{r} 
-hw4_categorical_plot(df, "PARENT1")
-hw4_target_table(df, "PARENT1")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### RED_CAR
-
-**Theoretical Effect**
-Urban legend says that red cars (espcially red sports cars) are more risky. Is that true?
-
-```{r} 
-hw4_categorical_plot(df, "RED_CAR")
-hw4_target_table(df, "RED_CAR")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### REVOKED
-
-**Theoretical Effect**
-If your license was revoked in the pat 7 years, you probably are a more risky driver.
-
-```{r}
-hw4_categorical_plot(df, "REVOKED")
-hw4_target_table(df, "REVOKED")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### SEX
-
-**Theoretical Effect**
-Urban legend says that women have less crashes then men.  Is that true?
-
-```{r}
-hw4_categorical_plot(df, "SEX")
-hw4_target_table(df, "SEX")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### TIF
-
-**Theoretical Effect**
-People who have been customers for a long time are usually more safe.
-
-```{r}
-hw4_plot(df, "TIF")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### TRAVTIME
-
-**Theoretical Effect**
-Log drives to work usually suggest greater risk
-
-```{r}
-hw4_plot(df, "TRAVTIME")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### URBANICITY
-
-**Theoretical Effect**
-Unknown
-
-```{r}
-hw4_categorical_plot(df, "URBANICITY")
-hw4_target_table(df, "URBANICITY")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-
-### YOJ
-
-**Theoretical Effect**
-People who stay at a job for a long time are usuaslly more safe
-
-```{r}
-hw4_plot(df, "YOJ")
-```
-
-**Observed Effect**
-INSERT TEXT HERE
-Looking at the data, we can see that there are some missing values:
+There are some missing values that we will need to deal with. There are also some values that seem invalid (i.e. -3 CAR_AGE).
 
@@
 
-```{r}
-missmap(df, main = "Missing vs Observed Values")
-```
-
-Specifically, there `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable are missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.
+There are `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable is missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.
 There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.
@@
 
-There are `r nrow(train[TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.
+There are `r sum(train$TARGET_FLAG)` out of `r nrow(train)` records in the training data set that have been in an accident.
 We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.
@@
 minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
-over_sample_train <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
+over_sample_train <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
   merge(train, .) %>%
@@
 ```
+
 The over sampled data frame has `r nrow(over_sample_train)` records, and as you can see in the following figure is now balanced:
@@
 over_sample_train %>% 
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
   ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +

### 2019-11-09 hw4/HW4_Mike.Rmd
Pulled visualizations out of HW4 and storedthem in HW4_Mike

@@
 Other variables that seems to come up is if the driver is in an urban location and possibly their home value and if it's a commercial vehicle
+
+```{r}
+hw4_plot <- function(df, var, strip_dollar_signs = FALSE, x_log_scale = FALSE){
+  if(strip_dollar_signs){
+    df$X <- strip_dollars(df[[var]])
+  } else {
+    df$X <- df[[var]]
+  }
+  if(x_log_scale){
+    df$X <- log10(df$X)
+  }
+
+  plot_1 <- ggplot(df, aes(X, color = as.factor(TARGET_FLAG))) +
+    geom_density() + 
+    scale_color_brewer(palette="Set1") +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+
+  plot_2 <- ggplot(df, aes(X, TARGET_AMT, color = as.factor(TARGET_FLAG))) +
+    geom_point() +
+    scale_color_brewer(palette="Set1") +
+    scale_y_log10(labels=comma) +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+
+  grid.arrange(plot_1, plot_2, ncol=2)
+}
+
+hw4_categorical_plot <- function(df, var, strip_dollar_signs = FALSE){
+  if(strip_dollar_signs){
+    df$X <- strip_dollars(df[[var]])
+  } else {
+    df$X <- as.factor(as.character(df[[var]]))
+  }
+
+  temp <- df %>%
+    group_by(X, TARGET_FLAG) %>%
+    tally()
+  temp <- df %>%
+    group_by(X) %>%
+    tally() %>%
+    rename(TOTAL = n) %>%
+    merge(temp) %>%
+    mutate(PERCENT = n / TOTAL)
+
+  plot_1 <-  temp %>%
+    ggplot(.) +
+    geom_bar(aes(X, PERCENT, fill = as.factor(TARGET_FLAG)), stat="identity") +
+    scale_fill_brewer(palette="Set1") +
+    theme_minimal() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+  plot_2 <- df %>%
+    filter(TARGET_FLAG == 1) %>%
+    ggplot(., aes(X, as.factor(TARGET_AMT), color = X)) +
+      geom_boxplot() +
+      #scale_y_log10() +
+      scale_color_brewer(palette="Dark2") + #Paired
+      theme_minimal() +
+      theme(legend.position = "none",
+            axis.title.x = element_blank(),
+            axis.title.y = element_blank())
+
+  grid.arrange(plot_1, plot_2, ncol=2)
+}
+
+hw4_target_table <- function(df, var, strip_dollar_signs = FALSE){
+  if(strip_dollar_signs){
+    df$X <- strip_dollars(df[[var]])
+  } else {
+    df$X <- df[[var]]
+  }
+  df %>%
+    group_by(X, TARGET_FLAG) %>%
+    tally() %>%
+    spread(TARGET_FLAG, n) %>%
+    rename(Variable = X) %>%
+    kable() %>%
+    kable_styling()
+}
+```
+
+### AGE
+
+**Theoretical Effect**
+Very young people tend to be risky.  Maybe very old people also.
+
+```{r} 
+hw4_plot(df, "AGE")
+```
+
+**Observed Effect**
+There is very little information encapsulated in the age of the person.
+
+### BLUEBOOK
+
+**Theoretical Effect**
+Unknown effect on probability of collision, but probably effects the payout if there is a crash.
+
+```{r} 
+hw4_plot(df, "BLUEBOOK", TRUE, TRUE)
+```
+
+**Observed Effect**
+The plot suggests that people with low value and high value vehicles are equally likely to be in a crash.  It also suggests the payout is not strongly correlated with the vehicle's value.
+
+### CAR_AGE
+
+**Theoretical Effect**
+Unknown effect on probability of collision, but probably effects the payout if there is a crash.
+
+```{r} 
+hw4_plot(df, "CAR_AGE")
+```
+
+**Observed Effect**
+The plot suggests that people with old or new vehicles are equally likely to be in a crash.
+It also suggests the payout is not strongly correlated with the vehicle's age.
+*There is an observation with a negative value that will need to be cleaned up.*
+
+### CAR_TYPE
+
+**Theoretical Effect**
+Unknown effect on probability of collision, but probably effects the payout if there is a crash.
+
+```{r} 
+hw4_categorical_plot(df, "CAR_TYPE")
+hw4_target_table(df, "CAR_TYPE")
+```
+
+**Observed Effect**
+The plot suggests that the type of vehicle has no effect on the probability of being in an accident, and little effect of the cost.
+
+### CAR_USE
+
+**Theoretical Effect**
+Commercian vehicles are driven more, so might increase probability of collision.
+
+```{r} 
+hw4_categorical_plot(df, "CAR_USE")
+hw4_target_table(df, "CAR_USE")
+```
+
+**Observed Effect**
+The plot suggests that the use does not increase or decrease the probability of being in an accident or the amount of the claim.
+
+### CLM_FREQ
+
+**Theoretical Effect**
+The more claims you filed in the past, the more you are likely to file in the future.
+
+```{r} 
+hw4_categorical_plot(df, "CLM_FREQ")
+hw4_target_table(df, "CLM_FREQ")
+```
+
+**Observed Effect**
+Looks like there are few people who signed up for the frequent accident club.  All jokes aside, the theoretical effect is not observed in the data.
+
+### EDUCATION
+
+**Theoretical Effect**
+Unknown effect, but in theory more educated people tend to drive more safely
+
+```{r} 
+hw4_categorical_plot(df, "EDUCATION")
+hw4_target_table(df, "EDUCATION")
+```
+
+**Observed Effect**
+The theoretical effect has no basis and is somewhat elitist.
+
+### HOMEKIDS
+
+**Theoretical Effect**
+Unknown effect
+
+```{r} 
+hw4_categorical_plot(df, "HOMEKIDS")
+hw4_target_table(df, "HOMEKIDS")
+```
+
+**Observed Effect**
+It's interesting to note that those with kids have a 0.34 probability of being in an accident.  Those without kids have a 0.22 probability of being in an accident.  This may be useful.
+
+### HOME_VAL
+
+**Theoretical Effect**
+In theory, home owners tend to drive more responsibly.
+
+```{r} 
+hw4_plot(df, "HOME_VAL", TRUE, TRUE)
+```
+
+**Observed Effect**
+There are a lot of zero values.  I am assuming that it is a zero because they are renters.
+
+### INCOME
+
+**Theoretical Effect**
+In theory, rich people trend to get into fewer crashes.
+
+```{r} 
+hw4_plot(df, "INCOME", TRUE, TRUE)
+```
+
+**Observed Effect**
+There are a lot of zero values.  I am assuming that it is a zero because they are renters.
+
+### JOB
+
+**Theoretical Effect**
+In theory, white collar jobs tend to be safer
+
+```{r} 
+hw4_categorical_plot(df, "JOB")
+hw4_target_table(df, "JOB")
+```
+
+**Observed Effect**
+There is a 0.21 probability for white collar job holders being in an accident.  there's a 0.34 probability for the blue collar jobs.  There may be some truth to the theoretical effect.
+
+### KIDSDRIV
+
+**Theoretical Effect**
+When teenagers drive your car, ou are more likely to get into crashes
+
+```{r} 
+hw4_categorical_plot(df, "KIDSDRIV")
+hw4_target_table(df, "KIDSDRIV")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### MSTATUS
+
+**Theoretical Effect**
+In theory, married people drive more safely
+
+```{r} 
+hw4_categorical_plot(df, "MSTATUS")
+hw4_target_table(df, "MSTATUS")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### MVR_PTS
+
+**Theoretical Effect**
+If you get lots of traffic tickets, you trend to get into more crashes
+
+```{r} 
+hw4_plot(df, "MVR_PTS")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### OLDCLAIM
+
+**Theoretical Effect**
+If your total payout over the past five years was high, you are more likely to get into crashes.
+
+```{r} 
+hw4_plot(df, "OLDCLAIM", TRUE, TRUE)
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### PARENT1
+
+**Theoretical Effect**
+Unknown effect
+
+```{r} 
+hw4_categorical_plot(df, "PARENT1")
+hw4_target_table(df, "PARENT1")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### RED_CAR
+
+**Theoretical Effect**
+Urban legend says that red cars (espcially red sports cars) are more risky. Is that true?
+
+```{r} 
+hw4_categorical_plot(df, "RED_CAR")
+hw4_target_table(df, "RED_CAR")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### REVOKED
+
+**Theoretical Effect**
+If your license was revoked in the pat 7 years, you probably are a more risky driver.
+
+```{r}
+hw4_categorical_plot(df, "REVOKED")
+hw4_target_table(df, "REVOKED")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### SEX
+
+**Theoretical Effect**
+Urban legend says that women have less crashes then men.  Is that true?
+
+```{r}
+hw4_categorical_plot(df, "SEX")
+hw4_target_table(df, "SEX")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### TIF
+
+**Theoretical Effect**
+People who have been customers for a long time are usually more safe.
+
+```{r}
+hw4_plot(df, "TIF")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### TRAVTIME
+
+**Theoretical Effect**
+Log drives to work usually suggest greater risk
+
+```{r}
+hw4_plot(df, "TRAVTIME")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### URBANICITY
+
+**Theoretical Effect**
+Unknown
+
+```{r}
+hw4_categorical_plot(df, "URBANICITY")
+hw4_target_table(df, "URBANICITY")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+
+### YOJ
+
+**Theoretical Effect**
+People who stay at a job for a long time are usuaslly more safe
+
+```{r}
+hw4_plot(df, "YOJ")
+```
+
+**Observed Effect**
+INSERT TEXT HERE
+Looking at the data, we can see that there are some missing values:

### 2019-11-09 hw4/HW4.Rmd
Added baseline model

@@
 ```
+
+## Model Creation
+
+### Baseline Model
+
+```{r}
+evaluate_model <- function(model, test_df){
+  test_df$yhat <- ifelse(predict.glm(model, test_df,"response") >= 0.5, 1, 0)
+  cm <- confusionMatrix(factor(test_df$yhat), factor(test$TARGET_FLAG), "1")
+  deviance <- model$deviance
+  r2 <- 1 - model$deviance / model$null.deviance
+
+  cat("F1 =", cm$byClass[7],"\nR2 =", r2, "\n\n")
+  print(cm)
+  return(cm)
+}
+```
+
+We will create a simple model to serve as the baseline.
+This model posits that drivers history is a good representation of their future.
+So drivers that have been in an accident (`OLDCLAIM` > 0 or `AVG_CLAIM` > 0) are more likely to be in an accident.
+Conversely those who haven't aren't likely to be in one in the future.
+
+```{r, null_prefix = TRUE, comment=NA}
+baseline_model <- over_sample_train %>%
+  mutate(PRIOR_ACCIDENT = ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)) %>%
+  glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), .)
+summary(baseline_model)
+
+results <- test %>%
+  mutate(PRIOR_ACCIDENT = ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)) %>%
+  evaluate_model(baseline_model, .)
+```
+
+This simple model has statistically significant predictors.
+Applying this model to the test data set indicates this simple model has a `r  round(unname(results$overall['Accuracy'])*100,1)`% accuracy rate.
+It correcly identified `r round(unname(results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(results$byClass['Specificity'])*100,1)`% of those without.
+It is better than flipping a coin but not by much.
+All future models must outpreform this baseline.

### 2019-11-10 hw4/HW4.Rmd
More work on HW4

@@
 
+## Feature Creation
+
+```{r}
+create_features <- function(d){
+  d %>%
+    mutate(PRIOR_ACCIDENT = ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)) %>%
+    mutate(COLLEGE_EDUCATED = ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)) %>%
+    mutate(URBAN_DRIVER = ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)) %>%
+    mutate(YOUNG_MALE = ifelse(SEX == "M" & AGE < 21, 1, 0)) %>%
+    mutate(RED_SPORTS_CAR = ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)) %>%
+    mutate(HAS_KIDS = ifelse(HOMEKIDS == 0, 0, 1)) %>%
+    mutate(KID_DRIVERS = ifelse(KIDSDRIV == 0, 0, 1))
+}
+
+over_sample_train <- create_features(over_sample_train)
+test <- create_features(test)
+```
+
 ## Model Creation
@@
 ```{r, null_prefix = TRUE, comment=NA}
-baseline_model <- over_sample_train %>%
-  mutate(PRIOR_ACCIDENT = ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)) %>%
-  glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), .)
+baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), over_sample_train)
 summary(baseline_model)
 
-results <- test %>%
-  mutate(PRIOR_ACCIDENT = ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)) %>%
-  evaluate_model(baseline_model, .)
+results <- evaluate_model(baseline_model, test)
 ```
@@
 All future models must outpreform this baseline.
+
+### Risk Taker Model
+
+For this model we are going assume that risk takers are more likely to be in an accident.
+To identify those who tend to be risk takers, we are going with the urban legend that those who drive red sports cars are risk takers.
+Also young men (16-20) also tend to be risk takers.
+
+```{r, null_prefix = TRUE, comment=NA}
+risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample_train)
+summary(risk_taker_model)
+
+risk_taker_results <- evaluate_model(risk_taker_model, test)
+```
+
+This model does not have statistically significant predictors, however it has a `r  round(unname(risk_taker_results$overall['Accuracy'])*100,1)`% accuracy rate.
+This is because it identified `r  round(unname(risk_taker_results$byClass['Specificity'])*100,1)`% of those who didn't have an accident (and they are more numerous).
+When looking at the model's performance on identifying the minority group of interest we see it correcly identified `r round(unname(risk_taker_results$byClass['Sensitivity'])*100,1)`% of the people with accidents.
+So while the accuracy rate out preformed the baseline model, this one will not be the model we will use.
+
+### Misc
+
+```{r, null_prefix = TRUE, comment=NA}
+model <- glm(TARGET_FLAG ~ CAR_USE, family = binomial(link = "logit"), over_sample_train)
+summary(model)
+
+model_results <- evaluate_model(model, test)
+```

### 2019-11-10 final project/Manuscript.tex
Added more lit review to manuscript

@@
 \item
-  C4.5 (Seewald, 2000);
+  C4.5 (Seewald, 2000), (Rickets, 2000);
 \item
@@
 \item
+  Fuzzy Classifier (János Abonyi, 2000);
+\item
+  Genetic Algorithms and Hill-climbers (Carter, 2000);
+\item
   Inductive Learning by Logic Minimization (ILLM) (Gamberger, 2000);
 \item
+  K-Means (Vesanto \& Sinkkonen, 2000);
+\item
+  KXEN (Bera \& Lamy, 2000);
+\item
   Mask Perceptron with Boosting (Leckie \& Ferra, 2000);
+\item
+  Midos Algorithm (Krogel, 2000);
 \item
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Bera}{}%
+Bera, M., \& Lamy, B. (2000). Kxen at coil challenge 2000. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/BERAPS~1.pdf}
+
+\leavevmode\hypertarget{ref-Carter}{}%
+Carter, J. (2000). Coil 2000 challenge submission: Genetic algorithms and hill-climbers. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/CARTER~1.pdf}
+
 \leavevmode\hypertarget{ref-Crocoll}{}%
@@
 
+\leavevmode\hypertarget{ref-Abonyi}{}%
+János Abonyi, H. R. (2000). A simple fuzzy classifier based on inconsistency analysis of labeled data. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/ABONYI~1.pdf}
+
 \leavevmode\hypertarget{ref-Kim}{}%
@@
 
+\leavevmode\hypertarget{ref-Krogel}{}%
+Krogel, M.-A. (2000). A data mining case study. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/KROGEL~1.pdf}
+
 \leavevmode\hypertarget{ref-Leckie}{}%
@@
 
+\leavevmode\hypertarget{ref-Rickets}{}%
+Rickets, P. (2000). CoIL challenge 2000 submission. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/RICKET~1.pdf}
+
 \leavevmode\hypertarget{ref-Seewald}{}%
 Seewald, A. (2000). CoIL challenge 2000 submitted solution. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/SEEWAL~1.pdf}
+
+\leavevmode\hypertarget{ref-Vesanto}{}%
+Vesanto, J., \& Sinkkonen, J. (2000). Submission for the coil challenge 2000. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/VESANT~1.pdf}
 
@@
 \section{Appendices}\label{appendices}}
-
-\begin{itemize}
-\tightlist
-\item
-  Supplemental tables and/or figures.
-\end{itemize}
 

### 2019-11-10 hw4/HW4.Rmd
Update to HW4

@@
 ```{r, null_prefix = TRUE, comment=NA}
-model <- glm(TARGET_FLAG ~ CAR_USE, family = binomial(link = "logit"), over_sample_train)
+#"KIDSDRIV"         "AGE"             
+#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
+#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
+#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
+#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
+#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
+#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
+#[36] "KID_DRIVERS"     
+model <- glm(TARGET_FLAG ~ KID_DRIVERS, family = binomial(link = "logit"), over_sample_train)
 summary(model)

### 2019-11-11 final project/Manuscript.Rmd
Updated literature review section of final paper

@@
 
-Participants used a variety of approaches in formulating their submissions including:
-
-- Classification Trees with Bagging [@White];
-
-- C4.5 [@Seewald];
-
-- Evolutionary Algorithm [@Koudijs];
-
-- Inductive Learning by Logic Minimization (ILLM) [@Gamberger];
-
-- Mask Perceptron with Boosting [@Leckie];
-
-- Naïve Bayes [@Elkan]; and
-
-- Neural Networks[@Crocoll], [@Kim];
-
-- Scoring System [@Lewandowski];
-
-- Support Vector Machines; and
-
-- XCS [@Greenyer].
+Participants used a variety of approaches in formulating their submissions including: Boosted Decision Tree [@McKone], Classification and Regressio Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].
 
@@
 [@Putten]
-
-<!--
-10 Gustavo Denicolay and Gaston Morales. CoIL Challenge 2000 Submission for the Description Task. [ps.gz] [pdf] - Not sure what they used
-
-12 János Abonyi, Hans Roubos. A Simple Fuzzy Classifier based on Inconsistency Analysis of Labeled Data. [ps.gz],[pdf]
-
-13 Jonathan Carter. Coil 2000 Challenge Submission: Genetic Algorithms and Hill-climbers. [ps.gz], [pdf]
-
-14 Juha Vesanto and Janne Sinkkonen. Submission for the CoIL Challenge 2000. [ps.gz], [pdf]
-
-15 Mark-André Krogel. A Data Mining Case Study. [ps.gz],[pdf]
-
-16 Michel Bera and Bertrand Lamy. Kxen at CoIL Challenge 2000. [ps.gz], [pdf]
-
-17 Paul Rickets. CoIL Challenge 2000 submission. [ps.gz], [pdf]
-
-18 Petr Mikšovský, Jirí Klema. COIL Challenge 2000. [ps.gz], [pdf]
-
-19 Petri Kontkanen. CoIL 2000 Submission. [ps.gz], [pdf]
-
-20 Philip Brierley. COIL 2000 Challenge: Characteristics of caravan insurance policy owners. [ps.gz] [pdf]
-
-21 Robert M. Simmonds. ACT Study report using Classification and Regression Tree (CART) Analysis. [ps.gz], [pdf]
-
-22 S. Sathiya Keerthi and Chong Jin Ong. Solution of the CoIL Challenge 2000 Task Using Support Vector Machines. [ps.gz], [pdf]
-
-23 Serhiy Shtovba, Yakiv Mashnitskiy. The Backpropagation Multilayer Feedforward Neural Network Based Competition Task Solution.
-
-24 Serhiy Shtovba. Phase Intervals and Genetic Algorithms Based Competition Task Solution. [ps.gz], [pdf]
-
-25 Jurgen A. Doornik and Steve Moyle. LOGIT Modelling. [ps.gz], [pdf]
-
-26 Thomas Martini Jorgensen and Christian Linneberg. Subspace Projections – an approach to variable selection and modeling. [ps.gz], [pdf]
-
-27 Tom McKone and Curt Stenger. COIL CHALLENGE 2000 Submission. [ps.gz], [pdf]
-
-28 Tomislav Šmuc. COIL 2000 Challenge Solution based on ILLM-SG Methodology. [ps.gz], [pdf]
-
-29 Uzay Kaymak and Magne Setnes. Target Selection based on fuzzy clustering: a volume prototype approach to CoIL Challenge 2000. [ps.gz], [pdf]
-
--->
 
@@
 # Appendices
-
+<!--
 - Supplemental tables and/or figures.
-
+-->
 ## R statistical programming code.

### 2019-11-11 final project/Manuscript.tex
Updated literature review section of final paper

@@
 
-Participants used a variety of approaches in formulating their submissions including:
-
-\begin{itemize}
-\item
-  Classification Trees with Bagging (White \& Liu, 2000);
-\item
-  C4.5 (Seewald, 2000), (Rickets, 2000);
-\item
-  Evolutionary Algorithm (Koudijs, 2000);
-\item
-  Fuzzy Classifier (János Abonyi, 2000);
-\item
-  Genetic Algorithms and Hill-climbers (Carter, 2000);
-\item
-  Inductive Learning by Logic Minimization (ILLM) (Gamberger, 2000);
-\item
-  K-Means (Vesanto \& Sinkkonen, 2000);
-\item
-  KXEN (Bera \& Lamy, 2000);
-\item
-  Mask Perceptron with Boosting (Leckie \& Ferra, 2000);
-\item
-  Midos Algorithm (Krogel, 2000);
-\item
-  Naïve Bayes (Elkan, 2000); and
-\item
-  Neural Networks(Crocoll, 2000), (Kim \& Street, 2000);
-\item
-  Scoring System (Lewandowski, 2000);
-\item
-  Support Vector Machines; and
-\item
-  XCS (Greenyer, 2000).
-\end{itemize}
+Participants used a variety of approaches in formulating their submissions including: Boosted Decision Tree (McKone \& Stenger, 2000), Classification and Regressio Tree (CART) (Simmonds, 2000), Classification Trees with Bagging (White \& Liu, 2000), C4.5 (Rickets, 2000; Seewald, 2000), Evolutionary Algorithm (Koudijs, 2000), Fuzzy Classifier (János Abonyi, 2000; Kaymak \& Setnes, 2000), Genetic Algorithms and Hill-climbers (Carter, 2000), Inductive Learning by Logic Minimization (ILLM) (Gamberger, 2000; Šmuc, 2000), Instance Based Reasoning (iBARET) (Mikšovský \& Klema, 2000), K-Means (Vesanto \& Sinkkonen, 2000), KXEN (Bera \& Lamy, 2000), LOGIT (Doornik \& Moyle, 2000), Mask Perceptron with Boosting (Leckie \& Ferra, 2000), Midos Algorithm (Krogel, 2000), N-Tuple Classifier (Jorgensen \& Linneberg, 2000), Naïve Bayes (Elkan, 2000; Kontkanen, 2000), Neural Networks(Brierley, 2000; Crocoll, 2000; Kim \& Street, 2000; Shtovba \& Mashnitskiy, 2000), Phase Intervals and Genetic Algorithms (Shtovba, 2000), Scoring System (Lewandowski, 2000), Support Vector Machines(Keerthi \& Ong, 2000), and XCS (Greenyer, 2000).
 
@@
 
+\leavevmode\hypertarget{ref-Brierley}{}%
+Brierley, P. (2000). COIL 2000 challenge: Characteristics of caravan insurance policy owners. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/BRIERL~1.pdf}
+
 \leavevmode\hypertarget{ref-Carter}{}%
@@
 Crocoll, W. M. (2000). Artificial neural network portion of coil study. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/CROCOL~1.pdf}
+
+\leavevmode\hypertarget{ref-Doornik}{}%
+Doornik, J. A., \& Moyle, S. (2000). LOGIT modelling. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/MOYLEP~1.pdf}
 
@@
 
+\leavevmode\hypertarget{ref-Jorgensen}{}%
+Jorgensen, T.
+M., \& Linneberg, C.
+(2000).
+Subspace projections -- an approach to variable selection and modeling.
+Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/JORGEN~1.pdf}
+
+\leavevmode\hypertarget{ref-Kaymak}{}%
+Kaymak, U., \& Setnes, M.
+(2000).
+Target selection based on fuzzy clustering: A volume prototype approach to coil challenge 2000.
+Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/KAYMAK~1.pdf}
+
+\leavevmode\hypertarget{ref-Keerthi}{}%
+Keerthi, S.
+S., \& Ong, C.
+J.
+(2000).
+Solution of the coil challenge 2000 task using support vector machines.
+Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/KEERTH~1.pdf}
+
 \leavevmode\hypertarget{ref-Kim}{}%
@@
 Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/STREET~1.pdf}
+
+\leavevmode\hypertarget{ref-Kontkanen}{}%
+Kontkanen, P. (2000). CoIL 2000 submission. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/KONTKA~1.pdf}
 
@@
 
+\leavevmode\hypertarget{ref-McKone}{}%
+McKone, T., \& Stenger, C. (2000). COIL challenge 2000 submission. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/MCKONE~1.pdf}
+
+\leavevmode\hypertarget{ref-Miksovsky}{}%
+Mikšovský, P., \& Klema, J. (2000). CoIL challenge 2000. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/MIKSOV~1.pdf}
+
 \leavevmode\hypertarget{ref-Putten}{}%
-Putten, P., Ruiter, M., \& Someren, M. (2000). CoIL challenge 2000 tasks and results: Predicting and explaining caravan policy ownership.
+Putten, P., Ruiter, M., \& Someren, M.
+(2000).
+CoIL challenge 2000 tasks and results: Predicting and explaining caravan policy ownership.
+Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf}
 
@@
 Seewald, A. (2000). CoIL challenge 2000 submitted solution. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/SEEWAL~1.pdf}
+
+\leavevmode\hypertarget{ref-Shtovba}{}%
+Shtovba, S. (2000). Phase intervals and genetic algorithms based competition task solution. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/SHTOVB~2.pdf}
+
+\leavevmode\hypertarget{ref-Shtovba_Mashnitskiy}{}%
+Shtovba, S., \& Mashnitskiy, Y.
+(2000).
+The backpropagation multilayer feedforward neural network based competition task solution.
+Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/SHTOVB~1.pdf}
+
+\leavevmode\hypertarget{ref-Simmonds}{}%
+Simmonds, R. M. (2000). ACT study report using classification and regression tree (cart) analysis. Retrieved from \url{http://www.liacs.nl/~putten/library/cc2000/SIMMON~1.pdf}
+
+\leavevmode\hypertarget{ref-Smuc}{}%
+Šmuc, T. (2000). COIL 2000 challenge solution based on illm-sg methodology. Retrieved from \url{http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/SMUCPS~1.pdf}
 

### 2019-11-11 final project/Manuscript.Rmd
More work on final project

@@
 <!--Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are.
-Explain how your investigation is similar or different to the state-of-theart.
+Explain how your investigation is similar or different to the state-of-the-art.
 Please cite the relevant papers where appropriate.-->
@@
 
+```{r}
+# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
+prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121),
+                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
+ggplot(data=prediction_scores, aes(x=score, y=n)) +
+  geom_bar(stat="identity") +
+  xlab("Policy Owners") +
+  ylab("Submissions") +
+  xlim(37, 122) +
+  ggtitle("Correct Predictions from CoIL 2000 Challenge Submissions")
+```
+
 The maximum number of policyowners that could be found was 238.
-The winning model [@Elkan] selected 121 policy owners.
-Random selection results in 42 policy owners.
-The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear!) policy owners.
-[@Putten]
+The submissions identified 95 policy owners on average.
+The winning model [@Elkan] identified 121 policy owners.
+Random selection results in identifying 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners.
+[@Putten].
+
+Our investigation is similar to @Doornik in that we used a LOGIT model. **Reproduce his model and describe the results**
 

### 2019-11-11 final project/Manuscript.tex
More work on final project

@@
 
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-1-1.pdf}
+
 The maximum number of policyowners that could be found was 238.
-The winning model (Elkan, 2000) selected 121 policy owners.
-Random selection results in 42 policy owners.
-The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear!) policy owners.
-(Putten et al., 2000)
+The submissions identified 95 policy owners on average.
+The winning model (Elkan, 2000) identified 121 policy owners.
+Random selection results in identifying 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners.
+(Putten et al., 2000).
 

### 2019-11-11 final project/Manuscript.Rmd
Updated Final Project with data dictionary

@@
 library("papaja")
-
+library(kableExtra)
 library(tidyverse)
@@
 # Appendices
+
 <!--
@@
 -->
+
 ## R statistical programming code.
@@
 </code>
+
+\newpage
+## Data Dictionary
+
+```{r, results='asis'}
+data_dictionary <- data.frame(
+  Name=c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN"),
+  Description = c("Customer Subtype", "Number of houses", "Avg size household", "Avg age", "Customer main type", "Roman catholic", "Protestant", "Other religion", "No religion", "Married", "Living together", "Other relation", "Singles", "Household without children", "Household with children", "High level education", "Medium level education", "Lower level education", "High status", "Entrepreneur", "Farmer", "Middle management", "Skilled labourers", "Unskilled labourers", "Social class A", "Social class B1", "Social class B2", "Social class C", "Social class D", "Rented house", "Home owners", "1 car", "2 cars", "No car", "National Health Service", "Private health insurance", "Income < 30.000", "Income 30-45.000", "Income 45-75.000", "Income 75-122.000", "Income >123.000", "Average income", "Purchasing power class", "Contribution private third party insurance", "Contribution third party insurance (firms)", "Contribution third party insurane (agriculture)", "Contribution car policies", "Contribution delivery van policies", "Contribution motorcycle/scooter policies", "Contribution lorry policies", "Contribution trailer policies", "Contribution tractor policies", "Contribution agricultural machines policies", "Contribution moped policies", "Contribution life insurances", "Contribution private accident insurance policies", "Contribution family accidents insurance policies", "Contribution disability insurance policies", "Contribution fire policies", "Contribution surfboard policies", "Contribution boat policies", "Contribution bicycle policies", "Contribution property insurance policies", "Contribution social security insurance policies", "Number of private third party insurance", "Number of third party insurance (firms)", "Number of third party insurane (agriculture)", "Number of car policies", "Number of delivery van policies", "Number of motorcycle/scooter policies", "Number of lorry policies", "Number of trailer policies", "Number of tractor policies", "Number of agricultural machines policies", "Number of moped policies", "Number of life insurances", "Number of private accident insurance policies", "Number of family accidents insurance policies", "Number of disability insurance policies", "Number of fire policies", "Number of surfboard policies", "Number of boat policies", "Number of bicycle policies", "Number of property insurance policies", "Number of social security insurance policies", "Number of mobile home policy"))
+#apa_table(data_dictionary, longtable = T, booktabs = T)
+knitr::kable(data_dictionary, "latex", longtable = T) %>%
+  kable_styling(latex_options = c("repeat_header"))
+```

### 2019-11-11 final project/Manuscript.tex
Updated Final Project with data dictionary

@@
 (Putten et al., 2000).
+
+Our investigation is similar to Doornik and Moyle (2000) in that we used a LOGIT model. \textbf{Reproduce his model and describe the results}
 
@@
 
+\newpage
+
+\hypertarget{data-dictionary}{%
+\subsection{Data Dictionary}\label{data-dictionary}}
+
+\begin{longtable}{l|l}
+\hline
+Name & Description\\
+\hline
+\endfirsthead
+\multicolumn{2}{@{}l}{\textit{(continued)}}\\
+\hline
+Name & Description\\
+\hline
+\endhead
+MOSTYPE & Customer Subtype\\
+\hline
+MAANTHUI & Number of houses\\
+\hline
+MGEMOMV & Avg size household\\
+\hline
+MGEMLEEF & Avg age\\
+\hline
+MOSHOOFD & Customer main type\\
+\hline
+MGODRK & Roman catholic\\
+\hline
+MGODPR & Protestant\\
+\hline
+MGODOV & Other religion\\
+\hline
+MGODGE & No religion\\
+\hline
+MRELGE & Married\\
+\hline
+MRELSA & Living together\\
+\hline
+MRELOV & Other relation\\
+\hline
+MFALLEEN & Singles\\
+\hline
+MFGEKIND & Household without children\\
+\hline
+MFWEKIND & Household with children\\
+\hline
+MOPLHOOG & High level education\\
+\hline
+MOPLMIDD & Medium level education\\
+\hline
+MOPLLAAG & Lower level education\\
+\hline
+MBERHOOG & High status\\
+\hline
+MBERZELF & Entrepreneur\\
+\hline
+MBERBOER & Farmer\\
+\hline
+MBERMIDD & Middle management\\
+\hline
+MBERARBG & Skilled labourers\\
+\hline
+MBERARBO & Unskilled labourers\\
+\hline
+MSKA & Social class A\\
+\hline
+MSKB1 & Social class B1\\
+\hline
+MSKB2 & Social class B2\\
+\hline
+MSKC & Social class C\\
+\hline
+MSKD & Social class D\\
+\hline
+MHHUUR & Rented house\\
+\hline
+MHKOOP & Home owners\\
+\hline
+MAUT1 & 1 car\\
+\hline
+MAUT2 & 2 cars\\
+\hline
+MAUT0 & No car\\
+\hline
+MZFONDS & National Health Service\\
+\hline
+MZPART & Private health insurance\\
+\hline
+MINKM30 & Income < 30.000\\
+\hline
+MINK3045 & Income 30-45.000\\
+\hline
+MINK4575 & Income 45-75.000\\
+\hline
+MINK7512 & Income 75-122.000\\
+\hline
+MINK123M & Income >123.000\\
+\hline
+MINKGEM & Average income\\
+\hline
+MKOOPKLA & Purchasing power class\\
+\hline
+PWAPART & Contribution private third party insurance\\
+\hline
+PWABEDR & Contribution third party insurance (firms)\\
+\hline
+PWAAND & Contribution third party insurane (agriculture)\\
+\hline
+PPERSAUT & Contribution car policies\\
+\hline
+PBESAUT & Contribution delivery van policies\\
+\hline
+PMOTSCO & Contribution motorcycle/scooter policies\\
+\hline
+PVRAAUT & Contribution lorry policies\\
+\hline
+PAANHANG & Contribution trailer policies\\
+\hline
+PTRACTOR & Contribution tractor policies\\
+\hline
+PWERKT & Contribution agricultural machines policies\\
+\hline
+PBROM & Contribution moped policies\\
+\hline
+PLEVEN & Contribution life insurances\\
+\hline
+PPERSONG & Contribution private accident insurance policies\\
+\hline
+PGEZONG & Contribution family accidents insurance policies\\
+\hline
+PWAOREG & Contribution disability insurance policies\\
+\hline
+PBRAND & Contribution fire policies\\
+\hline
+PZEILPL & Contribution surfboard policies\\
+\hline
+PPLEZIER & Contribution boat policies\\
+\hline
+PFIETS & Contribution bicycle policies\\
+\hline
+PINBOED & Contribution property insurance policies\\
+\hline
+PBYSTAND & Contribution social security insurance policies\\
+\hline
+AWAPART & Number of private third party insurance\\
+\hline
+AWABEDR & Number of third party insurance (firms)\\
+\hline
+AWALAND & Number of third party insurane (agriculture)\\
+\hline
+APERSAUT & Number of car policies\\
+\hline
+ABESAUT & Number of delivery van policies\\
+\hline
+AMOTSCO & Number of motorcycle/scooter policies\\
+\hline
+AVRAAUT & Number of lorry policies\\
+\hline
+AAANHANG & Number of trailer policies\\
+\hline
+ATRACTOR & Number of tractor policies\\
+\hline
+AWERKT & Number of agricultural machines policies\\
+\hline
+ABROM & Number of moped policies\\
+\hline
+ALEVEN & Number of life insurances\\
+\hline
+APERSONG & Number of private accident insurance policies\\
+\hline
+AGEZONG & Number of family accidents insurance policies\\
+\hline
+AWAOREG & Number of disability insurance policies\\
+\hline
+ABRAND & Number of fire policies\\
+\hline
+AZEILPL & Number of surfboard policies\\
+\hline
+APLEZIER & Number of boat policies\\
+\hline
+AFIETS & Number of bicycle policies\\
+\hline
+AINBOED & Number of property insurance policies\\
+\hline
+ABYSTAND & Number of social security insurance policies\\
+\hline
+CARAVAN & Number of mobile home policy\\
+\hline
+\end{longtable}
+
 \end{document}

### 2019-11-12 hw4/HW4.Rmd

--- 2019-11-10 hw4/HW4.Rmd
@@
 library(RColorBrewer)
+library(ROCR)
 # Thank you Stack Overflow!
@@
 
-We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.
-Each record has two response variables.
-The first is a binary flag where one means a person was in a car crash and zero means they were not.
-There are `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` rows with a one flag and `r df %>% filter(TARGET_FLAG == 0) %>% nrow()` with a zero flag.
+We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.  Each record has two response variables.
+
+The first objective is to train a logistic regression classifier to predict if a person was in a car crash.
+The response variable here is `TARGET_FLAG` respresenting whether a person had an accident (1) or did not have one (0).
 
@@
 
-The first objective is to train a logistic regression classifier to predict if a person was in a car crash.
-The second reponse variable is the amount it will cost if the person crashes their car.
+The second objective will be to train a regression model to predict the cost of a crash, if one occurred.
+The second reponse variable, `TARGET_VALUE`, is the amount it will cost if the person crashes their car.
 The value is zero if the person did not crash their car.
-
-The second objective will be to train a regression model to predict the cost of a crash, if one occurred.
 
@@
 
+df$TARGET_FLAG <- factor(df$TARGET_FLAG)
+
 df <- df %>%
@@
 ```{r}
-feature_creation <- function(d){
+# Function to add features
+
+create_features <- function(d){
   d %>%
-    rowwise() %>%
     mutate(LOG_INCOME = log(INCOME + 1),
            LOG_HOME_VAL = log(HOME_VAL + 1),
-           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0)) %>%
-    ungroup()
+           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0),
+           PRIOR_ACCIDENT = factor(ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)),
+           COLLEGE_EDUCATED = factor(ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)),
+           URBAN_DRIVER = factor(ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)),
+           YOUNG_MALE = factor(ifelse(SEX == "M" & AGE < 25, 1, 0)),
+           YOUNG = factor(ifelse(AGE < 25, 1, 0)),
+           RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
+           HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
+           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)))
 }
 
-df <- feature_creation(df)
-evaluation <- feature_creation(evaluation)
+df <- create_features(df)
+evaluation <- create_features(evaluation)
 ```
@@
   labs(x="TARGET_FLAG", y="# Observations")
-```
 
-## Feature Creation
-
-```{r}
-create_features <- function(d){
-  d %>%
-    mutate(PRIOR_ACCIDENT = ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)) %>%
-    mutate(COLLEGE_EDUCATED = ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)) %>%
-    mutate(URBAN_DRIVER = ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)) %>%
-    mutate(YOUNG_MALE = ifelse(SEX == "M" & AGE < 21, 1, 0)) %>%
-    mutate(RED_SPORTS_CAR = ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)) %>%
-    mutate(HAS_KIDS = ifelse(HOMEKIDS == 0, 0, 1)) %>%
-    mutate(KID_DRIVERS = ifelse(KIDSDRIV == 0, 0, 1))
-}
-
-over_sample_train <- create_features(over_sample_train)
-test <- create_features(test)
+over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
 ```
@@
 To identify those who tend to be risk takers, we are going with the urban legend that those who drive red sports cars are risk takers.
-Also young men (16-20) also tend to be risk takers.
+Also young men (16-24) also tend to be risk takers.
 
@@
 
-This model does not have statistically significant predictors, however it has a `r  round(unname(risk_taker_results$overall['Accuracy'])*100,1)`% accuracy rate.
+This model only has one statistically significant predictor, however it has a `r  round(unname(risk_taker_results$overall['Accuracy'])*100,1)`% accuracy rate.
 This is because it identified `r  round(unname(risk_taker_results$byClass['Specificity'])*100,1)`% of those who didn't have an accident (and they are more numerous).
@@
 
+### Traditional Model
+
+Traditional wisdom holds that there are a few tried and true predictors of someone's risk of an automobile accident.
+Namely: age (under 25 vs.
+25+), marital status, driving record, distance to work, and the driver's sex (which may or may not be proper to use, but seems to have held throughout the years).
+
+```{r}
+traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
+                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
+             family = binomial(link = "logit"), over_sample_train)
+
+summary(traditional_model)
+
+model_results <- evaluate_model(traditional_model, test)
+```
+
+### Traditional Model with Cross-Validation
+
+Here we try to use cross-validation techniques to improve the traditional model.
+We go back to the raw training data, because it is typically not-advised to use data that is oversampled prior to cross-validation.
+
+```{r}
+# Get complete cases
+cases <- train %>%
+  select(YOUNG,MSTATUS,PRIOR_ACCIDENT,SEX,REVOKED,MVR_PTS, TRAVTIME, CAR_USE) %>%
+  complete.cases()
+
+temp <- train[cases,]
+
+# Use 5-fold cross-validation
+train_control <- trainControl(method="cv",number=5, sampling="up")
+
+traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
+                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
+                        method = "glm", family = "binomial", 
+                        data = temp, trControl = train_control)
+
+traditional_cv
+```
+
+```{r}
+# Evaluating the model
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = 
+predict(traditional_cv,newdata=test,type="prob"))
+
+confusionMatrix(eval$predicted, eval$actual, positive = "1")
+
+pred <- prediction(eval$prob.1, eval$actual)
+auc <- performance(pred, measure = "auc")@y.values[[1]]
+perf <- performance(pred,"tpr","fpr")
+plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
+```
+
 ### Misc
@@
 ```{r, null_prefix = TRUE, comment=NA}
-model <- glm(TARGET_FLAG ~ CAR_USE, family = binomial(link = "logit"), over_sample_train)
+#"KIDSDRIV"         "AGE"             
+#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
+#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
+#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
+#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
+#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
+#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
+#[36] "KID_DRIVERS"     
+model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
+             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
+             family = binomial(link = "logit"), over_sample_train)
 summary(model)

### 2019-11-12 final project/Manuscript.tex
Updated final project's data dictionary

--- 2019-11-11 final project/Manuscript.tex
@@
 
+\newpage
+
+\hypertarget{data-dictionary}{%
+\subsection{Data Dictionary}\label{data-dictionary}}
+
+\begin{center}
+\begin{ThreePartTable}
+\begin{longtable}{ll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:unnamed-chunk-3}Data Dictionary}\\
+\toprule
+Name & \multicolumn{1}{c}{Description}\\
+\midrule
+MOSTYPE & Customer Subtype\\
+MAANTHUI & Number of houses\\
+MGEMOMV & Avg size household\\
+MGEMLEEF & Avg age\\
+MOSHOOFD & Customer main type\\
+MGODRK & Roman catholic\\
+MGODPR & Protestant\\
+MGODOV & Other religion\\
+MGODGE & No religion\\
+MRELGE & Married\\
+MRELSA & Living together\\
+MRELOV & Other relation\\
+MFALLEEN & Singles\\
+MFGEKIND & Household without children\\
+MFWEKIND & Household with children\\
+MOPLHOOG & High level education\\
+MOPLMIDD & Medium level education\\
+MOPLLAAG & Lower level education\\
+MBERHOOG & High status\\
+MBERZELF & Entrepreneur\\
+MBERBOER & Farmer\\
+MBERMIDD & Middle management\\
+MBERARBG & Skilled labourers\\
+MBERARBO & Unskilled labourers\\
+MSKA & Social class A\\
+MSKB1 & Social class B1\\
+MSKB2 & Social class B2\\
+MSKC & Social class C\\
+MSKD & Social class D\\
+MHHUUR & Rented house\\
+MHKOOP & Home owners\\
+MAUT1 & 1 car\\
+MAUT2 & 2 cars\\
+MAUT0 & No car\\
+MZFONDS & National Health Service\\
+MZPART & Private health insurance\\
+MINKM30 & Income < 30.000\\
+MINK3045 & Income 30-45.000\\
+MINK4575 & Income 45-75.000\\
+MINK7512 & Income 75-122.000\\
+MINK123M & Income >123.000\\
+MINKGEM & Average income\\
+MKOOPKLA & Purchasing power class\\
+PWAPART & Contribution private third party insurance\\
+PWABEDR & Contribution third party insurance (firms)\\
+PWAAND & Contribution third party insurane (agriculture)\\
+PPERSAUT & Contribution car policies\\
+PBESAUT & Contribution delivery van policies\\
+PMOTSCO & Contribution motorcycle/scooter policies\\
+PVRAAUT & Contribution lorry policies\\
+PAANHANG & Contribution trailer policies\\
+PTRACTOR & Contribution tractor policies\\
+PWERKT & Contribution agricultural machines policies\\
+PBROM & Contribution moped policies\\
+PLEVEN & Contribution life insurances\\
+PPERSONG & Contribution private accident insurance policies\\
+PGEZONG & Contribution family accidents insurance policies\\
+PWAOREG & Contribution disability insurance policies\\
+PBRAND & Contribution fire policies\\
+PZEILPL & Contribution surfboard policies\\
+PPLEZIER & Contribution boat policies\\
+PFIETS & Contribution bicycle policies\\
+PINBOED & Contribution property insurance policies\\
+PBYSTAND & Contribution social security insurance policies\\
+AWAPART & Number of private third party insurance\\
+AWABEDR & Number of third party insurance (firms)\\
+AWALAND & Number of third party insurane (agriculture)\\
+APERSAUT & Number of car policies\\
+ABESAUT & Number of delivery van policies\\
+AMOTSCO & Number of motorcycle/scooter policies\\
+AVRAAUT & Number of lorry policies\\
+AAANHANG & Number of trailer policies\\
+ATRACTOR & Number of tractor policies\\
+AWERKT & Number of agricultural machines policies\\
+ABROM & Number of moped policies\\
+ALEVEN & Number of life insurances\\
+APERSONG & Number of private accident insurance policies\\
+AGEZONG & Number of family accidents insurance policies\\
+AWAOREG & Number of disability insurance policies\\
+ABRAND & Number of fire policies\\
+AZEILPL & Number of surfboard policies\\
+APLEZIER & Number of boat policies\\
+AFIETS & Number of bicycle policies\\
+AINBOED & Number of property insurance policies\\
+ABYSTAND & Number of social security insurance policies\\
+CARAVAN & Number of mobile home policy\\
+\bottomrule
+\end{longtable}
+\end{ThreePartTable}
+\end{center}
+
 \end{document}

### 2019-11-13 hw4/HW_4_Corey.Rmd
Renamed file

--- 2019-11-12 hw4/HW4.Rmd
@@
 # A Prefix nulling hook.
-
 # Make sure to keep the default for normal processing.
 default_output_hook <- knitr::knit_hooks$get("output")
-
 # Output hooks handle normal R console output.
 knitr::knit_hooks$set( output = function(x, options) {
-
   comment <- knitr::opts_current$get("comment")
@@
   }
-
   default_output_hook( x, options )
-
 })
@@
 evaluation <- read.csv("./data/insurance-evaluation-data.csv")
-
 strip_dollars <- function(x){
@@
 }
-
 fix_data_types <- function(messy_df){
@@
 }
-
 na_bad_values <- function(messy_df){
@@
 }
-
 df$TARGET_FLAG <- factor(df$TARGET_FLAG)
-
 df <- df %>%
@@
 # Function to add features
-
 create_features <- function(d){
@@
 }
-
 df <- create_features(df)
 evaluation <- create_features(evaluation)
+
+df$years <- ifelse(df$TIF >5,5,df$TIF)
+
+df$ClaimFreqPeryear <- df$CLM_FREQ/df$years
+
+df$ClaimAMTPerYear <- df$OLDCLAIM/df$years
+
+df$years <- NULL
 ```
@@
 
-There are `r sum(train$TARGET_FLAG)` out of `r nrow(train)` records in the training data set that have been in an accident.
+There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.
 We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.
@@
   labs(x="TARGET_FLAG", y="# Observations")
-
 over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
@@
 summary(baseline_model)
-
 results <- evaluate_model(baseline_model, test)
@@
 summary(risk_taker_model)
-
 risk_taker_results <- evaluate_model(risk_taker_model, test)
@@
              family = binomial(link = "logit"), over_sample_train)
-
 summary(traditional_model)
-
 model_results <- evaluate_model(traditional_model, test)
@@
   complete.cases()
-
 temp <- train[cases,]
-
 # Use 5-fold cross-validation
 train_control <- trainControl(method="cv",number=5, sampling="up")
-
 traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
@@
                         data = temp, trControl = train_control)
-
 traditional_cv
@@
 predict(traditional_cv,newdata=test,type="prob"))
-
 confusionMatrix(eval$predicted, eval$actual, positive = "1")
-
 pred <- prediction(eval$prob.1, eval$actual)
@@
 #[36] "KID_DRIVERS"     
-model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
-             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
+model <- glm(TARGET_FLAG ~ KID_DRIVERS + MSTATUS + INCOME +
+             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +ClaimAMTPerYear+ClaimFreqPeryear,
              family = binomial(link = "logit"), over_sample_train)
 summary(model)
+model_results <- evaluate_model(model, test)
 
-model_results <- evaluate_model(model, test)
 ```
+
+```{r}
+test$Prediction <- predict.glm(model, newdata=test, type="response")
+
+test$Prediction_Bool <- ifelse(test$Prediction >.5, '1','0')
+
+# Evaluating the model
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = test$Prediction_Bool, prob = 
+predict(traditional_cv,newdata=test,type="prob"))
+confusionMatrix(eval$predicted, eval$actual, positive = "1")
+pred <- prediction(eval$prob.1, eval$actual)
+auc <- performance(pred, measure = "auc")@y.values[[1]]
+perf <- performance(pred,"tpr","fpr")
+plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
+```
+
+#Regression
+
+```{r}
+model <- lm(TARGET_AMT ~ KID_DRIVERS + MSTATUS + INCOME +
+             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +ClaimAMTPerYear+ClaimFreqPeryear,
+             family = poisson, train)
+
+summary(model)
+# model_results <- evaluate_model(model, test)
+
+test$Prediction_Amount <- predict(model, newdata=test, type="response")
+
+cases <- test %>%
+  complete.cases()
+test <- test[cases,]
+
+cor(test$Prediction_Amount,test$TARGET_AMT)
+
+ggplot(test, aes(x=Prediction_Amount, y=TARGET_AMT)) + 
+  geom_point()+
+  geom_smooth(method=lm)
+```

### 2019-11-14 final project/Manuscript.Rmd

--- 2019-11-11 final project/Manuscript.Rmd
@@
 
+```{r data}
+df <- read.delim("data/ticdata2000.txt", header = FALSE)
+names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
+               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
+               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
+               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
+               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
+               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
+               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
+               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
+               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
+               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
+               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
+               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
+               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
+               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
+               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
+               "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
+eval <- read.delim("data/ticeval2000.txt", header = FALSE)
+temp <- read.delim("data/tictgts2000.txt", header = FALSE)
+eval$CARAVAN <- temp$V1
+names(eval) <- names(df)
+```
 # Introduction
@@
 
+```{r}
+doornik <- function(df){
+  df <- df %>%
+    mutate(PBRAND_2_or_6 = ifelse(PBRAND == 2 | PBRAND == 6, 1, 0),
+           PBRAND_3_or_4 = ifelse(PBRAND == 3 | PBRAND == 4, 1, 0),
+           PPLEZIER_gt_0 = ifelse(PPLEZIER > 0, 1, 0),
+           PPERSAUT_6 = ifelse(PPERSAUT == 6, 1, 0))
+  z <- -4.1 - (0.13 * df$MOPLLAAG) - (0.3 * df$MINK123M) + (0.17 * df$MINKGEM) + (0.54 * (df$ALEVEN + df$AFIETS + df$ABYSTAND)) - df$PBRAND_2_or_6 + (0.8 * df$PBRAND_3_or_4) + (2.4 * df$PPLEZIER_gt_0) + (1.5 * df$PPERSAUT_6) + (0.28 * (df$PWAOREG - df$PLEVEN))
+  p_caravan <- exp(z) / (1 + exp(z))
+  #p_caravan
+  ifelse(p_caravan > 0.5, 1, 0)
+}
+
+eval$doornik <- doornik(eval)
+```
+
 # Methodology
@@
 </code>
+
+\newpage
+## Data Dictionary
+
+```{r, results='asis'}
+data_dictionary <- data.frame(
+  Name=c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN"),
+  Description = c("Customer Subtype", "Number of houses", "Avg size household", "Avg age", "Customer main type", "Roman catholic", "Protestant", "Other religion", "No religion", "Married", "Living together", "Other relation", "Singles", "Household without children", "Household with children", "High level education", "Medium level education", "Lower level education", "High status", "Entrepreneur", "Farmer", "Middle management", "Skilled labourers", "Unskilled labourers", "Social class A", "Social class B1", "Social class B2", "Social class C", "Social class D", "Rented house", "Home owners", "1 car", "2 cars", "No car", "National Health Service", "Private health insurance", "Income < 30.000", "Income 30-45.000", "Income 45-75.000", "Income 75-122.000", "Income >123.000", "Average income", "Purchasing power class", "Contribution private third party insurance", "Contribution third party insurance (firms)", "Contribution third party insurane (agriculture)", "Contribution car policies", "Contribution delivery van policies", "Contribution motorcycle/scooter policies", "Contribution lorry policies", "Contribution trailer policies", "Contribution tractor policies", "Contribution agricultural machines policies", "Contribution moped policies", "Contribution life insurances", "Contribution private accident insurance policies", "Contribution family accidents insurance policies", "Contribution disability insurance policies", "Contribution fire policies", "Contribution surfboard policies", "Contribution boat policies", "Contribution bicycle policies", "Contribution property insurance policies", "Contribution social security insurance policies", "Number of private third party insurance", "Number of third party insurance (firms)", "Number of third party insurane (agriculture)", "Number of car policies", "Number of delivery van policies", "Number of motorcycle/scooter policies", "Number of lorry policies", "Number of trailer policies", "Number of tractor policies", "Number of agricultural machines policies", "Number of moped policies", "Number of life insurances", "Number of private accident insurance policies", "Number of family accidents insurance policies", "Number of disability insurance policies", "Number of fire policies", "Number of surfboard policies", "Number of boat policies", "Number of bicycle policies", "Number of property insurance policies", "Number of social security insurance policies", "Number of mobile home policy"))
+#knitr::kable(data_dictionary, "latex", longtable = T) %>%
+#  kable_styling(latex_options = c("repeat_header"))
+apa_table(data_dictionary, longtable = T, booktabs = T, caption = "Data Dictionary")
+```

### 2019-11-14 hw4/HW4.Rmd
Minor change
Update HW4.Rmd
Fixed misc model and added OLS models

@@
 
-## Model Creation
+## Model Creation - Logistic
 
@@
 
+Here we add meaningful variables to the model until we have a robust model with a decent AUC value.
+
 ```{r, null_prefix = TRUE, comment=NA}
-#"KIDSDRIV"         "AGE"             
-#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
-#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
-#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
-#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
-#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
-#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
-#[36] "KID_DRIVERS"     
-model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
-             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
+model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + LOG_INCOME +
+             SEX + CAR_USE + CAR_AGE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +
+               TIF + TRAVTIME + AGE,
              family = binomial(link = "logit"), over_sample_train)
@@
 ```
+
+```{r}
+# Evaluating the model
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(model,newdata=test,type="response"), prob = 
+predict(model,newdata=test,type="response"))
+
+pred <- prediction(eval$predicted, eval$actual)
+auc <- performance(pred, measure = "auc")@y.values[[1]]
+perf <- performance(pred,"tpr","fpr")
+plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
+```
+
+## Model Creation - Linear
+
+First we need to remove the zero values from our train and test sets, as we are only interested in cases where there was an accident claim.
+
+```{r}
+train <- train %>% filter(TARGET_AMT > 0)
+test <- test %>% filter(TARGET_AMT > 0)
+```
+
+### Baseline Model
+
+As with the logistic model, we start with a baseline model using the `BLUEBOOK` variable as the sole predictor. All other models will need to beat this baseline.
+
+```{r, null_prefix = TRUE, comment=NA}
+
+baseline_model <- lm(TARGET_AMT ~ BLUEBOOK, train)
+summary(baseline_model)
+```
+
+This model is rather poor, though the predictor is statistically significant in the model. We can build on this model and see if we can build a more robust version.
+
+### Car Model
+
+Here we look at the type of car the person drives, it's age, and how it is used.
+
+```{r}
+car_model <- lm(TARGET_AMT ~ CAR_AGE + CAR_TYPE + CAR_USE, data = train)
+summary(car_model)
+```
+
+Here we see that none of these predictors are significant in this model. In fact, taken one at a time, none are significant individually either.
+
+### Personal Model
+
+Here we use variables pertaining to the driver's personal life: family, income, driving record, and work.
+
+```{r}
+personal_model <- lm(TARGET_AMT ~ MSTATUS + INCOME + KIDSDRIV +
+                       HOME_VAL + AGE + REVOKED + MVR_PTS +
+                       HOMEKIDS, data= train)
+
+summary(personal_model)
+```
+
+Again, we see a very poor model. No variables are statistically significant here (though 2 are close).
+
+### Additive Model
+
+Here we start with `BLUEBOOK` from the baseline model and add to it until we get a decent model.
+
+```{r}
+add_model <- lm(TARGET_AMT ~ BLUEBOOK + CLM_FREQ, data = train)
+
+summary(add_model)
+```

### 2019-11-15 hw4/HW4.Rmd
More work on HW4

--- 2019-11-12 hw4/HW4.Rmd
@@
 library(ROCR)
+library(corrplot)
 # Thank you Stack Overflow!
@@
 set.seed(42)
-train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
+# ADAM THIS BROKE YOUR MODEL
+#train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
+train_index <- createDataPartition(df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
 train <- df[train_index,]
@@
 
-```{r}
+```{r, eval=FALSE}
 # Evaluating the model
-eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = 
-predict(traditional_cv,newdata=test,type="prob"))
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = predict(traditional_cv,newdata=test,type="prob"))
 
@@
 
-### Misc
+### Manual Model
 
@@
 ```
+
+## Amount Model
+
+To select build a model of how much the claim amount will be we need to subset the training data to only include those that have accidents.
+There will be `r train %>% filter(TARGET_FLAG == 1) %>% nrow()` data points in this set.
+
+```{r}
+amt_train <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  select(-INDEX, -TARGET_FLAG)
+amt_test <- test %>%
+  filter(TARGET_FLAG == 1)
+```
+
+```{r}
+M <- amt_train %>%
+  purrr::keep(is.numeric) %>%
+  #mutate(TARGET_AMT = log(TARGET_AMT)) %>%
+  na.omit() %>%
+  cor(.)
+corrplot(M, type = "upper")
+```
+
+It looks like `BLUEBOOK` is the only numeric variable that correlates with the claim amount.  We will use it as the baseline and see if we can improve this models.
+
+```{r}
+amt_fit <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
+summary(amt_fit)
+```

### 2019-11-15 hw4/HW4_Mike.Rmd
Tinkering

--- 2019-11-08 hw4/HW4_Mike.Rmd
@@
 ---
-title: "Mike's Scratchpad for DATA 621 Homework #4"
-author: "Mike Silva"
+title: "DATA 621 Homework #4"
+author: "Critical Thinking Group 3"
 date: "`r Sys.Date()`"
@@
 library(RColorBrewer)
-library(FFTrees)
+library(ROCR)
+library(corrplot)
 # Thank you Stack Overflow!
@@
 
-## Data Cleanup
-
-The dollar values have dollar signs.
-These will be stripped out and converted from a character to numeric.
-LOG10 and LN transformations of all dollar values will be preformed.
-Categorical data will be translated into dummy variables.
-
 ```{r}
@@
 
+## Introduction
+
+We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.  Each record has two response variables.
+
+The first objective is to train a logistic regression classifier to predict if a person was in a car crash.
+The response variable here is `TARGET_FLAG` respresenting whether a person had an accident (1) or did not have one (0).
+
+```{r, echo=FALSE}
+df %>% 
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
+  ggplot(aes(x=TARGET_FLAG,fill=TARGET_FLAG)) +
+  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  labs(x="TARGET_FLAG", y="# Observations")
+```
+
+The second objective will be to train a regression model to predict the cost of a crash, if one occurred.
+The second reponse variable, `TARGET_VALUE`, is the amount it will cost if the person crashes their car.
+The value is zero if the person did not crash their car.
+
+```{r, echo=FALSE}
+df %>% filter(TARGET_FLAG == 1) %>%
+  ggplot(aes(x=TARGET_AMT)) + geom_density() +
+  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
+  geom_label(aes(x=mean(TARGET_AMT),y=1,label="mu"),parse=T) +
+  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
+  geom_label(aes(x=median(TARGET_AMT),y=.5,label="median")) +
+  scale_x_log10(labels=comma) + theme_light() +
+  labs(title="TARGET_AMT Density Plot", caption="x-axis is log 10 scale",
+       y="Density", x="LOG(TARGET_AMT)")
+```
+
+Looking at the distribution of the `TARGET_AMT` variable, we can see that the variable is considerably right-skewed. Thus, a LOG transform might be best here.
+
+## Data Exploration
+
+We will first look at the summary statistics for the data
+
 ```{r}
-to_dummies <- function(df, var, drop = FALSE){
-  temp <- data.frame(INDEX = df$INDEX, CAT = df[[var]]) %>%
-    mutate(CAT = as.character(CAT),
-           VAL = 1) %>%
-    spread(CAT, VAL, fill = 0)
-  new_names <- c()
-  for (n in names(temp)){
-    if(n != "INDEX"){
-      n <- paste0(var,"_",n)
-      n <- gsub(" ", "_", n)
-    }
-    new_names <- c(new_names, n)
-  }
-  names(temp) <- new_names
+df %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
 
-  if(drop){
-    drops <- c(var)
-    df <- df[ , !(names(df) %in% drops)]
+There are some missing values that we will need to deal with. There are also some values that seem invalid (i.e. -3 CAR_AGE).
+
+## Data Preparation
+
+### Fix Data Types
+
+There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.
+We will do this to both the `df` and `evaluation` data frames.
+There are also some invalid data that will be changed to NAs.
+
+```{r}
+strip_dollars <- function(x){
+  x <- as.character(x)
+  x <- gsub(",", "", x)
+  x <- gsub("\\$", "", x)
+  as.numeric(x)
 }
 
-  df %>%
-    merge(temp)
-}
-
-get_driver_type <- function(x, y){
-  if(is.na(x) | is.na(y)){
-    return(NA)
-  }
-  if(y < 40){
-    driver_type <- "Young"
-  } else if (y < 70){
-    driver_type <- "Middle Age "
-  } else {
-    driver_type <- "Older"
-  }
-  if( x == "M"){
-    return(paste(driver_type, "Male"))
-  } else {
-    return(paste(driver_type, "Female"))
-  }
-}
-
-temp <- df %>%
+fix_data_types <- function(messy_df){
+  messy_df %>%
     rowwise() %>%
@@
            OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
-  mutate(JOB = ifelse(str_length(JOB) > 0, JOB, NA),
-         CAR_AGE = ifelse(str_length(CAR_AGE) > 0, CAR_AGE, NA),
-         EDUCATION = ifelse(EDUCATION == "<High_School", "LT_High_School", EDUCATION),
-         URBANICITY = ifelse(URBANICITY == "Highly Urban/ Urban", "Urban", "Rural")) %>%
-  mutate(DRIVER_TYPE = get_driver_type(SEX, AGE)) %>%
-  ungroup() %>%
-  na.omit() %>%
-  to_dummies("PARENT1", T) %>%
-  to_dummies("KIDSDRIV", T) %>%
-  to_dummies("HOMEKIDS", T) %>%
-  to_dummies("MSTATUS", T) %>%
-  to_dummies("SEX", T) %>%
-  to_dummies("EDUCATION", T) %>%
-  to_dummies("JOB", T) %>%
-  to_dummies("CAR_USE", T) %>%
-  to_dummies("CAR_TYPE", T) %>%
-  to_dummies("RED_CAR", T) %>%
-  to_dummies("REVOKED", T) %>%
-  to_dummies("MVR_PTS", T) %>%
-  to_dummies("URBANICITY", T) %>%
-  to_dummies("DRIVER_TYPE", T) %>%
+    ungroup()
+}
+
+na_bad_values <- function(messy_df){
+  messy_df %>%
     rowwise() %>%
-  mutate(BLUEBOOK = ifelse(BLUEBOOK > 0, BLUEBOOK, 0)) %>%
-  # mutate(LOG10_INCOME = log10(INCOME + 1),
-  #        LN_INCOME = log(INCOME + 1),
-  #        LOG10_HOME_VAL = log10(HOME_VAL + 1),
-  #        LN_HOME_VAL = log(HOME_VAL + 1),
-  #        LOG10_BLUEBOOK = log10(BLUEBOOK + 1),
-  #        LN_BLUEBOOK = log(BLUEBOOK + 1),
-  #        LOG10_OLDCLAIM = log10(OLDCLAIM + 1),
-  #        LN_OLDCLAIM = log(OLDCLAIM + 1)) %>%
-  mutate(PRIOR_CLAIM = ifelse(CLM_FREQ + OLDCLAIM > 0, 1,0)) %>%
+    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
     ungroup()
+}
+
+df$TARGET_FLAG <- factor(df$TARGET_FLAG)
+
+df <- df %>%
+  fix_data_types() %>%
+  na_bad_values()
+evaluation <- evaluation %>%
+  fix_data_types() %>%
+  na_bad_values()
+```
+
+Now that we have fixed the variables, we can look at a sumamry of the data:
+
+```{r}
+df %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Fix Missing Values
+
+There are `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable is missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.
+There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.
+
+### Feature Creation
+
+We will create a log transformed income and home value feature.  We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.
+
+```{r}
+# Function to add features
+
+create_features <- function(d){
+  d %>%
+    mutate(LOG_INCOME = log(INCOME + 1),
+           LOG_HOME_VAL = log(HOME_VAL + 1),
+           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0),
+           PRIOR_ACCIDENT = factor(ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)),
+           COLLEGE_EDUCATED = factor(ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)),
+           URBAN_DRIVER = factor(ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)),
+           YOUNG_MALE = factor(ifelse(SEX == "M" & AGE < 25, 1, 0)),
+           YOUNG = factor(ifelse(AGE < 25, 1, 0)),
+           RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
+           HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
+           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)))
+}
+
+df <- create_features(df)
+evaluation <- create_features(evaluation)
+```
+
+### Creating Training/Test Data Sets
+
+Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`). We'll use a 70-30 split between train and test, respectively.
+
+```{r}
+set.seed(42)
+# ADAM THIS BROKE YOUR MODEL
+#train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
+train_index <- createDataPartition(df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
+train <- df[train_index,]
+test <- df[-train_index,]
+```
+
+There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.
+We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.
+
+```{r}
+over_sample <- function(df, predictors){
+  vars <- c("INDEX", "TARGET_FLAG", predictors)
+  df <- df[vars] %>%
+    na.omit()
+  set.seed(42)
+  minority <- nrow(df[df$TARGET_FLAG == 1,])
+  majority <- nrow(df[df$TARGET_FLAG == 0,])
+  diff <- majority - minority
+  minority_index <- df[df$TARGET_FLAG == 1,]$INDEX
+  over_sample_df <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
+    merge(df, .) %>%
+    bind_rows(df)
+  return(over_sample_df)
+}
+
+evaluate_model <- function(model, test_df){
+  test_df$yhat <- ifelse(predict.glm(model, test_df,"response") >= 0.5, 1, 0)
+  cm <- confusionMatrix(factor(test_df$yhat), factor(test$TARGET_FLAG), "1")
+  deviance <- model$deviance
+  r2 <- 1 - model$deviance / model$null.deviance
+
+  cat("F1 =", cm$byClass[7],"\nR2 =", r2, "\n\n")
+  print(cm)
+  return(cm)
+}
+
+```
+
+## Model Creation
+
+### Baseline Model
+
+We will create a simple model to serve as the baseline.
+This model posits that drivers history is a good representation of their future.
+So drivers that have been in an accident (`OLDCLAIM` > 0 or `AVG_CLAIM` > 0) are more likely to be in an accident.
+Conversely those who haven't aren't likely to be in one in the future.
+
+```{r, null_prefix = TRUE, comment=NA}
+temp <- over_sample(train, c("PRIOR_ACCIDENT"))
+baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), temp)
+summary(baseline_model)
+
+results <- evaluate_model(baseline_model, test)
+```
+
+This simple model has statistically significant predictors.
+Applying this model to the test data set indicates this simple model has a `r  round(unname(results$overall['Accuracy'])*100,1)`% accuracy rate.
+It correcly identified `r round(unname(results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(results$byClass['Specificity'])*100,1)`% of those without.
+It is better than flipping a coin but not by much.
+All future models must outpreform this baseline.
+
+### Risk Taker Model
+
+For this model we are going assume that risk takers are more likely to be in an accident.
+To identify those who tend to be risk takers, we are going with the urban legend that those who drive red sports cars are risk takers.
+Also young men (16-24) also tend to be risk takers.
+
+```{r, null_prefix = TRUE, comment=NA}
+risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample(train, c("RED_SPORTS_CAR", "YOUNG_MALE")))
+summary(risk_taker_model)
+
+risk_taker_results <- evaluate_model(risk_taker_model, test)
+```
+
+This model only has one statistically significant predictor, however it has a `r  round(unname(risk_taker_results$overall['Accuracy'])*100,1)`% accuracy rate.
+This is because it identified `r  round(unname(risk_taker_results$byClass['Specificity'])*100,1)`% of those who didn't have an accident (and they are more numerous).
+When looking at the model's performance on identifying the minority group of interest we see it correcly identified `r round(unname(risk_taker_results$byClass['Sensitivity'])*100,1)`% of the people with accidents.
+So while the accuracy rate out preformed the baseline model, this one will not be the model we will use.
+
+### Traditional Model
+
+Traditional wisdom holds that there are a few tried and true predictors of someone's risk of an automobile accident.
+Namely: age (under 25 vs.
+25+), marital status, driving record, distance to work, and the driver's sex (which may or may not be proper to use, but seems to have held throughout the years).
+
+```{r}
+traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
+                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
+             family = binomial(link = "logit"), over_sample(train, c("YOUNG", "MSTATUS", "PRIOR_ACCIDENT", "SEX", "REVOKED", "MVR_PTS", "TRAVTIME", "CAR_USE")))
+
+summary(traditional_model)
+
+model_results <- evaluate_model(traditional_model, test)
+```
+
+### Traditional Model with Cross-Validation
+
+Here we try to use cross-validation techniques to improve the traditional model.
+We go back to the raw training data, because it is typically not-advised to use data that is oversampled prior to cross-validation.
+
+```{r}
+# Get complete cases
+cases <- train %>%
+  select(YOUNG,MSTATUS,PRIOR_ACCIDENT,SEX,REVOKED,MVR_PTS, TRAVTIME, CAR_USE) %>%
+  complete.cases()
+
+temp <- train[cases,]
+
+# Use 5-fold cross-validation
+train_control <- trainControl(method="cv",number=5, sampling="up")
+
+traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
+                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
+                        method = "glm", family = "binomial", 
+                        data = temp, trControl = train_control)
+
+traditional_cv
+```
+
+```{r, eval=FALSE}
+# Evaluating the model
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = predict(traditional_cv,newdata=test,type="prob"))
+
+confusionMatrix(eval$predicted, eval$actual, positive = "1")
+
+pred <- prediction(eval$prob.1, eval$actual)
+auc <- performance(pred, measure = "auc")@y.values[[1]]
+perf <- performance(pred,"tpr","fpr")
+plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
+```
+
+### Manual Model
+
+```{r, null_prefix = TRUE, comment=NA}
+#"KIDSDRIV"         "AGE"             
+#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
+#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
+#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
+#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
+#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
+#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
+#[36] "KID_DRIVERS"     
+model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
+             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
+             family = binomial(link = "logit"), over_sample(train, c("PRIOR_ACCIDENT", "KID_DRIVERS", "MSTATUS", "INCOME", "SEX", "CAR_USE", "COLLEGE_EDUCATED", "REVOKED", "URBAN_DRIVER")))
+summary(model)
+
+model_results <- evaluate_model(model, test)
+```
+
+## Amount Model
+
+To select build a model of how much the claim amount will be we need to subset the training data to only include those that have accidents.
+There will be `r train %>% filter(TARGET_FLAG == 1) %>% nrow()` data points in this set.
+
+```{r}
+amt_train <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  select(-INDEX, -TARGET_FLAG)
+amt_test <- test %>%
+  filter(TARGET_FLAG == 1)
 ```
@@
 ```{r}
-set.seed(42)
-minority <- nrow(temp[temp$TARGET_FLAG == 1,])
-majority <- nrow(temp[temp$TARGET_FLAG == 0,])
-diff <- majority - minority
-minority_index <- temp[temp$TARGET_FLAG == 1,]$INDEX
-over_sample_df <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
-  merge(temp, .) %>%
-  bind_rows(temp)
-
-train_index <- createDataPartition(over_sample_df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
-train <- over_sample_df[train_index,]
-test <- over_sample_df[-train_index,]
+M <- amt_train %>%
+  purrr::keep(is.numeric) %>%
+  #mutate(TARGET_AMT = log(TARGET_AMT)) %>%
+  na.omit() %>%
+  cor(.)
+corrplot(M, type = "upper")
 ```
 
+It looks like `BLUEBOOK` is the only numeric variable that correlates with the claim amount.  We will use it as the baseline and see if we can improve this models.
+
 ```{r}
-fft_train <- train %>%
-  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
-fft_test <- test %>%
-  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
-fit <- FFTrees(TARGET_FLAG ~ ., fft_train, fft_test)
-plot(fit)
+amt_fit <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
+summary(amt_fit)
 ```
-The FFTree suggests that past history of filing claims is a good predictor if someone will file a claim in the future.  This could be our baseline model criteria.
-
-Other variables that seems to come up is if the driver is in an urban location and possibly their home value and if it's a commercial vehicle

### 2019-11-15 hw4/HW4.Rmd
Work arround PR 21

@@
 ```
+```{r}
+# Evaluating the model
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(model,newdata=test,type="response"), prob = 
+predict(model,newdata=test,type="response"))
+
+pred <- prediction(eval$predicted, eval$actual)
+auc <- performance(pred, measure = "auc")@y.values[[1]]
+perf <- performance(pred,"tpr","fpr")
+plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
+```
+
+## Model Creation - Linear
+
+First we need to remove the zero values from our train and test sets, as we are only interested in cases where there was an accident claim.
+
+```{r}
+train <- train %>% filter(TARGET_AMT > 0)
+test <- test %>% filter(TARGET_AMT > 0)
+```
+
+### Baseline Model
+
+As with the logistic model, we start with a baseline model using the `BLUEBOOK` variable as the sole predictor. All other models will need to beat this baseline.
+
+```{r, null_prefix = TRUE, comment=NA}
+
+baseline_model <- lm(TARGET_AMT ~ BLUEBOOK, train)
+summary(baseline_model)
+```
+
+This model is rather poor, though the predictor is statistically significant in the model. We can build on this model and see if we can build a more robust version.
+
+### Car Model
+
+Here we look at the type of car the person drives, it's age, and how it is used.
+
+```{r}
+car_model <- lm(TARGET_AMT ~ CAR_AGE + CAR_TYPE + CAR_USE, data = train)
+summary(car_model)
+```
+
+Here we see that none of these predictors are significant in this model. In fact, taken one at a time, none are significant individually either.
+
+### Personal Model
+
+Here we use variables pertaining to the driver's personal life: family, income, driving record, and work.
+
+```{r}
+personal_model <- lm(TARGET_AMT ~ MSTATUS + INCOME + KIDSDRIV +
+                       HOME_VAL + AGE + REVOKED + MVR_PTS +
+                       HOMEKIDS, data= train)
+
+summary(personal_model)
+```
+
+Again, we see a very poor model. No variables are statistically significant here (though 2 are close).
+
+### Additive Model
+
+Here we start with `BLUEBOOK` from the baseline model and add to it until we get a decent model.
+
+```{r}
+add_model <- lm(TARGET_AMT ~ BLUEBOOK + CLM_FREQ, data = train)
+
+summary(add_model)
+```
 

### 2019-11-15 hw4/HW4_Mike.Rmd
More tinkering

@@
 ```{r}
-amt_fit <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
-summary(amt_fit)
+rural_fit <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 0) %>%
+  lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + TRAVTIME, .)
+urban_fit <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 1) %>%
+  lm(TARGET_AMT ~ BLUEBOOK, .)
+summary(urban_fit)
 ```
+
+```{r}
+M <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 1) %>%
+  purrr::keep(is.numeric) %>%
+  na.omit() %>%
+  cor(.)
+corrplot(M, type = "upper")
+```
+
+```{r}
+evaluation$TARGET_FLAG <- ifelse(predict.glm(model, evaluation, "response") >= 0.5, 1, 0)
+# We assume that everyone with a TARGET_FLAG = 0 has a TARGET_AMT as zero.  
+# Urban drivers will use one model, rural another
+evaluation$urban_yhat <- predict(urban_fit, evaluation)
+evaluation$rural_yhat <- predict(rural_fit, evaluation)
+evaluation <- evaluation %>%
+  mutate(TARGET_AMT = ifelse(URBAN_DRIVER == 1, urban_yhat, rural_yhat)) %>%
+  mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT))
+```

### 2019-11-15 hw4/HW4_Mike.Rmd
More work

@@
     highlight: kate
-    toc_depth: 3
+    toc_depth: 4
     code_folding: "hide"
@@
 There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.
+We will fill in the missing data with the median value.
+
+```{r}
+df <- df %>% 
+   mutate_at(vars(c("CAR_AGE", "YOJ", "AGE", "INCOME", "HOME_VAL")), ~ifelse(is.na(.), median(., na.rm = TRUE), .))
+```
 
@@
 evaluate_model <- function(model, test_df){
-  test_df$yhat <- ifelse(predict.glm(model, test_df,"response") >= 0.5, 1, 0)
+  test_df$yhat <- ifelse(predict.glm(model, test_df, "response") >= 0.5, 1, 0)
   cm <- confusionMatrix(factor(test_df$yhat), factor(test$TARGET_FLAG), "1")
@@
   print(cm)
+
+  eval <- data.frame(actual = test_df$TARGET_FLAG, 
+                     predicted = test_df$yhat, 
+                     prob = predict(model, test_df))
+
+  pred <- prediction(eval$prob, eval$actual)
+  auc <- performance(pred, measure = "auc")@y.values[[1]]
+  perf <- performance(pred, "tpr", "fpr")
+  plot(perf,main="ROC Curve", sub = paste0("AUC: ", round(auc, 3)))
+
   return(cm)
@@
 
-### Baseline Model
+### Classification Model
+
+#### Baseline Model
 
@@
 
-### Risk Taker Model
+#### Risk Taker Model
 
@@
 
-### Traditional Model
+#### Traditional Model
 
@@
 
-### Traditional Model with Cross-Validation
+This model has statistically significant predictors.
+Applying this model to the test data set indicates this simple model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.
+It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.
+This model out preforms the baseline model.
+
+#### Traditional Model with Cross-Validation
 
@@
 
-```{r, eval=FALSE}
+```{r}
 # Evaluating the model
@@
 
-### Manual Model
+#### Manual Model
 
@@
 
-## Amount Model
+## Linear Claim Model
 
@@
 
+#### Baseline Model
+
+```{r}
+baseline_lm <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
+summary(baseline_lm)
+```
+
+#### Rural Model
+```{r}
+M <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 0) %>%
+  purrr::keep(is.numeric) %>%
+  na.omit() %>%
+  cor(.)
+corrplot(M, type = "upper")
+```
+
 ```{r}
@@
   lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + TRAVTIME, .)
-urban_fit <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 1) %>%
-  lm(TARGET_AMT ~ BLUEBOOK, .)
-summary(urban_fit)
+summary(rural_fit)
 ```
+
+#### Urban Model
 
@@
 ```{r}
+urban_fit <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 1) %>%
+  lm(TARGET_AMT ~ BLUEBOOK, .)
+summary(urban_fit)
+```
+
+## Predictions
+
+```{r}
 evaluation$TARGET_FLAG <- ifelse(predict.glm(model, evaluation, "response") >= 0.5, 1, 0)

### 2019-11-15 hw4/HW4.Rmd
Expanded exploratory analysis

--- 2019-11-15 hw4/HW4.Rmd
@@
 library(scales)
+library(purrr)
 library(RColorBrewer)
@@
 # A Prefix nulling hook.
-
+options(scipen = 999)
 # Make sure to keep the default for normal processing.
@@
 
-We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.  Each record has two response variables.
-
+We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.
+Each record has two response variables.
 The first objective is to train a logistic regression classifier to predict if a person was in a car crash.
@@
 df %>% filter(TARGET_FLAG == 1) %>%
-  ggplot(aes(x=TARGET_AMT)) + geom_density() +
+  ggplot(aes(x=TARGET_AMT)) + 
+  geom_density() +
   geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
@@
 
-## Data Exploration
+## Data Preparation & Exploration
 
@@
 There are some missing values that we will need to deal with. There are also some values that seem invalid (i.e. -3 CAR_AGE).
-
-## Data Preparation
 
@@
 
-Now that we have fixed the variables, we can look at a sumamry of the data:
+Now that we have corrected the invalid variables, we can look at a sumamry of the data:
 
@@
 
-We will create a log transformed income and home value feature.  We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.
+We will create a log transformed income and home value feature.
+We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.
+We will also flag the `TARGET_AMT` observations that are outliers.
+These would be cases where the vehicle is totaled and are qualitatively different than the minor accidents.
 
@@
 # Function to add features
+
+outlier <- min(boxplot(df[df$TARGET_FLAG==1,]$TARGET_AMT, plot=FALSE)$out)
 
@@
            HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
-           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)))
+           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)),
+           TARGET_AMT_OUTLIER = ifelse(TARGET_AMT < outlier, 0, 1)) %>%
+    select(-URBANICITY)
 }
@@
 
-Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`). We'll use a 70-30 split between train and test, respectively.
+#### For Classifier Model
+
+Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`) for the classifier. We'll use a 70-30 split between train and test, respectively.
 
@@
 set.seed(42)
-# ADAM THIS BROKE YOUR MODEL
-#train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
 train_index <- createDataPartition(df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
@@
 There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.
-We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.
+We want to correct the imbalance in the dataset by over sampling this group so our classifier will do a better job identifying this minority group.
 
@@
 
-## Model Creation
+#### For Linear Regression Model
 
-### Baseline Model
+Since the linear regression model's goal is to predict the cost of the claim we will subset the `r nrow(df)` records for those involved in an accident.
+We will split these `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` records into training and test sets using the same 70-30 split.
+
+```{r}
+set.seed(42)
+accidents <- df %>%
+  filter(TARGET_FLAG == 1)
+amt_train_index <- createDataPartition(accidents$TARGET_AMT, p = .7, list = FALSE, times = 1)
+amt_train <- accidents[amt_train_index,]
+amt_test <- accidents[-amt_train_index,]
+```
+
+There are `r nrow(amt_train)` out of `r nrow(accidents)` records in the training data set.
+
+### Exploratory Data Analysis
+
+In exploring the data we are looking for two things.
+We are looking for variables that will help use divide the data into those who have been in an accident and those who have not.
+We are also looking for variables that are linearly correlated with the claim amount so it may be used as a predictor for the linear regression model.
+We will be looking at both training set for the two model types.
+
+First we will examine the numeric variables found in the oversampled classification data set:
+
+```{r}
+plot_vars <- c("TARGET_FLAG", names(keep(over_sample_train, is.numeric)))
+
+over_sample_train[plot_vars] %>%
+  select(-INDEX, -TARGET_AMT) %>%
+  gather(variable, value, -TARGET_FLAG) %>%
+  ggplot(., aes(TARGET_FLAG, value, color=TARGET_FLAG)) + 
+  geom_boxplot() +
+  scale_color_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+There are a few variables that seem to have a difference between the two groups.
+For example `CLM_FREQ` might be useful in differentiating.
+There isn't much difference between a lot of groups which is somewhat understandable.
+
+Next we will look at the catagorical variables in the oversampled classifcation data set.
+These charts look to see if a variable can be uses to distinguish those that have had an accident (blue) from those who haven't (red).
+
+```{r}
+plot_vars <- names(keep(over_sample_train, is.factor))
+
+temp <- over_sample_train[plot_vars] %>%
+  gather(variable, value, -TARGET_FLAG) %>%
+  group_by(TARGET_FLAG, variable, value) %>%
+  tally()
+
+temp %>%
+  group_by(variable, value) %>%
+  summarise(total = sum(n)) %>%
+  merge(temp,.) %>%
+  mutate(percent = n / total) %>%
+  ggplot(., aes(value, percent, fill=TARGET_FLAG)) + 
+  geom_col() +
+  scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+The `URBAN_DRIVER` is much more likely to be in an accident than the rural driver.
+The `YOUNG` and those with a `PRIOR_ACCIDENT` are also more likely to be in an accident.
+There doesn't seem to be much evidence for a difference between the sexes.
+
+Now let's focus on the relationship between these variables and the size of the claim.
+As a reminder this data only includes those who have been in an accident.
+For starters we will look at the distribution of the claims for those in an accident.
+
+```{r}
+ggplot(amt_train, aes(x=TARGET_AMT)) + 
+  geom_density() +
+  theme_light() +
+  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
+  geom_label(aes(x=25000, y=0.00015, label=paste("mean =", round(mean(TARGET_AMT),0)))) +
+  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
+  geom_label(aes(x=25000, y=0.00010, label=paste("median = ", round(median(TARGET_AMT), 0)))) +
+  labs(title="TARGET_AMT Density Plot", y="Density", x="TARGET_AMT")
+```
+
+As was previously noted this distribution has a long tail.
+The mean payout is $`r round(mean(amt_train$TARGET_AMT), 0)` and the median is $`r round(median(amt_train$TARGET_AMT), 0)`.
+The median and mean are higher, of course for those observations we classified as outliers.
+The outlier cutoff point is $`r round(outlier, 0)`.
+
+```{r}
+amt_train %>%
+  mutate(TARGET_AMT_OUTLIER = ifelse(TARGET_AMT_OUTLIER == 1, "Yes", "No")) %>%
+  group_by(TARGET_AMT_OUTLIER) %>%
+  summarise(Mean = mean(TARGET_AMT),
+            Median = median(TARGET_AMT)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+Now to look at the data for the linear regression model.  We are looking for good predictors of the claim amount.  We will look at the numeric variables scatter and correlation plots:
+
+```{r}
+amt_train %>%
+  keep(is.numeric) %>%
+  gather(variable, value, -TARGET_AMT) %>%
+  ggplot(., aes(value, TARGET_AMT)) + 
+  geom_point() +
+  scale_color_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+```{r}
+M <- amt_train %>%
+  select(-INDEX) %>%
+  keep(is.numeric) %>%
+  cor(.)
+corrplot(M, type = "upper")
+
+```
+
+These two points highlight that there isn't a really strong correlation between most of the predictors and the claim amount.
+The only one that is strong is the outlier flag that we created.
+Let's finish our exploration by looking at the categoricial variables.
+We are removing the `TARGET_AMT_OUTLIERS` from the plot to increase their readability.
+
+```{r}
+plot_vars <- c("TARGET_AMT", "TARGET_AMT_OUTLIER", names(keep(amt_train, is.factor)))
+
+amt_train[plot_vars] %>% 
+  filter(TARGET_AMT_OUTLIER == 0) %>%
+  gather(variable, value, -TARGET_AMT) %>%
+  ggplot(., aes(value, TARGET_AMT)) + 
+  geom_boxplot() +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+These boxplots show that there isn't really much difference in the claim amounts between these groups.
+
+#### Summary of Findings
+
+Through this analysis we have found that there is a lot of noise associated with this signal.
+When it comes the the classification task, most variables offer a similar picture.
+This is probably due to the random nature of vehicle accidents.
+There are a few variables that may be helpful in building our models.
+
+The claim prediciton task will definately be more of a challenge.  Little is correlated with the claim amounts.  The model fit will not be as strong as the classification models.
+
+## Model Creation & Evaluation
+
+Now that we have created datasets and explored them, it is time to build some predictive models and evaluate them.
+
+For the classification models we will look at the prediction accuracy on the test set.
+Especially with the accuracy at discovering those that are in an accident.
+Accurate prediction of this minority class indicates a good model.
 
@@
 ```
+
+### Classification Model
+
+#### Baseline Model
 
@@
 ```{r}
-traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
-                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
-             family = binomial(link = "logit"), over_sample_train)
-
+traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX + REVOKED + MVR_PTS + TRAVTIME + CAR_USE, family = binomial(link = "logit"), over_sample_train)
 summary(traditional_model)
-
 model_results <- evaluate_model(traditional_model, test)
@@
 
-### Traditional Model with Cross-Validation
+This model has statistically significant predictors.
+Applying this model to the test data set indicates this model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.
+It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.
+This model out preforms the baseline model.
+
+#### Traditional Model with Cross-Validation
 
@@
   complete.cases()
-
 temp <- train[cases,]
@@
 # Use 5-fold cross-validation
-train_control <- trainControl(method="cv",number=5, sampling="up")
-
-traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
-                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
-                        method = "glm", family = "binomial", 
-                        data = temp, trControl = train_control)
+train_control <- trainControl(method = "cv", number = 5, sampling = "up")
+traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX + REVOKED + MVR_PTS + TRAVTIME + CAR_USE, method = "glm", family = "binomial", data = temp, trControl = train_control)
 
@@
 
-### Manual Model
+#### Alternate Traditional Model
+
+This model is an alternate to the traditional model.
 
 ```{r, null_prefix = TRUE, comment=NA}
-#"KIDSDRIV"         "AGE"             
-#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
-#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
-#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
-#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
-#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
-#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
-#[36] "KID_DRIVERS"     
-model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
-             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
-             family = binomial(link = "logit"), over_sample_train)
+model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME + SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER, family = binomial(link = "logit"), over_sample_train)
 summary(model)
-
 model_results <- evaluate_model(model, test)
@@
 
-## Amount Model
+This model's predictors are all statistically significant.
+Applying this model to the test data set indicates this model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.
+It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.
+This model also out preforms the baseline model.
+Now that we have a few viable classification models, we can turn our attention to the linear claim model.
 
-To select build a model of how much the claim amount will be we need to subset the training data to only include those that have accidents.
-There will be `r train %>% filter(TARGET_FLAG == 1) %>% nrow()` data points in this set.
+### Linear Claim Model
+
+#### Baseline Model
+
+This baseline model is that the amount of the claim is based off of the value of the vehicle.
+More expensive vehicles should be more costly to repair than less expensive vehicles.
+It is overly simplistic but will serve as a baseline.
 
 ```{r}
-amt_train <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  select(-INDEX, -TARGET_FLAG)
-amt_test <- test %>%
-  filter(TARGET_FLAG == 1)
+baseline_lm <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
+summary(baseline_lm)
+```
+
+While this predictor is statistically significant and positive, the model's overall explanitory power is weak.  Let's see how it preformed on the test set:
+
+```{r}
+data.frame(yhat = predict(baseline_lm, amt_test), actual = amt_test$TARGET_AMT) %>%
+  ggplot(., aes(actual, yhat)) +
+  geom_point()
+```
+
+#### Outlier Model
+
+We previously flagged the outliers in the data set.  IF we can build another classification model, how well will it preform?
+
+```{r}
+outlier_lm <- lm(TARGET_AMT ~ TARGET_AMT_OUTLIER, amt_train)
+summary(outlier_lm)
+```
+
+This model is unfair as it is predicting the outcomes based on a predictor derived from the outcome.
+Unsuprisingly this model does a really good job.
+It has an adjusted $R^2$ of `r summary(outlier_lm)$adj.r.squared`.
+
+The reason we explored this model is to see if we has a reason to build another classifier.
+If we build a second classifier that can predict this variable would it be worth it.
+Let's see how the model preforms on the test set:
+
+```{r}
+temp <- data.frame(yhat = predict(outlier_lm, amt_test), actual = amt_test$TARGET_AMT, TARGET_AMT_OUTLIER = amt_test$TARGET_AMT_OUTLIER) 
+temp %>%
+  ggplot(., aes(actual, yhat, color=as.factor(TARGET_AMT_OUTLIER))) +
+  geom_point() +
+  scale_color_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") 
 ```
@@
 ```{r}
-M <- amt_train %>%
+temp %>%
+  mutate(error = yhat - actual) %>%
+  mutate(`error %` = (error / actual)*100) %>%
+  group_by(TARGET_AMT_OUTLIER) %>%
+  summarise(error = mean(error),
+            `error %` = mean(`error %`)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+It understates the claim prediciton by about $95 for the small claims and $1,100 for the large claims.
+The error on the is about 55% of the estimate for the small claims and 31% for the large claims.
+In summary it is roughly right.
+It would preform a lot better than the baseline model.
+The table below offers the similar metrics:
+
+```{r}
+data.frame(yhat = predict(baseline_lm, amt_test), actual = amt_test$TARGET_AMT, TARGET_AMT_OUTLIER = amt_test$TARGET_AMT_OUTLIER) %>%
+  mutate(error = yhat - actual) %>%
+  mutate(`error %` = (error / actual)*100) %>%
+  group_by(TARGET_AMT_OUTLIER) %>%
+  summarise(error = mean(error),
+            `error %` = mean(`error %`)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+The challenge will be to create a classifier.  Let's explore the correlations to see what is correlated:
+
+```{r, null_prefix = TRUE}
+amt_train %>%
+  keep(is.numeric)%>%
+  select(-TARGET_AMT_OUTLIER, -INDEX) %>%
+  cor(., amt_train$TARGET_AMT_OUTLIER)
+
+```
+
+#### Rural Model
+```{r}
+M <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 0) %>%
   purrr::keep(is.numeric) %>%
-  #mutate(TARGET_AMT = log(TARGET_AMT)) %>%
   na.omit() %>%
@@
 
-It looks like `BLUEBOOK` is the only numeric variable that correlates with the claim amount.  We will use it as the baseline and see if we can improve this models.
+```{r}
+rural_fit <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 0) %>%
+  lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + TRAVTIME, .)
+summary(rural_fit)
+```
+
+#### Urban Model
 
 ```{r}
-amt_fit <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
-summary(amt_fit)
+M <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 1) %>%
+  purrr::keep(is.numeric) %>%
+  na.omit() %>%
+  cor(.)
+corrplot(M, type = "upper")
 ```
+
+```{r}
+urban_fit <- train %>%
+  filter(TARGET_FLAG == 1) %>%
+  filter(URBAN_DRIVER == 1) %>%
+  lm(TARGET_AMT ~ BLUEBOOK, .)
+summary(urban_fit)
+```
+
+## Predictions
+
+```{r}
+predictions <- function(df, classifier, linear_model_1, linear_model_2){
+  df$TARGET_FLAG <- ifelse(predict.glm(classifier, df, "response") >= 0.5, 1, 0)
+  # We assume that everyone with a TARGET_FLAG = 0 has a TARGET_AMT as zero.  
+  # We then refine it with the two linear models
+  df$model_1_yhat <- predict(linear_model_1, df)
+  df$model_2_yhat <- predict(linear_model_2, df)
+  df <- df %>%
+    mutate(TARGET_AMT = ifelse(URBAN_DRIVER == 1, model_1_yhat, model_2_yhat)) %>%
+    mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT))
+  return(df)
+}
+
+evaluation <- predictions(evaluation, model, urban_fit, rural_fit)
+```

### 2019-11-15 hw4/HW4.Rmd
More work

@@
 corrplot(M, type = "upper")
+```
 
+```{r}
+amt_train %>%
+  keep(is.numeric)%>%
+  select(-INDEX, -TARGET_AMT) %>%
+  cor(., amt_train$TARGET_AMT) %>%
+  kable() %>%
+  kable_styling()
 ```
@@
 
-The challenge will be to create a classifier.  Let's explore the correlations to see what is correlated:
+The challenge will be to create a classifier.
+There are `r nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])` out of `r nrow(amt_train)` that are outliers.
+Let's create another balanced data set and see if we can build a classifier.
 
-```{r, null_prefix = TRUE}
-amt_train %>%
+```{r}
+set.seed(42)
+minority <- nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])
+majority <- nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 0,])
+diff <- majority - minority
+minority_index <- amt_train[amt_train$TARGET_AMT_OUTLIER == 1,]$INDEX
+over_sample_train_2 <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
+  merge(amt_train, .) %>%
+  bind_rows(amt_train)
+```
+
+Let's explore the correlations to see what is correlated:
+
+```{r}
+over_sample_train_2 %>%
   keep(is.numeric)%>%
-  select(-TARGET_AMT_OUTLIER, -INDEX) %>%
-  cor(., amt_train$TARGET_AMT_OUTLIER)
+  select(-TARGET_AMT_OUTLIER, -INDEX, -TARGET_AMT) %>%
+  cor(., over_sample_train_2$TARGET_AMT_OUTLIER) %>%
+  kable() %>%
+  kable_styling()
+```
 
+```{r}
+M <- over_sample_train_2 %>%
+  select(-INDEX, -TARGET_AMT) %>%
+  keep(is.numeric) %>%
+  cor(.)
+corrplot(M, type = "upper")
 ```

### 2019-11-15 hw4/HW4_Mike.Rmd
More work

@@
 There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.
-We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.
+We want to correct the imbalance in the dataset by over sampling this group so our classifier will do a better job identifying this minority group.
 
 ```{r}
-over_sample <- function(df, predictors){
-  vars <- c("INDEX", "TARGET_FLAG", predictors)
-  df <- df[vars] %>%
-    na.omit()
 set.seed(42)
-  minority <- nrow(df[df$TARGET_FLAG == 1,])
-  majority <- nrow(df[df$TARGET_FLAG == 0,])
+minority <- nrow(train[train$TARGET_FLAG == 1,])
+majority <- nrow(train[train$TARGET_FLAG == 0,])
 diff <- majority - minority
-  minority_index <- df[df$TARGET_FLAG == 1,]$INDEX
-  over_sample_df <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
-    merge(df, .) %>%
-    bind_rows(df)
-  return(over_sample_df)
-}
+minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
+over_sample_train <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
+  merge(train, .) %>%
+  bind_rows(train)
+```
 
-evaluate_model <- function(model, test_df){
-  test_df$yhat <- ifelse(predict.glm(model, test_df, "response") >= 0.5, 1, 0)
-  cm <- confusionMatrix(factor(test_df$yhat), factor(test$TARGET_FLAG), "1")
+The over sampled data frame has `r nrow(over_sample_train)` records, and as you can see in the following figure is now balanced:
+
+```{r, echo=FALSE}
+over_sample_train %>% 
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
+  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
+  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  labs(x="TARGET_FLAG", y="# Observations")
+
+over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
+```
+
+#### For Linear Regression Model
+
+Since the linear regression model's goal is to predict the cost of the claim we will subset the `r nrow(df)` records for those involved in an accident.
+We will split these `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` records into training and test sets using the same 70-30 split.
+
+```{r}
+set.seed(42)
+accidents <- df %>%
+  filter(TARGET_FLAG == 1)
+amt_train_index <- createDataPartition(accidents$TARGET_AMT, p = .7, list = FALSE, times = 1)
+amt_train <- accidents[amt_train_index,]
+amt_test <- accidents[-amt_train_index,]
+```
+
+There are `r nrow(amt_train)` out of `r nrow(accidents)` records in the training data set.
+
+### Exploratory Data Analysis
+
+In exploring the data we are looking for two things.
+We are looking for variables that will help use divide the data into those who have been in an accident and those who have not.
+We are also looking for variables that are linearly correlated with the claim amount so it may be used as a predictor for the linear regression model.
+We will be looking at both training set for the two model types.
+
+First we will examine the numeric variables found in the oversampled classification data set:
+
+```{r}
+plot_vars <- c("TARGET_FLAG", names(keep(over_sample_train, is.numeric)))
+
+over_sample_train[plot_vars] %>%
+  select(-INDEX, -TARGET_AMT) %>%
+  gather(variable, value, -TARGET_FLAG) %>%
+  ggplot(., aes(TARGET_FLAG, value, color=TARGET_FLAG)) + 
+  geom_boxplot() +
+  scale_color_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+There are a few variables that seem to have a difference between the two groups.
+For example `CLM_FREQ` might be useful in differentiating.
+There isn't much difference between a lot of groups which is somewhat understandable.
+
+Next we will look at the catagorical variables in the oversampled classifcation data set.
+These charts look to see if a variable can be uses to distinguish those that have had an accident (blue) from those who haven't (red).
+
+```{r}
+plot_vars <- names(keep(over_sample_train, is.factor))
+
+temp <- over_sample_train[plot_vars] %>%
+  gather(variable, value, -TARGET_FLAG) %>%
+  group_by(TARGET_FLAG, variable, value) %>%
+  tally()
+
+temp %>%
+  group_by(variable, value) %>%
+  summarise(total = sum(n)) %>%
+  merge(temp,.) %>%
+  mutate(percent = n / total) %>%
+  ggplot(., aes(value, percent, fill=TARGET_FLAG)) + 
+  geom_col() +
+  scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+The `URBAN_DRIVER` is much more likely to be in an accident than the rural driver.
+The `YOUNG` and those with a `PRIOR_ACCIDENT` are also more likely to be in an accident.
+There doesn't seem to be much evidence for a difference between the sexes.
+
+Now let's focus on the relationship between these variables and the size of the claim.
+As a reminder this data only includes those who have been in an accident.
+For starters we will look at the distribution of the claims for those in an accident.
+
+```{r}
+ggplot(amt_train, aes(x=TARGET_AMT)) + 
+  geom_density() +
+  theme_light() +
+  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
+  geom_label(aes(x=25000, y=0.00015, label=paste("mean =", round(mean(TARGET_AMT),0)))) +
+  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
+  geom_label(aes(x=25000, y=0.00010, label=paste("median = ", round(median(TARGET_AMT), 0)))) +
+  labs(title="TARGET_AMT Density Plot", y="Density", x="TARGET_AMT")
+```
+
+As was previously noted this distribution has a long tail.
+The mean payout is $`r round(mean(amt_train$TARGET_AMT), 0)` and the median is $`r round(median(amt_train$TARGET_AMT), 0)`.
+The median and mean are higher, of course for those observations we classified as outliers.
+The outlier cutoff point is $`r round(outlier, 0)`.
+
+```{r}
+amt_train %>%
+  mutate(TARGET_AMT_OUTLIER = ifelse(TARGET_AMT_OUTLIER == 1, "Yes", "No")) %>%
+  group_by(TARGET_AMT_OUTLIER) %>%
+  summarise(Mean = mean(TARGET_AMT),
+            Median = median(TARGET_AMT)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+Now to look at the data for the linear regression model.  We are looking for good predictors of the claim amount.  We will look at the numeric variables scatter and correlation plots:
+
+```{r}
+amt_train %>%
+  keep(is.numeric) %>%
+  gather(variable, value, -TARGET_AMT) %>%
+  ggplot(., aes(value, TARGET_AMT)) + 
+  geom_point() +
+  scale_color_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+```{r}
+M <- amt_train %>%
+  select(-INDEX) %>%
+  keep(is.numeric) %>%
+  cor(.)
+corrplot(M, type = "upper")
+```
+
+```{r}
+amt_train %>%
+  keep(is.numeric)%>%
+  select(-INDEX, -TARGET_AMT) %>%
+  cor(., amt_train$TARGET_AMT) %>%
+  kable() %>%
+  kable_styling()
+```
+
+These two points highlight that there isn't a really strong correlation between most of the predictors and the claim amount.
+The only one that is strong is the outlier flag that we created.
+Let's finish our exploration by looking at the categoricial variables.
+We are removing the `TARGET_AMT_OUTLIERS` from the plot to increase their readability.
+
+```{r}
+plot_vars <- c("TARGET_AMT", "TARGET_AMT_OUTLIER", names(keep(amt_train, is.factor)))
+
+amt_train[plot_vars] %>% 
+  filter(TARGET_AMT_OUTLIER == 0) %>%
+  gather(variable, value, -TARGET_AMT) %>%
+  ggplot(., aes(value, TARGET_AMT)) + 
+  geom_boxplot() +
+  theme_light() +
+  theme(legend.position = "none") +
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+These boxplots show that there isn't really much difference in the claim amounts between these groups.
+
+#### Summary of Findings
+
+Through this analysis we have found that there is a lot of noise associated with this signal.
+When it comes the the classification task, most variables offer a similar picture.
+This is probably due to the random nature of vehicle accidents.
+There are a few variables that may be helpful in building our models.
+
+The claim prediciton task will definately be more of a challenge.  Little is correlated with the claim amounts.  The model fit will not be as strong as the classification models.
+
+## Model Creation & Evaluation
+
+Now that we have created datasets and explored them, it is time to build some predictive models and evaluate them.
+
+For the classification models we will look at the prediction accuracy on the test set.
+Especially with the accuracy at discovering those that are in an accident.
+Accurate prediction of this minority class indicates a good model.
+
+```{r}
+evaluate_model <- function(model, test_df, target = "TARGET_FLAG", threshold = 0.5){
+  test_df$yhat <- ifelse(predict.glm(model, test_df, "response") >= threshold, 1, 0)
+  cm <- confusionMatrix(factor(test_df$yhat), factor(test_df[[target]]), "1")
   deviance <- model$deviance
@@
 }
-
 ```
-
-## Model Creation
 
@@
 ```{r, null_prefix = TRUE, comment=NA}
-temp <- over_sample(train, c("PRIOR_ACCIDENT"))
-baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), temp)
+baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), over_sample_train)
 summary(baseline_model)
-
 results <- evaluate_model(baseline_model, test)
@@
 ```{r, null_prefix = TRUE, comment=NA}
-risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample(train, c("RED_SPORTS_CAR", "YOUNG_MALE")))
+risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample_train)
 summary(risk_taker_model)
-
 risk_taker_results <- evaluate_model(risk_taker_model, test)
@@
 ```{r}
-traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
-                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
-             family = binomial(link = "logit"), over_sample(train, c("YOUNG", "MSTATUS", "PRIOR_ACCIDENT", "SEX", "REVOKED", "MVR_PTS", "TRAVTIME", "CAR_USE")))
-
+traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX + REVOKED + MVR_PTS + TRAVTIME + CAR_USE, family = binomial(link = "logit"), over_sample_train)
 summary(traditional_model)
-
 model_results <- evaluate_model(traditional_model, test)
@@
 This model has statistically significant predictors.
-Applying this model to the test data set indicates this simple model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.
+Applying this model to the test data set indicates this model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.
 It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.
@@
 
-#### Manual Model
+#### Alternate Traditional Model
+
+This model is an alternate to the traditional model.
 
 ```{r, null_prefix = TRUE, comment=NA}
-#"KIDSDRIV"         "AGE"             
-#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
-#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
-#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
-#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
-#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
-#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
-#[36] "KID_DRIVERS"     
-model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
-             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
-             family = binomial(link = "logit"), over_sample(train, c("PRIOR_ACCIDENT", "KID_DRIVERS", "MSTATUS", "INCOME", "SEX", "CAR_USE", "COLLEGE_EDUCATED", "REVOKED", "URBAN_DRIVER")))
+model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME + SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER, family = binomial(link = "logit"), over_sample_train)
 summary(model)
-
 model_results <- evaluate_model(model, test)
@@
 
-## Linear Claim Model
+This model's predictors are all statistically significant.
+Applying this model to the test data set indicates this model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.
+It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.
+This model also out preforms the baseline model.
+Now that we have a few viable classification models, we can turn our attention to the linear claim model.
 
-To select build a model of how much the claim amount will be we need to subset the training data to only include those that have accidents.
-There will be `r train %>% filter(TARGET_FLAG == 1) %>% nrow()` data points in this set.
+### Linear Claim Model
+
+#### Baseline Model
+
+Given the unusual distribution one posible solution would be to predict the claim is the median value.
+This is a gross simplification but will serve as the baseline for the linear models to compete against.
+The way we are looking at how well these models preform is calculating the prediction error as a percentage of the actual and then averaging it across the data set.
 
 ```{r}
-amt_train <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  select(-INDEX, -TARGET_FLAG)
-amt_test <- test %>%
-  filter(TARGET_FLAG == 1)
+the_median <- median(amt_train$TARGET_AMT)
+
+evaluate_lm <- function(yhat, actual){
+  data.frame(yhat = yhat, actual = actual) %>%
+    mutate(error = yhat - actual) %>%
+    mutate(error_percent = error / actual) %>%
+    summarise(`Average Error` = mean(error), `Average Percent` = mean(error_percent)) %>%
+    kable() %>%
+    kable_styling()
+}
+
+evaluate_lm(c(the_median), amt_test$TARGET_AMT)
+```
+
+#### Simple Linear Model
+
+This simple linerar model is that the amount of the claim is based off of the value of the vehicle.
+More expensive vehicles should be more costly to repair than less expensive vehicles.
+It is overly simplistic.
+
+```{r}
+baseline_lm <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
+summary(baseline_lm)
+```
+
+While this predictor is statistically significant and positive, the model's overall explanitory power is weak.  Let's see how it preformed on the test set:
+
+```{r}
+evaluate_lm(predict(baseline_lm, amt_test), amt_test$TARGET_AMT)
+```
+
+#### Outlier Model
+
+We previously flagged the outliers in the data set.  IF we can build another classification model, how well will it preform?
+
+```{r}
+outlier_lm <- lm(TARGET_AMT ~ TARGET_AMT_OUTLIER, amt_train)
+summary(outlier_lm)
+```
+
+This model is unfair as it is predicting the outcomes based on a predictor derived from the outcome.
+Unsuprisingly this model does a really good job.
+It has an adjusted $R^2$ of `r summary(outlier_lm)$adj.r.squared`.
+
+The reason we explored this model is to see if we has a reason to build another classifier.
+If we build a second classifier that can predict this variable would it be worth it.
+Let's see how the model preforms on the test set:
+
+```{r}
+evaluate_lm(predict(outlier_lm, amt_test), amt_test$TARGET_AMT)
 ```
@@
 ```{r}
-M <- amt_train %>%
-  purrr::keep(is.numeric) %>%
-  #mutate(TARGET_AMT = log(TARGET_AMT)) %>%
-  na.omit() %>%
+temp <- data.frame(yhat = predict(outlier_lm, amt_test), actual = amt_test$TARGET_AMT, TARGET_AMT_OUTLIER = amt_test$TARGET_AMT_OUTLIER) 
+temp %>%
+  ggplot(., aes(actual, yhat, color=as.factor(TARGET_AMT_OUTLIER))) +
+  geom_point() +
+  scale_color_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") 
+```
+
+```{r}
+temp %>%
+  mutate(error = yhat - actual) %>%
+  mutate(`error %` = (error / actual)*100) %>%
+  group_by(TARGET_AMT_OUTLIER) %>%
+  summarise(error = mean(error),
+            `error %` = mean(`error %`)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+It understates the claim prediciton by about $95 for the small claims and $1,100 for the large claims.
+The error on the is about 55% of the estimate for the small claims and 31% for the large claims.
+In summary it is roughly right.
+It would preform a lot better than the baseline model.
+The table below offers the similar metrics:
+
+```{r}
+data.frame(yhat = predict(baseline_lm, amt_test), actual = amt_test$TARGET_AMT, TARGET_AMT_OUTLIER = amt_test$TARGET_AMT_OUTLIER) %>%
+  mutate(error = yhat - actual) %>%
+  mutate(`error %` = (error / actual)*100) %>%
+  group_by(TARGET_AMT_OUTLIER) %>%
+  summarise(error = mean(error),
+            `error %` = mean(`error %`)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+The challenge will be to create a classifier.
+There are `r nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])` out of `r nrow(amt_train)` that are outliers.
+Let's create another balanced data set and see if we can build a classifier.
+
+```{r}
+set.seed(42)
+minority <- nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])
+majority <- nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 0,])
+diff <- majority - minority
+minority_index <- amt_train[amt_train$TARGET_AMT_OUTLIER == 1,]$INDEX
+over_sample_train_2 <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
+  merge(amt_train, .) %>%
+  bind_rows(amt_train)
+```
+
+Let's explore the correlations to see what is correlated:
+
+```{r}
+over_sample_train_2 %>%
+  keep(is.numeric)%>%
+  select(-TARGET_AMT_OUTLIER, -INDEX, -TARGET_AMT) %>%
+  cor(., over_sample_train_2$TARGET_AMT_OUTLIER) %>%
+  kable() %>%
+  kable_styling()
+```
+
+```{r}
+M <- over_sample_train_2 %>%
+  select(-INDEX, -TARGET_AMT) %>%
+  keep(is.numeric) %>%
   cor(.)
@@
 
-It looks like `BLUEBOOK` is the only numeric variable that correlates with the claim amount.  We will use it as the baseline and see if we can improve this models.
-
-#### Baseline Model
-
-```{r}
-baseline_lm <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
-summary(baseline_lm)
+```{r, null_prefix = TRUE, comment=NA}
+outlier_model <- glm(TARGET_AMT_OUTLIER ~ BLUEBOOK + MVR_PTS + JOB, family = binomial(link = "logit"), over_sample_train_2)
+summary(outlier_model)
+outlier_model_results <- evaluate_model(outlier_model, test, "TARGET_AMT_OUTLIER", 0.5)
 ```
@@
 
+```{r}
+temp <- over_sample_train_2 %>%
+  select(-TARGET_FLAG, - TARGET_AMT)
+fft <- FFTrees(TARGET_AMT_OUTLIER~., temp)
+fft
+```
+
 ## Predictions
@@
 ```{r}
-evaluation$TARGET_FLAG <- ifelse(predict.glm(model, evaluation, "response") >= 0.5, 1, 0)
+predictions <- function(df, classifier, linear_model_1, linear_model_2){
+  df$TARGET_FLAG <- ifelse(predict.glm(classifier, df, "response") >= 0.5, 1, 0)
   # We assume that everyone with a TARGET_FLAG = 0 has a TARGET_AMT as zero.  
-# Urban drivers will use one model, rural another
-evaluation$urban_yhat <- predict(urban_fit, evaluation)
-evaluation$rural_yhat <- predict(rural_fit, evaluation)
-evaluation <- evaluation %>%
-  mutate(TARGET_AMT = ifelse(URBAN_DRIVER == 1, urban_yhat, rural_yhat)) %>%
+  # We then refine it with the two linear models
+  df$model_1_yhat <- predict(linear_model_1, df)
+  df$model_2_yhat <- predict(linear_model_2, df)
+  df <- df %>%
+    mutate(TARGET_AMT = ifelse(URBAN_DRIVER == 1, model_1_yhat, model_2_yhat)) %>%
     mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT))
+  return(df)
+}
+
+evaluation <- predictions(evaluation, model, urban_fit, rural_fit)
 ```

### 2019-11-16 hw4/HW4_Adam.Rmd
Prep for sync

--- 2019-11-15 hw4/HW4_Mike.Rmd
@@
 library(ROCR)
-library(corrplot)
 # Thank you Stack Overflow!
@@
 set.seed(42)
-# ADAM THIS BROKE YOUR MODEL
-#train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
-train_index <- createDataPartition(df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
+train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
 train <- df[train_index,]
@@
 ```{r}
-over_sample <- function(df, predictors){
-  vars <- c("INDEX", "TARGET_FLAG", predictors)
-  df <- df[vars] %>%
-    na.omit()
 set.seed(42)
-  minority <- nrow(df[df$TARGET_FLAG == 1,])
-  majority <- nrow(df[df$TARGET_FLAG == 0,])
+minority <- nrow(train[train$TARGET_FLAG == 1,])
+majority <- nrow(train[train$TARGET_FLAG == 0,])
 diff <- majority - minority
-  minority_index <- df[df$TARGET_FLAG == 1,]$INDEX
-  over_sample_df <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
-    merge(df, .) %>%
-    bind_rows(df)
-  return(over_sample_df)
-}
+minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
+over_sample_train <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
+  merge(train, .) %>%
+  bind_rows(train)
+```
 
+The over sampled data frame has `r nrow(over_sample_train)` records, and as you can see in the following figure is now balanced:
+
+```{r, echo=FALSE}
+over_sample_train %>% 
+  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
+  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
+  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  labs(x="TARGET_FLAG", y="# Observations")
+
+over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
+```
+
+## Model Creation - Logistic
+
+### Baseline Model
+
+```{r}
 evaluate_model <- function(model, test_df){
@@
 }
-
 ```
-
-## Model Creation
-
-### Baseline Model
 
@@
 ```{r, null_prefix = TRUE, comment=NA}
-temp <- over_sample(train, c("PRIOR_ACCIDENT"))
-baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), temp)
+baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), over_sample_train)
 summary(baseline_model)
@@
 ```{r, null_prefix = TRUE, comment=NA}
-risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample(train, c("RED_SPORTS_CAR", "YOUNG_MALE")))
+risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample_train)
 summary(risk_taker_model)
@@
                            REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
-             family = binomial(link = "logit"), over_sample(train, c("YOUNG", "MSTATUS", "PRIOR_ACCIDENT", "SEX", "REVOKED", "MVR_PTS", "TRAVTIME", "CAR_USE")))
+             family = binomial(link = "logit"), over_sample_train)
 
@@
 
-```{r, eval=FALSE}
+```{r}
 # Evaluating the model
-eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = predict(traditional_cv,newdata=test,type="prob"))
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = 
+predict(traditional_cv,newdata=test,type="prob"))
 
@@
 
-### Manual Model
+### Misc
+
+Here we add meaningful variables to the model until we have a robust model with a decent AUC value.
 
 ```{r, null_prefix = TRUE, comment=NA}
-#"KIDSDRIV"         "AGE"             
-#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
-#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
-#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
-#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
-#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
-#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
-#[36] "KID_DRIVERS"     
-model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
-             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
-             family = binomial(link = "logit"), over_sample(train, c("PRIOR_ACCIDENT", "KID_DRIVERS", "MSTATUS", "INCOME", "SEX", "CAR_USE", "COLLEGE_EDUCATED", "REVOKED", "URBAN_DRIVER")))
+model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + LOG_INCOME +
+             SEX + CAR_USE + CAR_AGE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +
+               TIF + TRAVTIME + AGE,
+             family = binomial(link = "logit"), over_sample_train)
 summary(model)
@@
 
-## Amount Model
+```{r}
+# Evaluating the model
+eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(model,newdata=test,type="response"), prob = 
+predict(model,newdata=test,type="response"))
 
-To select build a model of how much the claim amount will be we need to subset the training data to only include those that have accidents.
-There will be `r train %>% filter(TARGET_FLAG == 1) %>% nrow()` data points in this set.
+pred <- prediction(eval$predicted, eval$actual)
+auc <- performance(pred, measure = "auc")@y.values[[1]]
+perf <- performance(pred,"tpr","fpr")
+plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
+```
+
+## Model Creation - Linear
+
+### Baseline Model
+
+As with the logistic model, we start with a baseline model using the `BLUEBOOK` variable as the sole predictor. All other models will need to beat this baseline.
+
+```{r, null_prefix = TRUE, comment=NA}
+
+baseline_model <- lm(log(TARGET_AMT+1) ~ BLUEBOOK -1, train)
+summary(baseline_model)
+```
+
+This model is rather poor, though the predictor is statistically significant in the model. We can build on this model and see if we can build a more robust version.
+
+### Car Model
+
+Here we look at the type of car the person drives, it's age, and how it is used.
 
 ```{r}
-amt_train <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  select(-INDEX, -TARGET_FLAG)
-amt_test <- test %>%
-  filter(TARGET_FLAG == 1)
+car_model <- lm(log(TARGET_AMT+1) ~ CAR_AGE + CAR_TYPE + CAR_USE, data = train)
+summary(car_model)
 ```
 
+Here we see that none of these predictors are significant in this model. In fact, taken one at a time, none are significant individually either.
+
+### Personal Model
+
+Here we use variables pertaining to the driver's personal life: family, income, driving record, and work.
+
 ```{r}
-M <- amt_train %>%
-  purrr::keep(is.numeric) %>%
-  #mutate(TARGET_AMT = log(TARGET_AMT)) %>%
-  na.omit() %>%
-  cor(.)
-corrplot(M, type = "upper")
+personal_model <- lm(log(TARGET_AMT+1) ~ MSTATUS + INCOME + KIDSDRIV +
+                       log(HOME_VAL+1) + AGE + REVOKED + MVR_PTS +
+                       HAS_KIDS, data= train)
+
+summary(personal_model)
 ```
 
-It looks like `BLUEBOOK` is the only numeric variable that correlates with the claim amount.  We will use it as the baseline and see if we can improve this models.
+Again, we see a very poor model. No variables are statistically significant here (though 2 are close).
+
+### Additive Model
+
+Here we start with `BLUEBOOK` from the baseline model and add to it until we get a decent model.
 
 ```{r}
-rural_fit <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 0) %>%
-  lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + TRAVTIME, .)
-urban_fit <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 1) %>%
-  lm(TARGET_AMT ~ BLUEBOOK, .)
-summary(urban_fit)
+add_model <- lm(1/sqrt(TARGET_AMT+1) ~ BLUEBOOK + CAR_AGE,
+                data = train)
+
+summary(add_model)
 ```
-
-```{r}
-M <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 1) %>%
-  purrr::keep(is.numeric) %>%
-  na.omit() %>%
-  cor(.)
-corrplot(M, type = "upper")
-```
-
-```{r}
-evaluation$TARGET_FLAG <- ifelse(predict.glm(model, evaluation, "response") >= 0.5, 1, 0)
-# We assume that everyone with a TARGET_FLAG = 0 has a TARGET_AMT as zero.  
-# Urban drivers will use one model, rural another
-evaluation$urban_yhat <- predict(urban_fit, evaluation)
-evaluation$rural_yhat <- predict(rural_fit, evaluation)
-evaluation <- evaluation %>%
-  mutate(TARGET_AMT = ifelse(URBAN_DRIVER == 1, urban_yhat, rural_yhat)) %>%
-  mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT))
-```

### 2019-11-25 final project/CoIL.R
More work on final

@@ -22,13 +22,14 @@ names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
                "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
                "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
                "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
-               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
-               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
-               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
-               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
-               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
-               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
-               "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
+               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM",
+               "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", 
+               "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND",
+               "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
+               "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT",
+               "ABROM", "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",
+               "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
+               "ABYSTAND", "CARAVAN")
 eval <- read.delim("data/ticeval2000.txt", header = FALSE)
 temp <- read.delim("data/tictgts2000.txt", header = FALSE)
 eval$CARAVAN <- temp$V1

### 2019-11-27 hw5/HW5.rmd

---
title: "Data-621-Project-5"
author: "Critical Thinking Group 3"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r}
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(purrr)
library(RColorBrewer)
library(ROCR)
library(corrplot)
```

```{r}

df <- read.csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw5/data/wine-training-data.csv")

hold_out_data <- read.csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw5/data/wine-evaluation-data.csv")
```

```{r}
str(df)
```
#Look at the Distribution of the Target Variable

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = df, aes(x = TARGET)) +
  geom_bar(width = 1, color = 'black',fill = I('orange'))+ggtitle("Number of Wine Cases Purchased")

```

## Training Test Split
```{r}
set.seed(42)
train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

## Lets Look at some decision Trees wooo!

```{r}
source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")

test <- RF_with_Nulls(train,test,"TARGET",.5,10,5,.005,4,1)

```

```{r}

df_complete <- df[complete.cases(df), ]

library(corrplot)
M <- cor(df_complete)
corrplot(M, method = "circle") #plot matrix
library("PerformanceAnalytics")
```

## Analyze Relationship Between Variables

### Strong Relationship between Expert Rating (STARS) and the number of cases purchased
```{r}
# Set color by cond
ggplot(df_complete, aes(x=as.factor(STARS), y=TARGET,color =as.factor(STARS))) + geom_boxplot()+
  labs(title = "Cases of Wine Purchased Compared to Expert Rating", x = "Expert Rating", y = "Cases Purchased", color = "Expert Rating")

```

```{r}
# Set color by cond
ggplot(df_complete, aes(x=as.factor(LabelAppeal), y=TARGET,color =as.factor(LabelAppeal))) + geom_boxplot()+
  labs(title = "Cases of Wine Purchased Compared to Label Appeal", x = "Label Appeal", y = "Cases Purchased", color = "LabelAppeal")

```

## Fitting a Poisson Model to the Data

```{r}
model <- glm(TARGET ~ LabelAppeal + STARS +AcidIndex+VolatileAcidity,
             family = poisson, train)

summary(model)

test$Prediction <- predict.glm(model, newdata=test, type="response")

cor(test$Prediction,test$TARGET)
```

### 2019-11-29 final project/Manuscript.Rmd

--- 2019-11-11 final project/Manuscript.Rmd
@@
 shorttitle        : " "
-author: 
-  - name          : "Corey Arnouts"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "119 W 31st St., New York, NY 10001"
-    email         : "Corey.Arnouts@spsmail.cuny.edu"
-  - name          : "Adam Douglas"
-    affiliation   : "1"
-  - name          : "Jason Givens-Doyle"
-    affiliation   : "1"
-  - name          : "Michael Silva"
-    affiliation   : "1"
-affiliation:
-  - id            : "1"
-    institution   : "MS in Data Science Students CUNY School of Professional Studies"
+author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
+affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
 abstract: |
@@
 output: 
+  word_document:
+    reference_docx: word-template.docx
   papaja::apa6_pdf:
@@
 library("papaja")
-
+library(kableExtra)
 library(tidyverse)
+library(pander)
 
@@
 
-```{r analysis-preferences}
+```{r analysis-preferences, echo=FALSE}
 # Seed for random number generation
@@
 
+```{r data, echo=FALSE}
+df <- read.delim("data/ticdata2000.txt", header = FALSE)
+names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
+               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
+               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
+               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
+               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
+               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
+               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
+               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
+               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
+               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
+               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
+               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
+               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
+               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
+               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
+               "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
+eval <- read.delim("data/ticeval2000.txt", header = FALSE)
+temp <- read.delim("data/tictgts2000.txt", header = FALSE)
+eval$CARAVAN <- temp$V1
+names(eval) <- names(df)
+```
 # Introduction
@@
 
-In total 147 participants registered and 43 submitted solutions [@Putten]. In this paper we set out to complete the first part of the COIL Challenge. **SUMARISE FINDINGS?**
+```{r, echo=FALSE}
+# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
+prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121),
+                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
+ggplot(data=prediction_scores, aes(x=score, y=n)) +
+  geom_bar(stat="identity") +
+  xlab("Number of Correct Predictions") +
+  ylab("CoIL Challenge Submissions") +
+  xlim(37, 122)
+```
+
+In total 147 participants registered and 43 submitted solutions [@Putten].
+The maximum number of policyowners that could be found was 238.
+The submissions identified 95 policy owners on average.
+The winning model [@Elkan] identified 121 policy owners.
+Random selection results in identifying 42 policy owners.
+The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners.
+[@Putten].
+
+In this paper we set out to complete the first part of the COIL Challenge. <!--**SUMARISE FINDINGS?**-->
 
@@
 
-Participants used a variety of approaches in formulating their submissions including: Boosted Decision Tree [@McKone], Classification and Regressio Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].
+Participants used a variety of approaches in formulating their submissions including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].
 
-```{r}
-# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
-prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121),
-                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
-ggplot(data=prediction_scores, aes(x=score, y=n)) +
-  geom_bar(stat="identity") +
-  xlab("Policy Owners") +
-  ylab("Submissions") +
-  xlim(37, 122) +
-  ggtitle("Correct Predictions from CoIL 2000 Challenge Submissions")
+<!--Our investigation is similar to @Doornik in that we used a LOGIT model. **Reproduce his model and describe the results**-->
+
+```{r, echo=FALSE}
+doornik <- function(df){
+  df <- df %>%
+    mutate(PBRAND_2_or_6 = ifelse(PBRAND == 2 | PBRAND == 6, 1, 0),
+           PBRAND_3_or_4 = ifelse(PBRAND == 3 | PBRAND == 4, 1, 0),
+           PPLEZIER_gt_0 = ifelse(PPLEZIER > 0, 1, 0),
+           PPERSAUT_6 = ifelse(PPERSAUT == 6, 1, 0))
+  z <- -4.1 - (0.13 * df$MOPLLAAG) - (0.3 * df$MINK123M) + (0.17 * df$MINKGEM) + (0.54 * (df$ALEVEN + df$AFIETS + df$ABYSTAND)) - df$PBRAND_2_or_6 + (0.8 * df$PBRAND_3_or_4) + (2.4 * df$PPLEZIER_gt_0) + (1.5 * df$PPERSAUT_6) + (0.28 * (df$PWAOREG - df$PLEVEN))
+  p_caravan <- exp(z) / (1 + exp(z))
+  #p_caravan
+  ifelse(p_caravan > 0.5, 1, 0)
+}
+
+eval$doornik <- doornik(eval)
 ```
-
-The maximum number of policyowners that could be found was 238.
-The submissions identified 95 policy owners on average.
-The winning model [@Elkan] identified 121 policy owners.
-Random selection results in identifying 42 policy owners.
-The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners.
-[@Putten].
-
-Our investigation is similar to @Doornik in that we used a LOGIT model. **Reproduce his model and describe the results**
 
@@
 # References
-```{r create_references}
+```{r create_references, echo=FALSE}
 r_refs(file = "Manuscript.bib")
@@
 -->
+
+## Data Dictionary
+
+```{r, results='asis', echo=FALSE}
+data_dictionary <- data.frame(
+  Name=c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN"),
+  Description = c("Customer Subtype", "Number of houses", "Avg size household", "Avg age", "Customer main type", "Roman catholic", "Protestant", "Other religion", "No religion", "Married", "Living together", "Other relation", "Singles", "Household without children", "Household with children", "High level education", "Medium level education", "Lower level education", "High status", "Entrepreneur", "Farmer", "Middle management", "Skilled labourers", "Unskilled labourers", "Social class A", "Social class B1", "Social class B2", "Social class C", "Social class D", "Rented house", "Home owners", "1 car", "2 cars", "No car", "National Health Service", "Private health insurance", "Income < 30.000", "Income 30-45.000", "Income 45-75.000", "Income 75-122.000", "Income >123.000", "Average income", "Purchasing power class", "Contribution private third party insurance", "Contribution third party insurance (firms)", "Contribution third party insurane (agriculture)", "Contribution car policies", "Contribution delivery van policies", "Contribution motorcycle/scooter policies", "Contribution lorry policies", "Contribution trailer policies", "Contribution tractor policies", "Contribution agricultural machines policies", "Contribution moped policies", "Contribution life insurances", "Contribution private accident insurance policies", "Contribution family accidents insurance policies", "Contribution disability insurance policies", "Contribution fire policies", "Contribution surfboard policies", "Contribution boat policies", "Contribution bicycle policies", "Contribution property insurance policies", "Contribution social security insurance policies", "Number of private third party insurance", "Number of third party insurance (firms)", "Number of third party insurane (agriculture)", "Number of car policies", "Number of delivery van policies", "Number of motorcycle/scooter policies", "Number of lorry policies", "Number of trailer policies", "Number of tractor policies", "Number of agricultural machines policies", "Number of moped policies", "Number of life insurances", "Number of private accident insurance policies", "Number of family accidents insurance policies", "Number of disability insurance policies", "Number of fire policies", "Number of surfboard policies", "Number of boat policies", "Number of bicycle policies", "Number of property insurance policies", "Number of social security insurance policies", "Number of mobile home policy"))
+pander(data_dictionary)
+```
+
+\newpage
+
 ## R statistical programming code.
@@
 <code>
-```{r, null_prefix = TRUE, comment=NA}
+```{r, null_prefix = TRUE, comment=NA, echo=FALSE}
 conn <- file("CoIL.r", open="r")

### 2019-11-29 hw4/HW4_Mike.Rmd

@@
 ```{r}
-df <- df %>% 
+fix_missing <- function(df) {
+  df %>% 
     mutate_at(vars(c("CAR_AGE", "YOJ", "AGE", "INCOME", "HOME_VAL")), ~ifelse(is.na(.), median(., na.rm = TRUE), .))
+}
+
+df <- fix_missing(df)
+evaluation <- fix_missing(evaluation)
 ```
@@
            YOUNG = factor(ifelse(AGE < 25, 1, 0)),
+           STUDENT = factor(ifelse(JOB == "Student", 1, 0)),
+           MVR_PTS_GT_0 = factor(ifelse(MVR_PTS > 0, 1, 0)),
            RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
@@
 
+By simply using the median to predict the values of the claims we would underestimate the actual value on average by $1,600 dollars.  On a percentage basis the error would be 50% of the actual.
+
 #### Simple Linear Model
@@
 ```
+
+Well the average error has decreased which is good but the average error share of the actual has doubled.  This is a garbage model.
 
@@
 
-```{r}
-temp <- data.frame(yhat = predict(outlier_lm, amt_test), actual = amt_test$TARGET_AMT, TARGET_AMT_OUTLIER = amt_test$TARGET_AMT_OUTLIER) 
-temp %>%
-  ggplot(., aes(actual, yhat, color=as.factor(TARGET_AMT_OUTLIER))) +
-  geom_point() +
-  scale_color_brewer(palette="Set1") +
-  theme_light() +
-  theme(legend.position = "none") 
-```
-
-```{r}
-temp %>%
-  mutate(error = yhat - actual) %>%
-  mutate(`error %` = (error / actual)*100) %>%
-  group_by(TARGET_AMT_OUTLIER) %>%
-  summarise(error = mean(error),
-            `error %` = mean(`error %`)) %>%
-  kable() %>%
-  kable_styling()
-```
-
-It understates the claim prediciton by about $95 for the small claims and $1,100 for the large claims.
-The error on the is about 55% of the estimate for the small claims and 31% for the large claims.
-In summary it is roughly right.
-It would preform a lot better than the baseline model.
-The table below offers the similar metrics:
-
-```{r}
-data.frame(yhat = predict(baseline_lm, amt_test), actual = amt_test$TARGET_AMT, TARGET_AMT_OUTLIER = amt_test$TARGET_AMT_OUTLIER) %>%
-  mutate(error = yhat - actual) %>%
-  mutate(`error %` = (error / actual)*100) %>%
-  group_by(TARGET_AMT_OUTLIER) %>%
-  summarise(error = mean(error),
-            `error %` = mean(`error %`)) %>%
-  kable() %>%
-  kable_styling()
-```
+The predictions on average understate the actual value by $140.
+The error as a percent of the actual is close to what we get with just using the median.
+The improvement really comes from using a higher value to estimate the claim for the outliers.
 
@@
 There are `r nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])` out of `r nrow(amt_train)` that are outliers.
-Let's create another balanced data set and see if we can build a classifier.
+Let's create another balanced data set and see if we can build another classifier.
 
@@
 
-```{r}
+```{r eval=FALSE, echo=FALSE}
 over_sample_train_2 %>%
@@
 
+It looks like the `BLUEBOOK` and `PARENT1` are the only variables that are correlated with this outcome.  So let's build the classifier.
+
 ```{r, null_prefix = TRUE, comment=NA}
-outlier_model <- glm(TARGET_AMT_OUTLIER ~ BLUEBOOK + MVR_PTS + JOB, family = binomial(link = "logit"), over_sample_train_2)
+outlier_model <- glm(TARGET_AMT_OUTLIER ~ BLUEBOOK + PARENT1, family = binomial(link = "logit"), over_sample_train_2)
 summary(outlier_model)
-outlier_model_results <- evaluate_model(outlier_model, test, "TARGET_AMT_OUTLIER", 0.5)
+outlier_model_results <- evaluate_model(outlier_model, test, "TARGET_AMT_OUTLIER")
 ```
 
-#### Rural Model
+Now to test this out.
+Based on our linear regression test set we will predict if the observation is likely to be an outlier or not.
+We will then estimate the `TARGET_AMT` using the linear model.
+We will then evaluate the results of the "model".
+
 ```{r}
-M <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 0) %>%
-  purrr::keep(is.numeric) %>%
-  na.omit() %>%
-  cor(.)
-corrplot(M, type = "upper")
+temp <- amt_test
+temp$TARGET_AMT_OUTLIER <- ifelse(predict.glm(outlier_model, amt_test, "response") >= 0.5, 1, 0)
+temp$TARGET_AMT_yhat <- predict(outlier_lm, amt_test)
+evaluate_lm(temp$TARGET_AMT_yhat, temp$TARGET_AMT)
 ```
 
-```{r}
-rural_fit <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 0) %>%
-  lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + TRAVTIME, .)
-summary(rural_fit)
-```
-
-#### Urban Model
-
-```{r}
-M <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 1) %>%
-  purrr::keep(is.numeric) %>%
-  na.omit() %>%
-  cor(.)
-corrplot(M, type = "upper")
-```
-
-```{r}
-urban_fit <- train %>%
-  filter(TARGET_FLAG == 1) %>%
-  filter(URBAN_DRIVER == 1) %>%
-  lm(TARGET_AMT ~ BLUEBOOK, .)
-summary(urban_fit)
-```
-
-```{r}
-temp <- over_sample_train_2 %>%
-  select(-TARGET_FLAG, - TARGET_AMT)
-fft <- FFTrees(TARGET_AMT_OUTLIER~., temp)
-fft
-```
+So through this somewhat convoluted approach we have done slightly better than using the median.
 
@@
 
+The final predictions will use the alternative traditional model to predict the `TARGET_FLAG` variable, then pass it through the second classifier which predicts the `TARGET_AMT_OUTLIER` flag.
+We then use the linear model to make the final amount prediction.
+
 ```{r}
-predictions <- function(df, classifier, linear_model_1, linear_model_2){
-  df$TARGET_FLAG <- ifelse(predict.glm(classifier, df, "response") >= 0.5, 1, 0)
+predictions <- function(df, classifier_1, linear_model, classifier_2){
+  df$TARGET_FLAG <- ifelse(predict.glm(classifier_1, df, "response") >= 0.5, 1, 0)
   # We assume that everyone with a TARGET_FLAG = 0 has a TARGET_AMT as zero.  
-  # We then refine it with the two linear models
-  df$model_1_yhat <- predict(linear_model_1, df)
-  df$model_2_yhat <- predict(linear_model_2, df)
+  # We then refine it with the outlier model.
+  df$TARGET_AMT_OUTLIER <- ifelse(predict.glm(classifier_2, df, "response") >= 0.5, 1, 0)
+  df$yhat <- predict(linear_model, df)
   df <- df %>%
-    mutate(TARGET_AMT = ifelse(URBAN_DRIVER == 1, model_1_yhat, model_2_yhat)) %>%
-    mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT))
+    mutate(TARGET_AMT = ifelse(TARGET_AMT_OUTLIER == 1, yhat, 0)) %>%
+    mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT)) %>%
+    select(-yhat, -TARGET_AMT_OUTLIER)
   return(df)
@@
 
-evaluation <- predictions(evaluation, model, urban_fit, rural_fit)
+evaluation <- predictions(evaluation, model, outlier_lm, outlier_model)
 ```

### 2019-11-30 hw5/HW5_Adam.Rmd
HW5
Added HW5_Adam.Rmd file

---
title: "Data-621-Project-5"
author: "Critical Thinking Group 3"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(purrr)
library(RColorBrewer)
library(ROCR)
library(corrplot)
```

## Importing Data

```{r}
# Read in training data
df <- read.csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw5/data/wine-training-data.csv")

# Read in evaluation data
hold_out_data <- read.csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw5/data/wine-evaluation-data.csv")
```

Datasets were provided for both training and evaluation. The training dataset consists of `r nrow(df)` observations while the evaluation dataset (which we will set aside for now) consists of `r nrow(hold_out_data)` observations.

```{r}
str(df)
summary(df)
```

Looking at our data, we have a good number of missing values in some variables. We'll try to reason a way to impute these values (if possible).

### Missing Values

#### STARS

First we start with the `STARS` variable. There are `r df %>% filter(is.na(STARS)) %>% nrow()` observations without a value populated. Looking at what is populated, the variable was imported as a numeric value, though it is used as a discrete variable. Since it isn't too much of a stretch to view rating as a numeric here (e.g. a 2 star wine is half as good as a 4 star wine), we'll keep the variable type as-is.

For the NA values, it makes sense that unrated wines would be somewhere in the middle of the pack or, possibly below it (after all, if it was such a good wine, why hasn't it come to the attention of experts to be rated?). The summary above shows a mean and median value at or close to 2, so we'll set our NAs to 2.

```{r}
df$STARS <- ifelse(is.na(df$STARS),2,df$STARS)
```

#### Alcohol

The `Alcohol` variable is a measure of the alcohol content. Looking at our summary, we see that there are some missing values. However, there are also some negative values - `r df %>% filter(Alcohol < 0) %>% nrow()` to be specific.

Obviously a wine cannot have a negative alcohol value, so we will assume data-entry issues caused the minus sign and change those to positive values. For missing values, we will assume they are alcohol-free variants (which do exist!) and set the value to zero.

```{r}
df$Alcohol <- ifelse(is.na(df$Alcohol),0,df$Alcohol)
df$Alcohol <- abs(df$Alcohol)
```

#### Others

For other variables with NA values such as: `ResidualSugar`, `Chlorides`, `FreeSulfurDioxide`, `TotalSulfurDioxide`, `pH`, and `Sulphates`, we will impute using the median value.

```{r}
df$ResidualSugar <- ifelse(is.na(df$ResidualSugar),
                           median(df$ResidualSugar,na.rm=T),
                           df$ResidualSugar)

df$Chlorides <- ifelse(is.na(df$Chlorides),
                           median(df$Chlorides,na.rm=T),
                           df$Chlorides)

df$FreeSulfurDioxide <- ifelse(is.na(df$FreeSulfurDioxide),
                           median(df$FreeSulfurDioxide,na.rm=T),
                           df$FreeSulfurDioxide)

df$TotalSulfurDioxide <- ifelse(is.na(df$TotalSulfurDioxide),
                           median(df$TotalSulfurDioxide,na.rm=T),
                           df$TotalSulfurDioxide)

df$pH <- ifelse(is.na(df$pH),
                           median(df$pH,na.rm=T),
                           df$pH)

df$Sulphates <- ifelse(is.na(df$Sulphates),
                      median(df$Sulphates,na.rm=T),
                      df$Sulphates)
```

## Data Exploration

### Look at the Distribution of the Target Variable

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = df, aes(x = factor(TARGET))) +
  geom_bar(width = 1, color = 'black',fill = I('orange')) +
  ggtitle("Number of Wine Cases Purchased") +
  scale_x_discrete(breaks = seq(0,8,1)) +
  theme_light() +
  theme(
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype=2),
        panel.grid.minor.y = element_blank()
        ) +
  labs(x="TARGET variable",y="Count")
```

Other than wines that have no cases sold, the distribution seems to be somewhat normal in shape.

### Training Test Split

Next we will look at splitting the data into a training and testing set at a standard 70-30 split between train and test:

```{r}
set.seed(42)
train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Decision Trees

Now that we have our testing set, we will use decision trees to help us decide which variables to potentially use in our linear model.

```{r, message=FALSE, results="hide"}
source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")

RF_with_Nulls(train,test,"TARGET",.5,10,5,.005,4,1)

```

We can also use the correlation between variables to try and deduce the best predictor variables to use in a linear model:

```{r, message=FALSE}
library(corrplot)
M <- cor(test)
corrplot(M, method = "circle") #plot matrix
```

Here we see the `TARGET` variable (which is our dependent variable) has some correlation with `LabelAppeal`, `AcidIndex`, and `STARS`. These variables also make appearances in the decision trees above, so they bear some investigation. Let's look at these relationships in turn.

### Analyze Relationship Between Variables

#### Variable STARS

```{r}
# Set color by cond
ggplot(df, aes(x=as.factor(STARS), y=TARGET,color =as.factor(STARS))) + geom_boxplot()+
  labs(title = "Cases of Wine Purchased Compared to Expert Rating", x = "Expert Rating", y = "Cases Purchased", color = "Expert Rating")

```

There is some clear correlation, as expected.

#### Variable LabelAppeal

```{r}
# Set color by cond
ggplot(df, aes(x=as.factor(LabelAppeal), y=TARGET,color =as.factor(LabelAppeal))) + geom_boxplot()+
  labs(title = "Cases of Wine Purchased Compared to Label Appeal Rating", x = "Label Rating", y = "Cases Purchased", color = "Label Rating")

```

Here we see a rather obvious correlation between cases purchased and the rating of the label. This will certainly be a variable of interest in our model.

#### Variable AcidIndex

```{r}
# Set color by cond
ggplot(df, aes(x=as.factor(AcidIndex), y=TARGET,color =as.factor(AcidIndex))) + geom_boxplot() +
  labs(title = "Cases of Wine Purchased Compared to Mean Acidity", x = "Acidity", y = "Cases Purchased", color = "Acidity")

```

Here we see a far less clear correlation, but there is still value to possibly including it in our model.

## Modeling

### Poisson Models

Since we are modeling counts, we will attempt a Poisson model first using the three variables we identified above:

```{r}
model <- glm(TARGET ~ LabelAppeal + STARS + AcidIndex,
             family = poisson, train)

summary(model)

test$Prediction <- predict.glm(model, newdata=test, type="response")
```

Here we have a correlation of `r round(cor(test$Prediction,test$TARGET,use="complete.obs"),2)` with our `TARGET` variable and a residual sum of squares (RSS) of `r sum(model$residuals^2)`.

Although a good start, this model doesn't fit all that well. Next, we try to add some additional variables:

```{r}
model2 <- glm(TARGET ~ LabelAppeal + STARS + AcidIndex + VolatileAcidity +
                TotalSulfurDioxide + FreeSulfurDioxide + Chlorides,
             family = poisson, train)

summary(model2)

test$Prediction2 <- predict.glm(model2, newdata=test, type="response")
```

This model has additional significant variables, we have the same correlation of `r round(cor(test$Prediction,test$TARGET,use="complete.obs"),2)` with our `TARGET` variable as in model #1 above, and the residual sum of squares (RSS) of `r sum(model$residuals^2)` is actually slightly worse.

### Diagnostics

We'll look at the residual values and see what they tell us about our models thus far. We'll plot the adjusted residuals: $\frac{\text{Observed} - \text{Predicted}}{\text{SD}(\text{Observed} - \text{Predicted}}$ versus the observed values and see if any pattern emerges.

```{r}
# Model 1
residModel1 <- data.frame(obs = test$TARGET, diff = (test$TARGET - test$Prediction))
residModel1$adjRes <- (residModel1$diff / sd(residModel1$diff))

# Model 2
residModel2 <- data.frame(obs = test$TARGET, diff = (test$TARGET - test$Prediction2))
residModel2$adjRes <- (residModel2$diff / sd(residModel2$diff))

residModel1 %>% ggplot(aes(x=factor(obs), y=adjRes)) + geom_boxplot() +
  labs(title="Adjusted Residuals", subtitle="Poisson Model #1",
       x="Observed Value",y="Adjusted Residual")

residModel2 %>% ggplot(aes(x=factor(obs), y=adjRes)) + geom_boxplot() +
  labs(title="Adjusted Residuals", subtitle="Poisson Model #2",
       x="Observed Value",y="Adjusted Residual")
```

There is an obvious pattern with our residuals. We tend to overestimate more on larger observed values of `TARGET`.

### 2019-12-03 final project/Manuscript_Corey.RMD
Added some simple correlation plots
I just added some simple correlation plots also let me know if you want to keep the variables names as the acronyms I just did it so they are more readable for the correlation plots

--- 2019-11-29 final project/Manuscript.Rmd
@@
 The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
-  <!-- https://tinyurl.com/ybremelq 
-    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
 
@@
   -->
+  <!-- https://tinyurl.com/ybremelq 
+    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
 keywords          : "CoIL Challenge, Logistic Regression"
@@
 output: 
+  html_document: default
   word_document:
@@
     #citation_package: natbib #biblatex
-  html_document: default
 ---
@@
 
+# Analysis
+
+```{r}
+colnames(df)<-c("Customer_Subtype","Number_of_houses", "Avg_size_household", "Avg_age", "Customer_main_type", "Roman_catholic", "Protestant","Other_religion", "No_religion", "Married", "Living_together", "Other_relation", "Singles", "Household_without_children", "Household_with_children", "High_level_education", "Medium_level_education", "Lower_level_education", "High_status", "Entrepreneur", "Farmer", "Middle_management", "Skilled_labourers", "Unskilled_labourers", "Social_class_A", "Social_class_B1", "Social_class_B2", "Social_class_C", "Social_class_D", "Rented_house", "Home_owners", "1_car", "2_cars", "No_car", "National_Health_Service", "Private_health_insurance", "Income_<_30.000", "Income_30-45.000", "Income_45-75.000", "Income_75-122.000", "Income_>123.000", "Average_income", "Purchasing_power_class", "Contribution_private_third_party_insurance", "Contribution_third_party_insurance_(firms)", "Contribution_third_party_insurane_(agriculture)", "Contribution_car_policies", "Contribution_delivery_van_policies", "Contribution_motorcycle/scooter_policies", "Contribution_lorry_policies", "Contribution_trailer_policies", "Contribution_tractor_policies", "Contribution_agricultural_machines_policies", "Contribution_moped_policies", "Contribution_life_insurances", "Contribution_private_accident_insurance_policies", "Contribution_family_accidents_insurance_policies", "Contribution_disability_insurance_policies", "Contribution_fire_policies", "Contribution_surfboard_policies", "Contribution_boat_policies", "Contribution_bicycle_policies", "Contribution_property_insurance_policies", "Contribution_social_security_insurance_policies", "Number_of_private_third_party_insurance", "Number_of_third_party_insurance_(firms)", "Number_of_third_party_insurane_(agriculture)", "Number_of_car_policies", "Number_of_delivery_van_policies", "Number_of_motorcycle/scooter_policies", "Number_of_lorry_policies", "Number_of_trailer_policies", "Number_of_tractor_policies", "Number_of_agricultural_machines_policies", "Number_of_moped_policies", "Number_of_life_insurances", "Number_of_private_accident_insurance_policies", "Number_of_family_accidents_insurance_policies", "Number_of_disability_insurance_policies", "Number_of_fire_policies", "Number_of_surfboard_policies", "Number_of_boat_policies", "Number_of_bicycle_policies", "Number_of_property_insurance_policies", "Number_of_social_security_insurance_policies", "Number_of_mobile_home_policy")
+
+colnames(eval)<-c("Customer_Subtype","Number_of_houses", "Avg_size_household", "Avg_age", "Customer_main_type", "Roman_catholic", "Protestant","Other_religion", "No_religion", "Married", "Living_together", "Other_relation", "Singles", "Household_without_children", "Household_with_children", "High_level_education", "Medium_level_education", "Lower_level_education", "High_status", "Entrepreneur", "Farmer", "Middle_management", "Skilled_labourers", "Unskilled_labourers", "Social_class_A", "Social_class_B1", "Social_class_B2", "Social_class_C", "Social_class_D", "Rented_house", "Home_owners", "1_car", "2_cars", "No_car", "National_Health_Service", "Private_health_insurance", "Income_<_30.000", "Income_30-45.000", "Income_45-75.000", "Income_75-122.000", "Income_>123.000", "Average_income", "Purchasing_power_class", "Contribution_private_third_party_insurance", "Contribution_third_party_insurance_(firms)", "Contribution_third_party_insurane_(agriculture)", "Contribution_car_policies", "Contribution_delivery_van_policies", "Contribution_motorcycle/scooter_policies", "Contribution_lorry_policies", "Contribution_trailer_policies", "Contribution_tractor_policies", "Contribution_agricultural_machines_policies", "Contribution_moped_policies", "Contribution_life_insurances", "Contribution_private_accident_insurance_policies", "Contribution_family_accidents_insurance_policies", "Contribution_disability_insurance_policies", "Contribution_fire_policies", "Contribution_surfboard_policies", "Contribution_boat_policies", "Contribution_bicycle_policies", "Contribution_property_insurance_policies", "Contribution_social_security_insurance_policies", "Number_of_private_third_party_insurance", "Number_of_third_party_insurance_(firms)", "Number_of_third_party_insurane_(agriculture)", "Number_of_car_policies", "Number_of_delivery_van_policies", "Number_of_motorcycle/scooter_policies", "Number_of_lorry_policies", "Number_of_trailer_policies", "Number_of_tractor_policies", "Number_of_agricultural_machines_policies", "Number_of_moped_policies", "Number_of_life_insurances", "Number_of_private_accident_insurance_policies", "Number_of_family_accidents_insurance_policies", "Number_of_disability_insurance_policies", "Number_of_fire_policies", "Number_of_surfboard_policies", "Number_of_boat_policies", "Number_of_bicycle_policies", "Number_of_property_insurance_policies", "Number_of_social_security_insurance_policies", "Number_of_mobile_home_policy","doornik")
+
+```
+
+# Relationships Between PolicyTypes
+```{r}
+
+policies <- df[ , grepl( "policies" , names( df ) ) ]
+policies <- policies[ , grepl( "Number" , names( policies ) ) ]
+
+policies2 <-sub("Number_of_", "", colnames(policies))
+policies3 <-sub("_policies", "", policies2)
+policies4 <-sub("insurance", "", policies3)
+policies5 <-sub("_", " ", policies4)
+policies6 <-sub("_", " ", policies5)
+
+colnames(policies)<- policies6
+
+cordf <- df %>% select(Social_class_A,Average_income)
+
+library(corrplot)
+M <- cor(policies)
+corrplot(M) #plot matrix
+
+```
+
+# Relationships Between Variables
+```{r}
+features <- df[ , !grepl( "policies" , names( df ) ) ]
+features <- features[ , !grepl( "Number" , names( features ) ) ]
+features <- features[ , !grepl( "Insurance" , names( features ) ) ]
+features <- features[ , !grepl( "insurance" , names( features ) ) ]
+features <- features[ , !grepl( "insurane" , names( features ) ) ]
+#features <- features[ , grepl( "education|income|Income" , names( features ) ) ]
+
+library(corrplot)
+M <- cor(features)
+corrplot(M,tl.cex = .5) #plot matrix
+
+features <- df[ , !grepl( "policies" , names( df ) ) ]
+features <- features[ , !grepl( "Number" , names( features ) ) ]
+features <- features[ , !grepl( "Insurance" , names( features ) ) ]
+features <- features[ , !grepl( "insurance" , names( features ) ) ]
+features <- features[ , !grepl( "insurane" , names( features ) ) ]
+features <- features[ , grepl( "education|income|Income" , names( features ) ) ]
+
+library(corrplot)
+M <- cor(features)
+corrplot(M,tl.cex = .5) #plot matrix
+
+```
+
 # Literature Review

### 2019-12-03 final project/Draft.Rmd
Initial commit of draft

---
title             : "A Logistic Regression Approach to CoIL Challenge 2000"
shorttitle        : " "
author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
abstract: |
  A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
  <!-- https://tinyurl.com/ybremelq 
    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  -->
keywords          : "CoIL Challenge, Logistic Regression"
wordcount         : "X"
bibliography      : ["Manuscript.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output: 
  word_document:
    reference_docx: word-template.docx
  papaja::apa6_pdf:
    #citation_package: natbib #biblatex
  html_document: default
---

```{r setup, include = FALSE}
# Documentation at https://crsh.github.io/papaja_man/
# If you have no use for TeX beyond rendering R Markdown documents, I recommend you use TinyTex.
# install.packages("tinytex")
# tinytex::install_tinytex()
if (!"papaja" %in% rownames(installed.packages())) {
  devtools::install_github("crsh/papaja")
}
library("papaja")
library(kableExtra)
library(tidyverse)
library(pander)
library(gridExtra)
library(corrplot)
library(ggmosaic)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                     x, perl = TRUE)
  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr("\\]", x)[1] - 1
    # Two cases: start or newline
    re <- paste0("^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub(re, rep,  x)
    re <- paste0("\\\n.{", align_index, "}\\]")
    rep <- paste0("\n", comment)
    x <- gsub(re, rep,  x)
  }

  default_output_hook(x, options)

})
knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
```

```{r analysis-preferences, echo=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r data, echo=FALSE}
prepare_data <- function(df){
  names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
               "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")

  MOSTYPE_labels <- c("1" = "1 High Income, expensive child",
                      "2" = "2 Very Important Provincials",
                      "3" = "3 High status seniors",
                      "4" = "4 Affluent senior apartments",
                      "5" = "5 Mixed seniors",
                      "6" = "6 Career and childcare",
                      "7" = "7 Dinki's (double income no kids)",
                      "8" = "8 Middle class families",
                      "9" = "9 Modern, complete families",
                      "10" = "10 Stable family",
                      "11" = "11 Family starters",
                      "12" = "12 Affluent young families",
                      "13" = "13 Young all american family",
                      "14" = "14 Junior cosmopolitan",
                      "15" = "15 Senior cosmopolitans",
                      "16" = "16 Students in apartments",
                      "17" = "17 Fresh masters in the city",
                      "18" = "18 Single youth",
                      "19" = "19 Suburban youth",
                      "20" = "20 Etnically diverse",
                      "21" = "21 Young urban have-nots",
                      "22" = "22 Mixed apartment dwellers",
                      "23" = "23 Young and rising",
                      "24" = "24 Young, low educated", 
                      "25" = "25 Young seniors in the city",
                      "26" = "26 Own home elderly",
                      "27" = "27 Seniors in apartments",
                      "28" = "28 Residential elderly",
                      "29" = "29 Porchless seniors: no front yard",
                      "30" = "30 Religious elderly singles",
                      "31" = "31 Low income catholics",
                      "32" = "32 Mixed seniors",
                      "33" = "33 Lower class large families",
                      "34" = "34 Large family, employed child",
                      "35" = "35 Village families",
                      "36" = "36 Couples with teens 'Married with children'",
                      "37" = "37 Mixed small town dwellers",
                      "38" = "38 Traditional families",
                      "39" = "39 Large religous families",
                      "40" = "40 Large family farms",
                      "41" = "41 Mixed rurals")

MGEMLEEF_labels <- c("1" = "20-30 years",
                     "2" = "30-40 years",
                     "3" = "40-50 years",
                     "4" = "50-60 years",
                     "5" = "60-70 years",
                     "6" = "70-80 years")

MOSHOOFD_labels <- c("1" = "Successful hedonists",
                     "2" = "Driven Growers",
                     "3" = "Average Family",
                     "4" = "Career Loners",
                     "5" = "Living well",
                     "6" = "Cruising Seniors",
                     "7" = "Retired and Religeous",
                     "8" = "Family with grown ups",
                     "9" = "Conservative families",
                     "10" = "Farmers")

MGODRK_labels <- c("0" = "0%",
                   "1" = "1 - 10%",
                   "2" = "11 - 23%",
                   "3" = "24 - 36%",
                   "4" = "37 - 49%",
                   "5" = "50 - 62%",
                   "6" = "63 - 75%",
                   "7" = "76 - 88%",
                   "8" = "89 - 99%",
                   "9" = "100%")

PWAPART_labels <- c("0" = "f 0",
                    "1" = "f 1 – 49",
                    "2" = "f 50 – 99",
                    "3" = "f 100 – 199",
                    "4" = "f 200 – 499",
                    "5" = "f 500 – 999",
                    "6" = "f 1000 – 4999",
                    "7" = "f 5000 – 9999",
                    "8" = "f 10.000 - 19.999",
                    "9" = "f 20.000 - ?")

df <- df %>%
  mutate(MOSTYPE = as.factor(MOSTYPE),
         MGEMLEEF = as.factor(MGEMLEEF),
         MOSHOOFD = as.factor(MOSHOOFD),
         MGODRK = as.factor(MGODRK),
         PWAPART = as.factor(PWAPART),
         CARAVAN = as.factor(CARAVAN)) %>%
  mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
         MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
         MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
         MGODRK = recode(MGODRK, !!!MGODRK_labels),
         PWAPART = recode(PWAPART, !!!PWAPART_labels))
}

eval <- read.delim("data/ticeval2000.txt", header = FALSE)
temp <- read.delim("data/tictgts2000.txt", header = FALSE)
eval$CARAVAN <- temp$V1
eval <- prepare_data(eval)
df <- prepare_data(read.delim("data/ticdata2000.txt", header = FALSE))
```

# Introduction
<!-- Describe the background and motivation of your problem.-->

Businesses use data science to extract insights from data.  It has many practical business applications.  Identifying households to include in a marketing campaign is one application.  One example using real world data is the Computational Intelligence and Learning (CoIL) Challenge. The CoIL Challenge competition was held from March 17 to May 8 in 2000.  The challenge is to:

1.  Identify potential customers for an insurance policy; and

2.  Provide a description of this customer base.

```{r, echo=FALSE}
# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83,
                                          86, 91, 94, 95, 96, 97, 99, 101, 103,
                                          104, 105, 106, 107, 108, 109, 110,
                                          111, 112, 115, 121),
                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3,
                                      1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1,
                                      1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Number of Correct Predictions") +
  ylab("CoIL Challenge Submissions") +
  xlim(37, 122)
```

In total 147 participants registered and 43 submitted solutions [@Putten]. The maximum number of policyowners that could be found was 238.  The submissions identified 95 policy owners on average.  The winning model [@Elkan] identified 121 policy owners.  Random selection results in identifying 42 policy owners. The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners. [@Putten].  In this paper we set out to complete the first part of the COIL Challenge. <!--**SUMARISE FINDINGS?**-->

# Exploratory Data Analysis

## Response Variable CARAVAN

```{r, echo=FALSE}
df %>%
  ggplot(aes(CARAVAN, fill=CARAVAN)) + 
  scale_fill_brewer(palette = "Set1") + 
  geom_bar() + 
  ylab("Count") + 
  theme(legend.position = "none")
```

The dataset is unbalanced.  There are only `r df %>% filter(CARAVAN == 1) %>% nrow()` cases that have purchased the mobile home insurance policy and `r df %>% filter(CARAVAN == 0) %>% nrow()` that have not.

## Explanatory Variables

### Categorical

```{r, echo=FALSE}
categorical_vars <- c("MOSTYPE", "MGEMLEEF", "MOSHOOFD", "MGODRK", "PWAPART")

for (var in categorical_vars){
  df$var <- df[,var]
  p <- ggplot(df) +
    geom_mosaic(aes(x = product(CARAVAN, var), fill=CARAVAN), na.rm=TRUE) +
    labs(x=var, y="CARAVAN") +
    theme(legend.position = "none") + 
    scale_fill_brewer(palette = "Set1")

  print(p)
}
```

### Numeric
```{r, echo=FALSE}
numeric_vars <- df %>%
  select(-CARAVAN, -MOSTYPE, -MGEMLEEF, -MOSHOOFD, -MGODRK, -PWAPART, -var) %>%
  names()

for (var in numeric_vars){
  df$var <- df[,var]
  p <- ggplot(df, aes(x = var, fill = CARAVAN, alpha = 0.5, color = CARAVAN)) +
    geom_density() +
    labs(x = var) +
    theme(legend.position = "none") + 
    scale_fill_brewer(palette = "Set1")
  print(p)
}

```

# Appendices

<!--
- Supplemental tables and/or figures.
-->

## Data Dictionary

```{r, results='asis', echo=FALSE}
data_dictionary <- data.frame(
  Name = c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
           "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV",
           "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD",
           "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD",
           "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD",
           "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART",
           "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM",
           "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT",
           "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM",
           "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
           "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR",
           "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG",
           "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG",
           "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
           "ABYSTAND", "CARAVAN"),
  Description = c("Customer Subtype", "Number of houses", "Avg size household",
                  "Avg age", "Customer main type", "Roman catholic",
                  "Protestant", "Other religion", "No religion", "Married",
                  "Living together", "Other relation", "Singles",
                  "Household without children", "Household with children",
                  "High level education", "Medium level education",
                  "Lower level education", "High status", "Entrepreneur",
                  "Farmer", "Middle management", "Skilled labourers",
                  "Unskilled labourers", "Social class A", "Social class B1",
                  "Social class B2", "Social class C", "Social class D",
                  "Rented house", "Home owners", "1 car", "2 cars", "No car",
                  "National Health Service", "Private health insurance",
                  "Income < 30.000", "Income 30-45.000", "Income 45-75.000",
                  "Income 75-122.000", "Income >123.000", "Average income",
                  "Purchasing power class",
                  "Contribution private third party insurance",
                  "Contribution third party insurance (firms)",
                  "Contribution third party insurane (agriculture)",
                  "Contribution car policies",
                  "Contribution delivery van policies",
                  "Contribution motorcycle/scooter policies",
                  "Contribution lorry policies",
                  "Contribution trailer policies",
                  "Contribution tractor policies",
                  "Contribution agricultural machines policies",
                  "Contribution moped policies",
                  "Contribution life insurances",
                  "Contribution private accident insurance policies",
                  "Contribution family accidents insurance policies",
                  "Contribution disability insurance policies",
                  "Contribution fire policies",
                  "Contribution surfboard policies",
                  "Contribution boat policies",
                  "Contribution bicycle policies",
                  "Contribution property insurance policies",
                  "Contribution social security insurance policies",
                  "Number of private third party insurance",
                  "Number of third party insurance (firms)",
                  "Number of third party insurane (agriculture)",
                  "Number of car policies",
                  "Number of delivery van policies",
                  "Number of motorcycle/scooter policies",
                  "Number of lorry policies",
                  "Number of trailer policies", "Number of tractor policies",
                  "Number of agricultural machines policies",
                  "Number of moped policies", "Number of life insurances",
                  "Number of private accident insurance policies",
                  "Number of family accidents insurance policies",
                  "Number of disability insurance policies",
                  "Number of fire policies", "Number of surfboard policies",
                  "Number of boat policies", "Number of bicycle policies",
                  "Number of property insurance policies",
                  "Number of social security insurance policies",
                  "Number of mobile home policy"))
pander(data_dictionary)
```

\newpage

## References

### 2019-12-03 final project/Draft.Rmd
Corrected typo

@@
                       "19" = "19 Suburban youth",
-                      "20" = "20 Etnically diverse",
+                      "20" = "20 Ethnically diverse",
                       "21" = "21 Young urban have-nots",

### 2019-12-04 hw5/HW5_Mike.Rmd
Initial Commit

---
title: "DATA 621 Homework #5"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(RColorBrewer)
library(corrplot)
library(gridExtra)
options(scipen = 999)
```

```{r data, echo=FALSE}
df <- read.csv("./data/wine-training-data.csv")
hold_out_data <- read.csv("./data/wine-evaluation-data.csv")
predictors <- names(df)[3:length(names(df))]
```

## Introduction

We have been given a dataset with information on `r nrow(df)` commercially available wines.  The variables in the dataset are mostly related to the chemical properties of the wine being sold.  A large wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.  The objective is explore, analyze and build a count regression model to predict the number of cases of wine that will be sold given certain properties of the wine.

## Data Exploration & Preparation

As previously stated we have `r nrow(df)` observations for developing the model.  The evaluation dataset (which we will set aside for now) consists of `r nrow(hold_out_data)` observations.  We will begin by taking a look at the modeling data:

```{r}
summary(df) %>%
  kable() %>%
  kable_styling()
```

Two things jump out about this dataset.  The first thing is that there are variables with negative values.  These negative values don't make sense in the context of chemical properties of wine.  After all, how can you have negative chlorides?  Second, there are quite a few records with missing values.  We will need to either impute values or drop the records with incomplete data.  These two issues will need to be corrected in the data.

### Training Test Split

Next we will look at splitting the data into a training and testing set at a standard 70-30 split between train and test:

```{r}
set.seed(42)
train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Correct Invalid Values

We will begin by correcting the invalid values.  We will assume the negative values were a data entry issue and a negative value was entered when a positive value was desired.  If these negative values were changed into positive numbers, they would be within the range of plausible values based on the other data.

```{r}
fix_negative_values <- function(df){
  df %>%
    mutate(FixedAcidity = abs(FixedAcidity),
           VolatileAcidity = abs(VolatileAcidity),
           CitricAcid = abs(CitricAcid),
           ResidualSugar = abs(ResidualSugar),
           Chlorides = abs(Chlorides),
           FreeSulfurDioxide = abs(FreeSulfurDioxide),
           TotalSulfurDioxide = abs(TotalSulfurDioxide),
           Sulphates = abs(Sulphates),
           Alcohol = abs(Alcohol))
}

train <- fix_negative_values(train)
test <- fix_negative_values(test)
hold_out_data <- fix_negative_values(hold_out_data)
```

Here is the summary of the corrected training data:

```{r}
summary(train) %>%
  kable() %>%
  kable_styling()
```

### Missing Values

```{r}
temp <- train
```

The `Alcohol`,  `ResidualSugar`, `Chlorides`, `FreeSulfurDioxide`, `TotalSulfurDioxide`, `pH`, and `Sulphates` have missing values.  We will impute using the median value.

There are `r df %>% filter(is.na(STARS)) %>% nrow()` observations without a `STARS`value populated. Looking at what is populated, the variable was imported as a numeric value, though it is used as a discrete variable. Since it isn't too much of a stretch to view rating as a numeric here (e.g. a 2 star wine is half as good as a 4 star wine), we'll keep the variable type as-is.

For the NA values, it makes sense that unrated wines would be somewhere in the middle of the pack or, possibly below it (after all, if it was such a good wine, why hasn't it come to the attention of experts to be rated?). The summary above shows a mean and median value at or close to 2, so we'll set our NAs to 2.

```{r, eval=FALSE, echo=FALSE}
set.seed(42)
knn <- train %>% knnImputation()
```

```{r}
impute_vars <- c("Alcohol",  "ResidualSugar", "Chlorides", "FreeSulfurDioxide", "TotalSulfurDioxide", "pH", "Sulphates")
fix_missing_values2 <- function(d){
  for(var in impute_vars){
    impute_me <- is.na(d[, var])
    d[impute_me, var] <- knn[impute_me, var] 
  }
  d
}

fix_missing_values <- function(df){
  df %>%
    mutate(
      ResidualSugar = ifelse(is.na(ResidualSugar), median(ResidualSugar, na.rm = T), ResidualSugar),
      Chlorides = ifelse(is.na(Chlorides), median(Chlorides, na.rm = T), Chlorides),
      FreeSulfurDioxide = ifelse(is.na(FreeSulfurDioxide), median(FreeSulfurDioxide, na.rm = T), FreeSulfurDioxide),
      TotalSulfurDioxide = ifelse(is.na(TotalSulfurDioxide), median(TotalSulfurDioxide, na.rm = T), TotalSulfurDioxide),
      pH = ifelse(is.na(pH), median(pH, na.rm = T), pH),
      Sulphates = ifelse(is.na(Sulphates), median(Sulphates, na.rm = T), Sulphates),
      Alcohol = ifelse(is.na(Alcohol), median(Alcohol, na.rm = T), Alcohol),
      STARS_imputed = ifelse(is.na(STARS), 1, 0),
      STARS = ifelse(is.na(STARS), 2, STARS)
    )
}
train <- fix_missing_values(train)
test <- fix_missing_values(test)
hold_out_data <- fix_missing_values(hold_out_data)
```

## Data Exploration

### Univariate Analysis

#### Response Variable

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = train, aes(x = factor(TARGET), fill = factor(TARGET))) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1") + 
  theme(legend.position = "none") +
  labs(x="Number of Wine Cases Purchased (TARGET)",y = "Count")
```

In order to model this as a Poisson process the mean should be equal to the variance

```{r}
train %>%
  summarise(mean = mean(TARGET), variance = var(TARGET)) %>%
  kable() %>%
  kable_styling()
```

They are not the same.  The implications of this requirement not being met is the the MLE $\beta$s are correct but the standard errors will be wrong.

#### Predictors

```{r}
dens <- lapply(predictors, FUN=function(var) {
  ggplot(train, aes_string(x = var)) + 
    geom_density(fill = "gray") +
    geom_vline(aes(xintercept = mean(train[, var])), color = "blue", size=1) +
    geom_vline(aes(xintercept = median(train[, var])), color = "red", size=1) +
    geom_vline(aes(xintercept = quantile(train[, var], 0.25)), linetype = "dashed", size = 0.5) + 
    geom_vline(aes(xintercept = quantile(train[, var], 0.75)), linetype = "dashed", size = 0.5)
  })
do.call(grid.arrange, args = c(dens, list(ncol = 3)))
```

```{r}
corrplot.mixed(cor(train))
```

### 2019-12-04 hw5/HW5_Mike.Rmd
Updated HW5 Mike

@@
 hold_out_data <- read.csv("./data/wine-evaluation-data.csv")
-predictors <- names(df)[3:length(names(df))]
+df <- df[,names(df)[2:length(names(df))]] # Drop the INDEX
+predictors <- names(df)[2:length(names(df))]
 ```
@@
 After all, how can you have negative chlorides?
-Second, there are quite a few records with missing values.
-We will need to either impute values or drop the records with incomplete data.
-These two issues will need to be corrected in the data.
 
-### Training Test Split
+Second, there are quite a few records with missing values.  We will need to either impute values or drop the records with incomplete data.  These two issues will need to be corrected in the data.
 
-Next we will look at splitting the data into a training and testing set at a standard 70-30 split between train and test:
-
-```{r}
-set.seed(42)
-train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
-train <- df[train_index,]
-test <- df[-train_index,]
-```
-
-### Correct Invalid Values
+### Correcting Invalid Values
 
@@
 }
-
-train <- fix_negative_values(train)
-test <- fix_negative_values(test)
+df <- fix_negative_values(df)
 hold_out_data <- fix_negative_values(hold_out_data)
@@
 
-Here is the summary of the corrected training data:
+### Missing Values
+
+The instructions indicate that sometimes, the fact that a variable is missing is actually predictive of the target.  Let's see what we can learn about `TARGET` from the missing values.
 
 ```{r}
-summary(train) %>%
+temp <- df %>%
+  select(TARGET, STARS, Alcohol, ResidualSugar, Chlorides, FreeSulfurDioxide, TotalSulfurDioxide, pH, Sulphates) %>%
+  gather("variable", "value", -TARGET) %>%
+  mutate(na = ifelse(is.na(value), "Yes", "No")) %>%
+  group_by(TARGET, na, variable) %>%
+  tally() %>%
+  ungroup()
+
+ temp <- temp %>%
+    group_by(variable, na) %>%
+    summarise(total = sum(n)) %>%
+    merge(temp) %>%
+    mutate(share = n / total)
+
+ ggplot(temp, aes(TARGET, share, fill = na)) +
+   geom_bar(stat = "identity", position = "dodge") +
+   scale_fill_brewer(palette = "Set1") +
+   facet_wrap(~variable, ncol = 4) + 
+   ylab("Percent of Group")
+```
+
+We learn that the observations that are missing `STARS` are much more likely to have a zero `TARGET`.  So let's look at how many stars wine with a zero `TARGET` have:
+
+```{r}
+df %>%
+  filter(TARGET == 0) %>%
+  select(STARS) %>%
+  na.omit() %>%
+  group_by(STARS) %>%
+  tally() %>%
   kable() %>%
@@
 
-### Missing Values
+Based on this it looks like we should assign 1 for all the missing `STARS`.
+
+There is little to no information encoded in the NAs for `Alcohol`,  `ResidualSugar`, `Chlorides`, `FreeSulfurDioxide`, `TotalSulfurDioxide`, `pH`, and `Sulphates`.
+We will impute using the median value.
 
 ```{r}
-temp <- train
-```
-
-The `Alcohol`,  `ResidualSugar`, `Chlorides`, `FreeSulfurDioxide`, `TotalSulfurDioxide`, `pH`, and `Sulphates` have missing values.  We will impute using the median value.
-
-There are `r df %>% filter(is.na(STARS)) %>% nrow()` observations without a `STARS`value populated.
-Looking at what is populated, the variable was imported as a numeric value, though it is used as a discrete variable.
-Since it isn't too much of a stretch to view rating as a numeric here (e.g.
-a 2 star wine is half as good as a 4 star wine), we'll keep the variable type as-is.
-
-For the NA values, it makes sense that unrated wines would be somewhere in the middle of the pack or, possibly below it (after all, if it was such a good wine, why hasn't it come to the attention of experts to be rated?).
-The summary above shows a mean and median value at or close to 2, so we'll set our NAs to 2.
-
-```{r, eval=FALSE, echo=FALSE}
-set.seed(42)
-knn <- train %>% knnImputation()
-```
-
-```{r}
-impute_vars <- c("Alcohol",  "ResidualSugar", "Chlorides", "FreeSulfurDioxide", "TotalSulfurDioxide", "pH", "Sulphates")
-fix_missing_values2 <- function(d){
-  for(var in impute_vars){
-    impute_me <- is.na(d[, var])
-    d[impute_me, var] <- knn[impute_me, var] 
-  }
-  d
-}
-
 fix_missing_values <- function(df){
@@
       STARS_imputed = ifelse(is.na(STARS), 1, 0),
-      STARS = ifelse(is.na(STARS), 2, STARS)
+      STARS = ifelse(is.na(STARS), 1, STARS)
     )
 }
-train <- fix_missing_values(train)
-test <- fix_missing_values(test)
+df <- fix_missing_values(df)
 hold_out_data <- fix_missing_values(hold_out_data)
@@
 
-## Data Exploration
+### Training Test Split
+
+Next we will look at splitting the data into a training and testing set at a standard 70-30 split between train and test:
+
+```{r}
+set.seed(42)
+train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
+train <- df[train_index,]
+test <- df[-train_index,]
+```
 
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
-ggplot(data = train, aes(x = factor(TARGET), fill = factor(TARGET))) +
-  geom_bar() +
+train %>%
+  group_by(TARGET) %>%
+  tally() %>%
+  ggplot(., aes(x = factor(TARGET), y = n, fill = factor(TARGET))) +
+  geom_bar(stat = "identity", position = "dodge") +
   scale_fill_brewer(palette = "Set1") + 
@@
 
-They are not the same.  The implications of this requirement not being met is the the MLE $\beta$s are correct but the standard errors will be wrong.
+They are not the same.  The implications of this requirement not being met is the the $\beta$s are correct but the standard errors will be wrong.
 
@@
 dens <- lapply(predictors, FUN=function(var) {
-  ggplot(train, aes_string(x = var)) + 
+  ggplot(df, aes_string(x = var)) + 
     geom_density(fill = "gray") +
@@
 
+Some of the predictors (the chemical composition variables) are right skewed.
+
+### Bivariate Analysis
+
+Now let's examine the correlations between the variables:
+
 ```{r}
@@
 ```
+
+Hmm.
+This suggests that the chemical makeup of the wine has little bearing on the number of cases sold.
+The label appeal, acidity index value and star rating have the strongest correlation.
+This also suggests that it may be worth the effort to do a better job filling in the missing values in the `STARS` variable.
+
+## Model Building & Evaluation
+
+We will begin by first constructing a count regression model based on the chemical properties of the wine.
+So we will be excluding the label, star rating and the acid index as it is a weighted average of the acidity variables.
+Based on our bivariate analysis we would not expect this model to preform well.
+Since the mean is not equal to the variance we will be using the quasi-Poisson model.
+
+```{r}
+train1 <- train %>%
+  select(-LabelAppeal, -AcidIndex, -STARS, -STARS_imputed)
+model1 <- glm(TARGET ~ ., family = quasipoisson, train1)
+summary(model1)
+```

### 2019-12-05 hw5/HW5_Mike.Rmd
More work

@@
 
+### Setting Up Evaluation
+
+In order to assess the models we will assume that a bottle of wine whosale costs $15.00 and there are 10 bottles in a case.
+We will assume that the `TARGET` is the maximum amount that will sell in the market.
+Any predictions below the `TARGET` represent potential revenue that is unrealized.
+Any predictions over the `TARGET` will not sell and represent a loss of $150.
+We will evaluate the model to see if it helps the manufacturers goal of maximizing revenue.
+
+```{r}
+evaluate_model <- function(model, test_df, yhat = FALSE){
+  price_per_bottle <- 15.00
+  bottles_per_case <- 10
+  case_price <- price_per_bottle * bottles_per_case
+  temp <- data.frame(yhat=c(0:8), TARGET = c(0:8), n=c(0))
+
+  if(yhat){
+    test_df$yhat <- yhat
+  } else {
+    test_df$yhat <- round(predict.glm(model, newdata=test_df, type="response"), 0)
+  }
+
+  test_df <- test_df %>%
+    group_by(yhat, TARGET) %>%
+    tally() %>%
+    mutate(accuracy = ifelse(yhat > TARGET, "Over", ifelse(yhat < TARGET, "Under", "Accurate"))) %>%
+    mutate(off_by = abs(yhat - TARGET)) %>%
+    mutate(revenue = ifelse(off_by == 0, 1, -1) * n * case_price)
+
+  rev_df <- test_df %>%
+    group_by(accuracy) %>%
+    summarise(revenue = sum(revenue)) %>%
+    spread(accuracy, revenue)
+
+  total_rev <- sum(test_df$revenue)
+
+  results <- test_df %>%
+    group_by(accuracy) %>%
+    summarise(n = sum(n))
+
+  confusion_matrix <- test_df %>%
+    bind_rows(temp) %>%
+    group_by(yhat, TARGET) %>%
+    summarise(n = sum(n)) %>%
+    spread(TARGET, n, fill = 0)
+
+  return(list("total_rev" = total_rev, "rev_df" = rev_df, "confusion_matrix" = confusion_matrix, "results" = results))
+}
+```
+
+### Chemical Property Model
+
 We will begin by first constructing a count regression model based on the chemical properties of the wine.
@@
 summary(model1)
+model1_results <- evaluate_model(model1, test)
 ```
+
+The following is a confusion matrix like examination of the model with the predicted cases going down the rows and the actual cases going across the columns.
+
+```{r}
+model1_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+The model is only accurate `r temp <- model1_results$results %>% filter(accuracy == "Accurate"); temp$n` times.
+It overestimates `r temp <- model1_results$results %>% filter(accuracy == "Over"); temp$n` times and underestimates `r temp <- model1_results$results %>% filter(accuracy == "Under"); temp$n` times.
+The following shows the revenue gained or lost due to the model's predictions:
+
+```{r}
+model1_results$rev_df %>%
+  kable() %>%
+  kable_styling()
+```
+
+This model would result in $`r model1_results$total_rev` in revenue.  The company would loose a lot of money if the model was implemented.
+
+### Second Chemical Property Model
+
+We will try one more chemical property model but only select the variables that were statistically significant in the preious model (`FixedAcidity`, `VolatileAcidity` and `Alcohol`).
+
+```{r}
+model2 <- glm(TARGET ~ FixedAcidity + VolatileAcidity + Alcohol, family = quasipoisson, train)
+summary(model2)
+model2_results <- evaluate_model(model2, test)
+```
+
+This model would generate $`r model2_results$total_rev` in revenue.
+This is a slight improvement but is still a horible model.
+Building a model based off of the chemical properties does not seem like a good solution.
+Let's turn our attention to the other variables with stronger correlations.
+
+### Preceived Quality Model
+
+This model relies on the consumer preception of quality based on the label, and the star ratings of the experts.
+
+```{r}
+model3 <- glm(TARGET ~ LabelAppeal + STARS, family = quasipoisson, train)
+summary(model3)
+model3_results <- evaluate_model(model3, test)
+```
+
+Here is a confusion matrix for this model:
+
+```{r}
+model3_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+The model is only accurate `r temp <- model3_results$results %>% filter(accuracy == "Accurate"); temp$n` times.
+It overestimates `r temp <- model3_results$results %>% filter(accuracy == "Over"); temp$n` times and underestimates `r temp <- model3_results$results %>% filter(accuracy == "Under"); temp$n` times.
+This is an improvement from just the chemical property models, however there is still a lot of error.
+The following shows the revenue gained or lost due to the model's predictions:
+
+```{r}
+model3_results$rev_df %>%
+  kable() %>%
+  kable_styling()
+```
+
+This model would result in $`r model3_results$total_rev` in revenue.
+
+### Perceived Quality Plus Model
+
+We will begin with the preceeding model but add in the Acid Index and the flag if the STARS was imputed.
+
+```{r}
+model4 <- glm(TARGET ~ LabelAppeal + STARS + AcidIndex + STARS_imputed, family = quasipoisson, train)
+summary(model4)
+model4_results <- evaluate_model(model4, test)
+```
+
+Let's see how the predictions line up with the reality:
+
+```{r}
+model4_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+The model is accurate `r temp <- model4_results$results %>% filter(accuracy == "Accurate"); temp$n` times.
+It overestimates `r temp <- model4_results$results %>% filter(accuracy == "Over"); temp$n` times and underestimates `r temp <- model4_results$results %>% filter(accuracy == "Under"); temp$n` times.
+The following shows the revenue gained or lost due to the model's predictions:
+
+```{r}
+model4_results$rev_df %>%
+  kable() %>%
+  kable_styling()
+```
+
+This model would result in $`r model4_results$total_rev` in revenue.  It is the best model so far but it still would loose a lot of money because it overestimates the number of cases to be sold.
+
+There are no zero predictions in this model which is not accurate.  We will try to address this in our last model
+
+### Adjusted Perceived Quality Plus Model
+
+We will start with the predictions from preceived quality plus model.  We will then manually override the the predictions that have the stars filled in with a predicted zero.
+
+```{r}
+temp <- test
+temp$yhat <- round(predict.glm(model4, newdata=temp, type="response"), 0)
+temp <- temp %>%
+  mutate(yhat = ifelse(STARS_imputed == 1, 0, yhat))
+model5_results <- evaluate_model(model4, test, temp$yhat)
+```
+
+```{r}
+model5_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+```{r}
+model5_results$rev_df %>%
+  kable() %>%
+  kable_styling()
+```

### 2019-12-05 hw5/HW5_Mike.Rmd
More work on Mike's HW5

--- 2019-12-04 hw5/HW5_Mike.Rmd
@@
 
+### Evaluation Approach
+
+In order to assess the models we will return the the original intent of the exercise: If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.
+We will evaluate the models based on the number of cases that are likely to be sold.
+We will assume that when the model overestimates the cases sold, these cases will never be sold and represent a loss.
+When the model underestimates there is also a loss but it is unseen.
+It is potential sales that are not realized.
+
+We will summarize the evaluation by first constructing a confusion matrix like summary looking at what our model predicts vs what actually occured.
+We will also summarize the total number of cases that would be sold, the net cases sold (after adjusting for the glut due to overestimating), and a final adjusted net which factors in "losses" from under estimation.
+
+```{r}
+
+evaluate_model <- function(model, test_df, yhat = FALSE){
+  temp <- data.frame(yhat=c(0:8), TARGET = c(0:8), n=c(0))
+
+  if(yhat){
+    test_df$yhat <- yhat
+  } else {
+    test_df$yhat <- round(predict.glm(model, newdata=test_df, type="response"), 0)
+  }
+
+  test_df <- test_df %>%
+    group_by(yhat, TARGET) %>%
+    tally() %>%
+    mutate(accuracy = ifelse(yhat > TARGET, "Over", ifelse(yhat < TARGET, "Under", "Accurate"))) %>%
+    mutate(cases_sold = ifelse(yhat > TARGET, TARGET, yhat) * n,
+           glut = ifelse(yhat > TARGET, yhat - TARGET, 0) * n,
+           missed_opportunity = ifelse(yhat < TARGET, TARGET - yhat, 0) * n) %>%
+    mutate(net_cases_sold = cases_sold - glut,
+           adj_net_cases_sold = cases_sold - glut - missed_opportunity)
+
+  results <- test_df %>%
+    group_by(accuracy) %>%
+    summarise(n = sum(n)) %>%
+    spread(accuracy, n)
+
+  accurate <- results$Accurate
+  over <- results$Over
+  under <- results$Under
+
+  cases_sold <- sum(test_df$cases_sold)
+  net_cases_sold <- sum(test_df$net_cases_sold)
+  adj_net_cases_sold <- sum(test_df$adj_net_cases_sold)
+  missed_opportunity <- sum(test_df$missed_opportunity)
+  glut <- sum(test_df$glut)
+
+  confusion_matrix <- test_df %>%
+    bind_rows(temp) %>%
+    group_by(yhat, TARGET) %>%
+    summarise(n = sum(n)) %>%
+    spread(TARGET, n, fill = 0)
+
+  return(list("confusion_matrix" = confusion_matrix, "results" = results, "df" = test_df, "accurate" = accurate, "over" = over, "under" = under, "cases_sold" = cases_sold, "net_cases_sold" = net_cases_sold, "adj_net_cases_sold" = adj_net_cases_sold, "glut" = glut, "missed_opportunity" = missed_opportunity))
+}
+```
+
+### Chemical Property Model
+
 We will begin by first constructing a count regression model based on the chemical properties of the wine.
@@
 summary(model1)
+model1_results <- evaluate_model(model1, test)
 ```
+
+The following is a confusion matrix like examination of the model with the predicted cases going down the rows and the actual cases going across the columns.
+
+```{r}
+model1_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+The model is only accurate `r model1_results$accurate` times.
+It overestimates `r model1_results$over` times and underestimates `r model1_results$under` times.
+If decisions were based off of this model, one would expect to sell `r model1_results$cases_sold`.
+If we subtract out the extra cases due to overestimation we would have `r model1_results$net_cases_sold` net cases sold.
+If we adjust this for the cases that could have been sold if the model didn’t underestimate we would have `r model1_results$adj_net_cases_sold` cases sold.
+
+### Second Chemical Property Model
+
+We will try one more chemical property model but only select the variables that were statistically significant in the preious model (`FixedAcidity`, `VolatileAcidity` and `Alcohol`).
+
+```{r}
+model2 <- glm(TARGET ~ FixedAcidity + VolatileAcidity + Alcohol, family = quasipoisson, train)
+summary(model2)
+model2_results <- evaluate_model(model2, test)
+```
+
+We would expect to have `r model2_results$cases_sold` cases sold using this model (with a net of `r model2_results$net_cases_sold` and adjusted net of `r model2_results$adj_net_cases_sold`).
+Looking at the confusion matrix we see this model is accurate only `r model2_results$accurate` out of `r nrow(test)` times.
+
+```{r}
+model2_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+The previous model is slightly better but is still a horible model.
+Building a model based off of the chemical properties does not seem like a good solution.
+Let's turn our attention to the other variables with stronger correlations.
+
+### Preceived Quality Model
+
+This model relies on the consumer preception of quality based on the label, and the star ratings of the experts.
+
+```{r}
+model3 <- glm(TARGET ~ LabelAppeal + STARS, family = quasipoisson, train)
+summary(model3)
+model3_results <- evaluate_model(model3, test)
+```
+
+Here is a confusion matrix for this model:
+
+```{r}
+model3_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+### Perceived Quality Plus Model
+
+We will begin with the preceeding model but add in the Acid Index and the flag if the STARS was imputed.
+
+```{r}
+model4 <- glm(TARGET ~ LabelAppeal + STARS + AcidIndex + STARS_imputed, family = quasipoisson, train)
+summary(model4)
+model4_results <- evaluate_model(model4, test)
+```
+
+Let's see how the predictions line up with the reality:
+
+```{r}
+model4_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+We would expect to have `r model3_results$cases_sold` net cases sold using this model (with a net of `r model3_results$net_cases_sold` and adjusted net of `r model3_results$adj_net_cases_sold`).
+Looking at the confusion matrix we see this model is accurate only `r model3_results$accurate` out of `r nrow(test)` times.
+
+There are no zero predictions in this model which is not accurate.  We will try to address this in our last model
+
+### Adjusted Perceived Quality Plus Model
+
+We will start with the predictions from preceived quality plus model.  We will then manually override the the predictions that have the stars filled in with a predicted zero.
+
+```{r}
+temp <- test
+temp$yhat <- round(predict.glm(model4, newdata=temp, type="response"), 0)
+temp <- temp %>%
+  mutate(yhat = ifelse(STARS_imputed == 1, 0, yhat))
+model5_results <- evaluate_model(model4, test, temp$yhat)
+```
+
+```{r}
+model5_results$confusion_matrix %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+We see that applying the adjustment based on if the stars are imputed increased the correct number of cases where it should be predicted as a zero.  It also introduced some underestimation.
+
+We would expect to have `r model4_results$cases_sold` net cases sold using this model (with a net of `r model4_results$net_cases_sold` and adjusted net of `r model4_results$adj_net_cases_sold`).
+Looking at the confusion matrix we see this model is accurate only `r model4_results$accurate` out of `r nrow(test)` times.
+
+## Model Selection
+
+The following is a summary of the five models
+
+<table>
+<thead>
+  <tr>
+    <th>Model</th>
+    <th>Number of Cases Sold</th>
+    <th>Number of Cases Overestimated to Sell</th>
+    <th>Missed Opportunities</th>
+    <th>Times Accurate</th>
+  </tr>
+</thead>
+<tbody>
+  <tr>
+    <td>Chemical Property Model 1</td>
+    <td>`r model1_results$cases_sold`</td>
+    <td>`r model1_results$glut`</td>
+    <td>`r model1_results$missed_opportunity`</td>
+    <td>`r model1_results$accurate`</td>
+  </tr>
+  <tr>
+    <td>Chemical Property Model 2</td>
+    <td>`r model2_results$cases_sold`</td>
+    <td>`r model2_results$glut`</td>
+    <td>`r model2_results$missed_opportunity`</td>
+    <td>`r model2_results$accurate`</td>
+  </tr>
+  <tr>
+    <td>Perceived Quality Model</td>
+    <td>`r model3_results$cases_sold`</td>
+    <td>`r model3_results$glut`</td>
+    <td>`r model3_results$missed_opportunity`</td>
+    <td>`r model3_results$accurate`</td>
+  </tr>
+  <tr>
+    <td>Perceived Quality Plus Model</td>
+    <td>`r model4_results$cases_sold`</td>
+    <td>`r model4_results$glut`</td>
+    <td>`r model4_results$missed_opportunity`</td>
+    <td>`r model4_results$accurate`</td>
+  </tr>
+  <tr>
+    <td>Adjusted Perceived Quality Plus Model</td>
+    <td>`r model5_results$cases_sold`</td>
+    <td>`r model5_results$glut`</td>
+    <td>`r model5_results$missed_opportunity`</td>
+    <td>`r model5_results$accurate`</td>
+  </tr>
+</tbody>
+</table>
+
+The `Perceived Quality Plus Model` has the most cases sold out of the models.
+However the `Adjusted Perceived Quality Plus Model` has the highest accuracy.
+One of the reasons why the `Perceived Quality Plus Model` predicted more cases sold is because we forced more zero predictions in the `Adjusted Perceived Quality Plus Model`.
+
+If we subtract out the waste from overestimation the `Adjusted Perceived Quality Plus Model` is the best model.
+This has real business consequences as having extra cases of wine on hand that don't sell becomes a real cost to a business.
+For these two reasons it will be the model of choice in producing estimates for the hold out datase.
+
+```{r}
+hold_out_data$TARGET <- round(predict.glm(model4, newdata=hold_out_data, type="response"), 0)
+hold_out_data <- hold_out_data %>%
+  mutate(TARGET = ifelse(STARS_imputed == 1, 0, TARGET))
+```

### 2019-12-06 final project/Draft.Rmd
More work on draft

@@
 library(ggmosaic)
+library(wesanderson)
 # Thank you Stack Overflow!
@@
 }
+df <- df %>% select(-var)
 ```
@@
 ### Numeric
+
 ```{r, echo=FALSE}
 numeric_vars <- df %>%
-  select(-CARAVAN, -MOSTYPE, -MGEMLEEF, -MOSHOOFD, -MGODRK, -PWAPART, -var) %>%
+  select(-CARAVAN, -MOSTYPE, -MGEMLEEF, -MOSHOOFD, -MGODRK, -PWAPART) %>%
   names()
@@
 
+df <- df %>% select(-var)
+```
+
+Here is a correlation heatmap.  Although you can't see the variable names, you get a general sense of how they are correlated:
+
+```{r, echo=FALSE}
+pal <- wes_palette("Zissou1", 100, type = "continuous")
+
+correlation_data <- df %>%
+  select(-MOSTYPE, -MGEMLEEF, -MOSHOOFD, -MGODRK, -PWAPART) %>%
+  mutate(CARAVAN = as.numeric(as.character(CARAVAN)))
+
+correlation_matrix <- round(cor(correlation_data),2) %>%
+  as.data.frame() %>%
+  rownames_to_column("Var1") %>%
+  gather(Var2, Correlation, -Var1) %>%
+  arrange(Var1, Var2)
+
+correlation_matrix$Var2 <- factor(correlation_matrix$Var2,levels=rev(unique(correlation_matrix$Var2)))
+
+ggplot(correlation_matrix, aes(x=Var1, y=Var2, fill=Correlation)) + 
+  geom_tile() +
+  scale_fill_gradientn(colours = pal) +
+   theme(axis.title.x=element_blank(),
+         axis.text.x=element_blank(),
+         axis.ticks.x=element_blank(),
+         axis.title.y=element_blank(),
+         axis.text.y=element_blank(),
+         axis.ticks.y=element_blank())
+```
+
+Most of the correlation heatmap is yellow which is has a correlation coeffient of zero.  Here's a density plot of the correlation coefficients:
+
+```{r, echo=FALSE}
+ggplot(correlation_matrix, aes(Correlation)) +
+  geom_density(fill="red", color="red")
 ```
@@
                   "Number of mobile home policy"))
-pander(data_dictionary)
+
+data_dictionary %>%
+  arrange(Name) %>%
+  pander()
 ```

### 2019-12-06 final project/Manuscript.Rmd
Minor changes

@@
 # tinytex::install_tinytex()
-if(!"papaja" %in% rownames(installed.packages())) devtools::install_github("crsh/papaja")
+if (!"papaja" %in% rownames(installed.packages())) {
+  devtools::install_github("crsh/papaja")
+}
 library("papaja")
@@
 # Output hooks handle normal R console output.
-knitr::knit_hooks$set( output = function(x, options) {
+knitr::knit_hooks$set(output = function(x, options) {
 
   comment <- knitr::opts_current$get("comment")
-  if( is.na(comment) ) comment <- ""
-  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+  if (is.na(comment)) comment <- ""
+  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                      x, perl = TRUE)
-  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
-  if( can_null && do_null ) {
+  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
+  if (can_null && do_null) {
     # By default R print output aligns at the right brace.
-    align_index <- regexpr( "\\]", x )[1] - 1
+    align_index <- regexpr("\\]", x)[1] - 1
     # Two cases: start or newline
-    re <- paste0( "^.{", align_index, "}\\]")
+    re <- paste0("^.{", align_index, "}\\]")
     rep <- comment
-    x <- gsub( re, rep,  x )
-    re <- paste0( "\\\n.{", align_index, "}\\]")
-    rep <- paste0( "\n", comment )
-    x <- gsub( re, rep,  x )
+    x <- gsub(re, rep,  x)
+    re <- paste0("\\\n.{", align_index, "}\\]")
+    rep <- paste0("\n", comment)
+    x <- gsub(re, rep,  x)
   }
 
-  default_output_hook( x, options )
+  default_output_hook(x, options)
 
 })
-knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
+knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
 ```
@@
 ```{r data, echo=FALSE}
-df <- read.delim("data/ticdata2000.txt", header = FALSE)
+prepare_data <- function(df){
   names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
@@
                "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
+
+  MOSTYPE_labels <- c("1" = "1 High Income, expensive child",
+                      "2" = "2 Very Important Provincials",
+                      "3" = "3 High status seniors",
+                      "4" = "4 Affluent senior apartments",
+                      "5" = "5 Mixed seniors",
+                      "6" = "6 Career and childcare",
+                      "7" = "7 Dinki's (double income no kids)",
+                      "8" = "8 Middle class families",
+                      "9" = "9 Modern, complete families",
+                      "10" = "10 Stable family",
+                      "11" = "11 Family starters",
+                      "12" = "12 Affluent young families",
+                      "13" = "13 Young all american family",
+                      "14" = "14 Junior cosmopolitan",
+                      "15" = "15 Senior cosmopolitans",
+                      "16" = "16 Students in apartments",
+                      "17" = "17 Fresh masters in the city",
+                      "18" = "18 Single youth",
+                      "19" = "19 Suburban youth",
+                      "20" = "20 Ethnically diverse",
+                      "21" = "21 Young urban have-nots",
+                      "22" = "22 Mixed apartment dwellers",
+                      "23" = "23 Young and rising",
+                      "24" = "24 Young, low educated", 
+                      "25" = "25 Young seniors in the city",
+                      "26" = "26 Own home elderly",
+                      "27" = "27 Seniors in apartments",
+                      "28" = "28 Residential elderly",
+                      "29" = "29 Porchless seniors: no front yard",
+                      "30" = "30 Religious elderly singles",
+                      "31" = "31 Low income catholics",
+                      "32" = "32 Mixed seniors",
+                      "33" = "33 Lower class large families",
+                      "34" = "34 Large family, employed child",
+                      "35" = "35 Village families",
+                      "36" = "36 Couples with teens 'Married with children'",
+                      "37" = "37 Mixed small town dwellers",
+                      "38" = "38 Traditional families",
+                      "39" = "39 Large religous families",
+                      "40" = "40 Large family farms",
+                      "41" = "41 Mixed rurals")
+
+MGEMLEEF_labels <- c("1" = "20-30 years",
+                     "2" = "30-40 years",
+                     "3" = "40-50 years",
+                     "4" = "50-60 years",
+                     "5" = "60-70 years",
+                     "6" = "70-80 years")
+
+MOSHOOFD_labels <- c("1" = "Successful hedonists",
+                     "2" = "Driven Growers",
+                     "3" = "Average Family",
+                     "4" = "Career Loners",
+                     "5" = "Living well",
+                     "6" = "Cruising Seniors",
+                     "7" = "Retired and Religeous",
+                     "8" = "Family with grown ups",
+                     "9" = "Conservative families",
+                     "10" = "Farmers")
+
+MGODRK_labels <- c("0" = "0%",
+                   "1" = "1 - 10%",
+                   "2" = "11 - 23%",
+                   "3" = "24 - 36%",
+                   "4" = "37 - 49%",
+                   "5" = "50 - 62%",
+                   "6" = "63 - 75%",
+                   "7" = "76 - 88%",
+                   "8" = "89 - 99%",
+                   "9" = "100%")
+
+PWAPART_labels <- c("0" = "f 0",
+                    "1" = "f 1 – 49",
+                    "2" = "f 50 – 99",
+                    "3" = "f 100 – 199",
+                    "4" = "f 200 – 499",
+                    "5" = "f 500 – 999",
+                    "6" = "f 1000 – 4999",
+                    "7" = "f 5000 – 9999",
+                    "8" = "f 10.000 - 19.999",
+                    "9" = "f 20.000 - ?")
+
+df <- df %>%
+  mutate(MOSTYPE = as.factor(MOSTYPE),
+         MGEMLEEF = as.factor(MGEMLEEF),
+         MOSHOOFD = as.factor(MOSHOOFD),
+         MGODRK = as.factor(MGODRK),
+         PWAPART = as.factor(PWAPART),
+         CARAVAN = as.factor(CARAVAN)) %>%
+  mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
+         MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
+         MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
+         MGODRK = recode(MGODRK, !!!MGODRK_labels),
+         PWAPART = recode(PWAPART, !!!PWAPART_labels))
+}
+
 eval <- read.delim("data/ticeval2000.txt", header = FALSE)
@@
 eval$CARAVAN <- temp$V1
-names(eval) <- names(df)
+eval <- prepare_data(eval)
+df <- prepare_data(read.delim("data/ticdata2000.txt", header = FALSE))
 ```
@@
 # Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
-prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121),
-                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
-ggplot(data=prediction_scores, aes(x=score, y=n)) +
-  geom_bar(stat="identity") +
+prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83,
+                                          86, 91, 94, 95, 96, 97, 99, 101, 103,
+                                          104, 105, 106, 107, 108, 109, 110,
+                                          111, 112, 115, 121),
+                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3,
+                                      1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1,
+                                      1))
+ggplot(data = prediction_scores, aes(x = score, y = n)) +
+  geom_bar(stat = "identity") +
   xlab("Number of Correct Predictions") +
@@
 [@Putten].
-
-In this paper we set out to complete the first part of the COIL Challenge. <!--**SUMARISE FINDINGS?**-->
+In this paper we set out to complete the first part of the COIL Challenge.
+<!--**SUMARISE FINDINGS?**-->
 
@@
 ```{r, echo=FALSE}
-doornik <- function(df){
+doornik <- function(df) {
   df <- df %>%
@@
 data_dictionary <- data.frame(
-  Name=c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN"),
-  Description = c("Customer Subtype", "Number of houses", "Avg size household", "Avg age", "Customer main type", "Roman catholic", "Protestant", "Other religion", "No religion", "Married", "Living together", "Other relation", "Singles", "Household without children", "Household with children", "High level education", "Medium level education", "Lower level education", "High status", "Entrepreneur", "Farmer", "Middle management", "Skilled labourers", "Unskilled labourers", "Social class A", "Social class B1", "Social class B2", "Social class C", "Social class D", "Rented house", "Home owners", "1 car", "2 cars", "No car", "National Health Service", "Private health insurance", "Income < 30.000", "Income 30-45.000", "Income 45-75.000", "Income 75-122.000", "Income >123.000", "Average income", "Purchasing power class", "Contribution private third party insurance", "Contribution third party insurance (firms)", "Contribution third party insurane (agriculture)", "Contribution car policies", "Contribution delivery van policies", "Contribution motorcycle/scooter policies", "Contribution lorry policies", "Contribution trailer policies", "Contribution tractor policies", "Contribution agricultural machines policies", "Contribution moped policies", "Contribution life insurances", "Contribution private accident insurance policies", "Contribution family accidents insurance policies", "Contribution disability insurance policies", "Contribution fire policies", "Contribution surfboard policies", "Contribution boat policies", "Contribution bicycle policies", "Contribution property insurance policies", "Contribution social security insurance policies", "Number of private third party insurance", "Number of third party insurance (firms)", "Number of third party insurane (agriculture)", "Number of car policies", "Number of delivery van policies", "Number of motorcycle/scooter policies", "Number of lorry policies", "Number of trailer policies", "Number of tractor policies", "Number of agricultural machines policies", "Number of moped policies", "Number of life insurances", "Number of private accident insurance policies", "Number of family accidents insurance policies", "Number of disability insurance policies", "Number of fire policies", "Number of surfboard policies", "Number of boat policies", "Number of bicycle policies", "Number of property insurance policies", "Number of social security insurance policies", "Number of mobile home policy"))
+  Name = c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
+           "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV",
+           "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD",
+           "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD",
+           "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD",
+           "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART",
+           "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM",
+           "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT",
+           "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM",
+           "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
+           "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR",
+           "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG",
+           "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG",
+           "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
+           "ABYSTAND", "CARAVAN"),
+  Description = c("Customer Subtype", "Number of houses", "Avg size household",
+                  "Avg age", "Customer main type", "Roman catholic",
+                  "Protestant", "Other religion", "No religion", "Married",
+                  "Living together", "Other relation", "Singles",
+                  "Household without children", "Household with children",
+                  "High level education", "Medium level education",
+                  "Lower level education", "High status", "Entrepreneur",
+                  "Farmer", "Middle management", "Skilled labourers",
+                  "Unskilled labourers", "Social class A", "Social class B1",
+                  "Social class B2", "Social class C", "Social class D",
+                  "Rented house", "Home owners", "1 car", "2 cars", "No car",
+                  "National Health Service", "Private health insurance",
+                  "Income < 30.000", "Income 30-45.000", "Income 45-75.000",
+                  "Income 75-122.000", "Income >123.000", "Average income",
+                  "Purchasing power class",
+                  "Contribution private third party insurance",
+                  "Contribution third party insurance (firms)",
+                  "Contribution third party insurane (agriculture)",
+                  "Contribution car policies",
+                  "Contribution delivery van policies",
+                  "Contribution motorcycle/scooter policies",
+                  "Contribution lorry policies",
+                  "Contribution trailer policies",
+                  "Contribution tractor policies",
+                  "Contribution agricultural machines policies",
+                  "Contribution moped policies",
+                  "Contribution life insurances",
+                  "Contribution private accident insurance policies",
+                  "Contribution family accidents insurance policies",
+                  "Contribution disability insurance policies",
+                  "Contribution fire policies",
+                  "Contribution surfboard policies",
+                  "Contribution boat policies",
+                  "Contribution bicycle policies",
+                  "Contribution property insurance policies",
+                  "Contribution social security insurance policies",
+                  "Number of private third party insurance",
+                  "Number of third party insurance (firms)",
+                  "Number of third party insurane (agriculture)",
+                  "Number of car policies",
+                  "Number of delivery van policies",
+                  "Number of motorcycle/scooter policies",
+                  "Number of lorry policies",
+                  "Number of trailer policies", "Number of tractor policies",
+                  "Number of agricultural machines policies",
+                  "Number of moped policies", "Number of life insurances",
+                  "Number of private accident insurance policies",
+                  "Number of family accidents insurance policies",
+                  "Number of disability insurance policies",
+                  "Number of fire policies", "Number of surfboard policies",
+                  "Number of boat policies", "Number of bicycle policies",
+                  "Number of property insurance policies",
+                  "Number of social security insurance policies",
+                  "Number of mobile home policy"))
 pander(data_dictionary)
@@
 ```{r, null_prefix = TRUE, comment=NA, echo=FALSE}
-conn <- file("CoIL.r", open="r")
+conn <- file("CoIL.r", open = "r")
 lines <- readLines(conn)
-for (i in 1:length(lines)){
+n_lines <- length(lines)
+for (i in 1:n_lines) {
   cat(lines[i], "\n")

### 2019-12-08 hw5/HW5_Corey.RMD

--- 2019-11-30 hw5/HW5_Adam.Rmd
@@
 Now that we have our testing set, we will use decision trees to help us decide which variables to potentially use in our linear model.
+The first five trees look at number of cases and the last five trees look at whether or not any cases were purchased
 
@@
 
-RF_with_Nulls(train,test,"TARGET",.5,10,5,.005,4,1)
+train$TARGETBOOL <- ifelse(train$TARGET > 0,1,0)
+test$TARGETBOOL <- ifelse(test$TARGET>0,1,0)
+
+RF_with_Nulls(train,test,"TARGET",.5,5,5,.005,4,1)
+train %>% select(-TARGET) %>% RF_with_Nulls(test,"TARGETBOOL",.5,5,5,.005,4,1)
 
@@
 
+### Try the Zero Inflated Model
+
+```{r}
+library(pscl)
+
+model3 <- zeroinfl(TARGET ~ LabelAppeal + STARS + AcidIndex + VolatileAcidity +
+                TotalSulfurDioxide + FreeSulfurDioxide + Chlorides+CitricAcid+Sulphates+ResidualSugar+pH|LabelAppeal + STARS + AcidIndex + VolatileAcidity +TotalSulfurDioxide + FreeSulfurDioxide + Chlorides+CitricAcid+Sulphates+ResidualSugar+pH, data = train, dist = "poisson")
+
+summary(model3)
+
+test$Prediction3 <- predict(model3, newdata=test, type="response")
+
+cor(test$Prediction3,test$TARGET)
+
+```
+
 ### Diagnostics
@@
 
+# Model 3
+residModel3 <- data.frame(obs = test$TARGET, diff = (test$TARGET - test$Prediction3))
+residModel3$adjRes <- (residModel3$diff / sd(residModel3$diff))
+
 residModel1 %>% ggplot(aes(x=factor(obs), y=adjRes)) + geom_boxplot() +
@@
        x="Observed Value",y="Adjusted Residual")
+
+residModel3 %>% ggplot(aes(x=factor(obs), y=adjRes)) + geom_boxplot() +
+  labs(title="Adjusted Residuals", subtitle="Poisson Model #3",
+       x="Observed Value",y="Adjusted Residual")
 ```

### 2019-12-09 final project/Draft.Rmd
Updated Draft.RMD
Added some correlation to the Draft

@@
 ---
-title             : "A Logistic Regression Approach to CoIL Challenge 2000"
-shorttitle        : " "
-author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
-affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
-abstract: |
-  A logistic regression based solution to the CoIL Challenge 2000 is described.
-The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
-  <!-- https://tinyurl.com/ybremelq 
-    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-  -->
-keywords          : "CoIL Challenge, Logistic Regression"
-wordcount         : "X"
-bibliography      : ["Manuscript.bib"]
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-documentclass     : "apa6"
-classoption       : "man"
+title: "A Logistic Regression Approach to CoIL Challenge 2000"
+author: "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
+affiliation: MS in Data Science Students CUNY School of Professional Studies
 output:
@@
     reference_docx: word-template.docx
-  papaja::apa6_pdf:
-    #citation_package: natbib #biblatex
+  pdf_document: default
+  papaja::apa6_pdf: null
   html_document: default
+classoption: man
+documentclass: apa6
+draft: no
+figurelist: no
+floatsintext: no
+footnotelist: no
+keywords: CoIL Challenge, Logistic Regression
+linenumbers: no
+mask: no
+bibliography: Manuscript.bib
+shorttitle: DRAFT SUBMISSION
+tablelist: no
+abstract: "A logistic regression based solution to the CoIL Challenge 2000 is described.
+  The challenge consists of correctly identifying potential customers for an insurance
+  \n  One sentence summarizing the main result (with the words \"**here we show**\"
+  or their equivalent).\n\nTwo or three sentences explaining what the **main result**
+  reveals in direct comparison to what was thought to be the case previously, or how
+  the  main result adds to previous knowledge.\n\nOne or two sentences to put the
+  results into a more **general context**.\n\nTwo or three sentences to provide a
+  **broader perspective**, readily comprehensible to a scientist in any discipline.\n-->\n"
+  product, and describing their characteristics.\n<!-- https://tinyurl.com/ybremelq
+wordcount: X
 ---
@@
          CARAVAN = as.factor(CARAVAN)) %>%
-  mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
-         MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
-         MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
-         MGODRK = recode(MGODRK, !!!MGODRK_labels),
-         PWAPART = recode(PWAPART, !!!PWAPART_labels))
+  mutate(MOSTYPE = recode_factor(MOSTYPE, !!!MOSTYPE_labels),
+         MGEMLEEF = recode_factor(MGEMLEEF, !!!MGEMLEEF_labels),
+         MOSHOOFD = recode_factor(MOSHOOFD, !!!MOSHOOFD_labels),
+         MGODRK = recode_factor(MGODRK, !!!MGODRK_labels),
+         PWAPART = recode_factor(PWAPART, !!!PWAPART_labels))
 }
@@
   ylab("CoIL Challenge Submissions") +
-  xlim(37, 122)
+  xlim(37, 122) +
+  labs(title="Frequency Distribution of Prediction Scores")
 ```
@@
 
+Our response variable here is the `CARAVAN` variable. It represents whether the person in question purchased an caravan insurance plan.
+
 ```{r, echo=FALSE}
@@
 
-The dataset is unbalanced.
-There are only `r df %>% filter(CARAVAN == 1) %>% nrow()` cases that have purchased the mobile home insurance policy and `r df %>% filter(CARAVAN == 0) %>% nrow()` that have not.
+The dataset is unbalanced as there are only `r df %>% filter(CARAVAN == 1) %>% nrow()` cases that have purchased the mobile home insurance policy and `r df %>% filter(CARAVAN == 0) %>% nrow()` that have not.
 
@@
   df$var <- df[,var]
+
   p <- ggplot(df) +
-    geom_mosaic(aes(x = product(CARAVAN, var), fill=CARAVAN), na.rm=TRUE) +
+    geom_mosaic(aes(x = product(CARAVAN,var), fill=CARAVAN),na.rm=TRUE) +
     labs(x=var, y="CARAVAN") +
     theme(legend.position = "none") + 
-    scale_fill_brewer(palette = "Set1")
+    scale_fill_brewer(palette = "Set1") +
+    theme(axis.text.x = element_text(angle=45,hjust=1))
 
@@
 ```
+
+Looking at our categorical variables, there is some variation in frequency of `CARAVAN` in some values of `MOSTYPE` and `MGEMLEEF`, but the rest have quite a bit less variation.
 
@@
 ```
+
+For numeric variables, we don't see any which can clearly seperate the `CARAVAN` variable values.
+
+### Correlation
 
@@
 
-Most of the correlation heatmap is yellow which is has a correlation coeffient of zero.  Here's a density plot of the correlation coefficients:
+Most of the correlation heatmap is yellow which is has a correlation coeffient of zero.
+There are a few variable combinations with correlation, however.
+Here's a density plot of the correlation coefficients:
 
@@
   geom_density(fill="red", color="red")
+```
+
+Next we will look at variable correlation with the outcome variable:
+
+```{r, echo=FALSE}
+numeric_vars <- df %>%
+  select(-CARAVAN, -MOSTYPE, -MGEMLEEF, -MOSHOOFD, -MGODRK, -PWAPART) %>%
+  names()
+
+corrs <- data.frame(CARAVAN = 1)
+
+for (var in numeric_vars){
+   t <- df %>% select(CARAVAN,var) %>% mutate(CARAVAN = as.numeric(CARAVAN))
+   corrs[,var] <- cor(t,use="complete.obs")[1,2]
+}
+
+corrs <- corrs %>% gather(key="variable",value="correlation",-CARAVAN) %>%
+  select(-CARAVAN) %>% mutate(positive = ifelse(correlation > 0,TRUE,FALSE))
+
+corrs[1:20,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
+  geom_col() + scale_y_continuous(limits=c(-1,1)) +
+  scale_fill_manual(values=c("#d13a26","#369970")) +
+  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
+  labs(title="Variable Correlations with Dependent Variable",
+       subtitle="Part 1")
+
+corrs[21:40,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
+  geom_col() + scale_y_continuous(limits=c(-1,1)) +
+  scale_fill_manual(values=c("#d13a26","#369970")) +
+  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
+  labs(title="Variable Correlations with Dependent Variable",
+       subtitle="Part 2")
+
+corrs[41:60,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
+  geom_col() + scale_y_continuous(limits=c(-1,1)) +
+  scale_fill_manual(values=c("#d13a26","#369970")) +
+  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
+  labs(title="Variable Correlations with Dependent Variable",
+       subtitle="Part 3")
+
+corrs[61:80,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
+  geom_col() + scale_y_continuous(limits=c(-1,1)) +
+  scale_fill_manual(values=c("#d13a26","#369970")) +
+  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
+  labs(title="Variable Correlations with Dependent Variable",
+       subtitle="Part 4")
+
+```
+
+None of the numeric variables, alone, seem to be well-correlated with the dependent variable. In fact, the correlation is quite poor across the board:
+
+```{r, echo = FALSE}
+summary(corrs$correlation)
 ```

### 2019-12-10 final project/Manuscript.Rmd
Additional work on final project

@@
 
-df <- df %>%
+df %>%
   mutate(MOSTYPE = as.factor(MOSTYPE),
@@
 <!--Our investigation is similar to @Doornik in that we used a LOGIT model. **Reproduce his model and describe the results**-->
-
-```{r, echo=FALSE}
-doornik <- function(df) {
-  df <- df %>%
-    mutate(PBRAND_2_or_6 = ifelse(PBRAND == 2 | PBRAND == 6, 1, 0),
-           PBRAND_3_or_4 = ifelse(PBRAND == 3 | PBRAND == 4, 1, 0),
-           PPLEZIER_gt_0 = ifelse(PPLEZIER > 0, 1, 0),
-           PPERSAUT_6 = ifelse(PPERSAUT == 6, 1, 0))
-  z <- -4.1 - (0.13 * df$MOPLLAAG) - (0.3 * df$MINK123M) + (0.17 * df$MINKGEM) + (0.54 * (df$ALEVEN + df$AFIETS + df$ABYSTAND)) - df$PBRAND_2_or_6 + (0.8 * df$PBRAND_3_or_4) + (2.4 * df$PPLEZIER_gt_0) + (1.5 * df$PPERSAUT_6) + (0.28 * (df$PWAOREG - df$PLEVEN))
-  p_caravan <- exp(z) / (1 + exp(z))
-  #p_caravan
-  ifelse(p_caravan > 0.5, 1, 0)
-}
-
-eval$doornik <- doornik(eval)
-```
 

### 2019-12-10 final project/CoIL.R

@@ -1,5 +1,6 @@
 # CoIL Challenge Source Code
 library(tidyverse)
+library(caret)
 # Download the data sets from UCI if they are not present
 url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/tic-mld/"
 files <- c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt")
@@ -11,26 +12,186 @@ for (file_name in files) {
     download.file(file_url, file_path)
   }
 }
-# Read in the data
-df <- read.delim("data/ticdata2000.txt", header = FALSE)
-names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
-               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
-               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
-               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
-               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
-               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
-               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
-               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
-               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
-               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM",
-               "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", 
-               "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND",
-               "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
-               "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT",
-               "ABROM", "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",
-               "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
[…]

### 2019-12-11 final project/CoIL.R
More work on the final

@@ -1,7 +1,8 @@
 # CoIL Challenge Source Code
 library(tidyverse)
 library(caret)
-# Download the data sets from UCI if they are not present
+
+## Download the data sets from UCI if they are not present
 url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/tic-mld/"
 files <- c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt")
 for (file_name in files) {
@@ -12,108 +13,117 @@ for (file_name in files) {
     download.file(file_url, file_path)
   }
 }
-# Read in and clean the data
+
+## Read in and clean the data
 prepare_data <- function(df){
-  names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
-                 "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
-                 "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
-                 "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
-                 "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
-                 "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
-                 "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
-                 "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
-                 "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
-                 "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
-                 "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
-                 "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
-                 "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
-                 "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
[…]

### 2019-12-11 final project/Manuscript.Rmd
Cleaned up Manuscript

@@
 ---
-title             : "A Logistic Regression Approach to CoIL Challenge 2000"
-shorttitle        : " "
-author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
-affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
-abstract: |
-  A logistic regression based solution to the CoIL Challenge 2000 is described.
-The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
-  <!-- https://tinyurl.com/ybremelq 
-    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-  -->
-keywords          : "CoIL Challenge, Logistic Regression"
-wordcount         : "X"
-bibliography      : ["Manuscript.bib"]
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-documentclass     : "apa6"
-classoption       : "man"
+title: "A Logistic Regression Approach to CoIL Challenge 2000"
+author: "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
+bibliography: "Manuscript.bib"
 output: 
@@
     reference_docx: word-template.docx
-  papaja::apa6_pdf:
-    #citation_package: natbib #biblatex
-  html_document: default
 ---
 
-```{r setup, include = FALSE}
-# Documentation at https://crsh.github.io/papaja_man/
-# If you have no use for TeX beyond rendering R Markdown documents, I recommend you use TinyTex.
-# install.packages("tinytex")
-# tinytex::install_tinytex()
-if (!"papaja" %in% rownames(installed.packages())) {
-  devtools::install_github("crsh/papaja")
-}
-library("papaja")
-library(kableExtra)
+```{r echo=FALSE, message=FALSE}
 library(tidyverse)
@@
 
-# Thank you Stack Overflow!
-# A Prefix nulling hook.
-
-# Make sure to keep the default for normal processing.
-default_output_hook <- knitr::knit_hooks$get("output")
-
-# Output hooks handle normal R console output.
-knitr::knit_hooks$set(output = function(x, options) {
-
-  comment <- knitr::opts_current$get("comment")
-  if (is.na(comment)) comment <- ""
-  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
-                     x, perl = TRUE)
-  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
-  if (can_null && do_null) {
-    # By default R print output aligns at the right brace.
-    align_index <- regexpr("\\]", x)[1] - 1
-    # Two cases: start or newline
-    re <- paste0("^.{", align_index, "}\\]")
-    rep <- comment
-    x <- gsub(re, rep,  x)
-    re <- paste0("\\\n.{", align_index, "}\\]")
-    rep <- paste0("\n", comment)
-    x <- gsub(re, rep,  x)
-  }
-
-  default_output_hook(x, options)
-
-})
-knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
-```
-
-```{r analysis-preferences, echo=FALSE}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
-
-```{r data, echo=FALSE}
 prepare_data <- function(df){
-  names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
-               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
-               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
-               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
-               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
-               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
-               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
-               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
-               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
-               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
-               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
-               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
-               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
-               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
-               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
+  names(df) <- c(
+    "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
+    "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN",
+    "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG",
+    "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA",
+    "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2",
+    "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
+    "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", 
+    "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT",  "PAANHANG",
+    "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG",
+    "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED",
+    "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
+    "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM",
+    "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
     "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
 
-  MOSTYPE_labels <- c("1" = "1 High Income, expensive child",
+  MOSTYPE_labels <- c(
+    "1" = "1 High Income, expensive child",
     "2" = "2 Very Important Provincials",
@@
 
-MGEMLEEF_labels <- c("1" = "20-30 years",
+  MGEMLEEF_labels <- c(
+    "1" = "20-30 years",
     "2" = "30-40 years",
@@
 
-MOSHOOFD_labels <- c("1" = "Successful hedonists",
+  MOSHOOFD_labels <- c(
+    "1" = "Successful hedonists",
     "2" = "Driven Growers",
@@
 
-MGODRK_labels <- c("0" = "0%",
+  MGODRK_labels <- c(
+    "0" = "0%",
     "1" = "1 - 10%",
@@
 
-PWAPART_labels <- c("0" = "f 0",
-                    "1" = "f 1 – 49",
-                    "2" = "f 50 – 99",
-                    "3" = "f 100 – 199",
-                    "4" = "f 200 – 499",
-                    "5" = "f 500 – 999",
-                    "6" = "f 1000 – 4999",
-                    "7" = "f 5000 – 9999",
-                    "8" = "f 10.000 - 19.999",
-                    "9" = "f 20.000 - ?")
+  PWAPART_labels <- c(
+    "0" = "f 0",
+    "1" = "f 1 - 49",
+    "2" = "f 50 - 99",
+    "3" = "f 100 - 199",
+    "4" = "f 200 - 499",
+    "5" = "f 500 - 999",
+    "6" = "f 1000 - 4999",
+    "7" = "f 5000 - 9999",
+    "8" = "f 10,000 - 19,999",
+    "9" = "f 20,000 - ?")
+
+  set_to_1 <- c(12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, 11)
 
   df %>%
+    mutate(CUSTOMER_TYPE = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
+    mutate(CUSTOMER_TYPE = as.factor(CUSTOMER_TYPE)) %>%
     mutate(MOSTYPE = as.factor(MOSTYPE),
@@
 ```
+
+# Abstract
+
+A logistic regression based solution to the CoIL Challenge 2000 is described.
+The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
+
+*Key words: CoIL Challenge, Logistic Regression*
+
 # Introduction
-<!-- Describe the background and motivation of your problem.-->
 
 Businesses use data science to extract insights from data.
-It has many practical business applications.
-Identifying households to include in a marketing campaign is one application.
-One example using real world data is the Computational Intelligence and Learning (CoIL) Challenge.
-The CoIL Challenge competition was held from March 17 to May 8 in 2000.
-The challenge is to:
-
-1.  Identify potential customers for an insurance policy; and
-
-2.  Provide a description of this customer base.
-
-```{r, echo=FALSE}
-# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
-prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83,
-                                          86, 91, 94, 95, 96, 97, 99, 101, 103,
-                                          104, 105, 106, 107, 108, 109, 110,
-                                          111, 112, 115, 121),
-                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3,
-                                      1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1,
-                                      1))
-ggplot(data = prediction_scores, aes(x = score, y = n)) +
-  geom_bar(stat = "identity") +
-  xlab("Number of Correct Predictions") +
-  ylab("CoIL Challenge Submissions") +
-  xlim(37, 122)
-```
-
-In total 147 participants registered and 43 submitted solutions [@Putten].
-The maximum number of policyowners that could be found was 238.
-The submissions identified 95 policy owners on average.
-The winning model [@Elkan] identified 121 policy owners.
-Random selection results in identifying 42 policy owners.
-The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners.
-[@Putten].
-In this paper we set out to complete the first part of the COIL Challenge.
-<!--**SUMARISE FINDINGS?**-->
+One pratical application is identifying households to include in a marketing campaign.
+In this paper we set out to identify potential customers for an insurance product using real world data from the Computational Intelligence and Learning (CoIL) Challenge.
 
@@
 
-Participants used a variety of approaches in formulating their submissions including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].
+Fourty-three other research teams have attempted to identify potential caravan insurance customers [@Putten].
+They used a variety of approaches including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].
 
-<!--Our investigation is similar to @Doornik in that we used a LOGIT model. **Reproduce his model and describe the results**-->
+```{r, echo=FALSE}
+# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
+prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
+ggplot(data = prediction_scores, aes(x = score, y = n)) +
+  geom_bar(stat = "identity", fill = "#377eb8") +
+  xlab("Number of Correct Predictions") +
+  ylab("CoIL Challenge Submissions") +
+  xlim(37, 122) +
+  scale_fill_brewer(palette = "Set1")
+```
+
+The maximum number of potential policy owners that could be found is 238.
+Previous researchers identified 95 policy owners on average.
+The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.
+A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].
 
@@
 
-<!--Discuss the key aspects of your problem, data set and regression model(s).
+<!-- Discuss the key aspects of your problem, data set and regression model(s).
 Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.
-
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-# Experimentation and Results
-
-<!--Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).
-
-We used `r cite_r("Manuscript.bib")` for all our analyses.
 -->
 
-# Discussion and Conclusions
-
-<!--Conclude your findings, limitations, and suggest areas for future work.-->
-
-\newpage
-
-# References
-```{r create_references, echo=FALSE}
-r_refs(file = "Manuscript.bib")
-```
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id = "refs"></div>
-\endgroup
-
-\newpage
-
-# Appendices
-
-<!--
-- Supplemental tables and/or figures.
--->
+# Appendix
 
@@
 </code>
+
+## References

### 2019-12-11 final project/CoIL.R

@@ -109,15 +109,15 @@ prepare_data <- function(df){
   
   PWAPART_labels <- c(
     "0" = "f 0",
-    "1" = "f 1 – 49",
-    "2" = "f 50 – 99",
-    "3" = "f 100 – 199",
-    "4" = "f 200 – 499",
-    "5" = "f 500 – 999",
-    "6" = "f 1000 – 4999",
-    "7" = "f 5000 – 9999",
-    "8" = "f 10.000 - 19.999",
-    "9" = "f 20.000 - ?")
+    "1" = "f 1 - 49",
+    "2" = "f 50 - 99",
+    "3" = "f 100 - 199",
+    "4" = "f 200 - 499",
+    "5" = "f 500 - 999",
+    "6" = "f 1000 - 4999",
+    "7" = "f 5000 - 9999",
+    "8" = "f 10,000 - 19,999",
+    "9" = "f 20,000 - ?")
   
   set_to_1 <- c(12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, 11)
   

### 2019-12-11 final project/Manuscript.Rmd
Added instructions

@@
 
+<!-- In the abstract, the authors should explain the major objective of the study in an objective section, explain how the study was done in a Methods section, describe the findings in a Results section, and report whether the major goal was met in a Conclusion.
+In general, the reader should ask, If I could not read the entire manuscript, would the abstract adequately summarize it?
+-->
 A logistic regression based solution to the CoIL Challenge 2000 is described.
@@
 # Introduction
+
+<!-- Readers need to understand all the reasons for performing the study, the assumptions underlying the methodology, and the nuances of the performance of the study.
+The purposes of the Introduction are, first, to provide the rationale for the study and, second, to explain the study's goals.
+The Introduction should include a problem statement that conveys the important issues and provides the context for the study.
+-->
 
@@
 -->
+
+## Experimentation and Results
+
+<!-- Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).
+
+Needs to systematically and clearly articulate the study findings.
+If the results are unclear, the reviewer must decide whether the analysis of the data was poorly executed or whether the Results section is poorly organized.
+-->
+
+## Discussion
+
+<!-- Should state whether their hypotheses were verified or proven untrue or, if no hypotheses were given, whether their research questions were answered.
+The authors should also comment on their results in light of previous studies and explain what differences (if any) exist between their findings and those reported by others and attempt to provide an explanation for the discrepancies.-->
+
+# Conclusions
+
+<!-- Conclude your findings, limitations, and suggest areas for future work. -->
 

### 2019-12-11 final project/CoIL.R
Minor changes

@@ -122,8 +122,8 @@ prepare_data <- function(df){
   set_to_1 <- c(12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, 11)
   
   df %>%
-    mutate(CUSTOMER_TYPE = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
-    mutate(CUSTOMER_TYPE = as.factor(CUSTOMER_TYPE)) %>%
+    mutate(LIKELY_CUSTOMERS = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
+    mutate(LIKELY_CUSTOMERS = as.factor(LIKELY_CUSTOMERS)) %>%
     mutate(MOSTYPE = as.factor(MOSTYPE),
            MGEMLEEF = as.factor(MGEMLEEF),
            MOSHOOFD = as.factor(MOSHOOFD),
@@ -154,26 +154,28 @@ up_train <- upSample(x = select(train, -CARAVAN),
                      y = train$CARAVAN,
                      yname = "CARAVAN") 
 
-# ## Looking for important variables
+## Looking for important variables
 # set.seed(42)
 # library(randomForest)
 # rf_fit <- randomForest(CARAVAN ~ ., up_train)
 # varImpPlot(rf_fit)
-# 
-# ## Find important customer types
-# crosstab <- up_train %>%
-#   select(CARAVAN, MOSTYPE) %>%
-#   table() %>%
-#   data.frame()
-# 
-# crosstab %>%
-#   group_by(MOSTYPE) %>%
-#   summarise(total = sum(Freq)) %>%
-#   merge(crosstab) %>%
-#   mutate(share = Freq / total) %>%
-#   filter(CARAVAN == 1, share > 0.5) %>%
-#   arrange(desc(share)) %>%
-#   select(MOSTYPE, share)
+
+## Find likely customer types
+MOSTYPE_crosstab <- up_train %>%
+  select(CARAVAN, MOSTYPE) %>%
+  table() %>%
+  data.frame()
+
+MOSTYPE_crosstab <- MOSTYPE_crosstab %>%
+  group_by(MOSTYPE) %>%
+  summarise(total = sum(Freq)) %>%
+  merge(MOSTYPE_crosstab) %>%
+  mutate(share = Freq / total) %>%
+  filter(CARAVAN == 1, share > 0.5) %>%
+  arrange(desc(share)) %>%
+  select(MOSTYPE, share)
+
+MOSTYPE_crosstab
 
[…]

### 2019-12-13 final project/Manuscript.Rmd
More work on final

@@
 ```{r echo=FALSE, message=FALSE}
-library(tidyverse)
+source("CoIL.R")
 library(pander)
-
-prepare_data <- function(df){
-  names(df) <- c(
-    "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
-    "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN",
-    "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG",
-    "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA",
-    "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2",
-    "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
-    "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", 
-    "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT",  "PAANHANG",
-    "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG",
-    "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED",
-    "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
-    "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM",
-    "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
-    "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
-
-  MOSTYPE_labels <- c(
-    "1" = "1 High Income, expensive child",
-    "2" = "2 Very Important Provincials",
-    "3" = "3 High status seniors",
-    "4" = "4 Affluent senior apartments",
-    "5" = "5 Mixed seniors",
-    "6" = "6 Career and childcare",
-    "7" = "7 Dinki's (double income no kids)",
-    "8" = "8 Middle class families",
-    "9" = "9 Modern, complete families",
-    "10" = "10 Stable family",
-    "11" = "11 Family starters",
-    "12" = "12 Affluent young families",
-    "13" = "13 Young all american family",
-    "14" = "14 Junior cosmopolitan",
-    "15" = "15 Senior cosmopolitans",
-    "16" = "16 Students in apartments",
-    "17" = "17 Fresh masters in the city",
-    "18" = "18 Single youth",
-    "19" = "19 Suburban youth",
-    "20" = "20 Ethnically diverse",
-    "21" = "21 Young urban have-nots",
-    "22" = "22 Mixed apartment dwellers",
-    "23" = "23 Young and rising",
-    "24" = "24 Young, low educated",
-    "25" = "25 Young seniors in the city",
-    "26" = "26 Own home elderly",
-    "27" = "27 Seniors in apartments",
-    "28" = "28 Residential elderly",
-    "29" = "29 Porchless seniors: no front yard",
-    "30" = "30 Religious elderly singles",
-    "31" = "31 Low income catholics",
-    "32" = "32 Mixed seniors",
-    "33" = "33 Lower class large families",
-    "34" = "34 Large family, employed child",
-    "35" = "35 Village families",
-    "36" = "36 Couples with teens 'Married with children'",
-    "37" = "37 Mixed small town dwellers",
-    "38" = "38 Traditional families",
-    "39" = "39 Large religous families",
-    "40" = "40 Large family farms",
-    "41" = "41 Mixed rurals")
-
-  MGEMLEEF_labels <- c(
-    "1" = "20-30 years",
-    "2" = "30-40 years",
-    "3" = "40-50 years",
-    "4" = "50-60 years",
-    "5" = "60-70 years",
-    "6" = "70-80 years")
-
-  MOSHOOFD_labels <- c(
-    "1" = "Successful hedonists",
-    "2" = "Driven Growers",
-    "3" = "Average Family",
-    "4" = "Career Loners",
-    "5" = "Living well",
-    "6" = "Cruising Seniors",
-    "7" = "Retired and Religeous",
-    "8" = "Family with grown ups",
-    "9" = "Conservative families",
-    "10" = "Farmers")
-
-  MGODRK_labels <- c(
-    "0" = "0%",
-    "1" = "1 - 10%",
-    "2" = "11 - 23%",
-    "3" = "24 - 36%",
-    "4" = "37 - 49%",
-    "5" = "50 - 62%",
-    "6" = "63 - 75%",
-    "7" = "76 - 88%",
-    "8" = "89 - 99%",
-    "9" = "100%")
-
-  PWAPART_labels <- c(
-    "0" = "f 0",
-    "1" = "f 1 - 49",
-    "2" = "f 50 - 99",
-    "3" = "f 100 - 199",
-    "4" = "f 200 - 499",
-    "5" = "f 500 - 999",
-    "6" = "f 1000 - 4999",
-    "7" = "f 5000 - 9999",
-    "8" = "f 10,000 - 19,999",
-    "9" = "f 20,000 - ?")
-
-  set_to_1 <- c(12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, 11)
-
-  df %>%
-    mutate(CUSTOMER_TYPE = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
-    mutate(CUSTOMER_TYPE = as.factor(CUSTOMER_TYPE)) %>%
-    mutate(MOSTYPE = as.factor(MOSTYPE),
-           MGEMLEEF = as.factor(MGEMLEEF),
-           MOSHOOFD = as.factor(MOSHOOFD),
-           MGODRK = as.factor(MGODRK),
-           PWAPART = as.factor(PWAPART),
-           CARAVAN = as.factor(CARAVAN)) %>%
-    mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
-           MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
-           MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
-           MGODRK = recode(MGODRK, !!!MGODRK_labels),
-           PWAPART = recode(PWAPART, !!!PWAPART_labels))
-}
-
-eval <- read.delim("data/ticeval2000.txt", header = FALSE)
-temp <- read.delim("data/tictgts2000.txt", header = FALSE)
-eval$CARAVAN <- temp$V1
-eval <- prepare_data(eval)
-df <- prepare_data(read.delim("data/ticdata2000.txt", header = FALSE))
+knitr::opts_chunk$set(comment = NA)
 ```
@@
 -->
-A logistic regression based solution to the CoIL Challenge 2000 is described.
+A logistic regression based solution to the CoIL Challenge 2000 is described in this paper.
 The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
+Models were trained on over sampled data.
+The model out preformed other's attemps.
 
@@
 ```{r, echo=FALSE}
-# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
+# Similar the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
 prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
@@
   xlab("Number of Correct Predictions") +
-  ylab("CoIL Challenge Submissions") +
-  xlim(37, 122) +
-  scale_fill_brewer(palette = "Set1")
+  ylab(element_blank()) + 
+  xlim(37, 240) +
+  annotate("text", x = 65, y = 4, label = "Previous Researchers") +
+  geom_vline(xintercept = 238, color = "#e41a1c", linetype="dashed") +
+  annotate("text", x = 220, y = 4, label = "Maximum") +
+  geom_vline(xintercept = final_model$correct, color = "#4daf4a") +
+  annotate("text", x = final_model$correct - 20, y = 4, label = "Our Model") +
+  ggtitle("Number of Customers Correctly Predicted")
 ```
@@
 
+## Correlation Coefficients for Variables of Interest
+```{r, echo=FALSE}
+vars <- c("CARAVAN", "PPERSAUT", "PBRAND", "APERSAUT", "LIKELY_CUSTOMERS")
+temp <- up_train %>%
+  mutate(CARAVAN = as.numeric(as.character(CARAVAN)),
+         LIKELY_CUSTOMERS = as.numeric(as.character(LIKELY_CUSTOMERS)))
+temp <- temp[vars]
+
+pander(cor(temp[vars], temp[vars]))
+```
+
 ## Data Dictionary
@@
 data_dictionary <- data.frame(
-  Name = c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
+  Name = c("CARAVAN", "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
            "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV",
@@
            "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
-           "ABYSTAND", "CARAVAN"),
-  Description = c("Customer Subtype", "Number of houses", "Avg size household",
+           "ABYSTAND", "LIKELY_CUSTOMERS"),
+  Description = c("Number of mobile home policy",
+                  "Customer Subtype", "Number of houses", "Avg size household",
                   "Avg age", "Customer main type", "Roman catholic",
@@
                   "Number of social security insurance policies",
-                  "Number of mobile home policy"))
-pander(data_dictionary)
+                  "MOSTYPE = 12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, or 11"))
+temp <- data_dictionary[data_dictionary$Name %in% vars,]
+row.names(temp) <- NULL
+pander(temp)
 ```
@@
 
+## Confusion Matrix and Statistics for our Model
+
+```{r, echo=FALSE}
+final_model$cm
+```
+
 ## R statistical programming code.
 
-<code>
 ```{r, null_prefix = TRUE, comment=NA, echo=FALSE}
@@
 ```
-</code>
+
+\newpage
 

### 2019-12-13 final project/CoIL.R

@@ -177,39 +177,27 @@ MOSTYPE_crosstab <- MOSTYPE_crosstab %>%
 
 MOSTYPE_crosstab
 
-## Model Building
-
-### Model 1 - Top 5 Important Variables from Random Forest
-model1 <- glm(CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT, 
-              family = binomial(link = "logit"),
-              up_train)
-
-### Model 2 - Likely Customers and Car Policies Contribution Level
-model2 <- glm(CARAVAN ~ LIKELY_CUSTOMERS + PPERSAUT,
-              family = binomial(link = "logit"),
-              up_train)
-
-## Model Evaluation
+## Model Building & Evaluation
 
 score_model <- function(model, data, threshold = 0.5, predictions = FALSE){
-  ## Provides model scoring data
-  #
-  # INPUTS
-  #
-  # model = logit model object
-  # data = data frame to make predictions for
-  # threshold (optional) = the cutpoint to assign a 1 or 0 response
-  # predictions (optional) = 1 or 0 you want to use for the predicitions
-  # 
-  # RETURNS (list)
-  #
-  # cm = Confusion Matrix output from caret
-  # correct = the number of correct CARAVAN = 1 predictions
-  # accuracy = the accuracy of the CARAVAN = 1 predictions
+## Provides model scoring data
+#
+# INPUTS
+#
+# model = logit model object
+# data = data frame to make predictions for
+# threshold (optional) = the cutpoint to assign a 1 or 0 response
+# predictions (optional) = 1 or 0 you want to use for the predicitions
+# 
+# RETURNS (list)
+#
+# cm = Confusion Matrix output from caret
+# correct = the number of correct CARAVAN = 1 predictions
+# accuracy = the accuracy of the CARAVAN = 1 predictions
   
   # Generate the predicted outcome
   if(!predictions){
-    glm_predictions <- predict.glm(model, data, "response")
+    glm_predictions <- suppressWarnings(predict.glm(model, data, "response"))
     predictions <- ifelse(glm_predictions >= threshold, 1, 0)
   }
   data$yhat <- predictions
@@ -230,12 +218,79 @@ score_model <- function(model, data, threshold = 0.5, predictions = FALSE){
   return(list("cm" = cm, "correct" = correct, "accuracy" = accuracy))
 }
 
-results1 <- score_model(model1, test)
-results2 <- score_model(model2, test)
[…]

### 2019-12-13 final project/Manuscript.Rmd
Began to flesh out methodology section

@@
 ```{r echo=FALSE, message=FALSE}
+temp <- read.delim("data/ticdata2000.txt", header = FALSE, nrows = 1)
+df_cols <- length(names(temp))
 source("CoIL.R")
@@
 knitr::opts_chunk$set(comment = NA)
+knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })
 ```
@@
 Models were trained on over sampled data.
-The model out preformed other's attemps.
+The model out preformed other's attemps at solving this classification problem.
 
@@
 In this paper we set out to identify potential customers for an insurance product using real world data from the Computational Intelligence and Learning (CoIL) Challenge.
+Specifically we are predicting if a customer is likely candidate for a caravan (mobile home/camper) insurance policy.
+This is particularly challenging because the data is imballanced (only `r sum(as.numeric(as.character(df$CARAVAN)))` of the `r nrow(df)` records for model training/testing are policy holders).
 
@@
 
-Fourty-three other research teams have attempted to identify potential caravan insurance customers [@Putten].
+Fourty-three other research teams have attempted to identify potential insurance policy customers [@Putten].
 They used a variety of approaches including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].
@@
 -->
+The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.  An evaluation dataset is provided with `r nrow(eval)` observations.
+
+**THERE ARE X that are categorical and X that are numeric.**  Most of the variables have little to no correlation with the variable of interest (CARAVAN).
 
@@
 -->
+
+We split the data into training/test sets using a 70/30 split.
+We corrected the imballance by oversampling the minority class (caravan policy holders).
+Given the large number of possible predictors we used a random forest to aid in variable selection.
+A logistic regression model was trained on the oversampled training set using the top five variables selected by the random forest (MOSTYPE, PPERSAUT, MOSHOOFD, PBRAND, APERSAUT).
 

### 2019-12-13 final project/Manuscript.Rmd
Added more to methodology and appendix

@@
 
+The MOSTYPE variable has 40 customer types.
+Not all customer types were statiscally significant predictors.
+We identified those with a probability greater than 0.5 to purchase the insurance product as the `LIKELY_CUSTOMER` variable.
+Our second model used LIKELY_CUSTOMER and PPERSAUT.
+
 ## Discussion
@@
 
+\newpage
+
 # Appendix
+
+## Probability of Purchasing Product (in blue) by Customer Type
+
+```{r, echo=FALSE, fig.width=10, fig.height=11}
+temp <- up_train %>%
+  group_by(MOSTYPE) %>%
+  tally() %>%
+  rename(total = n)
+
+temp <- up_train %>%
+  filter(CARAVAN == 1) %>%
+  group_by(MOSTYPE) %>%
+  tally() %>%
+  merge(temp, ., all.x = TRUE) %>%
+  mutate(share_1 = ifelse(is.na(n), 0, n / total)) %>%
+  mutate(order = row_number(share_1)) %>%
+  select(-n) %>%
+  merge(temp)
+
+temp <- up_train %>%
+  group_by(CARAVAN, MOSTYPE) %>%
+  tally() %>%
+  merge(temp) %>%
+  mutate(share = n / total)
+
+ggplot(temp, aes(x = reorder(MOSTYPE, order), y = share, fill = CARAVAN))+
+  geom_bar(stat="identity") +
+  scale_fill_brewer(palette = "Set1") +
+  coord_flip() +
+  xlab(element_blank()) +
+  ylab(element_blank()) +
+  geom_hline(yintercept = 0.5, color = "white", linetype = "dashed") +
+  theme_minimal() +
+  theme(legend.position = "none") 
+```
+
+\newpage
 

### 2019-12-14 final project/Manuscript.Rmd
More work on the final project

@@
 
+The maximum number of potential policy owners that could be found is 238.
+Previous researchers identified 95 policy owners on average.
+The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.
+A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].
+With the benefit of this finding, we set out to create a simple logistic regression model that preforms just as well or better than the original CoIL Challenge cohort.
+The model that we found outpreformed the orignal researcher's model in correctly predicting the customers that would purchase the insurance policy.
+
 ```{r, echo=FALSE}
@@
   ylab(element_blank()) + 
-  xlim(37, 240) +
+  xlim(37, 250) +
   annotate("text", x = 65, y = 4, label = "Previous Researchers") +
@@
 ```
-
-The maximum number of potential policy owners that could be found is 238.
-Previous researchers identified 95 policy owners on average.
-The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.
-A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].
 
@@
 We identified those with a probability greater than 0.5 to purchase the insurance product as the `LIKELY_CUSTOMER` variable.
-Our second model used LIKELY_CUSTOMER and PPERSAUT.
+A figure in the appendix shows all 40 customer types and the probability that they purched the insurance policy.
+Our second model only used LIKELY_CUSTOMER and PPERSAUT.
+Although it only has two variables, the model preformed well.
+
+In evaluating the models we examined we focused on the specificity.
+The goal of the CoIL challenge was to accurately predict those would would purchase the insurance policy, so focusing on the model's specificity was the best evaluation metric.
+
+In order to get a better sense of how well the model generalizes, we repeatedly retrained and evaluated the model using diffrent samplings of the dataset.
 
@@
 ```{r, echo=FALSE}
-vars <- c("CARAVAN", "PPERSAUT", "PBRAND", "APERSAUT", "LIKELY_CUSTOMERS")
+vars <- c("CARAVAN", "MOSTYPE", "PPERSAUT", "PBRAND", "MOSHOOFD", "APERSAUT", "LIKELY_CUSTOMERS")
 temp <- up_train %>%
@@
 
-pander(cor(temp[vars], temp[vars]))
+numeric_vars <- unlist(lapply(temp, is.numeric))
+
+pander(cor(temp[numeric_vars], temp[numeric_vars]))
 ```
 
-## Data Dictionary
+## Data Dictionary for Variables of Interest
 

### 2019-12-14 final project/CoIL.R

@@ -193,7 +193,7 @@ score_model <- function(model, data, threshold = 0.5, predictions = FALSE){
 #
 # cm = Confusion Matrix output from caret
 # correct = the number of correct CARAVAN = 1 predictions
-# accuracy = the accuracy of the CARAVAN = 1 predictions
+# specificity = the specificity of the CARAVAN = 1 predictions
   
   # Generate the predicted outcome
   if(!predictions){
@@ -211,11 +211,11 @@ score_model <- function(model, data, threshold = 0.5, predictions = FALSE){
            yhat == CARAVAN) %>%
     nrow(.)
   
-  # Get the accuracy of the model's CARAVAN = 1 Predictions
-  accuracy <- correct / nrow(data[data$CARAVAN == 1,])
+  # Get the specificity of the model's CARAVAN = 1 Predictions
+  specificity <- correct / nrow(data[data$CARAVAN == 1,])
   
   # Return the data as a list
-  return(list("cm" = cm, "correct" = correct, "accuracy" = accuracy))
+  return(list("cm" = cm, "correct" = correct, "specificity" = specificity))
 }
 
 robust_results <- function(model_formula, n_tries = 250){
@@ -230,7 +230,7 @@ robust_results <- function(model_formula, n_tries = 250){
 #
 # seed = randome number seed
 # correct = the number of correct CARAVAN = 1 predictions
-# accuracy = the accuracy of the CARAVAN = 1 predictions
+# specificity = the specificity of the CARAVAN = 1 predictions
   
   # Convert the formula from a string
   model_formula <- as.formula(model_formula)
@@ -256,7 +256,7 @@ robust_results <- function(model_formula, n_tries = 250){
         # Store the results
         temp <- data.frame("seed" = seed,
                            "correct" = results$correct,
-                           "accuracy" = results$accuracy)
+                           "specificity" = results$specificity)
         if(exists("the_results")){
           the_results <- bind_rows(the_results, temp)
         } else {
@@ -276,21 +276,21 @@ model1 <- glm(CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT,
               family = binomial(link = "logit"),
               up_train)
 model1_results <- score_model(model1, test)
-model1_results$accuracy
+model1_results$specificity
 # TODO: Uncomment these lines
[…]

### 2019-12-15 final project/Draft.Rmd

--- 2019-12-06 final project/Draft.Rmd
@@
 The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
-  <!-- https://tinyurl.com/ybremelq 
-    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-  -->
 keywords          : "CoIL Challenge, Logistic Regression"

### 2019-12-15 final project/Manuscript.Rmd

--- 2019-12-13 final project/Manuscript.Rmd
@@
 library(pander)
-knitr::opts_chunk$set(comment = NA)
+knitr::opts_chunk$set(comment = NA, echo = FALSE)
 knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })
@@
 
-```{r, echo=FALSE}
+The maximum number of potential policy owners that could be found is 238.
+Previous researchers identified 95 policy owners on average.
+The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.
+It was a Naïve Bayes, suggesting that probabilities of some of the variables will be useful in identifing potential customers.
+
+A meta analysis of the initial researchers found that simpler algorithms tended to outpreform more complicated ones [@Putten].
+With the benefit of these findings, we set out to create a simple logistic regression model that preforms as well or better than the original CoIL Challenge cohort.
+In the end, our model outpreformed the orignal researcher's model in correctly predicting the customers that would purchase the insurance policy.
+
+```{r}
 # Similar the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
@@
   geom_bar(stat = "identity", fill = "#377eb8") +
-  xlab("Number of Correct Predictions") +
+  xlab(element_blank()) +
   ylab(element_blank()) + 
-  xlim(37, 240) +
+  xlim(37, 250) +
   annotate("text", x = 65, y = 4, label = "Previous Researchers") +
@@
 
-The maximum number of potential policy owners that could be found is 238.
-Previous researchers identified 95 policy owners on average.
-The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.
-A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].
-
 # Methodology
@@
 -->
-The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.  An evaluation dataset is provided with `r nrow(eval)` observations.
-
-**THERE ARE X that are categorical and X that are numeric.**  Most of the variables have little to no correlation with the variable of interest (CARAVAN).
+The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.
+An evaluation dataset is provided with `r nrow(eval)` observations.
+Five of the predictors are categorical and the remainder are numeric.
+Most of the predictors have little to no correlation with the variable of interest (CARAVAN).
 
@@
 Not all customer types were statiscally significant predictors.
-We identified those with a probability greater than 0.5 to purchase the insurance product as the `LIKELY_CUSTOMER` variable.
-Our second model used LIKELY_CUSTOMER and PPERSAUT.
+We identified those in the oversampled dataset with a probability greater than 0.5 to purchase the insurance product as the LIKELY_CUSTOMER variable.
+A figure in the appendix shows all 40 customer types and the probability that they purched the insurance policy in the oversampled dataset.
+
+We fit our second model to the LIKELY_CUSTOMER and PPERSAUT variables.
+Although it only has two variables, the model preformed well.
+This seemed in-line with expectations that simple models preform best.
+
+In evaluating the models we examined we focused on the specificity.
+The goal of the CoIL challenge was to accurately predict those would would purchase the insurance policy, so focusing on the model's specificity was the best evaluation metric.
+In order to get a better sense of how well the model generalizes, we repeatedly retrained and evaluated the model using diffrent samplings of the training dataset.
+The following figure summarizes the distribution of the specificty the models produced on the test dataset:
+
+```{r}
+#temp <- rbind(data.frame("Model" = c("1"), "Specificity" = model1_robust_results$specificity),
+#              data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))%>%
+temp <- rbind(data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))%>%
+  mutate(Model = as.factor(Model))
+
+ggplot(temp, aes(x = Specificity, fill = Model, color = Model)) +
+  geom_density(alpha = 0.5) +
+  ylab("Density") +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  #geom_vline(xintercept = mean(model1_robust_results$specificity), linetype="dashed", color = "#e41a1c") +
+  geom_vline(xintercept = mean(model2_robust_results$specificity), linetype="dashed", color = "#377eb8")
+```
 
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=11}
+```{r, fig.width=10, fig.height=11}
 temp <- up_train %>%
@@
 ## Correlation Coefficients for Variables of Interest
-```{r, echo=FALSE}
-vars <- c("CARAVAN", "PPERSAUT", "PBRAND", "APERSAUT", "LIKELY_CUSTOMERS")
+```{r}
+vars <- c("CARAVAN", "MOSTYPE", "PPERSAUT", "PBRAND", "MOSHOOFD", "APERSAUT", "LIKELY_CUSTOMERS")
 temp <- up_train %>%
@@
 
-pander(cor(temp[vars], temp[vars]))
+numeric_vars <- unlist(lapply(temp, is.numeric))
+
+pander(cor(temp[numeric_vars], temp[numeric_vars]))
 ```
 
-## Data Dictionary
+## Data Dictionary for Variables of Interest
 
-```{r, results='asis', echo=FALSE}
+```{r, results='asis'}
 data_dictionary <- data.frame(
@@
 
-```{r, echo=FALSE}
+```{r}
 final_model$cm
@@
 
-```{r, null_prefix = TRUE, comment=NA, echo=FALSE}
+```{r}
 conn <- file("CoIL.r", open = "r")

### 2019-12-17 final project/CoIL.R
added Driven growers feature

@@ -271,6 +271,16 @@ robust_results <- function(model_formula, n_tries = 250){
   return(the_results)
 }
 
+## Final Bit of Feature Engineering
+
+df$Driven_Growers <- ifelse(df$MOSHOOFD == "Driven Growers",1,0)
+test$Driven_Growers <- ifelse(test$MOSHOOFD == "Driven Growers",1,0)
+up_train$Driven_Growers <- ifelse(up_train$MOSHOOFD == "Driven Growers",1,0)
+eval$Driven_Growers <- ifelse(eval$MOSHOOFD == "Driven Growers",1,0)
+
+
+
+
 ### Model 1 - Top 5 Important Variables from Random Forest
 model1 <- glm(CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT, 
               family = binomial(link = "logit"),
@@ -281,16 +291,17 @@ model1_results$specificity
 #model1_robust_results <- robust_results("CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT")
 #summary(model1_robust_results$specificity)
 
-### Model 2 - Likely Customers and Car Policies Contribution Level
-model2 <- glm(CARAVAN ~ LIKELY_CUSTOMERS + PPERSAUT,
+### Model 2 - Likely Customers and Car Policies Contribution Level and whether or not they are a driven grower
+model2 <- glm(CARAVAN ~ LIKELY_CUSTOMERS + PPERSAUT+Driven_Growers,
               family = binomial(link = "logit"),
               up_train)
 model2_results <- score_model(model2, test)
 model2_results$specificity
-model2_robust_results <- robust_results("CARAVAN ~ LIKELY_CUSTOMERS + PPERSAUT")
+model2_robust_results <- robust_results("CARAVAN ~ LIKELY_CUSTOMERS + PPERSAUT+Driven_Growers")
 summary(model2_robust_results$specificity)
 
 ## Final Model Accuracy
 final_model <- score_model(model2, eval)
 final_model$correct
 final_model$specificity
+

### 2019-12-17 final project/Manuscript.Rmd
Moved Driven Growers to 3rd model

@@
 ```{r}
-#temp <- rbind(data.frame("Model" = c("1"), "Specificity" = model1_robust_results$specificity),
-#              data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))%>%
-temp <- rbind(data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))%>%
+temp <- rbind(data.frame("Model" = c("1"), "Specificity" = model1_robust_results$specificity),
+              data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))
+temp <- 
+  rbind(temp, data.frame("Model" = c("3"), "Specificity" = model3_robust_results$specificity))%>%
   mutate(Model = as.factor(Model))
 
-ggplot(temp, aes(x = Specificity, fill = Model, color = Model)) +
-  geom_density(alpha = 0.5) +
+ggplot(temp, aes(x = Specificity, fill = Model)) +
+  geom_density(alpha = 0.5, color = NA) +
   ylab("Density") +
@@
   theme(legend.position = "bottom") +
-  #geom_vline(xintercept = mean(model1_robust_results$specificity), linetype="dashed", color = "#e41a1c") +
-  geom_vline(xintercept = mean(model2_robust_results$specificity), linetype="dashed", color = "#377eb8")
+  geom_vline(xintercept = mean(model1_robust_results$specificity), linetype="dashed", color = "#e41a1c") +
+  geom_vline(xintercept = mean(model2_robust_results$specificity), linetype="dashed", color = "#377eb8") +
+  geom_vline(xintercept = mean(model3_robust_results$specificity), linetype="dashed", color = "#4daf4a") 
 ```
@@
 ```{r}
-vars <- c("CARAVAN", "MOSTYPE", "PPERSAUT", "PBRAND", "MOSHOOFD", "APERSAUT", "LIKELY_CUSTOMERS")
+vars <- c("CARAVAN", "MOSTYPE", "PPERSAUT", "PBRAND", "MOSHOOFD", "APERSAUT", "LIKELY_CUSTOMERS", "DRIVEN_GROWERS")
 temp <- up_train %>%
@@
            "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
-           "ABYSTAND", "LIKELY_CUSTOMERS"),
+           "ABYSTAND", "LIKELY_CUSTOMERS", "DRIVEN_GROWERS"),
   Description = c("Number of mobile home policy",
@@
                   "Number of social security insurance policies",
-                  "MOSTYPE = 12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, or 11"))
+                  "MOSTYPE = 12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, or 11",
+                  "MOSHOOFD = 2"))
 temp <- data_dictionary[data_dictionary$Name %in% vars,]

### 2019-12-17 final project/CoIL.R

@@ -124,6 +124,8 @@ prepare_data <- function(df){
   df %>%
     mutate(LIKELY_CUSTOMERS = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
     mutate(LIKELY_CUSTOMERS = as.factor(LIKELY_CUSTOMERS)) %>%
+    mutate(DRIVEN_GROWERS = ifelse(MOSHOOFD == "2", 1, 0)) %>%
+    mutate(DRIVEN_GROWERS = as.factor(DRIVEN_GROWERS)) %>%
     mutate(MOSTYPE = as.factor(MOSTYPE),
            MGEMLEEF = as.factor(MGEMLEEF),
            MOSHOOFD = as.factor(MOSHOOFD),
@@ -271,16 +273,6 @@ robust_results <- function(model_formula, n_tries = 250){
   return(the_results)
 }
 
-## Final Bit of Feature Engineering
-
-df$Driven_Growers <- ifelse(df$MOSHOOFD == "Driven Growers",1,0)
-test$Driven_Growers <- ifelse(test$MOSHOOFD == "Driven Growers",1,0)
-up_train$Driven_Growers <- ifelse(up_train$MOSHOOFD == "Driven Growers",1,0)
-eval$Driven_Growers <- ifelse(eval$MOSHOOFD == "Driven Growers",1,0)
-
-
-
-
 ### Model 1 - Top 5 Important Variables from Random Forest
 model1 <- glm(CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT, 
               family = binomial(link = "logit"),
@@ -288,20 +280,29 @@ model1 <- glm(CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT,
 model1_results <- score_model(model1, test)
 model1_results$specificity
 # TODO: Uncomment these lines
-#model1_robust_results <- robust_results("CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT")
-#summary(model1_robust_results$specificity)
+model1_robust_results <- robust_results("CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT")
+summary(model1_robust_results$specificity)
 
-### Model 2 - Likely Customers and Car Policies Contribution Level and whether or not they are a driven grower
[…]

### 2019-12-18 final project/CoIL.R
Added down sampled

@@ -220,17 +220,18 @@ score_model <- function(model, data, threshold = 0.5, predictions = FALSE){
   return(list("cm" = cm, "correct" = correct, "specificity" = specificity))
 }
 
-robust_results <- function(model_formula, n_tries = 250){
+robust_results <- function(model_formula, correction = "upSample", n_tries = 250){
 ## Trains and evaluates the model multiple times
 #
 # INPUTS
 #
 # model_formula = The formula for the logit model
+# correction (optional) = Correct for imbalanced data (i.e. upSample, downSample, none)
 # n_tries (optional) =  The number of runs (250 default)
 # 
 # RETURNS (data.frame)
 #
-# seed = randome number seed
+# seed = random number seed
 # correct = the number of correct CARAVAN = 1 predictions
 # specificity = the specificity of the CARAVAN = 1 predictions
   
@@ -245,10 +246,20 @@ robust_results <- function(model_formula, n_tries = 250){
         train_index <- createDataPartition(df$CARAVAN, p = .7, list = FALSE)
         train <- df[train_index,]
         test_df <- df[-train_index,]
-        # Correct the data imbalance through over sampling
-        training_df <- upSample(x = select(train, -CARAVAN), 
-                                y = train$CARAVAN, 
-                                yname = "CARAVAN")
+        if(correction == "upSample"){
+          # Correct the data imbalance through over sampling
+          training_df <- upSample(x = select(train, -CARAVAN), 
+                                  y = train$CARAVAN, 
+                                  yname = "CARAVAN")
+        } else if(correction == "downSample"){
+          # Correct the data imbalance through under sampling
+          training_df <- downSample(x = select(train, -CARAVAN), 
+                                    y = train$CARAVAN, 
+                                    yname = "CARAVAN")
+        } else {
+          # No correction
+          training_df <- train
+        }
         # Build the model
         model <- glm(model_formula, 
                      family = binomial(link = "logit"), 
@@ -279,7 +290,6 @@ model1 <- glm(CARAVAN ~ MOSTYPE + PPERSAUT + MOSHOOFD + PBRAND + APERSAUT,
               up_train)
 model1_results <- score_model(model1, test)
 model1_results$specificity
-# TODO: Uncomment these lines
[…]

### 2019-12-18 final project/Manuscript.Rmd
Started adding to discussion and conclusion sections

@@
 -->
-A logistic regression based solution to the CoIL Challenge 2000 is described in this paper.
+
+This paper describes a logistic regression based solution to the CoIL Challenge 2000.
 The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
@@
 -->
+
 The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.
@@
 
+The third model was fit the same as the second model other than the driven growers variable, which is a variable derived from MOSHOOFD variable, specifically it is when the MOSHOOFD variable is equal to 2, which means the specific customer's main type is "Driven Grower" hence the variable name.
+We chose this because when running decision trees we noticed this specific factor of MOSHOOFD variable often stood out in the decision trees.
+
 In evaluating the models we examined we focused on the specificity.
@@
 In order to get a better sense of how well the model generalizes, we repeatedly retrained and evaluated the model using diffrent samplings of the training dataset.
-The following figure summarizes the distribution of the specificty the models produced on the test dataset:
+The following figure summarizes the distribution of the specificty the models produced on the test dataset.
+The dashed line is how our model preformed:
 
@@
   theme(legend.position = "bottom") +
-  geom_vline(xintercept = mean(model1_robust_results$specificity), linetype="dashed", color = "#e41a1c") +
-  geom_vline(xintercept = mean(model2_robust_results$specificity), linetype="dashed", color = "#377eb8") +
-  geom_vline(xintercept = mean(model3_robust_results$specificity), linetype="dashed", color = "#4daf4a") 
+  geom_vline(xintercept = mean(model1_results$specificity), linetype="dashed", color = "#e41a1c") +
+  geom_vline(xintercept = mean(model2_results$specificity), linetype="dashed", color = "#377eb8") +
+  geom_vline(xintercept = mean(model3_results$specificity), linetype="dashed", color = "#4daf4a") 
 ```
@@
 
+Our model outpreforms the original cohort of CoIL Challenge researchers.  We found that one
+
 # Conclusions
@@
 <!-- Conclude your findings, limitations, and suggest areas for future work. -->
+
+This paper set out to define a model that was effective at identifying potential customers for an insurance policy.
+We found that a simple logistic regression model with three variables (two which we derived) outpreformed previous research.
+
+Areas for future work include building an ensemble model using the specification of the third model.  The specificity distribution charts suggests that such an ensemble could preform even better.
 
@@
 
+## Model Summary
+
+```{r}
+summary(model3)
+```
+
+\newpage
+
 ## Confusion Matrix and Statistics for our Model
